Progressive overfilling of readily releasable pool underlies short-term facilitation at recurrent excitatory synapses in layer 2/3 of the rat prefrontal cortex

Version 1

Abstract

Abstract

Short-term facilitation of recurrent excitatory synapses within the cortical network has been proposed to support persistent activity during working memory tasks, yet the underlying mechanisms remain poorly understood. We characterized short-term plasticity at the local excitatory synapses in layer 2/3 of the rat medial prefrontal cortex and studied its presynaptic mechanisms. Low-frequency stimulation induced slowly developing facilitation, whereas high-frequency stimulation initially induced strong depression followed by rapid facilitation. This non-monotonic delayed facilitation after a brief depression resulted from a high vesicular fusion probability and slow activation of Ca2+-dependent vesicle replenishment, which led to the overfilling of release sites beyond their basal occupancy. Pharmacological and gene knockdown (KD) experiments revealed that the facilitation was mediated by phospholipase C/diacylglycerol signaling and synaptotagmin 7 (Syt7). Notably, Syt7 KD abolished facilitation and slowed the refilling rate of vesicles with high fusion probability. Furthermore, Syt7 deficiency in layer 2/3 pyramidal neurons impaired the acquisition of trace fear memory and reduced c-Fos activity. In conclusion, Ca2+- and Syt7-dependent overfilling of release sites mediates synaptic facilitation at L2/3 recurrent excitatory synapses and contributes to temporal associative learning.

Introduction

Synapses in the mammalian central nervous system undergo activity-dependent modulation of synaptic weight during repetitive stimulation, known as short-term synaptic plasticity (STP). STP is mediated by presynaptic mechanisms that regulate the release probability (pr), size of the readily releasable pool (RRP), and kinetics of vesicle replenishment of RRP (Zucker and Regehr, 2002; Jackman and Regehr, 2017; Neher and Brose, 2018). Short-term depression has been observed at synapses with a high baseline pr, primarily because of the rapid depletion of releasable vesicles during high-frequency stimulation (HFS) (Lin et al., 2022). Conversely, vesicle depletion is less pronounced at synapses with a low baseline pr, and cumulative increases in pr during HFS are considered to mediate short-term facilitation (Rozov et al., 2001; Pan and Zucker, 2009; Aldahabi et al., 2024). Central synapses exhibit a wide spectrum of STP depending on different combinations of presynaptic factors mediating short-term depression and facilitation.

Previously, facilitation was modeled as an increase in pr after an action potential (AP) and its slow decay during the inter-spike interval (ISI), resulting in a cumulative increase in pr during HFS (Markram et al., 1998; Dittman et al., 2000). Recent studies have suggested that the number of docking (or release) sites in an active zone is limited and only partially occupied by releasable vesicles in the resting states (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Therefore, pr is determined by the vesicular fusion probability (pv) and release site occupancy (pocc). However, distinguishing the individual contributions of pv and pocc to pr is challenging (Silva et al., 2021; Neher, 2024). It is also unclear whether facilitation is mediated by activity-dependent increases in the pv, pocc, or both.

Two recent models for vesicle priming and release view the ‘priming’ as a two-step reversible process that leads to a release-ready state. The release-ready vesicles could be vesicles in a tightly docked state (TS) or those occupying specialized docking sites (DS), and are supplied from vesicles in a loosely docked state (LS) or those occupying a distinct replacement sites (RS), which are called LS/TS model ((Neher, 2024); (Aldahabi et al., 2024)) and RS/DS model ((Miki et al., 2016); (Silva et al., 2021)), respectively. Common to both models is the assumption that release-ready vesicles constitute only a fraction of all docked vesicles under resting conditions and that this fraction is up- and down-regulated by synaptic activity.

Synaptotagmin-7 (Syt7) was recently found to play a crucial role in facilitation at different types of central synapses (Jackman et al., 2016; Chen et al., 2017; Martinetti et al., 2022). In addition to its role in facilitation, Syt7 accelerates refilling of the RRP after depletion and mediates asynchronous release (Bacaj et al., 2013; Liu et al., 2014; Jackman et al., 2016; Tawfik et al., 2021). Syt7-dependent synaptic facilitation is interpreted as an activity-dependent increase in pv (Turecek et al., 2016; Jackman and Regehr, 2017; Turecek et al., 2017; Norman et al., 2023). However, it is unclear whether Syt7 contributes to facilitation through a Ca2+-dependent increase in pocc (i.e., overfilling). Overfilling has been proposed as a mechanism supporting facilitation at the Drosophila neuromuscular junction (Kobbersmed et al., 2020), yet whether this process also occurs at the mammalian central synapses mediated by Syt7 has not been studied.

Facilitation at recurrent excitatory synapses in cortical recurrent networks may play a crucial role in maintaining working memory (Mongillo et al., 2008; Mongillo et al., 2012; Hansel and Mato, 2013). Although layer 2/3 (L2/3) pyramidal cells (PCs) are highly interconnected and represent task-relevant information during a delay period in working memory tasks (Fujisawa et al., 2008; Ozdemir et al., 2020), most studies on local excitatory recurrent synapses in the neocortex have focused on layer 5 (L5) exhibiting short-term depression (Reyes and Sakmann, 1999; Hempel et al., 2000; Yoon et al., 2020). Further, the synaptic properties of excitatory synapses in L2/3 remain poorly understood.

In the present study, we found that HFS of local excitatory synapses in L2/3 of the medial prefrontal cortex (mPFC) induced initial strong depression followed by delayed facilitation, whereas low-frequency stimulation induced slow monotonous facilitation. This unique form of STP could be explained by 1) high pv of RRP vesicles at rest and throughout a train stimulation; 2) the low resting vesicular occupancy of RRP and its activity-dependent increase. These conditions led to delayed facilitation through progressive overfilling of release sites during HFS with high pv vesicles beyond the resting occupancy. Syt7 knockdown (KD) in L2/3 PCs completely abolished these synaptic enhancements by inhibiting the Ca2+-dependent increase in release site occupancy. Finally, behavioral tests for trace fear conditioning showed that Syt7 KD impaired the acquisition of trace fear memory and reduced c-Fos expression in the mPFC. Collectively, these results support that Ca2+- and Syt7-dependent overfilling of release sites mediates synaptic facilitation at L2/3 recurrent excitatory synapses and contributes to temporal associative learning.

Results

Frequency dependence of STP at local excitatory synapses in L2/3 of the prelimbic cortex

We examined STP profiles at recurrent excitatory synapses between L2/3 PCs and at synapses from L2/3-PCs to fast-spiking interneurons (FSINs) in the prelimbic area of the rat mPFC. The PCs and FSINs were identified based on their distinct intrinsic properties and morphologies (Fig. S1). To stimulate presynaptic PCs, L2/3 PCs were transfected with a plasmid encoding oChIEF using in utero electroporation (IUE; Fig. 1A; Lee et al., 2024). Photostimulation of the oChIEF-expressing cell soma at 40 Hz consistently evoked single APs with high fidelity throughout the 600-pulse trains (Fig. S2A-B), which was consistent with little use-dependent inactivation of oChIEF (Lin et al., 2009; Hass and Glickfeld, 2016).

Frequency-dependence of short-term synaptic plasticity at prelimbic L2/3 excitatory synapses.

(A) In utero electroporation following injection of the plasmid (CAG-oChIEF-tdTomato) into the ventricle of an embryo (E17.5). (B) Representative images showing specific expression of oChIEF-tdTomato in L2/3 pyramidal cells after IUE. The red fluorescence of tdTomato clearly visualizes oChIEF-expressing cell bodies in L2/3 and axons in L5. Scale bar: 1 mm, 100 μM, 50 μM from left to right. (C) Recording schematic showing photostimulation of oChIEF-expressing axon fibers of transfected PCs (filled triangles) and a whole-cell recording from non-transfected PCs (empty triangles) or FSINs (empty circles). A collimated DMD-coupled LED was used to confine the area of excitation (typically 3–4 μm in diameter) to a small region (blue circle) near the soma. (D, E) Representative traces for EPSCs averaged over 10 trials at each frequency (left) and average amplitudes of baseline-normalized EPSCs (right) during 20-pulse trains at frequencies from 5 to 40 Hz at PC-PC (D; n = 12, 9, 21, 10 cells for 5, 10, 20, 40 Hz, respectively) and PC-FSIN (E; n = 8, 9, 10, 10) synapses. Each data point was normalized to the average of the first EPSC. (F) Baseline-normalized amplitudes of steady-state EPSCs (EPSCss; black symbols) and synaptic efficacy (EPSCss × f; red symbols) as a function of stimulation frequency (f) at PC–PC synapses. EPSCss was measured from the average of last 5 EPSCs from the 20-pulse trains (n = 12, 9, 21, 10). (G) Paired pulse ratio (PPR) as a function of inter-spike intervals (n = 25, 9, 25, 22, 43 for 200, 100, 50, 25, 20 ms ISI, respectively) at PC-PC synapses. (H) EPSCss and synaptic efficacy at PC-FSIN (n = 8, 9, 10, 10). (I) PPR at PC-FSIN (n = 8, 9, 10, 15, 34). Gray symbols, individual data.

Given that IUE transfected approximately 10-20% of L2/3 PCs (Fig. 1A), we made a whole-cell patch on a non-transfected L2/3 PC or an FSIN (Fig. 1C). To stimulate a minimal number of excitatory axon fibers, EPSCs were evoked using minimal optical stimulation (Fig. S3A; 470 nm, 3-4 μm in radius and 3-5 ms in duration). EPSCs evoked by minimal stimulation were examined to determine the uniformity of the kinetic properties (Fig. S3B-C). Moreover, we reassessed the kinetics of EPSCs during trains post hoc to ensure uniformity. EPSCs were evoked by 20-pulse trains at 4 different frequencies (5, 10, 20, and 40 Hz) to cover the range of firing rates of PCs observed in vivo during working memory tasks (Baeg et al., 2003). Excitatory synapses from PCs onto both other PCs and FSINs (PC–PC and PC–FSIN) showed strong (∼2-fold enhancement) and lasting (several seconds) short-term facilitation at 20 Hz and lower frequencies (Fig. 1D-E). The PC-PC synapses exhibited stronger facilitation than did the PC-FSIN synapses.

Steady-state EPSCs at the PC-PC synapses were not significantly different from 5 to 20 Hz (Fig. 1F). Frequency invariance at the PC-PC synapses suggests that Ca2+-dependent synaptic facilitation may counteract vesicle depletion in a frequency-dependent manner (Turecek et al., 2017; Lin et al., 2022). Despite facilitation at all frequencies, the paired pulse ratio (PPR) rapidly decreased as the ISI decreased (Fig. 1G, I). At 40 Hz stimulation, both PC-PC and PC-FSIN synapses initially underwent strong depression then followed by facilitation in a few stimuli (Fig. 1Dd, Ed). Such delayed facilitation after paired pulse depression (PPD) has not been previously observed in central synapses except for a slight increase in EPSC at the cerebellar synapses under artificially depolarized conditions (Pulido and Marty, 2018).

We confirmed that optically evoked EPSCs were strictly dependent on AP generation (Fig. S4A), suggesting that EPSCs are evoked by optical stimulation of presynaptic axon fibers rather than by direct depolarization of presynaptic boutons. Additionally, dual patch-clamp techniques showed that the STP of optically evoked EPSCs did not differ from the electrically evoked STP of EPSCs at the PC-FSIN synapses (Fig. S4B). Finally, we confirmed that AMPA receptor (AMPAR) desensitization was not responsible for PPD at 40 Hz, ruling out the possible effects of postsynaptic factors on STP (Fig. S5A), and that AMPAR saturation was not significant during facilitation (Fig. S5B).

Delayed facilitation results from slow activation of Ca2+-dependent vesicle replenishment at a constantly high vesicular fusion probability

The strong PPD at 40 Hz suggests a high vesicular fusion probability (pv) and likely tight coupling of the releasable vesicles to a Ca2+ source. To test this hypothesis, we examined the effects of EGTA, a slow calcium chelator, on synaptic transmission at the PC-PC synapses. The baseline EPSC amplitude (EPSC1) was not affected by incubation of the slice in the bath solution containing 50 μM EGTA-AM for longer than 20 min, supporting tight coupling. Moreover, the PPR was decreased to 0.19, and subsequent facilitation slowed but was still distinct (Fig. 2A-B). These findings indicate that the initial pv should be higher than 0.81 considering vesicle recruitment during ISI. The effects of EGTA-AM were confirmed in the same cell by measuring the baseline EPSC and PPR before and after applying EGTA-AM to the bath. EPSC1 remained unchanged, while PPR was reduced by half (Fig. S6).

Delayed facilitation results from slow activation of Ca2+-dependent vesicle replenishment at a constantly high vesicular fusion probability

(A) Representative traces (left) and mean baseline-normalized amplitudes (right) of EPSCs evoked by 30-pulse trains at 40 Hz in control (n = 21; black) and in the presence of 50 μM EGTA-AM (n = 14; green). (B) Mean values for the first EPSC amplitude (EPSC1, left) and PPR (right) from the experiments displayed in (A). Gray symbols, individual data. (C) Plot of rate constants for short-term facilitation (kSTF) as a function of stimulation frequency (fstim), showing a linear relationship. The linear regression line (black) is shown fitted to kSTF values, estimated from Figure 1. (D) Representative EPSC traces (left) and average of baseline-normalized EPSCs (right) evoked by 12-pulse stimulation at 5 Hz, followed by 40 Hz 7-pulse train (n = 12). Note that slowly developing facilitation was converted to rapid facilitation after strong PPD. (E) Mean values for PPR at 40 Hz. The baseline PPR was reproduced from Figure 1G and the PPR during 5 Hz train was calculated as (13th EPSC) / (12th EPSC). Gray symbols, individual data. All statistical data are represented as mean ± S.E.M.; n.s. = not significant; **, P<0.01; unpaired t-test.

Notably, the application of EGTA-AM markedly slowed facilitation at 40 Hz (Fig. 2A,C). The frequency invariance of steady-state EPSCs (Fig. 1F) and the slower facilitation at 40 Hz in the presence of EGTA (Fig. 2A) imply a Ca2+-dependent increase in the refilling rate of the RRP and/or an increase in the pv of reluctant vesicles in light of previous studies (Hosoi et al., 2007; Liu et al., 2014; Ritzau-Jost et al., 2014; Turecek et al., 2016; Ritzau-Jost et al., 2018; Kusick et al., 2020). Moreover, facilitation was accelerated at higher stimulation frequencies. The plot of the rate constant for the increase in EPSCs (denoted as kSTF, the rate constant for the development of short-term facilitation) as a function of the stimulation frequency (fstim) revealed a linear relationship (Fig. 2C). kSTF in the presence of EGTA-AM was distinctly below this relationship, supporting the Ca2+-dependent acceleration of facilitation.

kSTF was also accelerated when the stimulation frequency was increased from 5 Hz to 40 Hz (Fig. 2D; 1.27 ± 0.18/s to 24.27 ± 3.44/s). Notably, the second 40-Hz train stimulation led to strong PPD, followed by accelerated facilitation (Fig. 2D,E). The strong PPD suggests that the vesicular fusion probability kept high during the 5-Hz stimulation, arguing against the possibility for an increase in pv. Both two-step sequential priming models, LS/TS and RS/DS models, are suitable to interprete our data and we will refer to the different populations of vesicles as TS and LS vesicles, respectively (Taschenberger et al., 2016; Neher and Brose, 2018; Lin et al., 2022; Neher, 2024). Within this framework, vesicles are are released from TS with high pv both in the resting state and during facilitation, and the delayed facilitation is seen as a result of Ca2+-dependent progressive increase in the occupancy of TS vesicles on docking sites.

Release sites have low baseline occupancy and these are increased during facilitation and post-tetanic augmentation

Given that pv was greater than 0.8, the two-fold increase in EPSC during 20 pulse train stimulations could not be explained by an increase in pv. Recent studies have shown that vesicle release sites (N) are limited in number and are partially occupied at rest (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Under such a high pv, two-fold synaptic facilitation should be attributed to an increase in pocc. The vesicle replenishment rate is accelerated during HFS in a Ca2+-dependent manner (Fig.2A-C; Hosoi et al., 2007). This may lead to the overfilling of release sites beyond their basal vesicular occupancy, which in turn may increase EPSC under high pv. To test this hypothesis, we determined the quantal parameters at the L2/3 local excitatory synapses. We first measured the quantal size from asynchronous release events after replacing extracellular Ca2+ with Sr2+, as previously described (Yoon et al., 2020) (Fig. S7A-B). The mean and coefficient of variance of the quantal sizes were 24 pA and 0.25 at PC-PC synapses, respectively, and were 30 pA and 0.31 at PC-FSIN synapses, (Fig. S7C).

Next, we performed a variance-mean (V-M) analysis of phasic EPSCs at excitatory synapses during train stimulation at 5 Hz and 40 Hz (Scheuss et al., 2002). Applying multiple-probability fluctuation analysis (MPFA) using intra-site quantal variability of 0.2 (Valera et al., 2012), the V-M plot was best fitted by the number of release sites (N) of 5.3 and the quantal size (q) of 25 pA at the PC-PC synapses (Fig. 3A). The pr for EPSC1 was calculated as 0.32. We performed the same analysis for the PC-FSIN synapses and found N = 5.3, q = 35 pA, and pr for EPSC1 = 0.31 (Fig. 3B). Our estimate for N was smaller but comparable to the estimate (6.9) at excitatory synapses in L2/3 of the mouse barrel cortex (Holler et al., 2021).

Low baseline occupancy of release sites and its increase during facilitation and post-tetanic augmentation

(A, B) Left, Representative EPSCs evoked by 5 and 40 Hz train stimulation (black, 5 Hz; purple, 40 Hz). Right, Variance-mean plots of EPSCs amplitude from averaged EPSCs recorded at PC-PC (A; n = 12, 17 for 5 and 40 Hz, respectively) and PC-FSIN (B; n = 8, 10) synapses. The data were fitted using multiple-probability fluctuation analysis (MPFA). Error bars are omitted for clarity. The 1st EPSC of 5 Hz train (broken line) was used to estimate the resting level of pocc. Filled circles were measured from post-tetanic augmented EPSCs (A, n = 12, 12, 9; B, n = 7, 7, 7 for 0.1, 0.2, 0.5 s IBIs, respectively). (C-E) Post-tetanic augmentation (PTA) experiments at PC-PC (top) and PC-FSIN (bottom) synapses. (C) Representative traces for EPSCs evoked by double 40 Hz train stimulations separated by 0.5 s. (D) Mean baseline-normalized amplitudes of EPSCs evoked by double 40 Hz trains at different inter-burst intervals (IBIs, 0.1, 0.2, 0.5, 1, 2, 5, 10 s). Upper, PC-PC synapse (n = 12, 12, 21, 11, 11, 16, 11 from short to long IBIs, respectively). Lower, PC-FSIN synapse (n = 10, 9, 8, 9, 9, 7, 9). (E) PTA time course. The baseline-normalized amplitudes of 1st EPSC from the 2nd train were plotted as a function of IBIs.

The pr estimate was clearly smaller than the possible minimum value for pv (0.81) estimated from the PPR data (Fig. 2B). Given that pr is a product of pocc and pv (Neher, 2024), a pr smaller than pv indicates the partial occupancy of release sites in the resting state. If synaptic facilitation were mediated by the recruitment of new release sites, it would be accompanied by an increase in the variance of augmented EPSCs (Valera et al., 2012; Kobbersmed et al., 2020). However, the augmented EPSCs during facilitation followed the same parabola in the V-M plot (Fig. 3A,B). Therefore, facilitation appears to be mediated by the progressive overfilling of a finite number of release sites beyond the basal occupancy. In the next section, we show that pv is close to unity based on a synaptic failure analysis (Fig. S8). Assuming this, the baseline pr at the first pulse estimated from the fitted parabolic curve can be largely regarded as a baseline pocc (c.a. 30%), and the two-fold increase in EPSCs during facilitation can be attributed to an increase of up to 60% in pocc.

We investigated the time course of recovery from facilitation and post-tetanic augmentation (PTA) by applying dual-train stimulations (30 pulses at 40 Hz) separated by different inter-burst intervals (IBIs; Fig. 3C,D). The initial EPSCs in the second train were enhanced by more than twofold, and followed by a strong PPD, suggesting that fusion probability stays high in the IBIs and thus synaptic transmission is carried by TS vesicles. The decay time course of PTA was characterized by two phases (initially fast and later slow; Fig. 3E), which may reflect a two-phase Ca2+ decay or two distinct recovery processes that occur sequentially. The variance of potentiated EPSCs measured at IBIs of 0.1, 0.2, and 0.5 s was not different from that of EPSCs at the peak of facilitation (filled circles, Fig. 3A-B). Given that the same RRP mediates both facilitation and augmentation and that post-tetanic EPSCs is carried by TS vesicles similar to the baseline EPSCs, this plot suggests that PTA, similar to short-term facilitation, is mediated by an increase in the TS vesicle occupancy.

Failure rate analysis suggests that vesicular fusion probability is closed to one

The vesicle dynamics of STP has been explained using a simple refilling model (Hosoi et al., 2007; Neher and Sakaba, 2008). This model assumes that the vesicles are reversibly docked to a limited number of release sites (N) with forward and reverse rate constants (denoted as k1 and b1, respectively). Under the framework of the simple refilling model (see Materials and Methods), the rate constant for the refilling of the RRP after depletion by the first pulse is the sum of the baseline k1 (k1) and b1. This is estimated as 23/s from the plot of PPR as a function of ISIs (Fig. S8A). From the baseline occupancy of 0.3 and pocc = k1 / (k1 + b1), we estimated k1 to be 6.9/s.

When paired APs were applied with an ISI of 20 ms, the failure rate at 1st pulse, P(F1), was 10.6%, and the probability of two consecutive failures, P(F1, F2), was 6.2% (Fig. S8A). From the equation P(F1) = (1 – pr)N, pr was calculated as 0.312 when N = 6, comparable to pr (= 0.32) estimated from the V-M analysis (Fig. 3A). We calculated P(F1, F2) under the framework of a simple refilling model, with pv and k1 set as free parameters and other parameters set according to the following relationships: P(F1) = (1 – pr)N, pr = pv ·pocc, and pocc = k1 / (k1 + b1).

The calculated values for P(F1, F2) and their difference from the observed value (6.2%) are shown in the plane of k1 vs. pv in Figure S8B-C. As previously mentioned, k1 was greater than 6/s. Under this high k1 condition, the difference between the calculated and observed values of P(F1,F2) was minimal when pv = 1 (Fig. S8C).

Pharmacological manipulations reveal specific molecular identities underlying vesicle loading processes

Our results suggest that the baseline occupancy of TS vesicles is low and that the activity-dependent progressive overfilling of docking sites with TS vesicles is responsible for short-term facilitation and augmentation. To elucidate the specific molecular link between synaptic enhancement and Ca2+-dependent overfilling, we examined the effects of several drugs known to affect vesicle dynamics. We first examined the roles of phospholipase C (PLC) and diacylglycerol (DAG) by applying 5 μM U73122, a PLC inhibitor, or 20 μM 1-oleoyl-2-acetyl-sn-glycerol (OAG), a DAG analog. The STP at 40 Hz and PTA at 0.5-s interval were examined before and after applying each drug to the bath. For both drugs, the facilitation and augmentation were reduced (Fig. 4A-B). U73122 did not affect the basal EPSC amplitude (EPSC1) but resulted in a stronger PPD, slower facilitation, and lower augmentation (Fig. 4A). These effects of U73122 suggest the involvement of Ca2+-induced PLC activation in progressive overfilling. In contrast, OAG increased EPSC1 while maintaining a pronounced PPD (Fig. 4B), indicating an increase in the TS vesicle pool size. Meanwhile, OAG slowed kSTF and reduced augmentation, indicating occlusion of activity-dependent synaptic facilitation. Phorbol esters and DAG are thought to accelerate priming through Munc13 activation (Rhee et al., 2002; Taschenberger et al., 2016; Aldahabi et al., 2022); they also shorten the bridges that interlink docked vesicles with the active zone (AZ) membrane (Papantoniou et al., 2023). Collectively, these suggest that PLC- and DAG-mediated signaling play a role in the overfilling of docking sites underlying short-term facilitation and augmentation.

Pharmacological experiments reveal specific molecular mechanisms underlying vesicle loading processes

(A-C) (a) Representative EPSC traces (left) and mean baseline-normalized EPSCs (right) evoked by double 40 Hz train stimulations separated by 0.5 s inter-burst interval (IBI) in control and in the presence of 5 μM U73122 (Aa, n = 7, orange), 20 μM OAG (Ba, n = 12, cyan) or 100 μM dynasore (Ca, n = 10, blue). (b) Mean values for baseline EPSCs (EPSC1, left) and augmentation (right) from the experiments shown in corresponding a panel. (Da) Representative EPSC traces evoked by paired pulses (left) and mean values for baseline EPSC amplitude (middle) and PPR (right) before and after applying 20 μM LatB (n = 6). (Db) Representative EPSC traces (left, upper) and average of normalized EPSCs (left, lower) evoked by double 40 Hz train stimulation separated by 0.5 s in control (n = 21) and in 20 μM LatB conditions (n = 16; pink). Right, Mean values for augmentation in control and LatB conditions. Gray symbols, individual data. All statistical data are represented as mean ± S.E.M., *, P<0.05; **, P<0.01; ***, P< 0.001; unpaired or paired t-test; n.s. = not significant.

Although the acceleration of Ca2+-dependent vesicle refilling supports rapid facilitation at 40 Hz, the later part of the delayed facilitation underwent a slight depression, which was more pronounced during the second burst (Fig. 3D, Fig. 4). Sustained synaptic transmission is limited by the spatial availability of docking sites because it is undermined by accumulation of exocytic remainings in the presynaptic AZ during HFS (Neher, 2010; Haucke et al., 2011). To verify this, we examined the effects of dynasore (100 μM), a specific inhibitor of endocytosis (Macia et al., 2006). The treatment with dynasore did not influence EPSC1 or the early phase of facilitation, but it exacerbated depression during the later phase and attenuated PTA (Fig. 4C). This suggests that late synaptic transmission during the 40-Hz train is limited by site clearance.

Given the involvement of actin polymerization in the multiple steps of vesicle docking and recycling, we examined the effects of 20 μM latrunculin B (LatB), an inhibitor of actin polymerization. The effects of LatB were complicated: (1) EPSC1 was decreased, whereas PPR was increased (Fig. 4Da) and (2) facilitation and augmentation were reduced (Fig. 4Db). The latter effect may result from defects in vesicle replenishment rates, either through the impaired physical movement of vesicles or disrupted site clearance (Sakaba and Neher, 2003; Lee et al., 2012; Hallermann and Silver, 2013; Miki et al., 2016), as Ca2+-dependent vesicle refilling mediates facilitation and augmentation (Figs. 2–3). Meanwhile, the former effect might be attributed to a shift in equilibrium in the priming states towards a loosely docked state (LS) at rest. It is unknown whether actin affects vesicular fusogenicity in neurons, although it facilitates multiple steps of vesicle fusion by enhancing membrane tension in neuroendocrine cells (Wu and Chan, 2022).

Short-term facilitation at both types of local excitatory synapses in L2/3 is abolished by Syt7 knockdown

We examined whether Syt7 mediates facilitation at the PC-PC and PC-FSIN synapses in L2/3 of the mPFC. To this end, oChIEF and shRNA against Syt7 (shSyt7) were co-expressed in L2/3 PCs using IUE. Syt7 KD resulted in a complete loss of facilitation at all tested stimulation frequencies (Fig. 5A-B). Short-term depression and paired-pulse depression were more pronounced as the stimulation frequency increased, and the frequency invariance of steady-state EPSC was abolished, implying little contribution of Ca2+-dependent vesicle recruitment (Fig. 5C-F). The lack of facilitation could not be attributed to the failure of APs in presynaptic axons, as optical stimulation elicited reliable APs in Syt7 KD PCs during the 40-Hz train (Fig. S2C). Syt7 KD had little effect on the basal properties, including the peak amplitude, rise/decay time, and time-to-peak, of single AP-induced EPSCs (Fig. S3D-E). Moreover, Syt7 KD did not significantly affect the intrinsic properties of neurons (Fig. S9).

STF at both types of local excitatory synapses is abolished by Syt7 KD

(A, B) Representative EPSC traces (left) and mean baseline-normalized amplitudes of EPSCs (right) evoked by 20-pulse trains at 5 to 40 Hz. STP was measured at PC-PC (A; n = 10, 9, 12, 9) and PC-FSIN (B; n = 9, 8, 10, 9) synapses, in which presynaptic Syt7 transcripts were depleted (Syt7 KD). Syt7 KD pyramidal cells are indicated as black triangles on the top. For comparison, STP in WT synapses is reproduced from Figure 1 (dotted lines). Same frequency color codes were used as in Figure 1. (C) Baseline-normalized amplitudes of steady-state EPSC (EPSCss; black symbols) and synaptic efficacy (EPSCss × f; red symbols) as a function of stimulation frequency (f) at PC-PC synapses. EPSCss was defined as the average of last 5 EPSC amplitudes from 20-pulse trains. (D) PPR as a function of inter-spike intervals (n = 10, 9, 12, 18, 26). (E) EPSCss and synaptic efficacy at PC-FSIN synapses. (F) PPR at PC-FSIN synapses (n = 9, 8, 10, 12, 24). Gray symbols, individual data.

To examine whether Syt7 KD had any nonautonomous effects on postsynaptic cells, we measured the quantal size (q) from asynchronous events in the presence of Sr2+ and found that the mean quantal size remained unaltered at Syt7 KD synapses (Fig. S7C). This indicated that Syt7 did not influence postsynaptic parameters, including the density and kinetic properties of AMPARs. To test the specificity of shSyt7, we examined whether co-expression of the shRNA-resistant form of Syt7 restored synaptic facilitation in both excitatory synapse types. Presynaptic overexpression of shRNA-resistant Syt7 under the CAG promoter rescued the effect of shRNA at both PC-PC and PC-FSIN synapses (Fig. S10). These results underscore the crucial role of presynaptic Syt7 in mediating short-term facilitation at local excitatory synapses in L2/3 of the mPFC.

Syt7 KD synapses exhibit complementary changes in the number of release sites and their vesicle occupancy

To estimate the quantal parameters in Syt7 KD synapses, we performed V-M analysis of EPSCs evoked by train stimulation (Fig. 6A-B). Fitting a parabola to the V-M plot revealed higher baseline pocc in Syt7 KD synapses than in wild-type (WT) synapses for both synapse types (KD, 0.68 and 0.42; WT, 0.32 and 0.31 for PC-PC and PC-FSIN synapses, respectively). The Pocc monotonously decreased during a train stimulation, as expected from short-term depression. Moreover, the number of release sites (N) was markedly lower in Syt7 KD synapses (N = 3.5) than in WT synapses (N = 5.3; Fig. 3A). Next, we examined the PTA induced by the same protocol as in Figure 3 (two 40-Hz train stimulations separated by variable IBIs) at Syt7 KD synapses (Fig. 6C). The 40 Hz train-induced synaptic depression quickly recovered during IBIs, but no PTA was observed at Syt7 KD synapses, indicating its crucial role in augmentation (Fig. 6D-E).

Syt7 KD synapses exhibit complementary changes in the number of release sites and their vesicle occupancy

(A, B) Left, Representative traces of EPSCs evoked by 5 and 40 Hz train stimulations (black, 5 Hz; purple, 40 Hz). Right, Variance-mean plots of EPSCs amplitude from averaged EPSCs recorded at PC-PC (A; n = 9, 15, 12, 10, 9) and PC-FSIN (B; n = 4, 14, 10, 9, 8) synapses in which presynaptic Syt7 has been knocked down. The data were fitted using MPFA and error bars are omitted for clarity. The mean 1st EPSC amplitude of 5 Hz train (dotted line) was used for estimation of baseline pocc. Broken lines indicate MPFA fits to variance-mean plot of WT synapses reproduced from Figure 2. (C-E) Recovery experiments at PC-PC (top) and PC-FSIN (bottom) synapses in which presynaptic Syt7 was knocked-down. (C) Representative EPSCs evoked by double 40 Hz train stimulations separated by 0.5 s. (D) Mean baseline-normalized amplitudes of EPSCs evoked by double 40 Hz trains at PC-PC (n = 12, 10, 9, 9, 10, 9, 8) and PC-FSIN (n = 10, 9, 10, 11, 8, 9, 10) synapses at different interburst intervals (IBIs). (E) PTA time course. Baseline-normalized amplitudes of 1st EPSC from the 2nd burst were plotted as a function of various IBIs. Dotted lines indicate augmented EPSCs in the WT reproduced from Figure 2.

TS vesicle recovery following depletion is accelerated by Syt7

Syt7 is recognized not only for its crucial role in facilitation, but also for mediating Ca2+- dependent replenishment of releasable vesicles (Liu et al., 2014; Tawfik et al., 2021). Given that facilitation at L2/3 excitatory synapses is driven by progressive overfilling (Figs. 2-3), the loss of facilitation at Syt7 KD synapses in the present study implies that Syt7 plays a key role in the Ca2+-dependent acceleration of refilling and overfilling of the docking sites with TS vesicles. To test this hypothesis, we examined the recovery kinetics of EPSCs after TS vesicles were depleted by 3 pulses at 40 Hz in wild-type (WT) and Syt7 KD synapses (Fig. 7). Strong PPD was observed not only in the first burst, but also in the second burst. This suggested that EPSCs in the second burst were mediated by TS vesicles, similar to those in the first burst (Fig. 7A,D). Therefore, EPSC1 in the second burst can be interpreted as a release from TS vesicles recovering during a given IBI. The recovery time course of EPSC1 in the second burst indicated full recovery of TS vesicles within 100 ms in both excitatory synapse types in WT rats (Fig. 7C,F). However, this recovery was remarkably slower in Syt7 KD synapses, requiring 5 s for full recovery, indicating that TS vesicle recovery was greatly accelerated by Syt7. Overall, synaptic facilitation and augmentation at local excitatory synapses in L2/3 can be ascribed to progressive activity-dependent increases in the occupancy of TS vesicles (Figs. 2-3), in which Syt7 plays a key role.

Recovery of TS vesicles following depletion is accelerated by Syt7

(A-C). Recovery experiments at PC-PC synapses in WT (A) and Syt7 KD (B). Mean baseline-normalized amplitudes of EPSCs evoked by two consecutive 3-pulse 40 Hz trains in WT (n = 16, 13, 11, 11, 11, 9 from short to long IBIs, respectively) and KD (n = 13, 11, 12, 8, 10, 11) synapses at different IBIs (0.1, 0.2, 0.5, 1, 2, 5 s). Inset, representative traces of EPSCs evoked by two consecutive 3-pulse 40 Hz train stimulations separated by 0.5 s. (D-F). Recovery experiments at PC-FSIN synapses in WT (D) and Syt7 KD (E). Mean baseline-normalized amplitudes of EPSCs evoked by two consecutive 3-pulse 40 Hz trains in WT (n = 9, 11, 11, 9, 10, 13) and KD (n = 9, 12, 9, 7, 10, 9) synapses at different IBIs. Inset, representative traces of EPSCs evoked by two consecutive 3-pulse 40 Hz train stimulations separated by 0.5 s. (C, F) Recovery time course. Baseline-normalized amplitudes of 1st EPSC from the 2nd burst were plotted as a function of various IBIs.

Behavioral consequences of Syt7 KD in L2/3 PCs of the mPFC

We previously proposed that the disparity in STP between the PC-PC and PC-IN synapses results in activity-dependent changes in the ratio of synaptic weights at the PC-PC and PC-IN synapses (Jee/Jie). This may in turn have a profound influence on persistent activity in a recurrent network by biasing the E-I balance (Yoon et al., 2020). Syt7 KD not only eliminated facilitation, but also the activity-dependent increase in the Jee/Jie ratio in the L2/3 network (Fig. S11). This suggests the possibility that Syt7 contributes to the persistent activity during working memory tasks in the L2/3 recurrent network. Trace fear conditioning (tFC) requires associative learning between an auditory cue (conditioned stimulus, CS) and temporally separate aversive events (unconditioned stimulus, US). The formation of trace fear memory requires the prelimbic area of the mPFC and hippocampus. mPFC activity during the trace interval is thought to temporarily hold CS information, enabling the network to associate temporally discontinuous events (i.e., temporal associative learning) (Gilmartin et al., 2013). Synaptic facilitation and augmentation are implicated in temporary memory retention in recurrent networks (Mongillo et al., 2008; Mongillo et al., 2012).

Given that Syt7 is crucial for synaptic facilitation (Fig. 5) and augmentation (Fig. 6) in prelimbic L2/3 recurrent excitatory synapses, we tested whether trace fear memory formation was impaired in rats with depleted Syt7 transcripts, specifically in the L2/3 PCs of the mPFC. Syt7-targeted shRNA (shSyt7) or scrambled shRNA (Scr) constructs were transfected bilaterally into the L2/3 PCs of the mPFC using IUE (Fig. 8A). Both the Scr-transfected control and shSyt7-transfected rats underwent tFC training (Fig. 8B). The following day, the rats were subjected to a tone test in which trace fear memory formation was assessed from the freezing response to a tone alone in a distinct context. Compared to the control group, the shSyt7-transfected group exhibited significantly lower levels of freezing behavior in response to auditory cues during the first four trials, suggesting an impairment in trace memory formation (Fig. 8C).

Behavioral effects of Syt7 deficiency in L2/3 PCs of the mPFC

(A) Top, Representative images showing bilateral expression of U6-GFP in L2/3 of PCs after IUE at E17.5. Scale bar: 1 mm, 100 μm. Bottom, Schematic of trace fear conditioning and extinction (tone test) protocol. (B) Freezing behavior of control (expressing scrambled shRNA, Scr) and Syt7 KD rats during acquisition of tFC. (C) Freezing ratio during tone tests on following days. Data are shown as average freezing during T1–T4, T5–T8, or T9–T12 (T, trials). The freezing on T1-T4 was significantly lower in KD rats suggesting that formation of trace memory was impaired in Syt7 KD rats (n = 10, 11 for Ctrl and KD, respectively; P = 0.0007, F(1, 19) = 16.43, two-way repeated measures ANOVA; P = 0.0064, 0.0211, 0.0064, 0.0064; Holm-Sidak test). (D) Representative images of c-Fos immunoreactivity in the prelimbic cortex of control or Syt7 KD rats 90 min after tFC acquisition. c-Fos (left, red, Cy5) and GAD67 (right, cyan, Cy3) were immunostained in the same brain slice expressing U6-GFP (middle; green). Scale bar, 50 μm. (E) Exemplar images of c-Fos positive neurons expressing GFP (top) or GAD67 (bottom) in control (left) or Syt7 KD rats (right). Scale bar, 50 μm. (F-H) Effects of Syt7 KD on c-Fos density (F) and percentage of c-Fos positive neurons co-labeled with GAD67 (G) or GFP (H) in L2/3 or L5 of prelimbic cortex (n = 9, 9 for Ctrl and KD, respectively). Open symbols, individual data. All statistical data are represented as mean ± S.E.M.; ****, P<0.0001; unpaired t-test.

The tone test was repeated the following day to test for extinction memory formation. Syt7 KD animals exhibited much less freezing to the CS (+1d) than did WT animals. Considering the Rescorla-Wagner model (Yau and McNally, 2023), our results suggest that the strength of CS-US associative memory was weaker in KD animals than in WT animals; thus, the acquisition of trace fear memory was impaired by Syt7 KD in the L2/3 PCs of the mPFC.

Next, we assessed the locomotor activity and anxiety levels of rats using the open field test (OFT) and elevated plus maze test (EPM). Scr controls and Syt7 KD rats showed similar exploratory patterns, displaying comparable total distance moved and time spent in the periphery of the OFT or the open/closed arms of the EPM (Fig. S12). Previous studies have suggested that the acquisition of trace fear memory requires elevated trace activity in prelimbic PCs (Gilmartin et al., 2013). Given that c-Fos, an immediate early gene, is upregulated following patterned activities in neurons, its expression is considered a reliable indicator of recent neuronal activation (Fields et al., 1997; Guzowski et al., 2005). To investigate the effect of Syt7 on neuronal activity in vivo during tFC, animals were sacrificed at the time of peak c-Fos protein expression, approximately 90 min after completion of the behavioral task (Arime and Akiyama, 2017).

To elucidate the activation patterns of specific neuronal populations, we performed immunohistochemical co-labeling of c-Fos and GAD67 (a GABAergic interneuron marker) upon tFC acquisition. The density of c-Fos(+) cells and the percentage of GAD67(+) cells among c-Fos(+) cells in all prelimbic layers were significantly lower in Syt7 KD rats than in control rats (Fig. 8F-G). Additionally, analysis of the proportion of active cells among transfected L2/3 PCs (c-Fos+/GFP+) revealed that Syt7 KD L2/3 PCs were significantly more active than Scr-transfected PCs (Fig. 8H). This unexpected hyperactivity of Syt7 KD PCs may be attributed to the high reciprocal connectivity between PCs and FSINs in the neocortex (Holmgren et al., 2003; Otsuka and Kawaguchi, 2009), as PCs with short-term depression would transmit less input to FSINs, which in turn would exert less feedback inhibition on reciprocally connected PCs. Collectively, these findings show that the slow activation of vesicle refilling supports various forms of facilitation at excitatory synapses in PFC L2/3, which is essential for neuronal activity during temporal associative learning.

Discussion

The present study characterized STP at local excitatory synapses in the L2/3 network of the rat mPFC at physiological extracellular [Ca2+]. These synapses displayed initial strong depression and then delayed facilitation at 40 Hz, while monotonous slowly developing facilitation was observed at lower frequencies. Our study suggests that such delayed facilitation after a brief depression results from a high pv (Figs. 2-3 and Fig. S8), along with Ca2+-dependent delayed acceleration of vesicle refilling and overfilling (Figs. 2–3). The two-fold increase in EPSC during 20 pulse train stimulations under the condition of such a high pv suggests a Ca2+-dependent increase in the vesicle replenishment rate during a train; this then leads to overfilling of release sites beyond their incomplete basal occupancy (Fig. 3A,B). The high pv and strong PPD did not result from artificial bouton stimulation, AMPAR desensitization, or any potential distortion caused by optogenetic methods (Figs. S4 and S5). Additionally, similar to short-term facilitation, PTA could be explained by an increase in the TS vesicle occupancy (Fig. 3C-E), consistent with previous studies showing post-tetanic increases in the pocc and/or vesicle pool size at various types of synapses (Lee et al., 2010; Vandael et al., 2020; Tran et al., 2023; Silva et al., 2024). These findings suggest that release sites are partially occupied at rest and that a progressive overfilling of release sites with TS vesicles underlies facilitation and augmentation at intracortical L2/3 excitatory synapses.

Progressive overfilling of docking sites

Conventional models for facilitation assume an increase in pv after an AP of RRP vesicles, which had been in a low_pv or reluctant state at rest (Dittman et al., 2000; Turecek et al., 2016). This facilitation model is unlikely to explain our results, because TS vesicles are predominantly released not only at rest but also in the facilitated state during or after train stimulation, as evidenced by the strong PPD (ISI, 25 ms) during a 5-Hz train (Fig. 2D-E) and at different intervals after a 40-Hz/30 pulse trains (Fig. 3D). Moreover, the PPD (ISI, 25 ms) remained pronounced during the early phase of facilitation (Fig. 7), arguing against the possibility of slowly increasing the pv of reluctant vesicles. These results suggested that facilitation was mediated by an activity-dependent progressive increase in occupancy of TS vesicles (TS occupancy). Facilitation through an increase in TS occupancy has been termed frequency facilitation, differentiating it from paired pulse facilitation that relies on a residual calcium-dependent transient increase in the pv of releasable vesicles (Neher, 2024).

Assuming an increase in TS occupancy, to attain a two-fold increase in EPSC (Fig. 1), the baseline occupancy of the docking sites by TS vesicles should be low, as estimated from the V-M plot (c.a. 30% in Fig. 3). However, it is unclear whether the remaining docking sites that are not occupied by TS vesicles are truly vacant or are occupied by reluctant LS vesicles that are not released by the AP or AP trains. TS vesicles can be supplied by the transformation of LS to TS vesicles, both of which reside at the same docking site (LS/TS model; reviewed in Neher, 2024). Alternatively, TS vesicles can be supplied from a distinct replacement pool (RS/DS model; reviewed in Silva et al., 2021). Further studies, including ultrastructural analyses of active zones, are required to address this issue.

Given the crucial role of Syt7 in facilitation (Fig. 5), the facilitation mechanisms are closely related to how Syt7 mediates synaptic facilitation. Based on the slow kinetics of Syt7, previous computational modeling studies have proposed that Syt7 may mediate short-term facilitation through an activity-dependent increase in pv (Turecek et al., 2016; Jackman and Regehr, 2017; Turecek et al., 2017; Norman et al., 2023). The present study, however, found that both facilitation and augmentation were mediated by overfilling of the release sites (Figs. 2–3). Moreover, Syt7-KD synapses displayed slower recovery of TS vesicles (Fig. 7). These results suggest that Syt7 is essential for activity-dependent overfilling of docking sites with TS vesicles. Whereas the baseline EPSC was not altered (Fig. S3), and complementary changes in the number of docking sites and their baseline occupancy were observed in Syt7 KD synapses (Fig. 6). These results raise a possibility that Syt7 may play a role in providing additional vacant docking sites that can be overfilled during facilitation. It remains to be elucidated whether the decrease in the number of docking sites at Syt7 KD synapses is related to the subcellular localization of Syt7 to the plasma membrane of the active zone (Sugita et al., 2001; Vevea et al., 2021).

STP model in light of known Ca2+ binding kinetics of Syt7

Syt7 acts as an upstream Ca2+ sensor that facilitates activity-dependent replenishment (Liu et al., 2014; Bacaj et al., 2015; Chen et al., 2017; Tawfik et al., 2021). The present study shows that EPSCs are mediated by TS vesicles both at rest and during activity-dependent facilitation. Releasable vesicles are homogeneous, and Syt7-dependent overfilling is responsible for the activity-dependent enhancement of EPSCs. Thus, we tested whether slowly developing facilitation could be replicated within the framework of a simple refilling model based on the known Ca2+ binding kinetics of Syt7 (Brandt et al., 2012). To this end, we made following assumptions: (1) a Ca2+-dependent increase in the forward refilling rate (k1) requires a full Ca2+-bound form of Syt7, to which multiple Ca2+ ions can bind (two or three Ca2+ ions on each C2A and/or C2B domain) (Brandt et al., 2012; Chon et al., 2015; Voleti et al., 2017); (2) sequential Ca2+ binding steps are synergistic through a cooperativity factor, b; (3) Syt7 senses AP-induced local [Ca2+] similar to that estimated at the calyx of Held (a Gaussian function with a full width at half maximum value of 0.2 ms and a peak of 40 μM) (Wang et al., 2008); and (4) residual [Ca2+] follows a mono-exponential function with an amplitude of 1 μM and decay time constant of 50 ms (Jackson and Redman, 2003).

Under these assumptions and the framework of the simple refilling model, slow and delayed facilitation could be replicated by modeling k1 as k1 = k1,b + K1,max × [Syt7: n Ca2+], where [Syt7: n Ca2+] represents the fraction of fully Ca2+-bound Syt7 (Fig. 9A-C). Syt7 was slowly activated over the course of the stimulus train, suggesting that the cooperative binding of multiple Ca2+ ions progressively increased the putative active form of Syt7 because of its slow membrane binding/unbinding kinetics. Nevertheless, the late phase of facilitation at high frequencies was predicted to be higher in the simulation than in the experiment. As expected, based on the notion that sustained release during HFS is limited by site clearance in the active zone (Fig. 4C), the model/data discrepancy became more pronounced at higher frequencies. Incorporating endocytic terms into the model may mitigate late-phase discrepancies.

STP model in light of known Ca2+ binding kinetics of Syt7

(A-D) Left, Schematic of allosteric calcium binding to Syt7. The number of Ca2+ bound to Syt7 was denoted as # in ‘S#’ in the reaction scheme. kon = 7/μM/s, koff = 10/s. Middle, Simulated changes of k1 (black) in response to 5 Hz train of Ca2+ transients (light blue traces). The priming step of the simple refilling model was assumed to be catalyzed by full Ca2+-bound from of Syt7. Accordingly, Ca2+-dependent increase in k1 was calculated as K1,max multiplied by a fraction of full Ca2+-bound form of Syt7. We assumed that local [Ca2+]i(t) follows a Gaussian function: (1/σ √2π) exp[-(t-tp)2/ 2σ2], in which tp = 0.25 ms and σ = 0.085 ms. Right, Fits of the Syt7 model to the STP data. To fit this model to the STP data, K1,max was set to 300/s (A), 220/s (B), and 180/s (C). Cooperativity factor (b) was set to 0.35 (A), 0.2 (B), and 0.05 (C). k1 is set to be constant for Syt7 KD (D). Same frequency color codes were used as in Figure 1.

Moreover, removal of the catalytic function of Syt7 in the model effectively accounted for the complete loss of facilitation observed in Syt7 KD (Fig. 9D).

Possible mechanisms underlying the depression-facilitation sequence

Cooperative binding of multiple Ca2+ ions to Syt7 predicts delayed activation and a progressive increase in vesicle refilling during train stimulations, resulting in delayed facilitation. Combining this with a high pv could replicate not only the slow facilitation at 5 Hz, but also the unique depression-facilitation sequence observed at 40 Hz (Fig. 9). As simulated by Pulido and Marty (2018), delayed facilitation can be reproduced by a high pv in combination with a low baseline occupancy of replacement sites and subsequent Ca2+-dependent refilling. However, whether this model is also able to replicate slow facilitation at low frequencies remains to be tested. An alternative, but not exclusive, possibility for delayed facilitation is that the presynaptic global Ca2+ itself may undergo a slow buildup during a train due to the saturation of endogenous Ca2+-binding proteins.

Facilitation requires not only Syt7, but also activation of PLC-DAG signaling

To gain insight into the specific molecular mechanisms underlying the vesicle-loading processes, we examined the effects of pharmacological inhibitors targeting PLC and DAG (Fig. 4). Both a PLC inhibitor and a DAG analog reduced facilitation and augmentation. This suggests that the progressive overfilling of docking sites with TS vesicles is supported by the PLC-DAG signaling pathway and probably downstream Munc13 (Rhee et al., 2002; Lou et al., 2008). Phorbol esters have been recently shown to increase the TS fraction of docked vesicles by enhancing vesicle priming without affecting Ca2+ transients or pv (Taschenberger et al., 2016; Aldahabi et al., 2022; Papantoniou et al., 2023; Aldahabi et al., 2024). In line with these findings, signaling involving PLC and DAG may support Ca2+-dependent overfilling of TS vesicles by enhancing molecular priming processes.

Roles of facilitation and augmentation in working memory and related tasks

Previous theoretical studies have proposed that working memory in a recurrent network can be maintained in the form of augmentation at synapses (Mongillo et al., 2008) and/or persistent activity of neurons in a memory ensemble. This is supported by the facilitation of recurrent excitatory synapses or disparity in STP between excitatory and inhibitory synapses. However, whether these short-term enhancements in synaptic efficacy contribute to working memory-related behaviors has not been examined. tFC is a temporal associative learning task requiring temporary retention of CS information during a trace period. The mPFC and hippocampus are crucial for tFC, and tFC is interfered with by a load of working memory and attention distraction, implying that the tFC shares the same neural substrate with other working memory tasks. In this study, tFC in WT and Syt7 KD animals revealed that Syt7 deficiency in the L2/3 pyramidal neurons of the mPFC impaired the acquisition of trace fear memory and reduced c-Fos expression associated with tFC training (Fig. 8). Given that synaptic enhancements in prelimbic L2/3 recurrent excitatory synapses can be attributed to an increase in pocc, these behavioral results suggest that presynaptic high Ca2+ transients are temporally memorized in the presynaptic terminal as an increase in TS occupancy.

Conclusions

Progressive overfilling of synaptic vesicles with a high fusion probability supports the facilitation and augmentation of recurrent excitatory synapses within the L2/3 network of the mPFC and is essential for temporal associative learning.

Materials and methods

Animals/Subjects

All experiments were carried out on Sprague-Dawley rats of either sex at p28-43. Rats were maintained in temperature-controlled rooms on a constant 12 h light/12 h dark cycle and were given access to water and food ad libitum. All animal procedures were approved and performed in accordance with the Institutional Animal Care and Use Committee at Seoul National University (SNU200522-1).

In utero electroporation (IUE)

Pregnant Sprague Dawley rats on embryonic day (E) 17.5 were deeply anesthetized with 5% (v/v) isoflurane for the duration of surgery. The uterine horns were exposed by laparotomy and wet with warm sterile phosphate-buffered saline (PBS). The plasmids (0.5-2 µg/µl) together with the 0.01% fast green dye were injected into either the left or the right lateral ventricle or bilateral ventricles of embryos through a thin glass capillary (Narishige PC-10). The embryo’s head was carefully held between the forceps-shaped electrodes of 10 mm diameter (CUY65010; NEPAGENE, Japan). Electroporation pulses (50 V for 50 ms) were delivered five times at intervals of 150 ms using a square-wave generator (ECM830; BEX, United States). The uterine horns were placed back into the abdominal cavity and abdominal muscles and skin were sutured. After the surgery, the animals were recovered under the infrared lamp. For knockdown experiments in slice electrophysiology, the shRNA constructs were co-electroporated with the pCAG-oChIEF-tdTomato, because of the consistent co-expression of the two constructs in a significant proportion of the transfected cells in vivo (Bony et al., 2013). Pups with robust fluorescence signals in the prefrontal cortex of both hemispheres were used for behavior experiments.

Preparation of vector constructs

For an optic stimulation, pCAG-oChIEF-tdTomato was transfected by IUE. Short hairpin RNA (shRNA) against Syt7 sequence (GATCTACCTGTCCTGGAAGAG) was expressed under the control of an U6 promoter of pAAV-U6-GFP vector (Cell Biolabs, #VPK-413). This shRNA was shown to effectively knockdown Syt7 in cell cultures and in vivo by previous reports from more than one research group (Bacaj et al., 2013; Li et al., 2017). Scrambled control sequence (TCGCATAGCGTATGCCGTT) was designed by shuffling the recognition region sequences and a BLAST analysis confirmed that the control shRNA sequence had no target gene. For rescue experiments, the cDNA for rat Syt7 (NM_021659) rendered resistant to the shRNA were inserted in the KD vector and their expression was driven by the CAG promoter; the vector also contained a self-cleaving P2A peptide sequence followed by EGFP, which ensures visualization of transfected cells. All the constructs above were verified by DNA sequencing. Plasmid DNA was purified and concentrated under endotoxin free conditions (Qiagen EndoFree Maxi Kit).

Slice preparation

Acute brain slices were prepared by using a VT1200S Vibratome (Leica Microsystems) after isoflurane anesthesia and decapitation. Coronal sections of the mPFC (300 μm in thickness) were cut in ice-cold artificial cerebrospinal fluid (aCSF) composed of (in mM) 125 NaCl, 3.2 KCl, 25 NaHCO3, 1.25 NaH2PO4, 20 D-glucose, 2 NaPyr, 1 MgCl2, and 1.3 CaCl2 (Yoon et al., 2020), bubbled with 95% O2 and CO2 (pH = 7.3; 310 mOsm). Slices were allowed to recover for a minimum of 30 minutes at 36 °C, and subsequently maintained at room temperature until recording. Composition of aCSF was identical throughout the preparation, incubation, and recording periods unless otherwise stated. For the recordings, the mPFC slices were continuously perfused with oxygenated aCSF at a flow rate of 1.5 ml/min, and the temperature was maintained at 31–32 °C using an in-line temperature controller (Warner Instruments).

Electrophysiology

Whole-cell patch-clamp recordings were performed from pyramidal cells and fast-spiking interneurons in layer 2/3 of the prelimbic mPFC. Neuronal subtypes were initially selected based on morphology under visual guidance by using an upright microscope equipped with infrared differential interference contrast (IR-DIC) optics (BX51WI, Olympus), and were further identified by their electrophysiological properties (see Supplementary Fig. 1; van Aerde and Feldmeyer, 2015; Tremblay et al., 2016). To analyze these intrinsic properties in current-clamp experiments, we measured the following parameters: (1) resting membrane potential (RMP), (2) input resistance (Rin), (3) sag ratio, and (4) the F-I curve, which plots firing frequencies (F) against the amplitude of injected currents (I). Both cell types exhibited a low sag ratio below 0.1, consistent with findings from other studies (van Aerde and Feldmeyer, 2015; Yoon et al., 2020). PCs and FSINs were easily distinguished by their unique firing frequency, Rin, spike adaptation, RMP and morphological characteristics including the shape of soma and the thickness of apical dendrites. For experiments involving oChIEF-expressing neurons in WT or Syt7 KD rats, whole-cell recordings were performed from the fluorescently labeled pyramidal neurons in layer 2/3 of the prelimbic cortex using suitable filter sets. Patch pipettes (tip resistance of 2–3 MΩ) were pulled from borosilicate glass and filled with (in mM) 130 K-gluconate, 8 KCl, 10 HEPES, 0.2 EGTA, 20 Na2Pcr, 0.3 Na2GTP, 4 MgATP. Intracellular solutions were adjusted to pH 7.25 and 300–310 mOsm. Neurons were voltage clamped at −78 mV and series resistance (Rs) were continuously monitored during the EPSCs recordings. Experiments were discarded before analysis if the Rs changed by 20% or larger deviation of baseline value. Measurements were not corrected for Rs compensation, bridge balance, or liquid junction potential. Electrophysiological data were low-pass-filtered at 1 kHz (Bessel) and acquired at 20 kHz using a Multiclamp 700B amplifier paired with Digidata 1440A digitizer and Clampex 10.2 software (Molecular Devices).

Synaptic stimulation

IUE resulted in sparse expression of channelrhodopsin (10-20%) (Fig. 1A). Exploiting the sparse expression, we employed a minimal optical stimulation method to activate a single or very few excitatory synapses in layer 2/3. The optic minimal stimulation was performed by our published protocol with minor modifications (Yoon et al., 2020). This approach aimed to minimize the risk of synaptic contamination, which could arise from engaging extra terminals that were initially inactive (depolarized below threshold) but became active by temporal summation during high-frequency trains. It also helped prevent the buildup of depolarization at the synaptic terminals, which might influence the probability of release, as noted by (Jackman et al., 2014). To achieve this, we employed a collimated digital micromirror device (DMD)-coupled LED (Polygon400; Mightex Systems) to confine 470 nm blue light to a small area with a radius of approximately 3–8 μm (typically 3–4 μm), as measured at the focal plane of a 60× water immersion objective [numerical aperture (NA) = 1.0; LUMPlanFL, Olympus] (Fig. S3A). To find the minimal stimulation area, the radius of illumination area was increased from 2 μm by 50% at each step (Fig. S3A), and selected the smallest illumination area that elicited EPSCs. Setting the duration of illumination between 3 and 5 ms (typically 5 ms), photostimulation onto oChIEF-expressing cells reliably induced action potentials (APs) during 600-pulse trains (see Fig. S2).

STP at various stimulus frequencies (5, 10, 20, and 40 Hz) was examined using 20-pulse trains. Trials of train stimulation were repeated with a prolonged interval of 30 s to avoid change in baseline properties. We observed no systematic changes in the amplitudes of the first EPSC of each train, nor did we find significant differences in the paired-pulse ratio (PPR) across stimulation trials in the same cell. For each stimulus frequency, approximately ten to twenty consecutive traces were collected and then averaged across trials to yield a mean EPSC trace. For recovery experiments, two consecutive trains of 3 or 30 stimuli at 40 Hz, with varying time intervals (0.1-10 s), were applied every 10 s or 60 s, respectively, at both excitatory synapses. For 600-stimulus trains at 10Hz, the trial was delivered only once for each cell to avoid possible after-effects caused by prolonged synaptic stimulation (Hirsch and Crepel, 1990). Prior to the stimulus trains, the EPSCs were normalized to the mean amplitude of the baseline. For the baseline measurements, paired pulses with various ISIs (20-200 ms) were delivered every 3 s (1/3 Hz) at least 20 times.

For all recordings, picrotoxin (PTX, 10 μM) was included in the recording solution to isolate glutamatergic responses. mPFC PC terminals were entirely glutamatergic as the application of 6-cyano-7-nitroquinoxaline-2,3-dione (CNQX, 20 μM), an AMPA receptor antagonist, completely abolished the monosynaptic events evoked by photostimulation. A low concentration of PTX was employed to minimize the relieving effects on spike suppression caused by blockade of GABAa receptors. We confirmed that no synaptic response was elicited by electrical stimulation in the presence of both 10 μM PTX and 20 μM CNQX.

In Fig. S4A, 1 µM TTX (1 mM stock solution) and 0.1 mM 4-AP (500 mM stock solution) were added to the bath solution to block action potentials and restore presynaptic glutamate release, respectively (Fig. S4A). In Fig. S5A, 50 µM of cyclothiazide, a positive allosteric modulator of the AMPA receptors, was included to minimize rapid desensitization of the postsynaptic receptors. In Fig. S5B, 0.5 mM kynurenate was added to bath solution to prevent possible saturation of AMPA receptors, especially during maximal synaptic facilitation. For pharmacological manipulations of synaptic vesicle dynamics (Fig. 3), drugs related to vesicle supply including U73122, OAG, LatB, EGTA-AM, and dynasore at indicated concentrations were added to the recording solution during baseline measurements or PTA/recovery experiments (IBI = 0.5 s). Chemicals were from Sigma-Aldrich, Merck, or Tocris.

Measurements of quantal size

Experiments for measuring asynchronous release were performed as previously described (Yoon et al., 2020). In brief, we replaced extracellular Ca2+ entirely with 2 mM Sr2+ to induce asynchronous release. A single stimulus was delivered at 3-second intervals over 100 repetitions, and asynchronous release events were detected during a 100-ms window starting 25 ms after stimulus onset for each trial. The same procedure was also executed under sham conditions (0% light intensity) to serve as a control, thus mitigating contamination from nonspecific spontaneous events and potential errors associated with in the detection algorithm. q was determined as the mean value from the first Gaussian function after fitting the EPSC distribution with either a single or double Gaussian function (Fig. S8).

Calculation of probability for double failure

The failure rate at the 1st pulse when initial RRP size (n1) = i can be calculated as: P(F1 | n1=i) = (1 - pv)i. The probability that initial RRP size is i was calculated as: P(n1=i) = C(N, i) * p i * (1-pocc)(N-i), where pocc = mean release site occupancy, and N = total number of release sites. Under the conditions of no vesicle release at the 1st pulse (F1) and n1 = i, the probability for the second failure is , where n2 = number of vesicles just before the 2nd pulse. When forward and backward rate constants are k1 and b1, respectively, P(n2|n1=i, F1) is calculated as follows: , where Δt = ISI. By the Bayes rule, Thus, the probability for two consecutive failures can be calculated as: P(F1) and N are observed values. The k1 and pv were set free variables. Other parameters were set according to following relationships: P(F1) = (1 – pr)N, pr = pv pocc, and pocc = k1 / (k1 + b1).

Behavioral tests

Trace fear conditioning

For trace fear conditioning (tFC) (Gilmartin et al., 2013), individual rats were placed in the training chamber (30.5 x 25.4 x 30.5 cm; Coulbourn Instruments) consisting of metal grid floor connected to an electrical stimulator (H10-11R-TC-SF, Coulbourn Instruments) for delivery of a footshock. After a 5 min habituation period, rats received five pairings of a 20 s tone conditioned stimulus (CS, 2.5 kHz, 75 dB), followed by empty 20s trace period and a unconditioned stimulus (US; 1 s, 0.7 mA footshock) with a pseudorandom inter-trial interval (ITI) of 240 ± 20 s. During this training session, rats learned to associate the auditory CS with the shock US. The next day, the rats were tested for memory of CS-US trace association (called tone test). The acquisition of fear learning, often indicated by the extent of freezing to the CS, was assessed in a novel chamber (a white hexagonal enclosure) in a dimly lit room. During this tone (or retrieval) test, rats were first allowed to explore the new context for 5 min (habituation), followed by twelve 20 s CS presentations with a variable ITI of 120 ± 20 s. This entire session was repeated on the next day for assessment of extinction learning. Identities of the chambers were also determined with the presence of distinct odors; the training and testing chambers were cleaned with 70% ethanol and 1% acetic acid respectively before and after each session. For all experimental sessions, the activity of animals in the chambers was recorded at 30 frames per second using the EthoVision XT (Noldus Information Technology, Wageningen, Netherlands), and stored as a video file. The freezing behavior, defined as behavioral immobility other than respiratory movement, was assessed using an open-source video analysis pipeline ezTrack (Pennington et al., 2019). The percentage of freezing was calculated as a total duration of freezing divided by the total duration of observation.

Elevated Plus Maze

Elevated plus maze (EPM) test was performed using a plus-shaped apparatus elevated 60 cm above the ground. The maze consisted of two open arms (50 x 10 x 0.5 cm (H)), two closed arms (50 x 10 x 40 cm (H)) and a center area (10 x 10 cm). Rats were individually placed on the center area, facing an open arm, and the path of the animal was recorded with a video camera for 5 min. We analyzed the total distance traveled for 5 min and time spent in each arm. Animals were tracked offline using the open-source tracking system ezTrack (Pennington et al., 2019).

Open field test

Open field exploration test (OFT) was performed in an open field apparatus in a dimly lit room. The open field consisted of a white plastic board (1.20 m in diameter) surrounded by white plastic walls (50 cm in height). Individual rats were placed in a center zone (0.6 m in diameter), and allowed to freely explore the apparatus for 10 min. Time spent in the center zone and total distance traveled were analyzed for initial 5 min. Animals were tracked offline using the open-source tracking system ezTrack (Pennington et al., 2019).

Histology and immunohistochemistry

At the end of the behavioral experiments, rats were anesthetized using isoflurane inhalation and transcardially perfused with PBS followed by 4% paraformaldehyde (PFA, T&I, Korea). Brains were extracted and post-fixed in 4% PFA solution for 24 h. Perfused brains were sliced at 100 μM for estimation of GFP expression or 50 μM for immunohistochemistry. For analysis of c-Fos expression, rats were returned to their home cages immediately after the training session and perfused approximately 90 minutes later (Arime and Akiyama, 2017). Brains were then cryopreserved, frozen, and sectioned coronally. Immunofluorescence staining was conducted on free-floating slices using a rabbit monoclonal anti c-fos (1:1000, Cell Signaling Technology, cat# 2250s) and mouse monoclonal anti-GAD67 (1:500, Millipore, cat# mab5406). Fluorescent secondary antibodies included Cy5-conjugated goat anti-rabbit IgG (1:1000, Abcam, cat# ab97077) and Cy3-conjugated goat anti-mouse IgG (1:1000, Abcam, cat# ab97035). Brain slices were permeabilized and blocked with PBS containing 0.5% Triton X-100 and 5% normal goat serum (NGS) for 1h at room temperature. Slices were incubated overnight at 4 °C with primary antibodies, followed by 4 h incubation with secondary antibodies at room temperature. Slices were then mounted with Antifading Mounting medium (Abcam, cat# ab104139). Confocal images were acquired using a Leica TCS SP8 microscope set to the same laser intensity and acquisition parameters to compare the immunosignals in sections from Scr and Syt7 KD groups. For cell counting and colocalization, fluorescent protein-positive cells were automatically quantified with spot-detection algorithms in Imaris 9.5 (Oxford instruments).

Data analysis

Data were analyzed and presented using ClampFit (Molecular Devices), Igor Pro (Wavemetrics), MATLAB (Mathworks), and Prism (GraphPad). Data points and error bars represent the mean ± standard error of the mean. All statistical analyses were conducted using two-tailed comparisons. Sample sizes and statistical tests for each group are detailed in the figure legends.

Numerical integration of the simple refilling model

This model posits reversible docking of vesicles at finite release sites (Nmax) with forward (k1) and reverse rate constants (b1) (Hosoi et al., 2007; Neher and Sakaba, 2008). According to this scheme, dn/dt = k1 u - b1 n - pv n δ(t - tAP), where u is the number of unoccupied sites (= Nmax - n); tAP is the timing of AP firing; and δ is the Dirac delta function. The pv was assumed to be unity (Fig. S8). The sum of basal k1 (= k1,b) and b1 was be estimated as 23/s from the dependence PPR on ISIs (Fig. S8Ab). From the baseline occupancy of 0.3 and pocc = k1 / (k1 + b1), we could estimate baseline k1 (k1,b) and b1 as 6.9/s, and 16.1/s, respectively. Ca2+ and Syt7-dependent increase of k was modeled as k = k + K [Syt7: n Ca2+], where [Syt7: n Ca2+] represents the fraction of full Ca2+-bound Syt7. For deterministic simulations, we time-integrated a set of differential equations for each model using Euler methods with a time step of 1 ms. All calculations were performed on the platform of Matlab (R2022b, Mathworks, USA).

eLife Assessment

This valuable work explores how synaptic activity encodes information during memory tasks. All reviewers agree that the quality of the work is high. Although experimental data do support the possibility that phospholipase diacylglycerol signaling and synaptotagmin 7 (Syt7) dynamically regulate the vesicle pool required for presynaptic release, concerns remain that the central finding of paired pulse depression at very short intervals was more likely caused by Ca2+ channel inactivation than pool depletion. Overall, this is a solid study although the results warrant consideration of alternative interpretations.

Reviewer #1 (Public review):

Shin et al. conduct extensive electrophysiological and behavioral experiments to study the mechanisms of short-term synaptic plasticity at excitatory synapses in layer 2/3 of the rat medial prefrontal cortex. The authors interestingly find that short-term facilitation is driven by progressive overfilling of the readily releasable pool, and that this process is mediated by phospholipase C/diacylglycerol signaling and synaptotagmin-7 (Syt7). Specifically, knockdown of Syt7 not only abolishes the refilling rate of vesicles with high fusion probability, but it also impairs the acquisition of trace fear memory.

Overall, the authors offer novel insight to the field of synaptic plasticity through well-designed experiments that incorporate a range of techniques.

Reviewer #2 (Public review):

Summary:

Shin et al aim to identify in a very extensive piece of work a mechanism that contributes to dynamic regulation of synaptic output in the rat cortex at the second time scale. This mechanism is related to a new powerful model is well versed to test if the pool of SV ready for fusion is dynamically scaled to adjust supply demand aspects. The methods applied are state-of-the-art and both address quantitative aspects with high signal to noise. In addition, the authors examine both excitatory output onto glutamatergic and GABAergic neurons, which provides important information on how general the observed signals are in neural networks, The results are compellingly clear and show that pool regulation may be predominantly responsible. Their results suggests that a regulation of release probability, the alternative contender for regulation, is unlikely to be involved in the observed short term plasticity behavior (but see below). Besides providing a clear analysis pof the underlying physiology, they test two molecular contenders for the observed mechanism by showing that loss of Synaptotagmin7 function and the role of the Ca dependent phospholipase activity seems critical for the short term plasticity behavior. The authors go on to test the in vivo role of the mechanism by modulating Syt7 function and examining working memory tasks as well as overall changes in network activity using immediate early gene activity. Finally, they model their data, providing strong support for their interpretation of TS pool occupancy regulation.

Strengths:

This is a very thorough study, addressing the research question from many different angles and the experimental execution is superb. The impact of the work is high, as it applies recent models of short term plasticity behavior to in vivo circuits further providing insights how synapses provide dynamic control to enable working memory related behavior through nonpermanent changes in synaptic output.

Weaknesses:

While this work is carefully examined and the results are presented and discussed in a detailed manner, the reviewer is still not fully convinced that regulation of release provability is not a putative contributor to the observed behavior. No additional work is needed but in the moment I am not convinced that changes in release probability are not in play. One solution may be to extend the discussion of changes in rules probability as an alternative.

Fig 3 I am confused about the interpretation of the Mean Variance analysis outcome. Since the data points follow the curve during induction of short term plasticity, aren't these suggesting that release probability and not the pool size increases? Related, to measure the absolute release probability and failure rate using the optogenetic stimulation technique is not trivial as the experimental paradigm bias the experiment to a given output strength, and therefore a change in release probability cannot be excluded.

Fig4B interprets the phorbol ester stimulation to be the result of pool overfilling, however, phorbol ester stimulation has also been shown to increase release probability without changing the size of the readily releasable pool. The high frequency of stimulation may occlude an increased paired pulse depression in presence of OAG, which others have interpreted in mammalian synapses as an increase in release probability.

The literature on Syt7 function is still quite controversial. An observation in the literature that loss of Syt7 function in the fly synapse leads to an increase of release probability. Thus the observed changes in short term plasticity characteristics in the Syt7 KD experiments may contain a release probability component. Can the authors really exclude this possibility? Figure 5 shows for the Syt7 KD group a very prominent depression of the EPSC/IPSC with the second stimulus, particularly for the short interpulse intervals, usually a strong sign of increased release probability, as lack of pool refilling can unlikely explain the strong drop in synaptic output.

Reviewer #3 (Public review):

Summary:

The report by Shin, Lee, Kim, and Lee entitled "Progressive overfilling of readily releasable pool underlies short-term facilitation at recurrent excitatory synapses in layer 2/3 of the rat prefrontal cortex" describes electrophysiological experiments of short-term synaptic plasticity during repetitive presynaptic stimulation at synapses between layer 2/3 pyramidal neurons and nearby target neurons. Manipulations include pharmacological inhibition of PLC and actin polymerization, activation of DAG receptors, and shRNA knockdown of Syt7. The results are interpreted as support for the hypothesis that synaptic vesicle release sites are vacant most of the time at resting synapses (i.e., p_occ is low) and that facilitation (and augmentation) components of short-term enhancement are caused by an increase in occupancy, presumably because of acceleration of the transition from not-occupied to occupied. The report additionally describes behavioural experiments where trace fear conditioning is degraded by knocking down syt7 in the same synapses.

Strengths:

The strength of the study is in the new information about short-term plasticity at local synapses in layer 2/3, and the major disruption of a memory task after eliminating short-term enhancement at only 15% of excitatory synapses in a single layer of a small brain region. The local synapses in layer 2/3 were previously difficult to study, but the authors have overcome a number of challenges by combining channel rhodopsins with in vitro electroporation, which is an impressive technical advance.

Weaknesses:

The question of whether or not short-term enhancement causes an increase in p_occ (i.e., "readily releasable pool overfilling") is important because it cuts to the heart of the ongoing debate about how to model short term synaptic plasticity in general. However, my opinion is that, in their current form, the results do not constitute strong support for an increase in p_occ, even though this is presented as the main conclusion. Instead, there are at least two alternative explanations for the results that both seem more likely. Neither alternative is acknowledged in the present version of the report.

The evidence presented to support overfilling is essentially two-fold. The first is strong paired pulse depression of synaptic strength when the interval between action potentials is 20 or 25 ms, but not when the interval is 50 ms. Subsequent stimuli at frequencies between 5 and 40 Hz then drive enhancement. The second is the observation that a slow component of recovery from depression after trains of action potentials is unveiled after eliminating enhancement by knocking down syt7. Of the two, the second is predicted by essentially all models where enhancement mechanisms operate independently of release site depletion - i.e., transient increases in p_occ, p_v, or even N - so isn't the sort of support that would distinguish the hypothesis from alternatives (Garcia-Perez and Wesseling, 2008, https://doi.org/10.1152/jn.01348.2007).

Regarding the paired pulse depression: The authors ascribe this to depletion of a homogeneous population of release sites, all with similar p_v. However, the details fit better with the alternative hypothesis that the depression is instead caused by quickly reversing inactivation of Ca2+ channels near release sites, as proposed by Dobrunz and Stevens to explain a similar phenomenon at a different type of synapse (1997, PNAS,https://doi.org/10.1113/jphysiol.1967.sp008367).

https://doi.org/10.1073/pnas.94.26.14843). The details that fit better with Ca2+ channel inactivation include the combination of the sigmoid time course of the recovery from depression (plotted backwards in Fig1G,I) and observations that EGTA (Fig2B) increases the paired-pulse depression seen after 25 ms intervals. That is, the authors ascribe the sigmoid recovery to a delay in the activation of the facilitation mechanism, but the increased paired pulse depression after loading EGTA indicates, instead, that the facilitation mechanism has already caused p_r to double within the first 25 ms (relative to the value if the facilitation mechanism was not active). Meanwhile, Ca2+ channel inactivation would be expected to cause a sigmoidal recovery of synaptic strength because of the sigmoidal relationship between Ca2+-influx and exocytosis (Dodge and Rahamimoff, 1967,

The Ca2+-channel inactivation hypothesis could probably be ruled in or out with experiments analogous to the 1997 Dobrunz study, except after lowering extracellular Ca2+ to the point where synaptic transmission failures are frequent. However, a possible complication might be a large increase in facilitation in low Ca2+ (Fig2B of Stevens and Wesseling, 1999, https://doi.org/10.1016/s0896-6273(00)80685-6).

On the other hand, even if the paired pulse depression is caused by depletion of release sites rather than Ca2+-channel inactivation, there does not seem to be any support for the critical assumption that all of the release sites have similar p_v. And indeed, there seems to be substantial emerging evidence from other studies for multiple types of release sites with 5 to 20-fold differences in p_v at a wide variety of synapse types (Maschi and Klyachko, eLife, 2020, https://doi.org/10.7554/elife.55210; Rodriguez Gotor et al, eLife, 2024, https://doi.org/10.7554/elife.88212 and refs. therein). If so, the paired pulse depression could be caused by depletion of release sites with high p_v, whereas the facilitation could occur at sites with much lower p_v that are still occupied. It might be possible to address this by eliminating assumptions about the distribution of p_v across release sites from the variance-mean analysis, but this seems difficult; simply showing how a few selected distributions wouldn't work - such as in standard multiple probability fluctuation analyses - wouldn't add much.

In any case, the large increase - often 10-fold or more - in enhancement seen after lowering Ca2+ below 0.25 mM at a broad range of synapses and neuro-muscular junctions noted above is a potent reason to be cautious about the LS/TS model. There is morphological evidence that the transitions from a loose to tight docking state (LS to TS) occur, and even that the timing is accelerated by activity. However, 10-fold enhancement would imply that at least 90 % of vesicles start off in the LS state, and this has not been reported. In addition, my understanding is that the reverse transition (TS to LS) is thought to occur within 10s of ms of the action potential, which is 10-fold too fast to account for the reversal of facilitation seen at the same synapses (Kusick et al, 2020, https://doi.org/10.1038/s41593-020-00716-1).

Individual points:

(1) An additional problem with the overfilling hypothesis is that syt7 knockdown increases the estimate of p_occ extracted from the variance-mean analysis, which would imply a faster transition from unoccupied to occupied, and would consequently predict faster recovery from depression. However, recovery from depression seen in experiments was slower, not faster. Meanwhile, the apparent decrease in the estimate of N extracted from the mean-variance analysis is not anticipated by the authors' model, but fits well with alternatives where p_v varies extensively among release sites because release sites with low p_v would essentially be silent in the absence of facilitation.

(2) Figure S4A: I like the TTX part of this control, but the 4-AP part needs a positive control to be meaningful (e.g., absence of TTX).

(3) Line 251: At least some of the previous studies that concluded these drugs affect vesicle dynamics used logic that was based on some of the same assumptions that are problematic for the present study, so the reasoning is a bit circular.

(4) Line 329 and Line 461: A similar problem with circularity for interpreting earlier syt7 studies.

Author response:

The following is the authors’ response to the original reviews

eLife Assessment

This valuable work explores how synaptic activity encodes information during memory tasks. All reviewers agree that the quality of the work is high. Although experimental data do support the possibility that phospholipase diacylglycerol signaling and synaptotagmin 7 (Syt7) dynamically regulate the vesicle pool required for presynaptic release, concerns remain that the central finding of paired pulse depression at very short intervals was more likely caused by Ca2+ channel inactivation than pool depletion. Overall, this is a solid study with valuable findings, but the results warrant consideration of alternative interpretations.

We greatly appreciate invaluable and constructive comments from Editors and Reviewers. We also thank for their time and patience. We are pleased for our manuscript to have been assessed valuable and solid.

One of the most critical concerns was a possible involvement of Ca2+ channel inactivation in the strong paired pulse depression (PPD). Meanwhile, we have measured total (free plus buffered) calcium increments induced by each of first four APs in 40 Hz trains at axonal boutons of prelimbic layer 2/3 pyramidal cells. We found that first four Ca2+ increments were not different from one another, arguing against possible contribution of Ca2+ channel inactivation to PPD. Please see our reply to the 2nd issue in the Weakness section of Reviewer #3.

The second critical issue was on the definition of ‘vesicular probability’. Previously, vesicular probability (pv) has been used with reference to the releasable vesicle pool which includes not only tightly docked vesicles but also reluctant vesicles. On the other hand, the meaning of pv in the present study is the release probability of tightly docked vesicles. We clarified this point in our replies to the 1st issues in the Weakness sections of Reviewer #2 and Reviewer #3.

We below described our point-by-point replies to the Reviewers’ comments.

Public Reviews:

Reviewer #1 (Public review):

Shin et al. conduct extensive electrophysiological and behavioral experiments to study the mechanisms of short-term synaptic plasticity at excitatory synapses in layer 2/3 of the rat medial prefrontal cortex. The authors interestingly find that short-term facilitation is driven by progressive overfilling of the readily releasable pool, and that this process is mediated by phospholipase C/diacylglycerol signaling and synaptotagmin-7 (Syt7). Specifically, knockdown of Syt7 not only abolishes the refilling rate of vesicles with high fusion probability, but it also impairs the acquisition of trace fear memory. Overall, the authors offer novel insight to the field of synaptic plasticity through well-designed experiments that incorporate a range of techniques.

Reviewer #2 (Public review):

Summary:

Shin et al aim to identify in a very extensive piece of work a mechanism that contributes to dynamic regulation of synaptic output in the rat cortex at the second time scale. This mechanism is related to a new powerful model is well versed to test if the pool of SV ready for fusion is dynamically scaled to adjust supply demand aspects. The methods applied are state-of-the-art and both address quantitative aspects with high signal to noise. In addition, the authors examine both excitatory output onto glutamatergic and GABAergic neurons, which provides important information on how general the observed signals are in neural networks, The results are compellingly clear and show that pool regulation may be predominantly responsible. Their results suggests that a regulation of release probability, the alternative contender for regulation, is unlikely to be involved in the observed short term plasticity behavior (but see below). Besides providing a clear analysis pof the underlying physiology, they test two molecular contenders for the observed mechanism by showing that loss of Synaptotagmin7 function and the role of the Ca dependent phospholipase activity seems critical for the short term plasticity behavior. The authors go on to test the in vivo role of the mechanism by modulating Syt7 function and examining working memory tasks as well as overall changes in network activity using immediate early gene activity. Finally, they model their data, providing strong support for their interpretation of TS pool occupancy regulation.

Strengths:

This is a very thorough study, addressing the research question from many different angles and the experimental execution is superb. The impact of the work is high, as it applies recent models of short term plasticity behavior to in vivo circuits further providing insights how synapses provide dynamic control to enable working memory related behavior through nonpermanent changes in synaptic output.

Weaknesses:

(1) While this work is carefully examined and the results are presented and discussed in a detailed manner, the reviewer is still not fully convinced that regulation of release provability is not a putative contributor to the observed behavior. No additional work is needed but in the moment I am not convinced that changes in release probability are not in play. One solution may be to extend the discussion of changes in release probability as an alternative.

Quantal content (m) depends on n * pv, where n = RRP size and pv =vesicular release probability. The value for pv critically depends on the definition of RRP size. Recent studies revealed that docked vesicles have differential priming states: loosely or tightly docked state (LS or TS, respectively). Because the RRP size estimated by hypertonic solution or long presynaptic depolarization is larger than that by back extrapolation of a cumulative EPSC plot (Moulder & Mennerick, 2005; Sakaba, 2006) in glutamatergic synapses, the former RRP (denoted as RRPhyper) may encompass not only AP-evoked fast-releasing vesicles (TS vesicle) but also reluctant vesicles (LS vesicles). Because we measured pv based on AP-evoked EPSCs such as strong paired pulse depression (PPD) and associated failure rates, pv in the present study denotes vesicular fusion probability of TS vesicles, not that of LS plus TS vesicles.

Recent studies suggest that release sites are not fully occupied by TS vesicles in the baseline (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Instead, the occupancy (pocc) by TS vesicles is subject to dynamic regulation by reversible rate constants (denoted by k1 and b1, respectively). The number of TS vesicles (n) can be factored into the number of release sites (N) and pocc, among which N is a fixed parameter but pocc depends on k1/(k1+b1) under the framework of the simple refilling model (see Methods). Because these refilling rate constants are regulated by Ca2+ (Hosoi, et al., 2008), pocc is not a fixed parameter. Therefore, release probability should be re-defined as pocc * pv. Given that N is fixed, the increase in release probability is a major player in STF. Our study asserts that STF by 2.3 times can be attributed to an increase in pocc rather than pv, because pv is close to unity (Fig. S8). Moreover, strong PPD was observed not only in the baseline but also at the early and in the middle of a train (Fig. 2 and 7) and during the recovery phase (Fig. 3), arguing against a gradual increase in pv of reluctant vesicles.

We imagine that the Reviewer meant vesicular release or fusion probability (pv) by ‘release probability’. If so, pv (of TS vesicles) cannot be a major player in STF, because the baseline pv is already higher than 0.8 even if it is most parsimoniously estimated (Fig. 2). Moreover, considering very high refilling rate (23/s), the high double failure rate cannot be explained without assuming that pv is close to unity (Fig. S8).

Conventional models for facilitation assume a post-AP residual Ca2+-dependent step increase in pv of RRP (Dittman et al., 2000) or reluctant vesicles (Turecek et al., 2016). Given that pv of TS vesicles is close to one, an increase in pv of TS vesicles cannot account for facilitation. The possibility for activity-dependent increase in fusion probability of LS vesicles (denoted as pv,LS) should be considered in two ways depending on whether LS and TS vesicles reside in distinct pools or in the same pool. Notably, strong PPD at short ISI implies that pv,LS is near zero at the resting state. Whereas LS vesicles do not contribute to baseline transmission, short-term facilitation (STF) may be mediated by cumulative increase in pv v,LS that reside in a distinct pool. Because the increase in pv,LS during facilitation recruits new release sites (increase in N), the variance of EPSCs should become larger as stimulation frequency increases, resulting in upward deviation from a parabola in the V-M plane, as shown in recent studies (Valera et al., 2012; Kobbersmed et al., 2020). This prediction is not compatible with our results of V-M analysis (Fig. 3), showing that EPSCs during STF fell on the same parabola regardless of stimulation frequencies. Therefore, it is unlikely that an increase in fusion probability of reluctant vesicles residing in a distinct release pool mediates STF in the present study.

For the latter case, in which LS and TS vesicles occupy in the same release sites, it is hard to distinguish a step increase in fusion probability of LS vesicles from a conversion of LS vesicles to TS. Nevertheless, our results do not support the possibility for gradual increase in pv,LS that occurs in parallel with STF. Strong PPD, indicative of high pv, was consistently found not only in the baseline (Fig. 2 and Fig. S6) but also during post-tetanic augmentation phase (Fig. 3D) and even during the early development of facilitation (Fig. 2D-E and Fig. 7), arguing against gradual increase in pv,LS. One may argue that STF may be mediated by a drastic step increase of pv,LS from zero to one, but it is not distinguishable from conversion of LS to TS vesicles.

To address the reviewer’s concern, we incorporated these perspectives into Discussion and further clarified the reasoning behind our conclusions.

References

Moulder KL, Mennerick S (2005) Reluctant vesicles contribute to the total readily releasable pool in glutamatergic hippocampal neurons. J Neurosci 25:3842–3850.

Sakaba, T (2006) Roles of the fast-releasing and the slowly releasing vesicles in synaptic transmission at the calyx of Held. J Neurosci 26(22): 5863-5871.

Please note that papers cited in the manuscript are not repeated here.

(2) Fig 3 I am confused about the interpretation of the Mean Variance analysis outcome. Since the data points follow the curve during induction of short term plasticity, aren't these suggesting that release probability and not the pool size increases? Related, to measure the absolute release probability and failure rate using the optogenetic stimulation technique is not trivial as the experimental paradigm bias the experiment to a given output strength, and therefore a change in release probability cannot be excluded.

Under the recent definition of release probability, it can be factored into pv and pocc, which are fusion probability of TS vesicles and the occupancy of release sites by TS vesicles, respectively. With this regard, our interpretation of the Variance-Mean results is consistent with conventional one: different data points along a parabola represent a change in release probability (= pocc x pv). Our novel finding is that the increase in release probability should be attributed to an increase in pocc, not to that in pv.

(3) Fig4B interprets the phorbol ester stimulation to be the result of pool overfilling, however, phorbol ester stimulation has also been shown to increase release probability without changing the size of the readily releasable pool. The high frequency of stimulation may occlude an increased paired pulse depression in presence of OAG, which others have interpreted in mammalian synapses as an increase in release probability.

To our experience in the calyx of Held synapses, OAG, a DAG analogue, increased the fast releasing vesicle pool (FRP) size (Lee JS et al., 2013), consistent with our interpretation (pool overfilling). Once the release sites are overfilled in the presence of OAG, it is expected that the maximal STF (ratio of facilitated to baseline EPSCs) becomes lower as long as the number of release sites (N) are limited. As aforementioned, the baseline pv is already close to one, and thus it cannot be further increased by OAG. Instead, the baseline pocc seems to be increased by OAG.

Reference

Lee JS, et al., Superpriming of synaptic vesicles after their recruitment to the readily releasable pool. Proc Natl Acad Sci U S A, 2013. 110(37): 15079-84.

(4) The literature on Syt7 function is still quite controversial. An observation in the literature that loss of Syt7 function in the fly synapse leads to an increase of release probability. Thus the observed changes in short term plasticity characteristics in the Syt7 KD experiments may contain a release probability component. Can the authors really exclude this possibility? Figure 5 shows for the Syt7 KD group a very prominent depression of the EPSC/IPSC with the second stimulus, particularly for the short interpulse intervals, usually a strong sign of increased release probability, as lack of pool refilling can unlikely explain the strong drop in synaptic output.

The reviewer raises an interesting point regarding the potential link between Syt7 KD and increased initial pv, particularly in light of observations in Drosophila synapses (Guan et al., 2020; Fujii et al., 2021), in which Syt7 mutants exhibited elevated initial pv. However, it is important to note that these findings markedly differ from those in mammalian systems, where the role of Syt7 in regulating initial pv has been extensively studied. In rodents, consistent evidence indicates that Syt7 does not significantly affect initial pv, as demonstrated in several studies (Jackman et al., 2016; Chen et al., 2017; Turecek and Regehr, 2018). Furthermore, in our study of excitatory synapses in the mPFC layer 2/3, we observed an initial pv already near its maximal level, approaching a value of 1. Consequently, it is unlikely that the loss of Syt7 could further elevate the initial pv. Instead, such effects are more plausibly explained by alternative mechanisms, such as alterations in vesicle replenishment dynamics, rather than a direct influence on pv.

References

Chen, C., et al., Triple Function of Synaptotagmin 7 Ensures Efficiency of High-Frequency Transmission at Central GABAergic Synapses. Cell Rep, 2017. 21(8): 2082-2089.

Fujii, T., et al., Synaptotagmin 7 switches short-term synaptic plasticity from depression to facilitation by suppressing synaptic transmission. Scientific reports, 2021. 11(1): 4059.

Guan, Z., et al., Drosophila Synaptotagmin 7 negatively regulates synaptic vesicle release and replenishment in a dosage-dependent manner. Elife, 2020. 9: e55443.

Jackman, S.L., et al., The calcium sensor synaptotagmin 7 is required for synaptic facilitation. Nature, 2016. 529(7584): 88-91.

Turecek, J. and W.G. Regehr, Synaptotagmin 7 mediates both facilitation and asynchronous release at granule cell synapses. Journal of Neuroscience, 2018. 38(13): 3240-3251.

Reviewer #3 (Public review):

Summary:

The report by Shin, Lee, Kim, and Lee entitled "Progressive overfilling of readily releasable pool underlies short-term facilitation at recurrent excitatory synapses in layer 2/3 of the rat prefrontal cortex" describes electrophysiological experiments of short-term synaptic plasticity during repetitive presynaptic stimulation at synapses between layer 2/3 pyramidal neurons and nearby target neurons. Manipulations include pharmacological inhibition of PLC and actin polymerization, activation of DAG receptors, and shRNA knockdown of Syt7. The results are interpreted as support for the hypothesis that synaptic vesicle release sites are vacant most of the time at resting synapses (i.e., p_occ is low) and that facilitation (and augmentation) components of short-term enhancement are caused by an increase in occupancy, presumably because of acceleration of the transition from not-occupied to occupied. The report additionally describes behavioural experiments where trace fear conditioning is degraded by knocking down syt7 in the same synapses.

Strengths:

The strength of the study is in the new information about short-term plasticity at local synapses in layer 2/3, and the major disruption of a memory task after eliminating short-term enhancement at only 15% of excitatory synapses in a single layer of a small brain region. The local synapses in layer 2/3 were previously difficult to study, but the authors have overcome a number of challenges by combining channel rhodopsins with in vitro electroporation, which is an impressive technical advance.

Weaknesses:

(1) The question of whether or not short-term enhancement causes an increase in p_occ (i.e., "readily releasable pool overfilling") is important because it cuts to the heart of the ongoing debate about how to model short term synaptic plasticity in general. However, my opinion is that, in their current form, the results do not constitute strong support for an increase in p_occ, even though this is presented as the main conclusion. Instead, there are at least two alternative explanations for the results that both seem more likely. Neither alternative is acknowledged in the present version of the report.

The evidence presented to support overfilling is essentially two-fold. The first is strong paired pulse depression of synaptic strength when the interval between action potentials is 20 or 25 ms, but not when the interval is 50 ms. Subsequent stimuli at frequencies between 5 and 40 Hz then drive enhancement. The second is the observation that a slow component of recovery from depression after trains of action potentials is unveiled after eliminating enhancement by knocking down syt7. Of the two, the second is predicted by essentially all models where enhancement mechanisms operate independently of release site depletion - i.e., transient increases in p_occ, p_v, or even N - so isn't the sort of support that would distinguish the hypothesis from alternatives (Garcia-Perez and Wesseling, 2008, https://doi.org/10.1152/jn.01348.2007).

The apparent discrepancy in interpretation of post-tetanic augmentation between the present and previous papers [Sevens Wesseling (1999), Garcia-Perez and Wesseling (2008)] is an important issue that should be clarified. We noted that different meanings of ‘vesicular release probability’ in these papers are responsible for the discrepancy. We added an explanation to Discussion on the difference in the meaning of ‘vesicular release probability’ between the present study and previous studies [Sevens Wesseling (1999), Garcia-Perez and Wesseling (2008)]. In summary, the pv in the present study was used for vesicular release probability of TS vesicles, while previous studies used it as vesicular release probability of vesicles in the RRP, which include LS and TS vesicles. Accordingly, pocc in the present study is the occupancy of release sites by TS vesicles.

Not only double failure rate but also other failure rates upon paired pulse stimulation were best fitted at pv close to 1 (Fig. S8 and associated text). Moreover, strong PPD, indicating release of vesicles with high pv, was observed not only at the beginning of a train but also in the middle of a 5 Hz train (Fig. 2D), during the augmentation phase after a 40 Hz train (Fig 3D), and in the recovery phase after three pulse bursts (Fig. 7). Given that pv is close to 1 throughout the EPSC trains and that N does not increase during a train (Fig. 3), synaptic facilitation can be attained only by the increase in pocc (occupancy of release sites by TS vesicles). In addition, it should be noted that Fig. 7 demonstrates strong PPD during the recovery phase after depletion of TS vesicles by three pulse bursts, indicating that recovered vesicles after depletion display high pv too. Knock-down of Syt7 slowed the recovery of TS vesicles after depletion of TS vesicles, highlighting that Syt7 accelerates the recovery of TS vesicles following their depletion.

As addressed in our reply to the first issue raised by Reviewer #2 and the third issue raised by Reviewer #3, our results do not support possibilities for recruitment of new release sites (increase in N) having low pv or for a gradual increase in pv of reluctant vesicles during short-term facilitation.

Following statement was added to Discussion in the revised manuscript

“Previous studies suggested that an increase in pv is responsible for post-tetanic augmentation (Stevens and Wesseling, 1999; Garcia-Perez and Wesseling, 2008) by observing invariance of the RRP size after tetanic stimulation. In these studies, the RRP size was estimated by hypertonic sucrose solution or as the sum of EPSCs evoked 20 Hz/60 pulses train (denoted as ‘RRPhyper’). Because reluctant vesicles (called LS vesicles) can be quickly converted to TS vesicles (16/s) and are released during a train (Lee et al., 2012), it is likely that the RRP size measured by these methods encompasses both LS and TS vesicles. In contrast, we assert high pv based on the observation of strong PPD and failure rates upon paired stimulations at ISI of 20 ms (Fig. 2 and Fig. S8). Given that single AP-induced vesicular release occurs from TS vesicles but not from LS vesicles, pv in the present study indicates the fusion probability of TS vesicles. From the same reasons, pocc denotes the occupancy of release sites by TS vesicles. Note that our study does not provide direct clue whether release sites are occupied by LS vesicles that are not tapped by a single AP, although an increase in the LS vesicle number may accelerate the recovery of TS vesicles. As suggested in Neher (2024), even if the number of LS plus TS vesicles are kept constant, an increase in pocc (occupancy by TS vesicles) would be interpreted as an increase in ‘vesicular release probability’ as in the previous studies (Stevens and Wesseling (1999); Garcia-Perez and Wesseling (2008)) as long as it was measured based on RRPhyper.”

(2) Regarding the paired pulse depression: The authors ascribe this to depletion of a homogeneous population of release sites, all with similar p_v. However, the details fit better with the alternative hypothesis that the depression is instead caused by quickly reversing inactivation of Ca2+ channels near release sites, as proposed by Dobrunz and Stevens to explain a similar phenomenon at a different type of synapse (1997, PNAS, https://doi.org/10.1073/pnas.94.26.14843). The details that fit better with Ca2+ channel inactivation include the combination of the sigmoid time course of the recovery from depression (plotted backwards in Fig1G,I) and observations that EGTA (Fig2B) increases the paired-pulse depression seen after 25 ms intervals. That is, the authors ascribe the sigmoid recovery to a delay in the activation of the facilitation mechanism, but the increased paired pulse depression after loading EGTA indicates, instead, that the facilitation mechanism has already caused p_r to double within the first 25 ms (relative to the value if the facilitation mechanism was not active). Meanwhile, Ca2+ channel inactivation would be expected to cause a sigmoidal recovery of synaptic strength because of the sigmoidal relationship between Ca2+-influx and exocytosis (Dodge and Rahamimoff, 1967, https://doi.org/10.1113/jphysiol.1967.sp008367).

The Ca2+-channel inactivation hypothesis could probably be ruled in or out with experiments analogous to the 1997 Dobrunz study, except after lowering extracellular Ca2+ to the point where synaptic transmission failures are frequent. However, a possible complication might be a large increase in facilitation in low Ca2+ (Fig2B of Stevens and Wesseling, 1999, https://doi.org/10.1016/s0896-6273(00)80685-6).

We appreciate the reviewer's thoughtful comment regarding the potential role of Ca2+ channel inactivation in the observed paired-pulse depression (PPD). As noted by the Reviewer, the Dobrunz and Stevens (1997) suggested that the high double failure rate at short ISIs in synapses exhibiting PPD can be attributed to Ca2+ channel inactivation. This interpretation seems to be based on a premise that the number of RRP vesicles are not varied trial-by-trial. The number of TS vesicles, however, can be dynamically regulated depending on the parameters k1 and b1, as shown in Fig. S8, implying that the high double failure rate at short ISIs cannot be solely attributed to Ca2+ channel inactivation. Nevertheless, we acknowledge the possibility that Ca2+ channel inactivation may contribute to PPD, and therefore, we have further investigated this possibility. Specifically, we measured action potential (AP)-evoked Ca2+ transients at individual axonal boutons of layer 2/3 pyramidal cells in the mPFC using two-dye ratiometry techniques. Our analysis revealed no evidence for Ca2+ channel inactivation during a 40 Hz train of APs. This finding indicates that voltage-gated Ca2+ channel inactivation is unlikely to contribute to the pronounced PPD.

Figure 2—figure supplement 2 shows how we measured the total Ca2+ increments at axonal boutons. First we estimated endogenous Ca2+-binding ratio from analyses of single AP-induced Ca2+ transients at different concentrations of Ca2+ indicator dye (panels A to E). And then, using the Ca2+ buffer properties, we converted free [Ca2+] amplitudes to total calcium increments for the first four AP-evoked Ca2+ transients in a 40 Hz train (panels G-I). We incorporated these results into the revised version of our manuscript to provide evidence against the Ca2+ channel inactivation.

(3) On the other hand, even if the paired pulse depression is caused by depletion of release sites rather than Ca2+-channel inactivation, there does not seem to be any support for the critical assumption that all of the release sites have similar p_v. And indeed, there seems to be substantial emerging evidence from other studies for multiple types of release sites with 5 to 20-fold differences in p_v at a wide variety of synapse types (Maschi and Klyachko, eLife, 2020, https://doi.org/10.7554/elife.55210; Rodriguez Gotor et al, eLife, 2024, https://doi.org/10.7554/elife.88212 and refs. therein). If so, the paired pulse depression could be caused by depletion of release sites with high p_v, whereas the facilitation could occur at sites with much lower p_v that are still occupied. It might be possible to address this by eliminating assumptions about the distribution of p_v across release sites from the variance-mean analysis, but this seems difficult; simply showing how a few selected distributions wouldn't work - such as in standard multiple probability fluctuation analyses - wouldn't add much.

We appreciate the reviewer’s insightful comments regarding the potential increase in pfusion of reluctant vesicles. It should be noted, however, that Maschi and Klyachko (2020) showed a distribution of release probability (pr) within a single active zone rather than a heterogeneity in pfusion of individual docked vesicles. Therefore both pocc and pv of TS vesicles would contribute to the pr distribution shown in Maschi and Klyachko (2020).

The Reviewer’s concern aligns closely with the first issue raised by Reviewer #2, to which we addressed in detail. Briefly, new release site may not be recruited during facilitation or post-tetanic augmentation, because variance of EPSCs during and after a train fell on the same parabola (Fig. 3). Secondly, strong PPD was observed not only in the baseline but also during early and late phases of facilitation, indicating that vesicles with very high pv contribute to EPSC throughout train stimulations (Fig. 2, 3, and 7). These findings argue against the possibilities for recruitment of new release sites harboring low pv vesicles and for a gradual increase in fusion probability of reluctant vesicles.

To address the reviewers’ concern, we incorporated the perspectives into Discussion and further clarified the reasoning behind our conclusions.

(4) In any case, the large increase - often 10-fold or more - in enhancement seen after lowering Ca2+ below 0.25 mM at a broad range of synapses and neuro-muscular junctions noted above is a potent reason to be cautious about the LS/TS model. There is morphological evidence that the transitions from a loose to tight docking state (LS to TS) occur, and even that the timing is accelerated by activity. However, 10-fold enhancement would imply that at least 90 % of vesicles start off in the LS state, and this has not been reported. In addition, my understanding is that the reverse transition (TS to LS) is thought to occur within 10s of ms of the action potential, which is 10-fold too fast to account for the reversal of facilitation seen at the same synapses (Kusick et al, 2020, https://doi.org/10.1038/s41593-020-00716-1).

As the Reviewer suggested, low external Ca2+ concentration can lower release probability (pr). Given that both pv and pocc are regulated by [Ca2+]i, low external [Ca2+] may affect not only pv but also pocc, both of which would contribute to low pr. Under such conditions, it would be plausible that the baseline pr becomes much lower than 0.1 due to low pv and pocc (for instance, pv decreases from 1 to 0.5, and pocc from 0.3 to 0.1, then pr = 0.05), and then pr (= pv x pocc) has a room for an increase by a factor of ten (0.5, for example) by short-term facilitation as cytosolic [Ca2+] accumulates during a train.

If pv is close to one, pr depends pocc, and thus facilitation depends on the number of TS vesicles just before arrival of each AP of a train. Thus, post-train recovery from facilitation would depend on restoration of equilibrium between TS and LS vesicles to the baseline. Even if transition between LS and TS vesicles is very fast (tens of ms), the equilibrium involved in de novo priming (reversible transitions between recycling vesicle pool and partially docked LS vesicles) seems to be much slower (13 s in Fig. 5A of Wu and Borst 1999). Thus, we can consider a two-step priming model (recycling pool -> LS -> TS), which is comprised of a slow 1st step (-> LS) and a fast 2nd step (-> TS). Under the framework of the two-step model, the slow 1st step (de novo priming step) is the rate limiting step regulating the development and recovery kinetics of facilitation. Given that on and off rate for Ca2+ binding to Syt7 is slow, it is plausible that Syt7 may contribute to short-term facilitation (STF) by Ca2+-dependent acceleration of the 1st step (as shown in Fig. 9). During train stimulation, the number of LS vesicles would slowly accumulate in a Syt7 and Ca2+-dependent manner, and this increase in LS vesicles would shift LS/TS equilibrium towards TS, resulting in STF. After tetanic stimulation, the recovery kinetics from facilitation would be limited by slow recovery of LS vesicles.

Reference

Wu, L.-G. and Borst J.G.G. (1999) The reduced release probability of releasable vesicles during recovery from short-term synaptic depression. Neuron, 23(4): 821-832.

Please note that papers cited in the manuscript are not repeated here.

Individual points:

(1) An additional problem with the overfilling hypothesis is that syt7 knockdown increases the estimate of p_occ extracted from the variance-mean analysis, which would imply a faster transition from unoccupied to occupied, and would consequently predict faster recovery from depression. However, recovery from depression seen in experiments was slower, not faster. Meanwhile, the apparent decrease in the estimate of N extracted from the mean-variance analysis is not anticipated by the authors' model, but fits well with alternatives where p_v varies extensively among release sites because release sites with low p_v would essentially be silent in the absence of facilitation.

Slower recovery from depression observed in the Syt7 knockdown (KD) synapses (Fig. 7) may results from a deficiency in activity-dependent acceleration of TS vesicle recovery. Although basal occupancy was higher in the Syt7 KD synapses, this does not indicate a faster activity-dependent recovery.

Higher baseline occupancy does not always imply faster recovery of PPR too. Actually PPR recovery was slower in Syt7 KD synapses than WT one (18.5 vs. 23/s). Under the framework of the simple refilling model (Fig. S8Aa), the baseline occupancy and PPR recovery rate are calculated as k1 / (k1 + b1) and (k1 + b1), respectively. The baseline occupancy depends on k1/b1, while the PPR recovery on absolute values of k1 and b1. Based on pocc and PPR recovery time constant of WT and KD synapses, we expect higher k1/b1 but lower values for (k1 + b1) in Syt7 KD synapses compared to WT ones.

Lower release sites (N) in Syt7-KD synapses was not anticipated. As you suggested, such low N might be ascribed to little recruitment of release sites during a train in KD synapses. But our results do not support this model. If silent release sites are recruited during a train, the variance should upwardly deviate from the parabola predicted under a fixed N (Valera et al., 2012; Kobbersmed et al. 2020). Our result was not the case (Fig. 3). In the first version of the manuscript, we have argued against this possibility in line 203-208.

As discussed in both the Results and Discussion sections, the baseline EPSC was unchanged by KD (Fig. S3) because of complementary changes in the number of docking sites and their baseline occupancy (Fig. 6). These findings suggest that Syt7 may be involved in maintaining additional vacant docking sites, which could be overfilled during facilitation. It remains to be determined whether the decrease in docking sites in Syt7 KD synapses is related to its specific localization of Syt7 at the plasma membrane of active zones, as proposed in previous studies (Sugita et al., 2001; Vevea et al., 2021).

(2) Figure S4A: I like the TTX part of this control, but the 4-AP part needs a positive control to be meaningful (e.g., absence of TTX).

The reason why we used 4-AP in the presence of TTX was to increase the length constant of axon fibers and to facilitate the conduction of local depolarization in the illumination area to axon terminals. The lack of EPSC in the presence of 4-AP and TTX indicates that illumination area is distant from axon terminals enough for optic stimulation-induced local depolarization not to evoke synaptic transmission. This methodology has been employed in previous studies including the work of Little and Carter (2013).

Reference

Little JP and Carter AG (2013) Synaptic mechanisms underlying strong reciprocal connectivity between the medial prefrontal cortex and basolateral amygdala. J Neurosci, 33(39): 15333-15342.

(3) Line 251: At least some of the previous studies that concluded these drugs affect vesicle dynamics used logic that was based on some of the same assumptions that are problematic for the present study, so the reasoning is a bit circular.

(4) Line 329 and Line 461: A similar problem with circularity for interpreting earlier syt7 studies.

(Reply to #3 and #4) We selected the target molecules as candidates based on their well-characterized roles in vesicle dynamics, and aimed to investigate what aspects of STP are affected by these molecules in our experimental context. For example, we could find that the baseline pocc and short-term facilitation (STF) are enhanced by the baseline DAG level and train stimulation-induced PLC activation, respectively. Notably, the effect of dynasore informed us that slow site clearing is responsible for the late depression of 40 Hz train EPSC. The knock-down experiments also provided us with information on the critical role of Syt7 in replenishment of TS vesicles. These approaches do not deviate from standard scientific reasoning but rather builds upon prior knowledge to formulate and test hypotheses.

Importantly, our conclusions do not rely solely on the assumption that altering the target molecule impacts synaptic transmission. Instead, our conclusions are derived from a comprehensive analysis of diverse outcomes obtained through both pharmacological and genetic manipulations. These interpretations align closely with prior literature, further validating our conclusions.

Therefore, the use of established studies to guide candidate selection and the consistency of our findings with existing knowledge do not represent a logical circularity but rather a reinforcement of the proposed mechanism through converging lines of evidence.

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

Comments:

(1) While the authors claim that Syt7-mediated facilitation is connected to the behavioral deficits they observed, this link is still somewhat speculative. This manuscript could benefit from further discussions of other alternative mechanisms to consider.

We added following statement to Discussion of the revised manuscript:

“The acquisition of trace fear memory was impaired by inhibition of persistent activity in mPFC during trace period (Gilmartin et al., 2013). The similar deficit observed in Syt7 KD animals is consistent with the hypothesis that STF provides bi-stable ensemble activity in a recurrent network (Mongillo et al., 2012). Nevertheless, alternative mechanisms may be responsible for the behavioral deficit. Not only recurrent network but also long-range loop between the mPFC and the mediodorsal (MD) thalamus play a critical role in maintaining persistent activity within the mPFC especially for a delay period longer than 10 s (Bolkan et al., 2017). Prefrontal L2/3 is heavily innervated by MD thalamus, and L2/3-PCs subsequently relay signals to L5 cortico-thalamic (CT) neurons (Collins et al., 2018). Given that L2/3 is an essential component of the PFC-thalamic loop, loss of STF at recurrent synapses between L2/3 PCs may lead to insufficient L2/3 inputs to L5 CT neurons and failure in the reverberant PFC-MD thalamic feedback loop. Therefore, not only L2/3 recurrent network but also its output to downstream network should be considered as a possible network mechanism underlying behavioral deficit caused by Syt7 KD L2/3.”

(2) The authors mention that Syt7 contributes to persistent activity during working memory tasks but focus on using only a trace fear conditioning task. However, it would be interesting to see if their results are generalizable to other working memory tasks (i.e. a delayed alternation task).

We thank to Reviewer for the insightful suggestion. Trace fear conditioning (tFC) shares behavioral properties with working memory (WM) tasks in that tFC is vulnerable to attentional distraction and to the load of WM task. In general WM tasks including delayed alternation tasks such as a T-maze task need persistent activity of ensemble neurons representing target-specific information among multiple choices. Different from such WM tasks, tFC is not appropriate to examine target-specific ensemble activity. Because it is not trivial to examine in vivo recordings in KD animals during delayed alternation tasks, it will be appropriate to study the effect of Syt7 KD in a separate study.

(3) The figure legend in Figure 6A and 6B mentions dotted lines and broken lines in the figure. However, this is confusing, and it is unclear as to what these lines are referring to in the figure.

To avoid the confusion in the figure legend for Figure 6A and 6B, we corrected “dotted line” to " vertical broken line", and “broken lines” to “dashed parabolas”.

(4) The manuscript can benefit from close reading and editing to catch typos and improve general readability (i.e. line 173: the word "are" is repeated twice).

We corrected typographical errors throughout the manuscript and carefully read the manuscript to improve readability. A revised version reflecting these corrections has been prepared and will be resubmitted for your consideration.

Reviewer #3 (Recommendations for the authors):

The points in this section are all minor.

(1) Line 44: Define release probability (p_r) more clearly. Authors use it to mean pv*pocc, but others routinely use it to mean pv*pocc*N.

We understand that the Reviewer meant “others routinely use it to mean pv”. At this statement, we meant conventional definition of release probability, which is release probability among vesicles of RRP. We think that it is not appropriate to re-define release probability as pv * pocc in this first paragraph of Introduction. Therefore we clarified this issue in Discussion as we mentioned in our reply to the 1st weakness issue raised by Reviewer #3.

(2) Line 82: For clarity, define better what recurrent excitatory synapses are. It seems that synapses between L2/3 PCs and local targets may all be recurrent?

Each of L2/3 and L5 of the prefrontal cortical layers harbors intralaminar recurrent excitatory synapses between pyramidal cells, called a recurrent network. Previous theoretical studies have proposed that a single layer recurrent network model can have bi-stable E/I balanced states (up- and down-states) if recurrent excitatory synapses display short-term facilitation (STF), and thus is able to temporally hold an information once external input shifts the network to the up-state. In this theory, synapses to local targets across layers are not considered and specific roles of L2/3 and L5 in working memory tasks are still elusive. For clarity, we added a statement at the beginning of the paragraph (line 82): “Each of layer 2/3 (L2/3) and layer 5 (L5) of neocortex displays intralaminar excitatory synapses between pyramidal cells comprising a recurrent network (Holmgren et al., 2003; Thomson and Lamy, 2007)”

(3) Cite earlier studies of short-term synaptic plasticity at synapses between L2/3 pyramidal neurons and local targets in mPFC. If there are none, take more explicit credit for being first.

As we mentioned in Introduction, previous studies on short-term plasticity (STP) at neocortical excitatory recurrent synapses have focused on synapses between L5 pyramidal cells (PCs) (Hemple et al. 2000; Wang et al. 2006; Morishima et al., 2011; Yoon et al., 2020). The local connectivity between L2/3 PCs in the somatosensory cortex has been elucidated by Homgren et al. (2003) and Ko et al. (2011). Although these study showed STP of EPSPs, it was at a fixed frequency or stimulus pattern at high external [Ca2+] (2 mM). There is a study on the frequency-dependence of STP of EPSP between L2/3-PCs (Feldmyer et al., 2006). Different from our study, Feldmyer et al., (2006) observed monotonous STD at all frequencies less than 50 Hz, but this study was done in the somatosensory cortex and at high external [Ca2+] (2 mM). To our knowledge, no previous study have investigated STP at recurrent excitatory synapses of L2/3 pyramidal cells of the mPFC especially at physiological external [Ca2+]. The present study, therefore, represents the first extensive investigation of STP at recurrent excitatory synapses in L2/3 of the mPFC under physiologically relevant external [Ca2+].

References

Feldmeyer D, Lubke J, Silver RA, Sakmann B (2002) Synaptic connections between layer 4 spiny neurone-layer 2/3 pyramidal cell pairs in juvenile rat barrel cortex: physiology and anatomy of interlaminar signalling within a cortical column. J Physiol 538:803-822.

Holmgren C, Harkany T, Svennenfors B, Zilberter Y (2003) Pyramidal cell communication within local networks in layer 2/3 of rat neocortex. J Physiol 551:139-153.

Ko H, Hofer SB, Pichler B, Buchanan KA, Sjöström PJ, Mrsic-Flogel TD (2011) Functional specificity of local synaptic connections in neocortical networks. Nature 473:87-91.

Morishima M, Morita K, Kubota Y, Kawaguchi Y (2011) Highly differentiated projection-specific cortical subnetworks. Journal of Neuroscience 31:10380-10391.

Wang Y, Markram H, Goodman PH, Berger TK, Ma J, Goldman-Rakic PS (2006) Heterogeneity in the pyramidal network of the medial prefrontal cortex. Nat Neurosci 9:534-542.

(4) I couldn't figure out the significance of Figure S3. Perhaps this could be explained better.

Optical minimal stimulation methods have not been previously documented in detail. This figure illustrates what parameters we should carefully examine in order to attain optical minimal stimulation, which hopefully stimulates a single afferent fiber. A single fiber stimulation by optical minimal stimulation is supported by the similarity of our estimate for the number of release sites (N) as the previous morphological estimate (Holler et al., 2021). For minimal stimulation, we used a collimated DMD-coupled LED was employed to restrict 470 nm illumination to a small and well-defined region within layer 2/3 of the prelimbic mPFC, and carefully adjusted the illumination radius such that one step smaller (by 1 μm) illumination results in failure to evoke EPSCs. Our typical illumination area ranged between 3–4 μm, as shown in Figure S3A. Under this minimal illumination area, we confirmed unimodal distributions for the EPSC parameters (amplitude, rise time, decay time and time to peak; Figure 3B-E). Otherwise, we excluded the recordings from analysis. We hope this explanation provides a clearer understanding of the figure's significance.

(5) Note that CTZ seems to alter p_r at some synapses.

We acknowledge that CTZ can increase release probability by blocking presynaptic K+ currents. Indeed, Ishikawa and Takahashi (2001) reported that CTZ slowed the repolarizing phase of presynaptic action potentials and the frequency of miniature EPSCs in the calyx synapses. Consistently, we observed a slight increase in the baseline EPSC amplitude, from 33.3 pA to 41.9 pA (p=0.045) following the application of 50 µM CTZ. However, given that vesicular release probability (pv) is already close to 1 at the synapse of our interest, we believe that the observed effect is more likely attributed to an increase in release sites occupancy (pocc), which would be reflected as an increase in miniature EPSC frequency in Ishikawa and Takahashi (2001). Given that PPR depends on pv rather than pocc, this increase in pocc would not critically change our conclusion that AMPA receptor desensitization is not responsible for the strong PPD.

Reference

Ishikawa, T., & Takahashi, T. (2001). Mechanisms underlying presynaptic facilitatory effect of cyclothiazide at the calyx of Held of juvenile rats. The Journal of Physiology, 533(2), 423-431.

(6) Figure 8B. The result in Figure 8C seems important, but I couldn't figure out why behaviour was not altered during the acquisition phase summarized in Figure 8B. Perhaps this could be explained more clearly for non-experts.

Little difference in freezing behavior during acquisition has been also observed when prelimbic persistent firing was optogenetically inhibited (Gilmartin, 2013). Not only CS (tone) but also other sensory inputs (visual and olfactory etc.) and the spatial context could be a cue predicting US (shock). Moreover, during the acquisition phase, the presence of the electric shock inherently induces a freezing response as a natural defensive behavior, which may obscure specific behavioral changes related to the associative learning process. Therefore, the freezing behavior during acquisition cannot be regarded as a sign for specific association of CS and US. Instead, on the next day, we specifically evaluated the CS-US association of the conditioned animals by measuring freezing behavior in response to CS in a distinct context. We explicitly documented little difference between WT and KD animals during the acquisition phase in the relevant paragraph (line 397).

Version 2

Abstract

Abstract

Short-term facilitation of recurrent excitatory synapses within the cortical network has been proposed to support persistent activity during working memory tasks, yet the underlying mechanisms remain poorly understood. We characterized short-term plasticity at the local excitatory synapses in layer 2/3 of the rat medial prefrontal cortex and studied its presynaptic mechanisms. Low-frequency stimulation induced slowly developing facilitation, whereas high-frequency stimulation initially induced strong depression followed by rapid facilitation. This non-monotonic delayed facilitation after a brief depression resulted from a high vesicular fusion probability and slow activation of Ca2+-dependent vesicle replenishment, which led to the overfilling of release sites beyond their basal occupancy. Pharmacological and gene knockdown (KD) experiments revealed that the facilitation was mediated by phospholipase C/diacylglycerol signaling and synaptotagmin 7 (Syt7). Notably, Syt7 KD abolished facilitation and slowed the refilling rate of vesicles with high fusion probability. Furthermore, Syt7 deficiency in layer 2/3 pyramidal neurons impaired the acquisition of trace fear memory and reduced c-Fos activity. In conclusion, Ca2+- and Syt7-dependent overfilling of release sites mediates synaptic facilitation at layer 2/3 recurrent excitatory synapses and contributes to temporal associative learning.

Introduction

Synapses in the mammalian central nervous system undergo activity-dependent modulation of synaptic weight during repetitive stimulation, known as short-term synaptic plasticity (STP). STP is mediated by presynaptic mechanisms that regulate the release probability (pr), size of the readily releasable pool (RRP), and kinetics of vesicle replenishment of RRP (Zucker and Regehr, 2002; Jackman and Regehr, 2017; Neher and Brose, 2018). Short-term depression has been observed at synapses with a high baseline pr, primarily because of the rapid depletion of releasable vesicles during high-frequency stimulation (HFS) (Lin et al., 2022). Conversely, vesicle depletion is less pronounced at synapses with a low baseline pr, and cumulative increases in pr during HFS are considered to mediate short-term facilitation (Rozov et al., 2001; Pan and Zucker, 2009; Aldahabi et al., 2024). Central synapses exhibit a wide spectrum of STP depending on different combinations of presynaptic factors mediating short-term depression and facilitation.

Previously, facilitation was modeled as an increase in pr after an action potential (AP) and its slow decay during the inter-spike interval (ISI), resulting in a cumulative increase in pr during HFS (Markram et al., 1998; Dittman et al., 2000). Recent studies have suggested that the number of docking (or release) sites in an active zone is limited and only partially occupied by releasable vesicles in the resting states (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Therefore, pr is determined by the vesicular fusion probability (pv) and release site occupancy (pocc). However, distinguishing the individual contributions of pv and pocc to pr is challenging (Silva et al., 2021; Neher, 2024). It is also unclear whether facilitation is mediated by activity-dependent increases in the pv, pocc, or both.

Two recent models for vesicle priming and release view the ‘priming’ as a two-step reversible process that leads to a release-ready state. The release-ready vesicles could be vesicles in a tightly docked state (TS) or those occupying specialized docking sites (DS), and are supplied from vesicles in a loosely docked state (LS) or those occupying a distinct replacement sites (RS), which are called LS/TS model (Aldahabi et al., 2024; Neher, 2024) and RS/DS model (Miki et al., 2016; Silva et al., 2021), respectively. Common to both models is the assumption that release-ready vesicles constitute only a fraction of all docked vesicles under resting conditions and that this fraction is up- and down-regulated by synaptic activity.

Synaptotagmin-7 (Syt7) was recently found to play a crucial role in facilitation at different types of central synapses (Jackman et al., 2016; Chen et al., 2017; Martinetti et al., 2022). In addition to its role in facilitation, Syt7 accelerates refilling of the RRP after depletion and mediates asynchronous release (Bacaj et al., 2013; Liu et al., 2014; Jackman et al., 2016; Tawfik et al., 2021). Syt7-dependent synaptic facilitation is interpreted as an activity-dependent increase in pv (Turecek et al., 2016; Jackman and Regehr, 2017; Turecek et al., 2017; Norman et al., 2023). However, it is unclear whether Syt7 contributes to facilitation through a Ca2+-dependent increase in pocc (i.e., overfilling). Overfilling has been proposed as a mechanism supporting facilitation at the Drosophila neuromuscular junction (Kobbersmed et al., 2020), yet whether this process also occurs at the mammalian central synapses mediated by Syt7 has not been studied.

Each of layer 2/3 (L2/3) and layer 5 (L5) of neocortex displays intralaminar excitatory synapses between pyramidal cells comprising a recurrent network (Holmgren et al., 2003; Thomson and Lamy, 2007). Short-term facilitation at recurrent excitatory synapses in a cortical network may play a crucial role in maintaining working memory (Mongillo et al., 2008; Mongillo et al., 2012; Hansel and Mato, 2013). Although L2/3 pyramidal cells (PCs) represent task-relevant information during a delay period in working memory tasks (Fujisawa et al., 2008; Ozdemir et al., 2020), most studies on local excitatory recurrent synapses in the neocortex have focused on L5 exhibiting short-term depression (Reyes and Sakmann, 1999; Hempel et al., 2000; Yoon et al., 2020). Further, the synaptic properties of excitatory synapses in L2/3 remain poorly understood.

In the present study, we found that HFS of local excitatory synapses in L2/3 of the medial prefrontal cortex (mPFC) induced initial strong depression followed by delayed facilitation, whereas low-frequency stimulation induced slow monotonous facilitation. This unique form of STP could be explained by 1) high pv of RRP vesicles at rest and throughout a train stimulation; 2) the low resting vesicular occupancy of RRP and its activity-dependent increase. These conditions led to delayed facilitation through progressive overfilling of release sites during HFS with high pv vesicles beyond the resting occupancy. Syt7 knockdown (KD) in L2/3 PCs completely abolished these synaptic enhancements by inhibiting the Ca2+-dependent increase in release site occupancy. Finally, behavioral tests for trace fear conditioning showed that Syt7 KD impaired the acquisition of trace fear memory and reduced c-Fos expression in the mPFC. Collectively, these results support that Ca2+- and Syt7-dependent overfilling of release sites mediates synaptic facilitation at L2/3 recurrent excitatory synapses and contributes to temporal associative learning.

Results

Frequency dependence of STP at local excitatory synapses in L2/3 of the prelimbic cortex

We examined STP profiles at recurrent excitatory synapses between L2/3 PCs and at synapses from L2/3-PCs to fast-spiking interneurons (FSINs) in the prelimbic area of the rat mPFC. The PCs and FSINs were identified based on their distinct intrinsic properties and morphologies (Figure 1—figure supplement 1). To stimulate presynaptic PCs, L2/3 PCs were transfected with a plasmid encoding oChIEF using in utero electroporation (IUE; Figure 1A; Lee et al., 2024). Photo-stimulation of the oChIEF-expressing cell soma at 40 Hz consistently evoked single APs with high fidelity throughout the 600-pulse trains (Figure 1—figure supplement 2A-B), which was consistent with little use-dependent inactivation of oChIEF (Lin et al., 2009; Hass and Glickfeld, 2016).

Frequency-dependence of short-term synaptic plasticity at prelimbic L2/3 excitatory synapses.

(A) In utero electroporation following injection of the plasmid (CAG-oChIEF-tdTomato) into the ventricle of an embryo (E17.5). (B) Representative images showing specific expression of oChIEF-tdTomato in L2/3 pyramidal cells after IUE. The red fluorescence of tdTomato clearly visualizes oChIEF-expressing cell bodies in L2/3 and axons in L5. Scale bar: 1 mm, 100 μm, 50 μm from left to right. (C) Recording schematic showing photostimulation of oChIEF-expressing axon fibers of transfected PCs (filled triangles) and a whole-cell recording from non-transfected PCs (empty triangles) or FSINs (empty circles). A collimated DMD-coupled LED was used to confine the area of excitation (typically 3–4 μm in diameter) to a small region (blue circle) near the soma. (D, E) Representative traces for EPSCs averaged over 10 trials at each frequency (left) and average amplitudes of baseline-normalized EPSCs (right) during 20-pulse trains at frequencies from 5 to 40 Hz at PC-PC (D; n = 12, 9, 21, 10 cells for 5, 10, 20, 40 Hz, respectively) and PC-FSIN (E; n = 8, 9, 10, 10) synapses. Each data point was normalized to the average of the first EPSC. (F) Baseline-normalized amplitudes of steady-state EPSCs (EPSCss; black symbols) and synaptic efficacy (EPSCss × f; red symbols) as a function of stimulation frequency (f) at PC–PC synapses. EPSCss was measured from the average of last 5 EPSCs from the 20-pulse trains (n = 12, 9, 21, 10). (G) Paired pulse ratio (PPR) as a function of inter-spike intervals (n = 25, 9, 25, 22, 43 for 200, 100, 50, 25, 20 ms ISI, respectively) at PC-PC synapses. (H) EPSCss and synaptic efficacy at PC-FSIN (n = 8, 9, 10, 10). (I) PPR at PC-FSIN (n = 8, 9, 10, 15, 34). Gray symbols, individual data.

Given that IUE transfected approximately 10-20% of L2/3 PCs (Figure 1A), we made a whole-cell patch on a non-transfected L2/3 PC or an FSIN (Figure 1C). To stimulate a minimal number of excitatory axon fibers, EPSCs were evoked using minimal optical stimulation (Figure 1— figure supplement 3A; 470 nm, 3-4 μm in radius and 3-5 ms in duration). EPSCs evoked by minimal stimulation were examined to determine the uniformity of the kinetic properties (Figure 1—figure supplement 3B-C). Moreover, we reassessed the kinetics of EPSCs during trains post hoc to ensure uniformity. EPSCs were evoked by 20-pulse trains at 4 different frequencies (5, 10, 20, and 40 Hz) to cover the range of firing rates of PCs observed in vivo during working memory tasks (Baeg et al., 2003). Excitatory synapses from PCs onto both other PCs and FSINs (PC–PC and PC–FSIN) showed strong (∼2-fold enhancement) and lasting (several seconds) short-term facilitation at 20 Hz and lower frequencies (Figure 1D-E). The PC-PC synapses exhibited stronger facilitation than did the PC-FSIN synapses.

Steady-state EPSCs at the PC-PC synapses were not significantly different from 5 to 20 Hz (Figure 1F). Frequency invariance at the PC-PC synapses suggests that Ca2+-dependent synaptic facilitation may counteract vesicle depletion in a frequency-dependent manner (Turecek et al., 2017; Lin et al., 2022). Despite facilitation at all frequencies, the paired pulse ratio (PPR) rapidly decreased as the ISI decreased (Figure 1G, I). At 40 Hz stimulation, both PC-PC and PC-FSIN synapses initially underwent strong depression then followed by facilitation in a few stimuli (Figure 1Dd, Ed). Such delayed facilitation after paired pulse depression (PPD) has not been previously observed in central synapses except for a slight increase in EPSC at the cerebellar synapses under artificially depolarized conditions (Pulido and Marty, 2018).

We confirmed that optically evoked EPSCs were strictly dependent on AP generation (Figure 1—figure supplement 4A), suggesting that EPSCs are evoked by optical stimulation of presynaptic axon fibers rather than by direct depolarization of presynaptic boutons. Additionally, dual patch-clamp techniques showed that the STP of optically evoked EPSCs did not differ from the electrically evoked STP of EPSCs at the PC-FSIN synapses (Figure 1—figure supplement 4B). Finally, we confirmed that AMPA receptor (AMPAR) desensitization was not responsible for PPD at 40 Hz, ruling out the possible effects of postsynaptic factors on STP (Figure 1—figure supplement 5A), and that AMPAR saturation was not significant during facilitation (Figure 1—figure supplement 5B).

Delayed facilitation results from slow activation of Ca2+-dependent vesicle replenishment at a constantly high vesicular fusion probability

The strong PPD at 40 Hz suggests a high vesicular fusion probability (pv) and likely tight coupling of the releasable vesicles to a Ca2+ source. To test this hypothesis, we examined the effects of EGTA, a slow calcium chelator, on synaptic transmission at the PC-PC synapses. The baseline EPSC amplitude (EPSC1) was not affected by incubation of the slice in the bath solution containing 50 μM EGTA-AM for longer than 20 min, supporting tight coupling. Moreover, the PPR was decreased to 0.19, and subsequent facilitation slowed but was still distinct (Figure 2A-B). These findings indicate that the initial pv should be higher than 0.81 considering vesicle recruitment during ISI. The effects of EGTA-AM were confirmed in the same cell by measuring the baseline EPSC and PPR before and after applying EGTA-AM to the bath. EPSC1 remained unchanged, while PPR was reduced by half (Figure 2—figure supplement 1).

Delayed facilitation results from slow activation of Ca2+-dependent vesicle replenishment at a constantly high vesicular fusion probability

(A) Representative traces (left) and mean baseline-normalized amplitudes (right) of EPSCs evoked by 30-pulse trains at 40 Hz in control (n = 21; black) and in the presence of 50 μM EGTA-AM (n = 14; green). (B) Mean values for the first EPSC amplitude (EPSC1, left) and PPR (right) from the experiments displayed in (A). Gray symbols, individual data. (C) Plot of rate constants for short-term facilitation (kSTF) as a function of stimulation frequency (fstim), showing a linear relationship. The linear regression line (black) is shown fitted to kSTF values, estimated from Figure 1. (D) Representative EPSC traces (left) and average of baseline-normalized EPSCs (right) evoked by 12-pulse stimulation at 5 Hz, followed by 40 Hz 7-pulse train (n = 12). Note that slowly developing facilitation was converted to rapid facilitation after strong PPD. (E) Mean values for PPR at 40 Hz. The baseline PPR was reproduced from Figure 1G and the PPR during 5 Hz train was calculated as (13th EPSC) / (12th EPSC). Gray symbols, individual data. All statistical data are represented as mean ± S.E.M.; n.s. = not significant; **, P<0.01; unpaired t-test.

Notably, the application of EGTA-AM markedly slowed facilitation at 40 Hz (Figure 2A, C). The frequency invariance of steady-state EPSCs (Figure 1F) and the slower facilitation at 40 Hz in the presence of EGTA (Figure 2A) imply a Ca2+-dependent increase in the refilling rate of the RRP and/or an increase in the pv of reluctant vesicles in light of previous studies (Hosoi et al., 2007; Liu et al., 2014; Ritzau-Jost et al., 2014; Turecek et al., 2016; Ritzau-Jost et al., 2018; Kusick et al., 2020). Moreover, facilitation was accelerated at higher stimulation frequencies. The plot of the rate constant for the increase in EPSCs (denoted as kSTF, the rate constant for the development of short-term facilitation) as a function of the stimulation frequency (fstim) revealed a linear relationship (Figure 2C). kSTF in the presence of EGTA-AM was distinctly below this relationship, supporting the Ca2+-dependent acceleration of facilitation.

kSTF was also accelerated when the stimulation frequency was increased from 5 Hz to 40 Hz (Figure 2D; 1.27 ± 0.18/s to 24.27 ± 3.44/s). Notably, the second 40-Hz train stimulation led to strong PPD, followed by accelerated facilitation (Figure 2D-E). The strong PPD suggests that the vesicular fusion probability kept high during the 5-Hz stimulation, arguing against the possibility for an increase in pv. Both two-step sequential priming models, LS/TS and RS/DS models, are suitable to interpret our data and we will refer to the different populations of vesicles as TS and LS vesicles, respectively (Taschenberger et al., 2016; Neher and Brose, 2018; Lin et al., 2022; Neher, 2024). Within this framework, vesicles are released from TS with high pv both in the resting state and during facilitation, and the delayed facilitation is seen as a result of Ca2+-dependent progressive increase in the occupancy of TS vesicles on docking sites.

Presynaptic calcium measurements at axonal boutons of L2/3 pyramidal cells

To test contribution of Ca2+ channel regulation to STP (Dobrunz et al., 1997; Jackman and Regehr, 2017), we estimated total amount of Ca2+ increments during 40 Hz AP trains at axonal boutons of L2/3 PCs in the mPFC. To this end, we measured AP-evoked Ca2+ transients (AP-CaTs) using two-dye ratiometry techniques, and estimated calcium binding ratio from AP-CaTs (Figure 2—figure supplement 2A-C). Measuring the amplitudes and decay time constants of AP-CaTs at different concentrations of Fluo-5F (150, 250 or 500 μM), and plotting them as a function of calcium binding ratio of Fluo-5F (κB), we estimated the amplitude of AP-CaTs (Aca = 1.16 μM), Ca2+ decay time constant (τc = 43 ms) in the absence exogenous Ca2+ buffer, and endogenous static Ca2+ binding ratio (κS = 96.4 or 77.3 estimated from Figure 2—figure supplement 2D-E, respectively; See Materials and Methods). The resting [Ca2+] ([Ca2+]rest) was measured as 50 nM (Figure 2—figure supplement 2F). Estimating the total Ca2+ increments for the first four AP-CaTs in a 40 Hz train according to Equation 2 (Materials and Methods), we found no evidence for inactivation or facilitation of calcium influx during a 40 Hz train (Figure 2—figure supplement 2G-I), arguing against the possibility that Ca2+ channel inactivation and facilitation contribute to strong PPD and delayed facilitation during a 40 Hz train.

Release sites have low baseline occupancy and these are increased during facilitation and post-tetanic augmentation

Given that pv was greater than 0.8, the two-fold increase in EPSC during 20 pulse train stimulations could not be explained by an increase in pv. Recent studies have shown that vesicle release sites (N) are limited in number and are partially occupied at rest (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Under such a high pv, two-fold synaptic facilitation should be attributed to an increase in pocc. The vesicle replenishment rate is accelerated during HFS in a Ca2+-dependent manner (Figure 2A-C; Hosoi et al., 2007). This may lead to the overfilling of release sites beyond their basal vesicular occupancy, which in turn may increase EPSC under high pv. To test this hypothesis, we determined the quantal parameters at the L2/3 local excitatory synapses. We first measured the quantal size from asynchronous release events after replacing extracellular Ca2+ with Sr2+, as previously described (Yoon et al., 2020) (Figure 3—figure supplement 1A-B). The mean and coefficient of variance of the quantal sizes were 24 pA and 0.25 at PC-PC synapses, respectively, and were 30 pA and 0.31 at PC-FSIN synapses (Figure 3—figure supplement 1C). Next, we performed a variance-mean (V-M) analysis of phasic EPSCs at excitatory synapses during train stimulation at 5 Hz and 40 Hz (Scheuss et al., 2002). Applying multiple-probability fluctuation analysis (MPFA) using intra-site quantal variability of 0.2 (Valera et al., 2012), the V-M plot was best fitted by the number of release sites (N) of 5.3 and the quantal size (q) of 25 pA at the PC-PC synapses (Figure 3A). The pr for EPSC1 was calculated as 0.32. We performed the same analysis for the PC-FSIN synapses and found N = 5.3, q = 35 pA, and pr for EPSC1 = 0.31 (Figure 3B). Our estimate for N was smaller but comparable to the estimate (6.9) at excitatory synapses in L2/3 of the mouse barrel cortex (Holler et al., 2021).

Low baseline occupancy of release sites and its increase during facilitation and post-tetanic augmentation

(A, B) Left, Representative EPSCs evoked by 5 and 40 Hz train stimulation (black, 5 Hz; purple, 40 Hz). Right, Variance-mean plots of EPSCs amplitude from averaged EPSCs recorded at PC-PC (A; n = 12, 17 for 5 and 40 Hz, respectively) and PC-FSIN (B; n = 8, 10) synapses. The data were fitted using multiple-probability fluctuation analysis (MPFA). Error bars are omitted for clarity. The 1st EPSC of 5 Hz train (broken line) was used to estimate the resting level of pocc. Filled circles were measured from post-tetanic augmented EPSCs (A, n = 12, 12, 9; B, n = 7, 7, 7 for 0.1, 0.2, 0.5 s IBIs, respectively). (C-E) Post-tetanic augmentation (PTA) experiments at PC-PC (top) and PC-FSIN (bottom) synapses. (C) Representative traces for EPSCs evoked by double 40 Hz train stimulations separated by 0.5 s. (D) Mean baseline-normalized amplitudes of EPSCs evoked by double 40 Hz trains at different inter-burst intervals (IBIs, 0.1, 0.2, 0.5, 1, 2, 5, 10 s). Upper, PC-PC synapse (n = 12, 12, 21, 11, 11, 16, 11 from short to long IBIs, respectively). Lower, PC-FSIN synapse (n = 10, 9, 8, 9, 9, 7, 9). (E) PTA time course. The baseline-normalized amplitudes of 1st EPSC from the 2nd train were plotted as a function of IBIs.

The pr estimate was clearly smaller than the possible minimum value for pv (0.81) estimated from the PPR data (Figure 2B). Given that pr is a product of pocc and pv (Neher, 2024), a pr smaller than pv indicates the partial occupancy of release sites in the resting state. If synaptic facilitation were mediated by the recruitment of new release sites, it would be accompanied by an increase in the variance of augmented EPSCs (Valera et al., 2012; Kobbersmed et al., 2020). However, the augmented EPSCs during facilitation followed the same parabola in the V-M plot (Figure 3A-B). Therefore, facilitation appears to be mediated by the progressive overfilling of a finite number of release sites beyond the basal occupancy. In the next section, we show that pv is close to unity based on a synaptic failure analysis (Figure 3—figure supplement 2). Assuming this, the baseline pr at the first pulse estimated from the fitted parabolic curve can be largely regarded as a baseline pocc (c.a. 30%), and the two-fold increase in EPSCs during facilitation can be attributed to an increase of up to 60% in pocc.

We investigated the time course of recovery from facilitation and post-tetanic augmentation (PTA) by applying dual-train stimulations (30 pulses at 40 Hz) separated by different inter-burst intervals (IBIs; Figure 3C-D). The initial EPSCs in the second train were enhanced by more than twofold, and followed by a strong PPD, suggesting that fusion probability stays high in the IBIs and thus synaptic transmission is carried by TS vesicles. The decay time course of PTA was characterized by two phases (initially fast and later slow; Figure 3E), which may reflect a two-phase Ca2+ decay or two distinct recovery processes that occur sequentially. The variance of potentiated EPSCs measured at IBIs of 0.1, 0.2, and 0.5 s was not different from that of EPSCs at the peak of facilitation (filled circles, Figure 3A-B). Given that the same RRP mediates both facilitation and augmentation and that post-tetanic EPSCs is carried by TS vesicles similar to the baseline EPSCs, this plot suggests that PTA, similar to short-term facilitation, is mediated by an increase in the TS vesicle occupancy.

Failure rate analysis suggests that vesicular fusion probability is close to one

The vesicle dynamics of STP has been explained using a simple refilling model (Hosoi et al., 2007; Neher and Sakaba, 2008). This model assumes that the vesicles are reversibly docked to a limited number of release sites (N) with forward and reverse rate constants (denoted as k1 and b1, respectively). Under the framework of the simple refilling model (see Materials and Methods), the rate constant for the refilling of the RRP after depletion by the first pulse is the sum of the baseline k1 and b1. This is estimated as 23/s from the plot of PPR as a function of ISIs (Figure 3—figure supplement 2A). From the baseline occupancy of 0.3 and pocc = k1 / (k1 + b1), we estimated k1 to be 6.9/s.

When paired APs were applied with an ISI of 20 ms, the failure rate at 1st pulse, P(F1), was 10.6%, and the probability of two consecutive failures, P(F1, F2), was 6.2% (Figure 3—figure supplement 2A). From the equation P(F1) = (1 – pr)N, pr was calculated as 0.312 when N = 6, comparable to pr (= 0.32) estimated from the V-M analysis (Figure 3A). We calculated P(F1, F2) under the framework of a simple refilling model, with pv and k1 set as free parameters and other parameters set according to the following relationships: P(F1) = (1 – pr)N, pr = pv ×pocc, and pocc = k1 / (k1 + b1). The calculated values for P(F1, F2) and their difference from the observed value (6.2%) are shown in the plane of k1 vs. pv in Figure 3—figure supplement 2B-C. Given that k1 was greater than 5/s, the difference between the calculated and observed values of P(F1,F2) was minimal when pv = 1 (Figure 3—figure supplement 2C).

To find optimal values for k1 and pv that best explain observed probabilities shown in Figure 3—figure supplement 2Ac, a cost function was implemented as the sum of squared errors between observed and predicted probability values for the four combinations of success and failure of 1st and 2nd EPSCs as described in Materials and Methods. The minimum of the cost function (5.63 x 10-5) was found at k1 = 5.21/s and pv = 0.999, which predicted P11 = 41.5%, P10 = 47.9%, P01 = 4.93%, and P00 = 5.67% (subscript 1 and 0 denote success and failure, respectively).

Pharmacological manipulations reveal specific molecular identities underlying vesicle loading processes

Our results suggest that the baseline occupancy of TS vesicles is low and that the activity-dependent progressive overfilling of docking sites with TS vesicles is responsible for short-term facilitation and augmentation. To elucidate the specific molecular link between synaptic enhancement and Ca2+-dependent overfilling, we examined the effects of several drugs known to affect vesicle dynamics. We first examined the roles of phospholipase C (PLC) and diacylglycerol (DAG) by applying 5 μM U73122, a PLC inhibitor, or 20 μM 1-oleoyl-2-acetyl-sn-glycerol (OAG), a DAG analog. The STP at 40 Hz and PTA at 0.5-s interval were examined before and after applying each drug to the bath. For both drugs, the facilitation and augmentation were reduced (Figure 4A-B). U73122 did not affect the basal EPSC amplitude (EPSC1) but resulted in a stronger PPD, slower facilitation, and lower augmentation (Figure 4A). These effects of U73122 suggest the involvement of Ca2+-induced PLC activation in progressive overfilling. In contrast, OAG increased EPSC1 while maintaining a pronounced PPD (Figure 4B), indicating an increase in the TS vesicle pool size. Meanwhile, OAG slowed kSTF and reduced augmentation, indicating occlusion of activity-dependent synaptic facilitation. Phorbol esters and DAG are thought to accelerate priming through Munc13 activation (Rhee et al., 2002; Taschenberger et al., 2016; Aldahabi et al., 2022); they also shorten the bridges that interlink docked vesicles with the active zone (AZ) membrane (Papantoniou et al., 2023). Collectively, these suggest that PLC- and DAG-mediated signaling play a role in the overfilling of docking sites underlying short-term facilitation and augmentation.

Pharmacological experiments reveal specific molecular mechanisms underlying vesicle loading processes

(A-C) (a) Representative EPSC traces (left) and mean baseline-normalized EPSCs (right) evoked by double 40 Hz train stimulations separated by 0.5 s inter-burst interval (IBI) in control and in the presence of 5 μM U73122 (Aa, n = 7, orange), 20 μM OAG (Ba, n = 12, cyan) or 100 μM dynasore (Ca, n = 10, blue). (b) Mean values for baseline EPSCs (EPSC1, left) and augmentation (right) from the experiments shown in corresponding a panel. (Da) Representative EPSC traces evoked by paired pulses (left) and mean values for baseline EPSC amplitude (middle) and PPR (right) before and after applying 20 μM LatB (n = 6). (Db) Representative EPSC traces (left, upper) and average of normalized EPSCs (left, lower) evoked by double 40 Hz train stimulation separated by 0.5 s in control (n = 21) and in 20 μM LatB conditions (n = 16; pink). Right, Mean values for augmentation in control and LatB conditions. Gray symbols, individual data. All statistical data are represented as mean ± S.E.M., *, P<0.05; **, P<0.01; ***, P< 0.001; unpaired or paired t-test; n.s. = not significant.

Although the acceleration of Ca2+-dependent vesicle refilling supports rapid facilitation at 40 Hz, the later part of the delayed facilitation underwent a slight depression, which was more pronounced during the second burst (Figure 3D and Figure 4). Sustained synaptic transmission is limited by the spatial availability of docking sites because it is undermined by accumulation of exocytic remainings in the presynaptic AZ during HFS (Neher, 2010; Haucke et al., 2011). To verify this, we examined the effects of dynasore (100 μM), a specific inhibitor of endocytosis (Macia et al., 2006). The treatment with dynasore did not influence EPSC1 or the early phase of facilitation, but it exacerbated depression during the later phase and attenuated PTA (Figure 4C). This suggests that late synaptic transmission during the 40-Hz train is limited by site clearance.

Given the involvement of actin polymerization in the multiple steps of vesicle docking and recycling, we examined the effects of 20 μM latrunculin B (LatB), an inhibitor of actin polymerization. The effects of LatB were complicated: (1) EPSC1 was decreased, whereas PPR was increased (Figure 4Da) and (2) facilitation and augmentation were reduced (Figure 4Db). The latter effect may result from defects in vesicle replenishment rates, either through the impaired physical movement of vesicles or disrupted site clearance (Sakaba and Neher, 2003; Lee et al., 2012; Hallermann and Silver, 2013; Miki et al., 2016), as Ca2+-dependent vesicle refilling mediates facilitation and augmentation (Figure 2 and Figure 3). Meanwhile, the former effect might be attributed to a shift in equilibrium in the priming states towards a loosely docked state (LS) at rest. It is unknown whether actin affects vesicular fusogenicity in neurons, although it facilitates multiple steps of vesicle fusion by enhancing membrane tension in neuroendocrine cells (Wu and Chan, 2022).

Short-term facilitation at both types of local excitatory synapses in L2/3 is abolished by Syt7 knockdown

We examined whether Syt7 mediates facilitation at the PC-PC and PC-FSIN synapses in L2/3 of the mPFC. To this end, oChIEF and shRNA against Syt7 (shSyt7) were co-expressed in L2/3 PCs using IUE. Syt7 KD resulted in a complete loss of facilitation at all tested stimulation frequencies (Figure 5A-B). Short-term depression and paired-pulse depression were more pronounced as the stimulation frequency increased, and the frequency invariance of steady-state EPSC was abolished, implying little contribution of Ca2+-dependent vesicle recruitment (Figure 5C-F). The lack of facilitation could not be attributed to the failure of APs in presynaptic axons, as optical stimulation elicited reliable APs in Syt7 KD PCs during the 40-Hz train (Figure 1—figure supplement 2C). Syt7 KD had little effect on the basal properties, including the peak amplitude, rise/decay time, and time-to-peak, of single AP-induced EPSCs (Figure 1—figure supplement 3D-E). Moreover, Syt7 KD did not significantly affect the intrinsic properties of neurons (Figure 5—figure supplement 1).

STF at both types of local excitatory synapses is abolished by Syt7 KD

(A, B) Representative EPSC traces (left) and mean baseline-normalized amplitudes of EPSCs (right) evoked by 20-pulse trains at 5 to 40 Hz. STP was measured at PC-PC (A; n = 10, 9, 12, 9) and PC-FSIN (B; n = 9, 8, 10, 9) synapses, in which presynaptic Syt7 transcripts were depleted (Syt7 KD). Syt7 KD pyramidal cells are indicated as black triangles on the top. For comparison, STP in WT synapses is reproduced from Figure 1 (dotted lines). Same frequency color codes were used as in Figure 1. (C) Baseline-normalized amplitudes of steady-state EPSC (EPSCss; black symbols) and synaptic efficacy (EPSCss × f; red symbols) as a function of stimulation frequency (f) at PC-PC synapses. EPSCss was defined as the average of last 5 EPSC amplitudes from 20-pulse trains. (D) PPR as a function of inter-spike intervals (n = 10, 9, 12, 18, 26). (E) EPSCss and synaptic efficacy at PC-FSIN synapses. (F) PPR at PC-FSIN synapses (n = 9, 8, 10, 12, 24). Gray symbols, individual data.

To examine whether Syt7 KD had any nonautonomous effects on postsynaptic cells, we measured the quantal size (q) from asynchronous events in the presence of Sr2+ and found that the mean quantal size remained unaltered at Syt7 KD synapses (Figure 3—figure supplement 1C). This indicated that Syt7 did not influence postsynaptic parameters, including the density and kinetic properties of AMPARs. To test the specificity of shSyt7, we examined whether co-expression of the shRNA-resistant form of Syt7 restored synaptic facilitation in both excitatory synapse types. Presynaptic overexpression of shRNA-resistant Syt7 under the CAG promoter rescued the effect of shRNA at both PC-PC and PC-FSIN synapses (Figure 5— figure supplement 2). These results underscore the crucial role of presynaptic Syt7 in mediating short-term facilitation at local excitatory synapses in L2/3 of the mPFC.

Syt7 KD synapses exhibit complementary changes in the number of release sites and their vesicle occupancy

To estimate the quantal parameters in Syt7 KD synapses, we performed V-M analysis of EPSCs evoked by train stimulation (Figure 6A-B). Fitting a parabola to the V-M plot revealed higher baseline pocc in Syt7 KD synapses than in wild-type (WT) synapses for both synapse types (KD, 0.68 and 0.42; WT, 0.32 and 0.31 for PC-PC and PC-FSIN synapses, respectively). The pocc monotonously decreased during a train stimulation, as expected from short-term depression. Moreover, the number of release sites (N) was markedly lower in Syt7 KD synapses (N = 3.5) than in WT synapses (N = 5.3; Figure 3A). Next, we examined the PTA induced by the same protocol as in Figure 3 (two 40-Hz train stimulations separated by variable IBIs) at Syt7 KD synapses (Figure 6C). The 40 Hz train-induced synaptic depression quickly recovered during IBIs, but no PTA was observed at Syt7 KD synapses, indicating its crucial role in augmentation (Figure 6D-E).

Syt7 KD synapses exhibit complementary changes in the number of release sites and their vesicle occupancy

(A, B) Left, Representative traces of EPSCs evoked by 5 and 40 Hz train stimulations (black, 5 Hz; purple, 40 Hz). Right, Variance-mean plots of EPSCs amplitude from averaged EPSCs recorded at PC-PC (A; n = 9, 15, 12, 10, 9) and PC-FSIN (B; n = 4, 14, 10, 9, 8) synapses in which presynaptic Syt7 has been knocked down. The data were fitted using MPFA and error bars are omitted for clarity. The mean 1st EPSC amplitude of 5 Hz train (vertical broken line) was used for estimation of baseline pocc. Dashed parabolas indicate MPFA fits to variance-mean plot of WT synapses reproduced from Figure 3. (C-E) Recovery experiments at PC-PC (top) and PC-FSIN (bottom) synapses in which presynaptic Syt7 was knocked-down. (C) Representative EPSCs evoked by double 40 Hz train stimulations separated by 0.5 s. (D) Mean baseline-normalized amplitudes of EPSCs evoked by double 40 Hz trains at PC-PC (n = 12, 10, 9, 9, 10, 9, 8) and PC-FSIN (n = 10, 9, 10, 11, 8, 9, 10) synapses at different interburst intervals (IBIs). (E) Recovery time course. Baseline-normalized amplitudes of 1st EPSC from the 2nd burst were plotted as a function of various IBIs. Dotted lines indicate augmented EPSCs in the WT reproduced from Figure 3.

TS vesicle recovery following depletion is accelerated by Syt7

Syt7 is recognized not only for its crucial role in facilitation, but also for mediating Ca2+- dependent replenishment of releasable vesicles (Liu et al., 2014; Tawfik et al., 2021). Given that facilitation at L2/3 excitatory synapses is driven by progressive overfilling (Figure 2 and Figure 3), the loss of facilitation at Syt7 KD synapses in the present study implies that Syt7 plays a key role in the Ca2+-dependent acceleration of refilling and overfilling of the docking sites with TS vesicles. To test this hypothesis, we examined the recovery kinetics of EPSCs after TS vesicles were depleted by 3 pulses at 40 Hz in wild-type (WT) and Syt7 KD synapses (Figure 7). Strong PPD was observed not only in the first burst, but also in the second burst. This suggested that EPSCs in the second burst were mediated by TS vesicles, similar to those in the first burst (Figure 7A, D). Therefore, EPSC1 in the second burst can be interpreted as a release from TS vesicles recovering during a given IBI. The recovery time course of EPSC1 in the second burst indicated full recovery of TS vesicles within 100 ms in both excitatory synapse types in WT rats (Figure 7C, F). However, this recovery was remarkably slower in Syt7 KD synapses, requiring 5 s for full recovery, indicating that TS vesicle recovery was greatly accelerated by Syt7. Overall, synaptic facilitation and augmentation at local excitatory synapses in L2/3 can be ascribed to progressive activity-dependent increases in the occupancy of TS vesicles (Figure 2 and Figure 3), in which Syt7 plays a key role.

Recovery of TS vesicles following depletion is accelerated by Syt7

(A-C). Recovery experiments at PC-PC synapses in WT (A) and Syt7 KD (B). Mean baseline-normalized amplitudes of EPSCs evoked by two consecutive 3-pulse 40 Hz trains in WT (n = 16, 13, 11, 11, 11, 9 from short to long IBIs, respectively) and KD (n = 13, 11, 12, 8, 10, 11) synapses at different IBIs (0.1, 0.2, 0.5, 1, 2, 5 s). Inset, representative traces of EPSCs evoked by two consecutive 3-pulse 40 Hz train stimulations separated by 0.2 s. (D-F). Recovery experiments at PC-FSIN synapses in WT (D) and Syt7 KD (E). Mean baseline-normalized amplitudes of EPSCs evoked by two consecutive 3-pulse 40 Hz trains in WT (n = 9, 11, 11, 9, 10, 13) and KD (n = 9, 12, 9, 7, 10, 9) synapses at different IBIs. Inset, representative traces of EPSCs evoked by two consecutive 3-pulse 40 Hz train stimulations separated by 0.2 s. (C, F) Recovery time course. Baseline-normalized amplitudes of 1st EPSC from the 2nd burst were plotted as a function of various IBIs.

Behavioral consequences of Syt7 KD in L2/3 PCs of the mPFC

We previously proposed that the disparity in STP between the PC-PC and PC-IN synapses results in activity-dependent changes in the ratio of synaptic weights at the PC-PC and PC-IN synapses (Jee/Jie). This may in turn have a profound influence on persistent activity in a recurrent network by biasing the E-I balance (Yoon et al., 2020). Syt7 KD not only eliminated facilitation, but also the activity-dependent increase in the Jee/Jie ratio in the L2/3 network (Figure 8—figure supplement 1). This suggests the possibility that Syt7 contributes to the persistent activity during working memory tasks in the L2/3 recurrent network. Trace fear conditioning (tFC) requires associative learning between an auditory cue (conditioned stimulus, CS) and temporally separate aversive events (unconditioned stimulus, US). The formation of trace fear memory requires the prelimbic area of the mPFC and hippocampus. mPFC activity during the trace interval is thought to temporarily hold CS information, enabling the network to associate temporally discontinuous events (i.e., temporal associative learning) (Gilmartin et al., 2013). Synaptic facilitation and augmentation are implicated in temporary memory retention in recurrent networks (Mongillo et al., 2008; Mongillo et al., 2012).

Given that Syt7 is crucial for synaptic facilitation (Figure 5) and augmentation (Figure 6) in prelimbic L2/3 recurrent excitatory synapses, we tested whether trace fear memory formation was impaired in rats with depleted Syt7 transcripts, specifically in the L2/3 PCs of the mPFC. Syt7-targeted shRNA (shSyt7) or scrambled shRNA (Scr) constructs were transfected bilaterally into the L2/3 PCs of the mPFC using IUE (Figure 8A). Both the Scr-transfected control and shSyt7-transfected rats underwent tFC training (Figure 8B). Consistent with Gilmartin et al. (2013), in which prefrontal persistent firing was optogenetically inhibited, WT and KD animals exhibited similar freezing behavior during this acquisition phase. The following day, the rats were subjected to a tone test in which trace fear memory formation was assessed from the freezing response to a tone alone in a distinct context. Compared to the control group, the shSyt7-transfected group exhibited significantly lower levels of freezing behavior in response to auditory cues during the first four trials, suggesting an impairment in trace memory formation (Figure 8C).

Behavioral effects of Syt7 deficiency in L2/3 PCs of the mPFC

(A) Top, Representative images showing bilateral expression of U6-GFP in L2/3 of PCs after IUE at E17.5. Scale bar: 1 mm, 100 μm. Bottom, Schematic of trace fear conditioning and extinction (tone test) protocol. (B) Freezing behavior of control (expressing scrambled shRNA, Scr) and Syt7 KD rats during acquisition of tFC. (C) Freezing ratio during tone tests on following days. Data are shown as average freezing during T1–T4, T5–T8, or T9–T12 (T, trials). The freezing on T1-T4 was significantly lower in KD rats suggesting that formation of trace memory was impaired in Syt7 KD rats (n = 10, 11 for Ctrl and KD, respectively; P = 0.0007, F(1, 19) = 16.43, two-way repeated measures ANOVA; P = 0.0064, 0.0211, 0.0064, 0.0064; Holm-Sidak test). (D) Representative images of c-Fos immunoreactivity in the prelimbic cortex of control or Syt7 KD rats 90 min after tFC acquisition. c-Fos (left, red, Cy5) and GAD67 (right, cyan, Cy3) were immunostained in the same brain slice expressing U6-GFP (middle; green). Scale bar, 50 μm. (E) Exemplar images of c-Fos positive neurons expressing GFP (top) or GAD67 (bottom) in control (left) or Syt7 KD rats (right). Scale bar, 50 μm. (F-H) Effects of Syt7 KD on c-Fos density (F) and percentage of c-Fos positive neurons co-labeled with GAD67 (G) or GFP (H) in L2/3 or L5 of prelimbic cortex (n = 9, 9 for Ctrl and KD, respectively). Open symbols, individual data. All statistical data are represented as mean ± S.E.M.; ****, P<0.0001; unpaired t-test.

The tone test was repeated the following day to test for extinction memory formation. Syt7 KD animals exhibited much less freezing to the CS (+1d) than did WT animals. Considering the Rescorla-Wagner model (Yau and McNally, 2023), our results suggest that the strength of CS-US associative memory was weaker in KD animals than in WT animals; thus, the acquisition of trace fear memory was impaired by Syt7 KD in the L2/3 PCs of the mPFC.

Next, we assessed the locomotor activity and anxiety levels of rats using the open field test (OFT) and elevated plus maze test (EPM). Scr controls and Syt7 KD rats showed similar exploratory patterns, displaying comparable total distance moved and time spent in the periphery of the OFT or the open/closed arms of the EPM (Figure 8—figure supplement 2). Previous studies have suggested that the acquisition of trace fear memory requires elevated trace activity in prelimbic PCs (Gilmartin et al., 2013). Given that c-Fos, an immediate early gene, is upregulated following patterned activities in neurons, its expression is considered a reliable indicator of recent neuronal activation (Fields et al., 1997; Guzowski et al., 2005). To investigate the effect of Syt7 on neuronal activity in vivo during tFC, animals were sacrificed at the time of peak c-Fos protein expression, approximately 90 min after completion of the behavioral task (Arime and Akiyama, 2017).

To elucidate the activation patterns of specific neuronal populations, we performed immunohistochemical co-labeling of c-Fos and GAD67 (a GABAergic interneuron marker) upon tFC acquisition. The density of c-Fos(+) cells and the percentage of c-Fos(+) cells among GAD67(+) cells in all prelimbic layers were significantly lower in Syt7 KD rats than in control rats (Figure 8F-G). Additionally, analysis of the proportion of active cells among transfected L2/3 PCs (c-Fos+/GFP+) revealed that Syt7 KD L2/3 PCs were significantly more active than Scr-transfected PCs (Figure 8H). This unexpected hyperactivity of Syt7 KD PCs may be attributed to the high reciprocal connectivity between PCs and FSINs in the neocortex (Holmgren et al., 2003; Otsuka and Kawaguchi, 2009), as PCs with short-term depression would transmit less input to FSINs, which in turn would exert less feedback inhibition on reciprocally connected PCs. Collectively, these findings show that the slow activation of vesicle refilling supports various forms of facilitation at excitatory synapses in PFC L2/3, which is essential for neuronal activity during temporal associative learning.

Discussion

The present study characterized STP at local excitatory synapses in the L2/3 network of the rat mPFC at physiological extracellular [Ca2+]. These synapses displayed initial strong depression and then delayed facilitation at 40 Hz, while monotonous slowly developing facilitation was observed at lower frequencies. Our study suggests that such delayed facilitation after a brief depression results from a high pv (Figure 2 and Figure 3 and Figure 3—figure supplement 2), along with Ca2+-dependent delayed acceleration of vesicle refilling and overfilling (Figure 2 and Figure 3). The two-fold increase in EPSC during 20 pulse train stimulations under the condition of such a high pv suggests a Ca2+-dependent increase in the vesicle replenishment rate during a train; this then leads to overfilling of release sites beyond their incomplete basal occupancy (Figure 3A-B). The high pv and strong PPD did not result from artificial bouton stimulation, AMPAR desensitization, or any potential distortion caused by optogenetic methods (Figure 1—figure supplements 4 and 5). Additionally, similar to short-term facilitation, PTA could be explained by an increase in the TS vesicle occupancy (Figure 3C-E), consistent with previous studies showing post-tetanic increases in the pocc and/or vesicle pool size at various types of synapses (Lee et al., 2010; Vandael et al., 2020; Tran et al., 2023; Silva et al., 2024). These findings suggest that release sites are partially occupied at rest and that a progressive overfilling of release sites with TS vesicles underlies facilitation and augmentation at intracortical L2/3 excitatory synapses.

Definition of vesicular release probability in the present and previous studies

Previous studies suggested that an increase in pv is responsible for post-tetanic augmentation (Stevens and Wesseling, 1999; Garcia-Perez and Wesseling, 2008) by observing invariance of the RRP size after tetanic stimulation. In these studies, the RRP size was estimated by hypertonic sucrose solution or as the sum of EPSCs evoked 20 Hz/60 pulses train (denoted as ‘RRPhyper’). Because reluctant vesicles (called LS vesicles) can be quickly converted to TS vesicles (16/s) and are released during a train (Lee et al., 2012), it is likely that the RRP size measured by these methods encompasses both LS and TS vesicles. In contrast, we assert high pv based on the observation of strong PPD and failure rates upon paired stimulations at ISI of 20 ms (Figure 2 and Figure 3—figure supplement 2). Given that single AP-induced vesicular release occurs from TS vesicles but not from LS vesicles, pv in the present study indicates the fusion probability of TS vesicles. From the same reasons, pocc denotes the occupancy of release sites by TS vesicles. Note that our study does not provide direct clue whether release sites are occupied by LS vesicles that are not tapped by a single AP, although an increase in the LS vesicle number may accelerate the recovery of TS vesicles. As suggested in Neher (2024), even if the number of LS plus TS vesicles are kept constant, an increase in pocc (occupancy by TS vesicles) would be interpreted as an increase in ‘vesicular release probability’ as in the previous studies (Stevens and Wesseling, 1999; Garcia-Perez and Wesseling, 2008) as long as it was measured based on RRPhyper.

Progressive overfilling of docking sites

Conventional models for facilitation assume an increase in pv after an AP of RRP vesicles, which had been in a low pv or reluctant state at rest (Dittman et al., 2000; Turecek et al., 2016). This facilitation model is unlikely to explain our results, because TS vesicles are predominantly released not only at rest but also in the facilitated state during or after train stimulation, as evidenced by the strong PPD (ISI, 25 ms) during a 5-Hz train (Figure 2D-E) and at different intervals after a 40-Hz/30 pulse trains (Figure 3D). Moreover, the PPD (ISI, 25 ms) remained pronounced during the early phase of facilitation (Figure 7), arguing against the possibility of slowly increasing the pv of reluctant vesicles. These results suggested that facilitation was mediated by an activity-dependent progressive increase in occupancy of TS vesicles (TS occupancy). Facilitation through an increase in TS occupancy has been termed frequency facilitation, differentiating it from paired pulse facilitation that relies on a residual calcium-dependent transient increase in the pv of releasable vesicles (Neher, 2024).

Assuming an increase in TS occupancy, to attain a two-fold increase in EPSC (Figure 1), the baseline occupancy of the docking sites by TS vesicles should be low, as estimated from the V-M plot (c.a. 30% in Figure 3). However, it is unclear whether the remaining docking sites that are not occupied by TS vesicles are truly vacant or are occupied by reluctant LS vesicles that are not released by the AP or AP trains. TS vesicles can be supplied by the transformation of LS to TS vesicles, both of which reside at the same docking site (LS/TS model; reviewed in Neher, 2024). Alternatively, TS vesicles can be supplied from a distinct replacement pool (RS/DS model; reviewed in Silva et al., 2021). Further studies, including ultrastructural analyses of active zones, are required to address this issue.

Given the crucial role of Syt7 in facilitation (Figure 5), the facilitation mechanisms are closely related to how Syt7 mediates synaptic facilitation. Based on the slow kinetics of Syt7, previous computational modeling studies have proposed that Syt7 may mediate short-term facilitation through an activity-dependent increase in pv (Turecek et al., 2016; Jackman and Regehr, 2017; Turecek et al., 2017; Norman et al., 2023). The present study, however, found that both facilitation and augmentation were mediated by overfilling of the release sites (Figure 2 and Figure 3). Moreover, Syt7-KD synapses displayed slower recovery of TS vesicles (Figure 7). These results suggest that Syt7 is essential for activity-dependent overfilling of docking sites with TS vesicles. Whereas the baseline EPSC was not altered (Figure 1—figure supplement 3), and complementary changes in the number of docking sites and their baseline occupancy were observed in Syt7 KD synapses (Figure 6). These results raise a possibility that Syt7 may play a role in providing additional vacant docking sites that can be overfilled during facilitation. It remains to be elucidated whether the decrease in the number of docking sites at Syt7 KD synapses is related to the subcellular localization of Syt7 to the plasma membrane of the active zone (Sugita et al., 2001; Vevea et al., 2021).

STP model in light of known Ca2+ binding kinetics of Syt7

Syt7 acts as an upstream Ca2+ sensor that facilitates activity-dependent replenishment (Liu et al., 2014; Bacaj et al., 2015; Chen et al., 2017; Tawfik et al., 2021). The present study shows that EPSCs are mediated by TS vesicles both at rest and during activity-dependent facilitation. Releasable vesicles are homogeneous, and Syt7-dependent overfilling is responsible for the activity-dependent enhancement of EPSCs. Thus, we tested whether slowly developing facilitation could be replicated within the framework of a simple refilling model based on the known Ca2+ binding kinetics of Syt7 (Brandt et al., 2012). To this end, we made following assumptions: (1) a Ca2+-dependent increase in the forward refilling rate (k1) requires a full Ca2+-bound form of Syt7, to which multiple Ca2+ ions can bind (two or three Ca2+ ions on each C2A and/or C2B domain) (Brandt et al., 2012; Chon et al., 2015; Voleti et al., 2017); (2) sequential Ca2+ binding steps are synergistic through a cooperativity factor, b; (3) Syt7 senses AP-induced local [Ca2+] similar to that estimated at the calyx of Held (a Gaussian function with a full width at half maximum value of 0.2 ms and a peak of 40 μM) (Wang et al., 2008); and (4) residual [Ca2+] follows a mono-exponential function with an amplitude of 1 μM and decay time constant of 50 ms (Jackson and Redman, 2003).

Under these assumptions and the framework of the simple refilling model, slow and delayed facilitation could be replicated by modeling k1 as k1 = k1,b + K1,max ξ [Syt7: n Ca2+], where [Syt7: n Ca2+] represents the fraction of fully Ca2+-bound Syt7 (Figure 9A-C). Syt7 was slowly activated over the course of the stimulus train, suggesting that the cooperative binding of multiple Ca2+ ions progressively increased the putative active form of Syt7 because of its slow membrane binding/unbinding kinetics. Nevertheless, the late phase of facilitation at high frequencies was predicted to be higher in the simulation than in the experiment. As expected, based on the notion that sustained release during HFS is limited by site clearance in the active zone (Figure 4C), the model/data discrepancy became more pronounced at higher frequencies. Incorporating endocytic terms into the model may mitigate late-phase discrepancies. Moreover, removal of the catalytic function of Syt7 in the model effectively accounted for the complete loss of facilitation observed in Syt7 KD (Figure 9D).

STP model in light of known Ca2+ binding kinetics of Syt7

(A-D) Left, Schematic of allosteric calcium binding to Syt7. The number of Ca2+ bound to Syt7 was denoted as # in ‘S#’ in the reaction scheme. kon = 7/μM/s, koff = 10/s. Middle, Simulated changes of k1 (black) in response to 5 Hz train of Ca2+ transients (light blue traces). The priming step of the simple refilling model was assumed to be catalyzed by full Ca2+-bound form of Syt7. Accordingly, Ca2+-dependent increase in k1 was calculated as K1,max multiplied by a fraction of full Ca2+-bound form of Syt7. We assumed that local [Ca2+]i(t) follows a Gaussian function: (1/σ √2π) exp[-(t-tp)2/ 2σ2], in which tp = 0.25 ms and σ = 0.085 ms. Right, Fits of the Syt7 model to the STP data. To fit this model to the STP data, K1,max was set to 300/s (A), 220/s (B), and 180/s (C). Cooperativity factor (b) was set to 0.35 (A), 0.2 (B), and 0.05 (C). k1 is set to be constant for Syt7 KD (D). Same frequency color codes were used as in Figure 1.

Conventional STF models involving cumulative increase in pv of reluctant vesicles

Conventional models for facilitation assume a post-AP residual Ca2+-dependent step increase in pv of RRP (Dittman et al., 2000) or reluctant vesicles (Turecek et al., 2016). Given that pv of TS vesicles is close to one in the present study, an increase in pv of TS vesicles cannot account for facilitation. The possibility for activity-dependent increase in fusion probability of LS vesicles (denoted as pv,LS) should be considered in two ways depending on whether LS and TS vesicles reside in distinct pools or in the same pool. Notably, strong PPD at short ISI implies that pv,LS is near zero at the resting state. Whereas LS vesicles do not contribute to baseline transmission, short-term facilitation (STF) may be mediated by cumulative increase in pv of LS vesicles that reside in a distinct pool. Because the increase in pv,LS during facilitation recruits new release sites (increase in N), the variance of EPSCs should become larger as stimulation frequency increases, resulting in upward deviation from a parabola in the V-M plane, as shown in recent studies (Valera et al., 2012; Kobbersmed et al., 2020). This prediction is not compatible with our results of V-M analysis (Figure 3), showing that EPSCs during STF fell on the same parabola regardless of stimulation frequencies. Therefore, it is unlikely that an increase in fusion probability of reluctant vesicles residing in a distinct release pool mediates STF in the present study. For the latter case, in which LS and TS vesicles occupy in the same release sites, it is hard to distinguish a step increase in fusion probability of LS vesicles from a conversion of LS vesicles to TS. Nevertheless, our results do not support the possibility for a gradual increase in pv,LS that occurs in parallel with STF. Strong PPD, indicative of high pv, was consistently found not only in the baseline (Figure 2 and Figure 2—figure supplement 1) but also during post-tetanic augmentation phase (Figure 3D) and even during the early development of facilitation (Figure 2D-E and Figure 7), arguing against the gradual increase in pv,LS. One may argue that STF may be mediated by a drastic step increase of pv,LS from zero to one, but it is not distinguishable from conversion of LS to TS vesicles.

Possible mechanisms underlying the depression-facilitation sequence

Cooperative binding of multiple Ca2+ ions to Syt7 predicts delayed activation and a progressive increase in vesicle refilling during train stimulations, resulting in delayed facilitation. Combining this with a high pv could replicate not only the slow facilitation at 5 Hz, but also the unique depression-facilitation sequence observed at 40 Hz (Figure 9). As simulated by Pulido and Marty (2018), delayed facilitation can be reproduced by a high pv in combination with a low baseline occupancy of replacement sites and subsequent Ca2+-dependent refilling. However, whether this model is also able to replicate slow facilitation at low frequencies remains to be tested. An alternative, but not exclusive, possibility for delayed facilitation is that the presynaptic global Ca2+ itself may undergo a slow buildup during a train due to the saturation of endogenous Ca2+-binding proteins.

Facilitation requires not only Syt7, but also activation of PLC-DAG signaling

To gain insight into the specific molecular mechanisms underlying the vesicle-loading processes, we examined the effects of pharmacological inhibitors targeting PLC and DAG (Figure 4). Both a PLC inhibitor and a DAG analog reduced facilitation and augmentation. This suggests that the progressive overfilling of docking sites with TS vesicles is supported by the PLC-DAG signaling pathway and probably downstream Munc13 (Rhee et al., 2002; Lou et al., 2008). Phorbol esters have been recently shown to increase the TS fraction of docked vesicles by enhancing vesicle priming without affecting Ca2+ transients or pv (Taschenberger et al., 2016; Aldahabi et al., 2022; Papantoniou et al., 2023; Aldahabi et al., 2024). In line with these findings, signaling involving PLC and DAG may support Ca2+-dependent overfilling of TS vesicles by enhancing molecular priming processes.

Roles of facilitation and augmentation in working memory and related tasks

Previous theoretical studies have proposed that working memory in a recurrent network can be maintained in the form of augmentation at synapses (Mongillo et al., 2008) and/or persistent activity of neurons in a memory ensemble (Mongillo et al., 2012). The latter can be supported by short-term facilitation of recurrent excitatory synapses or disparity in STP between excitatory and inhibitory synapses (Mongillo et al., 2012; Yoon et al., 2020). However, whether these short-term enhancements in synaptic efficacy contribute to working memory-related behaviors has not been examined. tFC is a temporal associative learning task requiring temporary retention of CS information during a trace period. The mPFC and hippocampus are crucial for tFC, and tFC is interfered with by a load of working memory and attention distraction, implying that the tFC shares the same neural substrate with other working memory tasks (Raybuck and Lattal, 2014). In the present study, tFC in WT and Syt7 KD animals revealed that Syt7 deficiency in the L2/3 PCs of the mPFC impaired the acquisition of trace fear memory and reduced c-Fos expression associated with tFC training (Figure 8).

The acquisition of trace fear memory was impaired by inhibition of persistent activity in mPFC during trace period (Gilmartin et al., 2013). The similar deficit observed in Syt7 KD animals is consistent with the hypothesis that STF provides bi-stable ensemble activity in a recurrent network (Mongillo et al., 2012). Nevertheless, alternative mechanisms may be responsible for the behavioral deficit. Not only recurrent network but also long-range loop between the mPFC and the mediodorsal (MD) thalamus play a critical role in maintaining persistent activity within the mPFC especially for a delay period longer than 10 s (Bolkan et al., 2017). Prefrontal L2/3 is heavily innervated by MD thalamus, and L2/3-PCs subsequently relay signals to L5 cortico-thalamic (CT) neurons (Collins et al., 2018). Given that L2/3 is an essential component of the PFC-thalamic loop, loss of STF at recurrent synapses between L2/3 PCs may lead to insufficient L2/3 inputs to L5 CT neurons and failure in the reverberant PFC-MD thalamic feedback loop. Therefore, not only L2/3 recurrent network but also its output to downstream network should be considered as a possible network mechanism underlying behavioral deficit caused by Syt7 KD L2/3.

In summary, activity-dependent synaptic facilitation and augmentation at prelimbic L2/3 recurrent excitatory synapses can be attributed to overfilling of release sites with TS vesicles. Given that the lack of short-term facilitation in L2/3 PCs impairs trace fear memory acquisition, temporary maintenance of CS information during trace period may be supported by an increase in the TS occupancy in presynaptic terminals of CS-representing neuronal ensembles.

Materials and Methods

Animals/Subjects

All experiments were carried out on Sprague-Dawley rats of either sex at p28-43. Rats were maintained in temperature-controlled rooms on a constant 12 h light/12 h dark cycle and were given access to water and food ad libitum. All animal procedures were approved and performed in accordance with the Institutional Animal Care and Use Committee at Seoul National University (SNU200522-1).

In utero electroporation (IUE)

Pregnant Sprague Dawley rats on embryonic day (E) 17.5 were deeply anesthetized with 5% (v/v) isoflurane for the duration of surgery. The uterine horns were exposed by laparotomy and wet with warm sterile phosphate-buffered saline (PBS). The plasmids (0.5-2 µg/µl) together with the 0.01% fast green dye were injected into either the left or the right lateral ventricle or bilateral ventricles of embryos through a thin glass capillary (Narishige PC-10). The embryo’s head was carefully held between the forceps-shaped electrodes of 10 mm diameter (CUY65010; NEPAGENE, Japan). Electroporation pulses (50 V for 50 ms) were delivered five times at intervals of 150 ms using a square-wave generator (ECM830; BEX, United States). The uterine horns were placed back into the abdominal cavity and abdominal muscles and skin were sutured. After the surgery, the animals were recovered under the infrared lamp. For knockdown experiments in slice electrophysiology, the shRNA constructs were co-electroporated with the pCAG-oChIEF-tdTomato, because of the consistent co-expression of the two constructs in a significant proportion of the transfected cells in vivo (Bony et al., 2013). Pups with robust fluorescence signals in the prefrontal cortex of both hemispheres were used for behavior experiments.

Preparation of vector constructs

For an optic stimulation, pCAG-oChIEF-tdTomato was transfected by IUE. Short hairpin RNA (shRNA) against Syt7 sequence (GATCTACCTGTCCTGGAAGAG) was expressed under the control of an U6 promoter of pAAV-U6-GFP vector (Cell Biolabs, #VPK-413). This shRNA was shown to effectively knockdown Syt7 in cell cultures and in vivo by previous reports from more than one research group (Bacaj et al., 2013; Li et al., 2017). Scrambled control sequence (TCGCATAGCGTATGCCGTT) was designed by shuffling the recognition region sequences and a BLAST analysis confirmed that the control shRNA sequence had no target gene. For rescue experiments, the cDNA for rat Syt7 (NM_021659) rendered resistant to the shRNA were inserted in the KD vector and their expression was driven by the CAG promoter; the vector also contained a self-cleaving P2A peptide sequence followed by EGFP, which ensures visualization of transfected cells. All the constructs above were verified by DNA sequencing. Plasmid DNA was purified and concentrated under endotoxin free conditions (Qiagen EndoFree Maxi Kit).

Slice preparation

Acute brain slices were prepared by using a VT1200S Vibratome (Leica Microsystems) after isoflurane anesthesia and decapitation. Coronal sections of the mPFC (300 μm in thickness) were cut in ice-cold artificial cerebrospinal fluid (aCSF) composed of (in mM) 125 NaCl, 3.2 KCl, 25 NaHCO3, 1.25 NaH2PO4, 20 D-glucose, 2 NaPyr, 1 MgCl2, and 1.3 CaCl2 (Yoon et al., 2020), bubbled with 95% O2 and CO2 (pH = 7.3; 310 mOsm). Slices were allowed to recover for a minimum of 30 minutes at 36 °C, and subsequently maintained at room temperature until recording. Composition of aCSF was identical throughout the preparation, incubation, and recording periods unless otherwise stated. For the recordings, the mPFC slices were continuously perfused with oxygenated aCSF at a flow rate of 1.5 ml/min, and the temperature was maintained at 31–32 °C using an in-line temperature controller (Warner Instruments).

Electrophysiology

Whole-cell patch-clamp recordings were performed from pyramidal cells and fast-spiking interneurons in layer 2/3 of the prelimbic mPFC. Neuronal subtypes were initially selected based on morphology under visual guidance by using an upright microscope equipped with infrared differential interference contrast (IR-DIC) optics (BX51WI, Olympus), and were further identified by their electrophysiological properties (see Figure 1—figure supplement 1; van Aerde and Feldmeyer, 2015; Tremblay et al., 2016). To analyze these intrinsic properties in current-clamp experiments, we measured the following parameters: (1) resting membrane potential (RMP), (2) input resistance (Rin), (3) sag ratio, and (4) the F-I curve, which plots firing frequencies (F) against the amplitude of injected currents (I). Both cell types exhibited a low sag ratio below 0.1, consistent with findings from other studies (van Aerde and Feldmeyer, 2015; Yoon et al., 2020). PCs and FSINs were easily distinguished by their unique firing frequency, Rin, spike adaptation, RMP and morphological characteristics including the shape of soma and the thickness of apical dendrites. For experiments involving oChIEF-expressing neurons in WT or Syt7 KD rats, whole-cell recordings were performed from the fluorescently labeled pyramidal neurons in layer 2/3 of the prelimbic cortex using suitable filter sets. Patch pipettes (tip resistance of 2–3 MΩ) were pulled from borosilicate glass and filled with (in mM) 130 K-gluconate, 8 KCl, 10 HEPES, 0.2 EGTA, 20 Na2Pcr, 0.3 Na2GTP, 4 MgATP. Intracellular solutions were adjusted to pH 7.25 and 300–310 mOsm. Neurons were voltage clamped at −78 mV and series resistance (Rs) were continuously monitored during the EPSCs recordings. Experiments were discarded before analysis if the Rs changed by 20% or larger deviation of baseline value. Measurements were not corrected for Rs compensation, bridge balance, or liquid junction potential. Electrophysiological data were low-pass-filtered at 1 kHz (Bessel) and acquired at 20 kHz using a Multiclamp 700B amplifier paired with Digidata 1440A digitizer and Clampex 10.2 software (Molecular Devices).

Synaptic stimulation

IUE resulted in sparse expression of channelrhodopsin (10-20%) (Figure 1A). Exploiting the sparse expression, we employed a minimal optical stimulation method to activate a single or very few excitatory synapses in layer 2/3. The optic minimal stimulation was performed by our published protocol with minor modifications (Yoon et al., 2020). This approach aimed to minimize the risk of synaptic contamination, which could arise from engaging extra terminals that were initially inactive (depolarized below threshold) but became active by temporal summation during high-frequency trains. It also helped prevent the buildup of depolarization at the synaptic terminals, which might influence the probability of release, as noted by (Jackman et al., 2014 ). To achieve this, we employed a collimated digital micromirror device (DMD)-coupled LED (Polygon400; Mightex Systems) to confine 470 nm blue light to a small area with a radius of approximately 3–8 μm (typically 3–4 μm), as measured at the focal plane of a 60× water immersion objective [numerical aperture (NA) = 1.0; LUMPlanFL, Olympus] (Figure 1—figure supplement 3A). To find the minimal stimulation area, the radius of illumination area was increased from 2 μm by 50% at each step (Figure 1—figure supplement 3A), and selected the smallest illumination area that elicited EPSCs. Setting the duration of illumination between 3 and 5 ms (typically 5 ms), photostimulation onto oChIEF-expressing cells reliably induced action potentials (APs) during 600-pulse trains (see Figure 1—figure supplement 2).

STP at various stimulus frequencies (5, 10, 20, and 40 Hz) was examined using 20-pulse trains. Trials of train stimulation were repeated with a prolonged interval of 30 s to avoid change in baseline properties. We observed no systematic changes in the amplitudes of the first EPSC of each train, nor did we find significant differences in the paired-pulse ratio (PPR) across stimulation trials in the same cell. For each stimulus frequency, approximately ten to twenty consecutive traces were collected and then averaged across trials to yield a mean EPSC trace. For recovery experiments, two consecutive trains of 3 or 30 stimuli at 40 Hz, with varying time intervals (0.1-10 s), were applied every 10 s or 60 s, respectively, at both excitatory synapses. For 600-stimulus trains at 10Hz, the trial was delivered only once for each cell to avoid possible after-effects caused by prolonged synaptic stimulation (Hirsch and Crepel, 1990). Prior to the stimulus trains, the EPSCs were normalized to the mean amplitude of the baseline. For the baseline measurements, paired pulses with various ISIs (20-200 ms) were delivered every 3 s (1/3 Hz) at least 20 times.

For all recordings, picrotoxin (PTX, 10 μM) was included in the recording solution to isolate glutamatergic responses. mPFC PC terminals were entirely glutamatergic as the application of 6-cyano-7-nitroquinoxaline-2,3-dione (CNQX, 20 μM), an AMPA receptor antagonist, completely abolished the monosynaptic events evoked by photostimulation. A low concentration of PTX was employed to minimize the relieving effects on spike suppression caused by blockade of GABAa receptors. We confirmed that no synaptic response was elicited by electrical stimulation in the presence of both 10 μM PTX and 20 μM CNQX.

In Figure 1—figure supplement 4A, 1 µM TTX (1 mM stock solution) and 0.1 mM 4-AP (500 mM stock solution) were added to the bath solution to block action potentials and restore presynaptic glutamate release, respectively (Figure 1—figure supplement 4). In Figure 1— figure supplement 5A, 50 µM of cyclothiazide, a positive allosteric modulator of the AMPA receptors, was included to minimize rapid desensitization of the postsynaptic receptors. In Figure 1—figure supplement 5B, 0.5 mM kynurenate was added to bath solution to prevent possible saturation of AMPA receptors, especially during maximal synaptic facilitation. For pharmacological manipulations of synaptic vesicle dynamics (Figure 4), drugs related to vesicle supply including U73122, OAG, LatB, EGTA-AM, and dynasore at indicated concentrations were added to the recording solution during baseline measurements or PTA/recovery experiments (IBI = 0.5 s). Chemicals were from Sigma-Aldrich, Merck, or Tocris.

Measurements of quantal size

Experiments for measuring asynchronous release were performed as previously described (Yoon et al., 2020). In brief, we replaced extracellular Ca2+ entirely with 2 mM Sr2+ to induce asynchronous release. A single stimulus was delivered at 3-second intervals over 100 repetitions, and asynchronous release events were detected during a 100-ms window starting 25 ms after stimulus onset for each trial. The same procedure was also executed under sham conditions (0% light intensity) to serve as a control, thus mitigating contamination from nonspecific spontaneous events and potential errors associated with in the detection algorithm. q was determined as the mean value from the first Gaussian function after fitting the EPSC distribution with either a single or double Gaussian function (Figure 3—figure supplement 1).

Estimation of Ca2+ kinetics at axonal boutons

Cytosolic [Ca2+] at single axonal boutons was estimated by employing the two-dye, dual-excitation wavelength technique as described by (Oheim et al., 1998). This method involves measuring the fluorescence ratio of a Ca2+-sensitive dye to that of a Ca2+-insensitive dye to correct for dye concentrations. We used Fluo-5F at concentrations of 150, 250, or 500 μM (dissociation constant, Kd = 2.3 μM) as the calcium indicator dye and Alexa Flour 555 (50 μM) as the reference dye. The fluorophores were from Invitrogen. Both dyes were co-loaded into presynaptic terminals of layer 2/3 pyramidal cells in mPFC through whole-cell patch clamp. Before data acquisition, the dyes were allowed to diffuse to terminal boutons for at least 30 minutes after the patch break-in, a duration sufficient for the equilibrium of dye concentrations between the soma and its axonal terminals.

Calcium imaging was performed by sequentially exciting Fluo-5F and AF555 at wavelengths of 473 and 559 nm, respectively, using a confocal laser-scanning microscope (FV1200; Olympus, Japan) equipped with a 60× water immersion objective (NA, 0.9; LUMPlanFl; Olympus, Tokyo, Japan). Once a single axonal bouton was identified, lines for scanning were drawn across boutons, oriented perpendicular to the axon (Figure 2—figure supplement 2A-B). After baseline fluorescence was measured, line scanning was proceeded while a single or train APs were induced by applying current pulses. Upon stimulation, green fluorescence but not red one increased along the vertical array of lines across the bouton (Figure 2—figure supplement 2B-C). Measurements were obtained under current-clamp conditions, with the resting potential regularly monitored. The scanning speed was adjusted to 600-700 Hz by controlling the line length. Line scan for measuring fluorescent transients in a bouton was repeated at least 3 times, and then averaged to enhance the signal-to-noise ratio. To mitigate photobleaching, minimal intensity of the laser power and maximal pinhole size were used. Emitted light was filtered through 490/590 or 575/675 bandpass filters before detection by photomultiplier tubes for green or red fluorescence, respectively.

Fluorescence measurements were converted to [Ca2+]i using the ratio of background-subtracted fluorescence (F = F0 - Fb) from two dyes. The fluorescence ratio (R = Fgreen/Fred) in data traces were converted to [Ca2+]i using the following equation: Calibration parameters were obtained using an in-cell calibration protocol (Helmchen, 2011), wherein the minimum (Rmin) and maximum ratio (Rmax) values were established through the use of intracellular solutions containing 10 mM EGTA or 10 mM [Ca2+], respectively.

When Ca2+ is extruded with a rate constant, γ, after an AP-induced short pulse of Ca2+ influx to a single compartment, [Ca2+]i initially rises to ACa and then undergoes a mono-exponential decay with a time constant (τCa). Under this framework, ACa and τCa depend on Ca2+ binding ratios of endo- and exogenous Ca2+ buffers (denoted as κS and κB, respectively) as following equations (Neher, 1995; Kim et al., 2003):

, where Δ[Ca2+]T is an increment in total [Ca2+] (free plus bound) elicited by an AP. Ca2+ binding ratio indicates the ratio of concentration changes in the Ca2+-buffer complex relative to free [Ca2+] change, and linearized κB is calculated from the dye concentration and the change from [Ca2+]1 to [Ca2+]2 as follows: The total dye concentration, [B]t, was taken as [Fluo-5F] in the patch pipette. With κB determined from Equation 4 and τCa from the decay of a Ca2+ transient, κS can be estimated by plotting τ versus κB and applying a linear fit to the plot, which gives the x-intercept as -(1 + κS). Once determined, κS can be used to calculate Δ[Ca2+]T.

Optimization of k1 and pv from failure rates

Let n0 = the number of docked SVs before arrival of a first AP; n1 = the number of docked SVs remaining after a first AP (i.e. n0 - n1 vesicles are released); n2 = the number of docked SVs just before arrival of a 2nd AP (ISI = 20 ms; Figure 3—figure supplement 2). If the number of release sites (N), forward and reverse rates for vesicle docking (k1 and b1, respectively), mean release site occupancy (pocc) and vesicular fusion probability upon an AP (pv) are given, we can predict P(n0, n1, n2), the probability for [n0, n1, n2], as following steps.

Probability for a synapse harbouring n0 docked SVs just before arrival of a first AP is calculated as:

Probability that n1 vesicles remain after a first AP when the initial number of docked SVs is n0 [P(n1|n0)] is calculated as:

When forward and backward rate constants are k1 and b1, respectively, probability that n2 vesicles are docked just before arrival of a 2nd AP when n1 SVs remain after a 1st AP (Δt = 20 ms) [P(n2 | n1)] is predicted as follows: Assuming that docking and undocking follow a Poisson process,

Combining probabilities calculated from above three steps, we can calculate P(n0, n1, n2):

Once P(n0, n1, n2) is given, we can predict probability for double failure (P00) and probability for the 1st success and the 2nd failure (P10) as:

Probability for the 1st failure and the 2nd success (P01) and probability for double success (P11) were calculated as:

, in which P(F1) and P(S1) are probability for the 1st failure and the 1st success, respectively, and these are calculated as: To optimize k1 and pv, we used the Matlab routine, fmincon, in which the sum of squared errors (= observed value minus predicted value) was set as a cost function to be minimized. Lower and upper bounds for k1 and pv were set as [4, 8] and [0.8, 1], respectively. Release probability (pr) was calculated from the observed 1st failure rate, P(F1), as N was assumed to be six as shown in Figure 3. Other parameters (b1, and pocc) were set according to following relationships:

Calculation of probability for double failure

Double failure probability, P00, can be more easily predicted. By the Bayes rule, , where F1 and F2 means 1st and 2nd failure. Moreover, and n0 = n1 because of the 1st failure. Thus, the probability for two consecutive failures can be calculated as: P(F1) and N are observed values. For calculation in Figure 3—figure supplement 2, k1 and pv were set free variables. Other parameters were set under the following constrains: P(F1) = (1 – pr)N, pr = pv pocc, and pocc = k1 / (k1 + b1).

Behavioral tests

Trace fear conditioning

For trace fear conditioning (tFC) (Gilmartin et al., 2013), individual rats were placed in the training chamber (30.5 x 25.4 x 30.5 cm; Coulbourn Instruments) consisting of metal grid floor connected to an electrical stimulator (H10-11R-TC-SF, Coulbourn Instruments) for delivery of a footshock. After a 5 min habituation period, rats received five pairings of a 20 s tone conditioned stimulus (CS, 2.5 kHz, 75 dB), followed by empty 20s trace period and a unconditioned stimulus (US; 1 s, 0.7 mA footshock) with a pseudorandom inter-trial interval (ITI) of 240 ± 20 s. During this training session, rats learned to associate the auditory CS with the shock US. The next day, the rats were tested for memory of CS-US trace association (called tone test). The acquisition of fear learning, often indicated by the extent of freezing to the CS, was assessed in a novel chamber (a white hexagonal enclosure) in a dimly lit room. During this tone (or retrieval) test, rats were first allowed to explore the new context for 5 min (habituation), followed by twelve 20 s CS presentations with a variable ITI of 120 ± 20 s. This entire session was repeated on the next day for assessment of extinction learning. Identities of the chambers were also determined with the presence of distinct odors; the training and testing chambers were cleaned with 70% ethanol and 1% acetic acid respectively before and after each session. For all experimental sessions, the activity of animals in the chambers was recorded at 30 frames per second using the EthoVision XT (Noldus Information Technology, Wageningen, Netherlands), and stored as a video file. The freezing behavior, defined as behavioral immobility other than respiratory movement, was assessed using an open-source video analysis pipeline ezTrack (Pennington et al., 2019). The percentage of freezing was calculated as a total duration of freezing divided by the total duration of observation.

Elevated Plus Maze

Elevated plus maze (EPM) test was performed using a plus-shaped apparatus elevated 60 cm above the ground. The maze consisted of two open arms (50 x 10 x 0.5 cm (H)), two closed arms (50 x 10 x 40 cm (H)) and a center area (10 x 10 cm). Rats were individually placed on the center area, facing an open arm, and the path of the animal was recorded with a video camera for 5 min. We analyzed the total distance traveled for 5 min and time spent in each arm. Animals were tracked offline using the open-source tracking system ezTrack (Pennington et al., 2019).

Open field test

Open field exploration test (OFT) was performed in an open field apparatus in a dimly lit room. The open field consisted of a white plastic board (1.20 m in diameter) surrounded by white plastic walls (50 cm in height). Individual rats were placed in a center zone (0.6 m in diameter), and allowed to freely explore the apparatus for 10 min. Time spent in the center zone and total distance traveled were analyzed for initial 5 min. Animals were tracked offline using the open-source tracking system ezTrack (Pennington et al., 2019).

Histology and immunohistochemistry

At the end of the behavioral experiments, rats were anesthetized using isoflurane inhalation and transcardially perfused with PBS followed by 4% paraformaldehyde (PFA, T&I, Korea). Brains were extracted and post-fixed in 4% PFA solution for 24 h. Perfused brains were sliced at 100 μm for estimation of GFP expression or 50 μm for immunohistochemistry. For analysis of c-Fos expression, rats were returned to their home cages immediately after the training session and perfused approximately 90 minutes later (Arime and Akiyama, 2017). Brains were then cryopreserved, frozen, and sectioned coronally. Immunofluorescence staining was conducted on free-floating slices using a rabbit monoclonal anti c-fos (1:1000, Cell Signaling Technology, cat# 2250s) and mouse monoclonal anti-GAD67 (1:500, Millipore, cat# mab5406). Fluorescent secondary antibodies included Cy5-conjugated goat anti-rabbit IgG (1:1000, Abcam, cat# ab97077) and Cy3-conjugated goat anti-mouse IgG (1:1000, Abcam, cat# ab97035). Brain slices were permeabilized and blocked with PBS containing 0.5% Triton X-100 and 5% normal goat serum (NGS) for 1h at room temperature. Slices were incubated overnight at 4 °C with primary antibodies, followed by 4 h incubation with secondary antibodies at room temperature. Slices were then mounted with Antifading Mounting medium (Abcam, cat# ab104139). Confocal images were acquired using a Leica TCS SP8 microscope set to the same laser intensity and acquisition parameters to compare the immunosignals in sections from Scr and Syt7 KD groups. For cell counting and colocalization, fluorescent protein-positive cells were automatically quantified with spot-detection algorithms in Imaris 9.5 (Oxford instruments).

Data analysis

Data were analyzed and presented using ClampFit (Molecular Devices), Igor Pro (Wavemetrics), MATLAB (Mathworks), and Prism (GraphPad). Data points and error bars represent the mean ± standard error of the mean. All statistical analyses were conducted using two-tailed comparisons. Sample sizes and statistical tests for each group are detailed in the figure legends.

Numerical integration of the simple refilling model

This model posits reversible docking of vesicles at finite release sites (Nmax) with forward (k1) and reverse rate constants (b1) (Hosoi et al., 2007; Neher and Sakaba, 2008). According to this scheme, dn/dt = k1 u - b1 n - pv n 8(t - tAP), where u is the number of unoccupied sites (= Nmax - n); tAP is the timing of AP firing; and δ is the Dirac delta function. The pv was assumed to be unity (Figure 3—figure supplement 2). The sum of basal k1 (= k1,b) and b1 was be estimated as 23/s from the dependence PPR on ISIs (Figure 3—figure supplement 2Ab). From the baseline occupancy of 0.3 and pocc = k1,b / (k1,b + b1), we could estimate baseline k1 and b1 as 6.9/s, and 16.1/s, respectively. Ca2+ and Syt7-dependent increase of k1 was modeled as k1 = k1,b + K1,max [Syt7: n Ca2+], where [Syt7: n Ca2+] represents the fraction of full Ca2+-bound Syt7. For deterministic simulations, we time-integrated a set of differential equations for each model using Euler methods with a time step of 1 ms. All calculations were performed on the platform of Matlab (R2022b, Mathworks, USA).

eLife Assessment

This valuable work explores how synaptic activity encodes information during memory tasks. All reviewers agree that the work is of very high quality and that the methodological approach is praiseworthy. Although the experimental data support the possibility that phospholipase diacylglycerol signaling and synaptotagmin 7 (Syt7) dynamically regulate the vesicle pool required for presynaptic release, concerns remain that the central finding of paired-pulse depression at very short intervals may be more likely due to Ca²⁺ channel inactivation rather than vesicle pool depletion. Overall, this is a solid study although the results warrant consideration of alternative interpretations.

Reviewer #1 (Public review):

Shin et al. conduct extensive electrophysiological and behavioral experiments to study the mechanisms of short-term synaptic plasticity at excitatory synapses in layer 2/3 of the rat medial prefrontal cortex. The authors interestingly find that short-term facilitation is driven by progressive overfilling of the readily releasable pool, and that this process is mediated by phospholipase C/diacylglycerol signaling and synaptotagmin-7 (Syt7). Specifically, knockdown of Syt7 not only abolishes the refilling rate of vesicles with high fusion probability, but it also impairs the acquisition of trace fear memory.

Overall, the authors offer novel insight to the field of synaptic plasticity through well-designed experiments that incorporate a range of techniques.

Comments on revisions:

The authors have adequately addressed my earlier comments and questions.

Reviewer #2 (Public review):

Summary:

Shin et al aim to identify in a very extensive piece of work a mechanism that contributes to dynamic regulation of synaptic output in the rat cortex at the second time scale. This mechanism is related to a new powerful model and is well versed to test if the pool of SV ready for fusion is dynamically scaled to adjust supply demand aspects. The methods applied are state-of-the-art and both address quantitative aspects with high signal to noise. In addition, the authors examine both excitatory output onto glutamatergic and GABAergic neurons, which provides important information on how general the observed signals are in neural networks. The results are compellingly clear and show that pool regulation may be predominantly responsible. Their results suggests that a regulation of release probability, the alternative contender for regulation, is unlikely to be involved in the observed short term plasticity behavior (but see below). Besides providing a clear analysis pof the underlying physiology, they test two molecular contenders for the observed mechanism by showing that loss of Synaptotagmin7 function and the role of the Ca dependent phospholipase activity seems critical for the short term plasticity behavior. The authors go on to test the in vivo role of the mechanism by modulating Syt7 function and examining working memory tasks as well as overall changes in network activity using immediate early gene activity. Finally, they model their data, providing strong support for their interpretation of TS pool occupancy regulation.

Strengths:

This is a very thorough study, addressing the research question from many different angles and the experimental execution is superb. The impact of the work is high, as it applies recent models of short term plasticity behavior to in vivo circuits further providing insights how synapses provide dynamic control to enable working memory related behavior through non-permanent changes in synaptic output.

Weaknesses:

While this work is carefully examined and the results are presented and discussed in a detailed manner, the reviewer is still not fully convinced that regulation of release probability is not a putative contributor to the observed behavior. No additional work is needed, but in the moment, I am not convinced that changes in release probability are not in play. One solution may be to extend the discussion of changes in rules probability as an alternative.

Fig 3. I am confused about the interpretation of the Mean Variance analysis outcome. Since the data points follow the curve during induction of short term plasticity, doesn't these suggests that release probability and not the pool size increases? Related, to measure the absolute release probability and failure rate using the optogenetic stimulation technique is not trivial as the experimental paradigm bias the experiment to a given output strength, and therefore a change in release probability cannot be excluded.

Fig. 4B interprets the phorbol ester stimulation to be the result of pool overfilling, however, phorbol ester stimulation has also been shown to increase release probability without changing the size of the readily releasable pool. The high frequency of stimulation may occlude a increased paired pulse depression in presence of OAG, that others have interpreted in mammalian synapses as an increase in release probability.

The literature on Syt7 function is still quite controversial. An observation in the literature that loss of Syt7 function in the fly synapse leads to an increase of release probability. Thus the observed changes in short term plasticity characteristics in the Syt7 KD experiments may contain a release probability component. Can the authors really exclude this possibility? Figure 5 shows for the Syt7 KD group a very prominent depression of the EPSC/IPSC with the second stimulus, particularly for the short interpulse intervals, usually a strong sign of increased release probability, as lack of pool refilling can unlikely explain the strong drop in synaptic output.

Comments on revisions:

I am satisfied with the reply of the authors and I do not have any further points of concern.

Reviewer #3 (Public review):

The results are consistent with the main claim that facilitation is caused by overfilling a readily releasable pool, but alternative interpretations continue to seem more likely, especially when the current results are taken together with previous studies. Key doubts could be resolved with a single straightforward experiment (see below).

The central issue is the interpretation of paired pulse depression that occurs when the interval between action potentials is 25 ms, but not when 50. To summarize: a similar phenomenon was observed at Schaffer collateral synapses (Dobrunz and Stevens, 1997), but was interpreted as evidence for a decrease in pv. Ca2+-channel inactivation was proposed as the mechanism, but this was not proven. The key point for evaluating the current study is that Dobrunz and Stevens specifically ruled out the kind of decrease in pocc that is the keystone premise of the current study because the depression occurred independently of whether or not the first action potential elicited exocytosis. Of course, the mechanism might be different at layer 2/3 cortical synapses. But, it seems reasonable to hope that the older hypothesis would be ruled out for the cortical synapses before concluding that the new hypothesis must be correct.

The old and new hypotheses could be distinguished from each other cleanly with a straightforward experiment. Most/maybe all central synapses strengthen a great amount when extracellular Ca2+ is increased from 1.3 to 2 mM, even when intracellular Ca2+ is buffered with EGTA. According to the authors' model, this is only possible when pv is low, and so could not occur at synapses between layer 2/3 neurons. Because of this, confirmation that increasing extracellular Ca2+ does not change synaptic strength would support the hypothesis that baseline pv is high, as the authors claim, and the support would be impressive because large changes have been seen at every other type of synapse where this has been studied (to my knowledge at least). In contrast, the Ca2+ imaging experiment that has been added to the new version of the manuscript does not address the central issue because a wide range of mechanisms could, in principle, decrease release without involving prior exocytosis or altering bulk Ca2+ signals, including: a small decrease in nano-domain Ca2+, which wouldn't be detected because nano-domains contribute a minuscule amount to the bulk signal during Ca2+-imaging; or even very fast activity-dependent undocking of synaptic vesicles, which was reported in the same Kusick et al, 2020 study that is central to the LS/TS terminology adopted by the authors.

Additional points:

(1) A new section in the Discussion (lines 458-475) suggests that previous techniques employed to show that augmentation and facilitation are caused by increases in pv did not have the resolution to distinguish between pv and pocc, but this is misleading. The confusion might be because the terminology has changed, but this is all the more reason to clarify this section. The previous evidence for increases in pv - and against increases in pocc - is as follows: The residual Ca2+ that drives augmentation decreases the latency between the onset of hypertonic solution and onset of the postsynaptic response by about 150 ms, which is large compared to the rise time of the response. The decrease indicates that the residual Ca2+ drives a decrease in the energy barrier that must be overcome before readily releasable vesicles can undergo exocytosis, which is precisely the type of mechanism that would enhance pv. In contrast, an increase in pocc could change the rise time, but not the latency. There is a small change in the rise time, but this could be caused by changes in either pv or pocc, and one of the studies (Garcia-Perez and Wesseling, 2008) showed that augmentation occluded facilitation, even at times when pocc was reduced by a factor of 3, which would seem to argue against parallel increases in both pv and pocc.

(2) Similar evidence from hypertonic stimulation indicates that Phorbol esters increase pv, but I am not aware of evidence ruling out a parallel increase in pocc.

Comments on revisions:

There are at least two straightforward ways to address the main concern.

The first would be experiments analogous to those in Dobrunz and Stevens that show that - unlike at Schaffer collateral synapses - paired pulse depression at L2/3 synapses requires neurotransmitter release. I proposed this in the first round, but realized since that a simpler and more powerful strategy would be to test directly that pv is/is-not near 1.0 in 1.2 mM Ca2+ simply by increasing to 2 mM Ca2+ (and showing that synaptic strength does-not/does change). This would be powerful because the increase in Ca2+ greatly increases synaptic strength at Schaffer collaterals by about 2.5-fold. Concerns about a confounding elevation in the basal intracellular Ca2+ concentration could be easily neutralized by pre-treating with EGTA-AM, which the authors have already done for other experiments.

Author response:

The following is the authors’ response to the previous reviews

Public Reviews:

Reviewer #1 (Public review):

Shin et al. conduct extensive electrophysiological and behavioral experiments to study the mechanisms of short-term synaptic plasticity at excitatory synapses in layer 2/3 of the rat medial prefrontal cortex. The authors interestingly find that short-term facilitation is driven by progressive overfilling of the readily releasable pool, and that this process is mediated by phospholipase C/diacylglycerol signaling and synaptotagmin-7 (Syt7). Specifically, knockdown of Syt7 not only abolishes the refilling rate of vesicles with high fusion probability, but it also impairs the acquisition of trace fear memory.

Overall, the authors offer novel insight to the field of synaptic plasticity through well-designed experiments that incorporate a range of techniques.

Comments on revisions:

The authors have adequately addressed my earlier comments and questions.

Reviewer #2 (Public review):

All the comments from Reviewer #2 are the same as her/his comments to our original manuscript. Therefore, we have already responded to all the following comments in the first revision. Here we described our additional responses to the same comments.

Summary:

Shin et al aim to identify in a very extensive piece of work a mechanism that contributes to dynamic regulation of synaptic output in the rat cortex at the second time scale. This mechanism is related to a new powerful model and is well versed to test if the pool of SV ready for fusion is dynamically scaled to adjust supply demand aspects. The methods applied are state-of-the-art and both address quantitative aspects with high signal to noise. In addition, the authors examine both excitatory output onto glutamatergic and GABAergic neurons, which provides important information on how general the observed signals are in neural networks. The results are compellingly clear and show that pool regulation may be predominantly responsible. Their results suggests that a regulation of release probability, the alternative contender for regulation, is unlikely to be involved in the observed short term plasticity behavior (but see below). Besides providing a clear analysis of the underlying physiology, they test two molecular contenders for the observed mechanism by showing that loss of Synaptotagmin7 function and the role of the Ca dependent phospholipase activity seems critical for the short term plasticity behavior. The authors go on to test the in vivo role of the mechanism by modulating Syt7 function and examining working memory tasks as well as overall changes in network activity using immediate early gene activity. Finally, they model their data, providing strong support for their interpretation of TS pool occupancy regulation.

Strengths:

This is a very thorough study, addressing the research question from many different angles and the experimental execution is superb. The impact of the work is high, as it applies recent models of short term plasticity behavior to in vivo circuits further providing insights how synapses provide dynamic control to enable working memory related behavior through non-permanent changes in synaptic output.

Weaknesses:

While this work is carefully examined and the results are presented and discussed in a detailed manner, the reviewer is still not fully convinced that regulation of release probability is not a putative contributor to the observed behavior. No additional work is needed, but in the moment, I am not convinced that changes in release probability are not in play. One solution may be to extend the discussion of changes in rules probability as an alternative.

As the Reviewer #3 suggested, we examined the dependence of EPSC amplitude on extracellular [Ca2+] ([Ca2+]o) in order to test our assertion that vesicular release probability (pv) is already saturated in resting conditions at L2/3 recurrent synapses. A three-fold increase is expected according to Dodge and Rahamimoff (1967), if resting pv has enough room to increase, when [Ca2+]o is elevated from 1.3 to 2.5 mM. We found an increase in the baseline EPSC amplitude only by 23%, and this change was not statistically significant, supporting our assertion.

Fig 3. I am confused about the interpretation of the Mean Variance analysis outcome. Since the data points follow the curve during induction of short term plasticity, doesn't these suggests that release probability and not the pool size increases?

We separated the conventional release probability into a multiplication of pv and pocc, in which pv = probability of TS vesicles and pocc = occupancy of release sites by TS vesicles. In this regard, the abscissa of V-M plot represents the conventional release probability. Because pv is close to unity, we interpreted a change along the abscissa as a change of pocc.

Related, to measure the absolute release probability and failure rate using the optogenetic stimulation technique is not trivial as the experimental paradigm bias the experiment to a given output strength, and therefore a change in release probability cannot be excluded.

We agree to this concern. Because EPSC data were obtained by optogenetic stimulation, it cannot be ruled out a possibility that optogenetic stimulation biased the release probability. Although we found that STP obtained by dual patch experiment was not different from that by optogenetic stimulation, it needs to confirm our conclusion using dual patch or other methods.

Fig. 4B interprets the phorbol ester stimulation to be the result of pool overfilling, however, phorbol ester stimulation has also been shown to increase release probability without changing the size of the readily releasable pool. The high frequency of stimulation may occlude an increased paired pulse depression in presence of OAG, that others have interpreted in mammalian synapses as an increase in release probability.

Provided that pv of TS vesicles is very high, the OAG-induced increase in EPSC1 and low STF and PTA are consistent with higher baseline pocc in PDBu conditions, while the number of docking sites is limited. It should be noted that previous PDBu-induced invariance of the RRP size is based on measuring the RRP size using hypertonic solution (Basu et al., 2007). Given that this sucrose method releases not only TS but also LS vesicles, the sucrose-based RRP size may not be affected by PDBu or OAG at L2/3 synapses too. Therefore, PDBu or OAG-induced increase in pocc (proportion of TS vesicles over LS+TS vesicles) would result in an increase in release probability without a change in the RRP size.

The literature on Syt7 function is still quite controversial. An observation in the literature that loss of Syt7 function in the fly synapse leads to an increase of release probability. Thus the observed changes in short term plasticity characteristics in the Syt7 KD experiments may contain a release probability component. Can the authors really exclude this possibility? Figure 5 shows for the Syt7 KD group a very prominent depression of the EPSC/IPSC with the second stimulus, particularly for the short interpulse intervals, usually a strong sign of increased release probability, as lack of pool refilling can unlikely explain the strong drop in synaptic output.

Comments on revisions:

I am satisfied with the reply of the authors and I do not have any further points of concern.

Reviewer #3 (Public review):

The results are consistent with the main claim that facilitation is caused by overfilling a readily releasable pool, but alternative interpretations continue to seem more likely, especially when the current results are taken together with previous studies. Key doubts could be resolved with a single straightforward experiment (see below).

The central issue is the interpretation of paired pulse depression that occurs when the interval between action potentials is 25 ms, but not when 50. To summarize: a similar phenomenon was observed at Schaffer collateral synapses (Dobrunz and Stevens, 1997), but was interpreted as evidence for a decrease in pv. Ca2+-channel inactivation was proposed as the mechanism, but this was not proven. The key point for evaluating the current study is that Dobrunz and Stevens specifically ruled out the kind of decrease in pocc that is the keystone premise of the current study because the depression occurred independently of whether or not the first action potential elicited exocytosis. Of course, the mechanism might be different at layer 2/3 cortical synapses. But, it seems reasonable to hope that the older hypothesis would be ruled out for the cortical synapses before concluding that the new hypothesis must be correct.

The old and new hypotheses could be distinguished from each other cleanly with a straightforward experiment. Most/maybe all central synapses strengthen a great amount when extracellular Ca2+ is increased from 1.3 to 2 mM, even when intracellular Ca2+ is buffered with EGTA. According to the authors' model, this is only possible when pv is low, and so could not occur at synapses between layer 2/3 neurons. Because of this, confirmation that increasing extracellular Ca2+ does not change synaptic strength would support the hypothesis that baseline pv is high, as the authors claim, and the support would be impressive because large changes have been seen at every other type of synapse where this has been studied (to my knowledge at least). In contrast, the Ca2+ imaging experiment that has been added to the new version of the manuscript does not address the central issue because a wide range of mechanisms could, in principle, decrease release without involving prior exocytosis or altering bulk Ca2+ signals, including: a small decrease in nano-domain Ca2+, which wouldn't be detected because nano-domains contribute a minuscule amount to the bulk signal during Ca2+-imaging; or even very fast activity-dependent undocking of synaptic vesicles, which was reported in the same Kusick et al, 2020 study that is central to the LS/TS terminology adopted by the authors.

Additional points:

(1) A new section in the Discussion (lines 458-475) suggests that previous techniques employed to show that augmentation and facilitation are caused by increases in pv did not have the resolution to distinguish between pv and pocc, but this is misleading. The confusion might be because the terminology has changed, but this is all the more reason to clarify this section. The previous evidence for increases in pv - and against increases in pocc - is as follows: The residual Ca2+ that drives augmentation decreases the latency between the onset of hypertonic solution and onset of the postsynaptic response by about 150 ms, which is large compared to the rise time of the response. The decrease indicates that the residual Ca2+ drives a decrease in the energy barrier that must be overcome before readily releasable vesicles can undergo exocytosis, which is precisely the type of mechanism that would enhance pv. In contrast, an increase in pocc could change the rise time, but not the latency. There is a small change in the rise time, but this could be caused by changes in either pv or pocc, and one of the studies (Garcia-Perez and Wesseling, 2008) showed that augmentation occluded facilitation, even at times when pocc was reduced by a factor of 3, which would seem to argue against parallel increases in both pv and pocc.

We greatly appreciate for pointing out our mis-understanding. We acknowledge that the post-tetanic acceleration of the latency in the hypertonicity-induced vesicle release may reflect a decrease in the activation energy barrier (ΔEa) for vesicle fusion resulting in an increase in fusion probability of TS vesicles (Stevens and Wesseling, 1999; Garcia-Perez and Wesseling, 2008). We agree that such latency changes are not easily explained by increases in pocc alone. Indeed, Taschenberger et al (2016) concluded that PTP is similar to the PDBu-induced increase in baseline EPSCs. Subsequently, Lin et al (2025) estimated PDBu-induced changes of TS vesicle pool size and p_fusion of TS vesicles (these correspond to pocc and pv in this study, respectively), and found that PDBu increases majorly the former (2 folds) and minorly the latter (1.3 folds). Although it has not been directly tested, it is possible that PTP increases pv. Accordingly, we corrected the first statement of the paragraph, and mentioned the possibility for a post-tetanic increase in pv of TS vesicles.

It should be noted, however, it is still puzzling what is represented by the acceleration of the latency in the hypertonicity-induced vesicle release. Schotten et al (2015) simulated how vesicle release rate is affected by reducing ΔEa for vesicle fusion. They found that a reduction of ΔEa resulted in increases in the peak amplitude and shorter time-to-peak of vesicle fusion, but did not accelerate the latency. Therefore, it remains to be clarified whether shorter latency can be regarded as lower activation barrier. Moreover, the sucrose-induced release rate is comparable with the vesicle recruitment rate (1-2/s; Neher, Neuron, 2008). This slowness of sucrose-induced vesicle release rate makes it difficult to distinguish the vesicle fusion rate from their priming rate.

(2) Similar evidence from hypertonic stimulation indicates that Phorbol esters increase pv, but I am not aware of evidence ruling out a parallel increase in pocc.

As noted above, none of known mechanisms can clearly explain the PDBu-induced shorter latency to hypertonicity-induced vesicle fusion (Schotten et al, 2015). Even if shorter latency reflects higher pv, it does not rule out a concurrent change in pocc. Supporting this notion, Lin et al. (2025) showed in the framework of the two state vesicle fusion model that PDBu application leads to a substantial increase in the number of TS vesicles (vesicles having high fusion propensity), with a moderate change in fusion probability (pfusion). In light of previous observation that high tonicity (500 or 1000 mOsm) did not alter the RRP size (Basu et al., 2007), the results of Lin et al. (2025) can be interpreted as an increase of ‘pocc’ in terms of the present study.

Reference:

Schotten et al. (2015). Additive effects on the energy barrier for synaptic vesicle fusion cause supralinear effects on the vesicle fusion rate. eLife 4:e05531.

Lin, K.-H., Ranjan, M., Lipstein, N., Brose, N., Neher, E., & Taschenberger, H. (2025). Number and relative abundance of synaptic vesicles in functionally distinct priming states determine synaptic strength and short-term plasticity. J. Physiology.

Comments on revisions:

There are at least two straightforward ways to address the main concern.

The first would be experiments analogous to those in Dobrunz and Stevens that show that - unlike at Schaffer collateral synapses - paired pulse depression at L2/3 synapses requires neurotransmitter release. I proposed this in the first round, but realized since that a simpler and more powerful strategy would be to test directly that pv is/is-not near 1.0 in 1.2 mM Ca2+ simply by increasing to 2 mM Ca2+ (and showing that synaptic strength does-not/does change). This would be powerful because the increase in Ca2+ greatly increases synaptic strength at Schaffer collaterals by about 2.5-fold. Concerns about a confounding elevation in the basal intracellular Ca2+ concentration could be easily neutralized by pre-treating with EGTA-AM, which the authors have already done for other experiments.

We thank to Reviewer #3 for suggesting an experiment for testing our assertion that the vesicular release probability (pv) is very high at layer 2/3 recurrent excitatory synapses. As the Reviewer recommended, we assessed EPSC changes induced by an increase in extracellular calcium concentration ([Ca2+]o). The results are added as Figure 3—figure supplement 3 to the revised manuscript.

Dodge and Rahamimoff (1967) discovered a fourth-power relationship between end-plate potential (EPP) and [Ca2+]o at a neuromuscular junction. More specifically they found

EPP amplitude µ ([Ca2+]o / (1 + [Ca2+]o /1.1 mM + [Ma2+]o /2.97 mM))4.

This equation nicely predicts the effects of high external calcium on EPSC amplitudes observed at the calyx synapses: a 2.6-fold increase of EPSC by changing [Ca2+]o from 1.25 to 2 mM (Thanawala and Regehr, 2013; predicted as 2.57); a 2.36-fold increase by changing [Ca2+] from 1.5 to 2 mM (Lin and Taschenberger, 2025; predicted as 2.16). In the framework of two-step priming model, Lin et al. (2015) estimated a 1.9-fold increase (from 0.22 to 0.42) in pv of TS vesicles and a 1.23-fold increase in the number of TS vesicles. It is clear that the increase in pv would be possible only if pv is not saturated, while the increase in the number of TS vesicles is still possible regardless of baseline pv of TS vesicles.

The Dodge and Rahamimoff’s equation predicts a 3.24-fold increase in baseline EPSC amplitude by elevating [Ca Ca2+]o from 1.3 mM to 2.5 mM at L2/3 synapses. Contrary to this prediction, our recordings revealed a 1.23 fold increase in baseline EPSC amplitude, and this change was not statistically significant.

Given the steep dependence of vesicle release on [Ca2+]o, this minimal increase strongly suggests that pv at L2/3 recurrent synapses is already near maximal at rest, limiting the dynamic range for further enhancement through increased calcium influx. Accordingly, we observed a small but statistically significant decrease in the paired-pulse ratio (PPR) at higher [Ca2+]o. Although this reduction in PPR might be indicative of increased pv, it is more consistent with a slight increase in pocc rather than a substantive increase in pv under the context of very high pv. Accordingly, Lin et al. (2025) recently estimated an increase in the TS vesicle subpool size as 1.23-fold by elevating [Ca2+]o under the framework of the two-step vesicle priming mode. Taken together, these findings suggest that an increase in the number of TS vesicles or pocc may contribute to both an increase in baseline EPSC amplitudes and a decrease in PPR.

Overall, our central claim that baseline pv is near maximal at L2/3 recurrent synapses is supported by 1) high baseline PPR; 2) insensitivity to EGTA-AM; 3) high double failure rate; 4) insensitivity to elevating [Ca2+]o. These data are difficult to reconcile with a model in which facilitation is mediated by Ca2+-dependent increases in pv. Instead, our results support a mechanism in which facilitation arises from changes in release site occupancy.

References

Dodge, F.A., & Rahamimoff, R. (1967). Co-operative action of calcium ions in transmitter release at the neuromuscular junction. J Physiol, 193(2), 419–432.

Thanawala, M.S., & Regehr, W.G. (2013). Presynaptic calcium influx controls neurotransmitter release in part by regulating the effective size of the readily releasable pool. J Neurosci, 33(11), 4625–4633.

Lin, K.-H., Ranjan, M., Lipstein, N., Brose, N., Neher, E., & Taschenberger, H. (2025). Number and relative abundance of synaptic vesicles in functionally distinct priming states determine synaptic strength and short-term plasticity. J. Physiology.

Neher E, Sakaba T (2008) Multiple Roles of Calcium Ions in the Regulation of Neurotransmitter Release. Neuron 59:861-872.

Version 3

Abstract

Abstract

Short-term facilitation of recurrent excitatory synapses within the cortical network has been proposed to support persistent activity during working memory tasks, yet the underlying mechanisms remain poorly understood. We characterized short-term plasticity at the local excitatory synapses in layer 2/3 of the rat medial prefrontal cortex and studied its presynaptic mechanisms. Low-frequency stimulation induced slowly developing facilitation, whereas high-frequency stimulation initially induced strong depression followed by rapid facilitation. This non-monotonic delayed facilitation after a brief depression resulted from a high vesicular fusion probability and slow activation of Ca2+-dependent vesicle replenishment, which led to the overfilling of release sites beyond their basal occupancy. Pharmacological and gene knockdown (KD) experiments revealed that the facilitation was mediated by phospholipase C/diacylglycerol signaling and synaptotagmin 7 (Syt7). Notably, Syt7 KD abolished facilitation and slowed the refilling rate of vesicles with high fusion probability. Furthermore, Syt7 deficiency in layer 2/3 pyramidal neurons impaired the acquisition of trace fear memory and reduced c-Fos activity. In conclusion, Ca2+- and Syt7-dependent overfilling of release sites mediates synaptic facilitation at layer 2/3 recurrent excitatory synapses and contributes to temporal associative learning.

Introduction

Synapses in the mammalian central nervous system undergo activity-dependent modulation of synaptic weight during repetitive stimulation, known as short-term synaptic plasticity (STP). STP is mediated by presynaptic mechanisms that regulate the release probability (pr), size of the readily releasable pool (RRP), and kinetics of vesicle replenishment of RRP (Zucker and Regehr, 2002; Jackman and Regehr, 2017; Neher and Brose, 2018). Short-term depression has been observed at synapses with a high baseline pr, primarily because of the rapid depletion of releasable vesicles during high-frequency stimulation (HFS) (Lin et al., 2022). Conversely, vesicle depletion is less pronounced at synapses with a low baseline pr, and cumulative increases in pr during HFS are considered to mediate short-term facilitation (Rozov et al., 2001; Pan and Zucker, 2009; Aldahabi et al., 2024). Central synapses exhibit a wide spectrum of STP depending on different combinations of presynaptic factors mediating short-term depression and facilitation.

Previously, facilitation was modeled as an increase in pr after an action potential (AP) and its slow decay during the inter-spike interval (ISI), resulting in a cumulative increase in pr during HFS (Markram et al., 1998; Dittman et al., 2000). Recent studies have suggested that the number of docking (or release) sites in an active zone is limited and only partially occupied by releasable vesicles in the resting states (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Therefore, pr is determined by the vesicular fusion probability (pv) and release site occupancy (pocc). However, distinguishing the individual contributions of pv and pocc to pr is challenging (Silva et al., 2021; Neher, 2024). It is also unclear whether facilitation is mediated by activity-dependent increases in the pv, pocc, or both.

Two recent models for vesicle priming and release view the ‘priming’ as a two-step reversible process that leads to a release-ready state. The release-ready vesicles could be vesicles in a tightly docked state (TS) or those occupying specialized docking sites (DS), and are supplied from vesicles in a loosely docked state (LS) or those occupying a distinct replacement sites (RS), which are called LS/TS model (Aldahabi et al., 2024; Neher, 2024) and RS/DS model (Miki et al., 2016; Silva et al., 2021), respectively. Common to both models is the assumption that release-ready vesicles constitute only a fraction of all docked vesicles under resting conditions and that this fraction is up- and down-regulated by synaptic activity.

Synaptotagmin-7 (Syt7) was recently found to play a crucial role in facilitation at different types of central synapses (Jackman et al., 2016; Chen et al., 2017; Martinetti et al., 2022). In addition to its role in facilitation, Syt7 accelerates refilling of the RRP after depletion and mediates asynchronous release (Bacaj et al., 2013; Liu et al., 2014; Jackman et al., 2016; Tawfik et al., 2021). Syt7-dependent synaptic facilitation is interpreted as an activity- dependent increase in pv (Turecek et al., 2016; Jackman and Regehr, 2017; Turecek et al., 2017; Norman et al., 2023). However, it is unclear whether Syt7 contributes to facilitation through a Ca2+-dependent increase in pocc (i.e., overfilling). Overfilling has been proposed as a mechanism supporting facilitation at the Drosophila neuromuscular junction (Kobbersmed et al., 2020), yet whether this process also occurs at the mammalian central synapses mediated by Syt7 has not been studied.

Each of layer 2/3 (L2/3) and layer 5 (L5) of neocortex displays intralaminar excitatory synapses between pyramidal cells comprising a recurrent network (Holmgren et al., 2003; Thomson and Lamy, 2007). Short-term facilitation at recurrent excitatory synapses in a cortical network may play a crucial role in maintaining working memory (Mongillo et al., 2008; Mongillo et al., 2012; Hansel and Mato, 2013). Although L2/3 pyramidal cells (PCs) represent task-relevant information during a delay period in working memory tasks (Fujisawa et al., 2008; Ozdemir et al., 2020), most studies on local excitatory recurrent synapses in the neocortex have focused on L5 exhibiting short-term depression (Reyes and Sakmann, 1999; Hempel et al., 2000; Yoon et al., 2020). Further, the synaptic properties of excitatory synapses in L2/3 remain poorly understood.

In the present study, we found that HFS of local excitatory synapses in L2/3 of the medial prefrontal cortex (mPFC) induced initial strong depression followed by delayed facilitation, whereas low-frequency stimulation induced slow monotonous facilitation. This unique form of STP could be explained by 1) high pv of RRP vesicles at rest and throughout a train stimulation; 2) the low resting vesicular occupancy of RRP and its activity-dependent increase. These conditions led to delayed facilitation through progressive overfilling of release sites during HFS with high pv vesicles beyond the resting occupancy. Syt7 knockdown (KD) in L2/3 PCs completely abolished these synaptic enhancements by inhibiting the Ca2+-dependent increase in release site occupancy. Finally, behavioral tests for trace fear conditioning showed that Syt7 KD impaired the acquisition of trace fear memory and reduced c-Fos expression in the mPFC. Collectively, these results support that Ca2+- and Syt7-dependent overfilling of release sites mediates synaptic facilitation at L2/3 recurrent excitatory synapses and contributes to temporal associative learning.

Results

Frequency dependence of STP at local excitatory synapses in L2/3 of the prelimbic cortex

We examined STP profiles at recurrent excitatory synapses between L2/3 PCs and at synapses from L2/3-PCs to fast-spiking interneurons (FSINs) in the prelimbic area of the rat mPFC. The PCs and FSINs were identified based on their distinct intrinsic properties and morphologies (Figure 1—figure supplement 1). To stimulate presynaptic PCs, L2/3 PCs were transfected with a plasmid encoding oChIEF using in utero electroporation (IUE; Figure 1A; Lee et al., 2024). Photo-stimulation of the oChIEF-expressing cell soma at 40 Hz consistently evoked single APs with high fidelity throughout the 600-pulse trains (Figure 1—figure supplement 2A-B), which was consistent with little use-dependent inactivation of oChIEF (Lin et al., 2009; Hass and Glickfeld, 2016).

Frequency-dependence of short-term synaptic plasticity at prelimbic L2/3 excitatory synapses.

(A) In utero electroporation following injection of the plasmid (CAG-oChIEF-tdTomato) into the ventricle of an embryo (E17.5). (B) Representative images showing specific expression of oChIEF-tdTomato in L2/3 pyramidal cells after IUE. The red fluorescence of tdTomato clearly visualizes oChIEF-expressing cell bodies in L2/3 and axons in L5. Scale bar: 1 mm, 100 μm, 50 μm from left to right. (C) Recording schematic showing photostimulation of oChIEF-expressing axon fibers of transfected PCs (filled triangles) and a whole-cell recording from non- transfected PCs (empty triangles) or FSINs (empty circles). A collimated DMD-coupled LED was used to confine the area of excitation (typically 3–4 μm in diameter) to a small region (blue circle) near the soma. (D, E) Representative traces for EPSCs averaged over 10 trials at each frequency (left) and average amplitudes of baseline-normalized EPSCs (right) during 20-pulse trains at frequencies from 5 to 40 Hz at PC-PC (D; n = 12, 9, 21, 10 cells for 5, 10, 20, 40 Hz, respectively) and PC-FSIN (E; n = 8, 9, 10, 10) synapses. Each data point was normalized to the average of the first EPSC. (F) Baseline-normalized amplitudes of steady-state EPSCs (EPSCss; black symbols) and synaptic efficacy (EPSCss × f; red symbols) as a function of stimulation frequency (f) at PC–PC synapses. EPSCss was measured from the average of last 5 EPSCs from the 20-pulse trains (n = 12, 9, 21, 10). (G) Paired pulse ratio (PPR) as a function of inter-spike intervals (n = 25, 9, 25, 22, 43 for 200, 100, 50, 25, 20 ms ISI, respectively) at PC-PC synapses. (H) EPSCss and synaptic efficacy at PC-FSIN (n = 8, 9, 10, 10). (I) PPR at PC-FSIN (n = 8, 9, 10, 15, 34). Gray symbols, individual data.

Given that IUE transfected approximately 10-20% of L2/3 PCs (Figure 1A), we made a whole- cell patch on a non-transfected L2/3 PC or an FSIN (Figure 1C). To stimulate a minimal number of excitatory axon fibers, EPSCs were evoked using minimal optical stimulation (Figure 1— figure supplement 3A; 470 nm, 3-4 μm in radius and 3-5 ms in duration). EPSCs evoked by minimal stimulation were examined to determine the uniformity of the kinetic properties (Figure 1—figure supplement 3B-C). Moreover, we reassessed the kinetics of EPSCs during trains post hoc to ensure uniformity. EPSCs were evoked by 20-pulse trains at 4 different frequencies (5, 10, 20, and 40 Hz) to cover the range of firing rates of PCs observed in vivo during working memory tasks (Baeg et al., 2003). Excitatory synapses from PCs onto both other PCs and FSINs (PC–PC and PC–FSIN) showed strong (∼2-fold enhancement) and lasting (several seconds) short-term facilitation at 20 Hz and lower frequencies (Figure 1D-E). The PC-PC synapses exhibited stronger facilitation than did the PC-FSIN synapses.

Steady-state EPSCs at the PC-PC synapses were not significantly different from 5 to 20 Hz (Figure 1F). Frequency invariance at the PC-PC synapses suggests that Ca2+-dependent synaptic facilitation may counteract vesicle depletion in a frequency-dependent manner (Turecek et al., 2017; Lin et al., 2022). Despite facilitation at all frequencies, the paired pulse ratio (PPR) rapidly decreased as the ISI decreased (Figure 1G, I). At 40 Hz stimulation, both PC-PC and PC-FSIN synapses initially underwent strong depression then followed by facilitation in a few stimuli (Figure 1Dd, Ed). Such delayed facilitation after paired pulse depression (PPD) has not been previously observed in central synapses except for a slight increase in EPSC at the cerebellar synapses under artificially depolarized conditions (Pulido and Marty, 2018).

We confirmed that optically evoked EPSCs were strictly dependent on AP generation (Figure 1—figure supplement 4A), suggesting that EPSCs are evoked by optical stimulation of presynaptic axon fibers rather than by direct depolarization of presynaptic boutons. Additionally, dual patch-clamp techniques showed that the STP of optically evoked EPSCs did not differ from the electrically evoked STP of EPSCs at the PC-FSIN synapses (Figure 1—figure supplement 4B). Finally, we confirmed that AMPA receptor (AMPAR) desensitization was not responsible for PPD at 40 Hz, ruling out the possible effects of postsynaptic factors on STP (Figure 1—figure supplement 5A), and that AMPAR saturation was not significant during facilitation (Figure 1—figure supplement 5B).

Delayed facilitation results from slow activation of Ca2+-dependent vesicle replenishment at a constantly high vesicular fusion probability

The strong PPD at 40 Hz suggests a high vesicular fusion probability (pv) and likely tight coupling of the releasable vesicles to a Ca2+ source. To test this hypothesis, we examined the effects of EGTA, a slow calcium chelator, on synaptic transmission at the PC-PC synapses. The baseline EPSC amplitude (EPSC1) was not affected by incubation of the slice in the bath solution containing 50 μM EGTA-AM for longer than 20 min, supporting tight coupling. Moreover, the PPR was decreased to 0.19, and subsequent facilitation slowed but was still distinct (Figure 2A-B). These findings indicate that the initial pv should be higher than 0.81 considering vesicle recruitment during ISI. The effects of EGTA-AM were confirmed in the same cell by measuring the baseline EPSC and PPR before and after applying EGTA-AM to the bath. EPSC1 remained unchanged, while PPR was reduced by half (Figure 2—figure supplement 1).

Delayed facilitation results from slow activation of Ca2+-dependent vesicle replenishment at a constantly high vesicular fusion probability

(A) Representative traces (left) and mean baseline-normalized amplitudes (right) of EPSCs evoked by 30-pulse trains at 40 Hz in control (n = 21; black) and in the presence of 50 μM EGTA-AM (n = 14; green). (B) Mean values for the first EPSC amplitude (EPSC1, left) and PPR (right) from the experiments displayed in (A). Gray symbols, individual data. (C) Plot of rate constants for short-term facilitation (kSTF) as a function of stimulation frequency (fstim), showing a linear relationship. The linear regression line (black) is shown fitted to kSTF values, estimated from Figure 1. (D) Representative EPSC traces (left) and average of baseline- normalized EPSCs (right) evoked by 12-pulse stimulation at 5 Hz, followed by 40 Hz 7-pulse train (n = 12). Note that slowly developing facilitation was converted to rapid facilitation after strong PPD. (E) Mean values for PPR at 40 Hz. The baseline PPR was reproduced from Figure 1G and the PPR during 5 Hz train was calculated as (13th EPSC) / (12th EPSC). Gray symbols, individual data. All statistical data are represented as mean ± S.E.M.; n.s. = not significant; **, P<0.01; unpaired t-test.

Notably, the application of EGTA-AM markedly slowed facilitation at 40 Hz (Figure 2A, C). The frequency invariance of steady-state EPSCs (Figure 1F) and the slower facilitation at 40 Hz in the presence of EGTA (Figure 2A) imply a Ca2+-dependent increase in the refilling rate of the RRP and/or an increase in the pv of reluctant vesicles in light of previous studies (Hosoi et al., 2007; Liu et al., 2014; Ritzau-Jost et al., 2014; Turecek et al., 2016; Ritzau-Jost et al., 2018; Kusick et al., 2020). Moreover, facilitation was accelerated at higher stimulation frequencies. The plot of the rate constant for the increase in EPSCs (denoted as kSTF, the rate constant for the development of short-term facilitation) as a function of the stimulation frequency (fstim) revealed a linear relationship (Figure 2C). kSTF in the presence of EGTA-AM was distinctly below this relationship, supporting the Ca2+-dependent acceleration of facilitation. kSTF was also accelerated when the stimulation frequency was increased from 5 Hz to 40 Hz (Figure 2D; 1.27 ± 0.18/s to 24.27 ± 3.44/s). Notably, the second 40-Hz train stimulation led to strong PPD, followed by accelerated facilitation (Figure 2D-E). The strong PPD suggests that the vesicular fusion probability kept high during the 5-Hz stimulation, arguing against the possibility for an increase in pv. Both two-step sequential priming models, LS/TS and RS/DS models, are suitable to interpret our data and we will refer to the different populations of vesicles as TS and LS vesicles, respectively (Taschenberger et al., 2016; Neher and Brose, 2018; Lin et al., 2022; Neher, 2024). Within this framework, vesicles are released from TS with high pv both in the resting state and during facilitation, and the delayed facilitation is seen as a result of Ca2+-dependent progressive increase in the occupancy of TS vesicles on docking sites.

Presynaptic calcium measurements at axonal boutons of L2/3 pyramidal cells

To test contribution of Ca2+ channel regulation to STP (Dobrunz et al., 1997; Jackman and Regehr, 2017), we estimated total amount of Ca2+ increments during 40 Hz AP trains at axonal boutons of L2/3 PCs in the mPFC. To this end, we measured AP-evoked Ca2+ transients (AP- CaTs) using two-dye ratiometry techniques, and estimated calcium binding ratio from AP-CaTs (Figure 2—figure supplement 2A-C). Measuring the amplitudes and decay time constants of AP-CaTs at different concentrations of Fluo-5F (150, 250 or 500 μM), and plotting them as a function of calcium binding ratio of Fluo-5F (κB), we estimated the amplitude of AP-CaTs (Aca = 1.16 μM), Ca2+ decay time constant (τc = 43 ms) in the absence exogenous Ca2+ buffer, and endogenous static Ca2+ binding ratio (κS = 96.4 or 77.3 estimated from Figure 2—figure supplement 2D-E, respectively; See Materials and Methods). The resting [Ca2+] ([Ca2+]rest) was measured as 50 nM (Figure 2—figure supplement 2F). Estimating the total Ca2+ increments for the first four AP-CaTs in a 40 Hz train according to Equation 2 (Materials and Methods), we found no evidence for inactivation or facilitation of calcium influx during a 40 Hz train (Figure 2—figure supplement 2G-I), arguing against the possibility that Ca2+ channel inactivation and facilitation contribute to strong PPD and delayed facilitation during a 40 Hz train.

Release sites have low baseline occupancy and these are increased during facilitation and post- tetanic augmentation

Given that pv was greater than 0.8, the two-fold increase in EPSC during 20 pulse train stimulations could not be explained by an increase in pv. Recent studies have shown that vesicle release sites (N) are limited in number and are partially occupied at rest (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Under such a high pv, two-fold synaptic facilitation should be attributed to an increase in pocc. The vesicle replenishment rate is accelerated during HFS in a Ca2+-dependent manner (Figure 2A-C; Hosoi et al., 2007). This may lead to the overfilling of release sites beyond their basal vesicular occupancy, which in turn may increase EPSC under high pv. To test this hypothesis, we determined the quantal parameters at the L2/3 local excitatory synapses. We first measured the quantal size from asynchronous release events after replacing extracellular Ca2+ with Sr2+, as previously described (Yoon et al., 2020) (Figure 3—figure supplement 1A-B). The mean and coefficient of variance of the quantal sizes were 24 pA and 0.25 at PC-PC synapses, respectively, and were 30 pA and 0.31 at PC-FSIN synapses (Figure 3—figure supplement 1C). Next, we performed a variance-mean (V-M) analysis of phasic EPSCs at excitatory synapses during train stimulation at 5 Hz and 40 Hz (Scheuss et al., 2002). Applying multiple-probability fluctuation analysis (MPFA) using intra-site quantal variability of 0.2 (Valera et al., 2012), the V-M plot was best fitted by the number of release sites (N) of 5.3 and the quantal size (q) of 25 pA at the PC-PC synapses (Figure 3A). The pr for EPSC1 was calculated as 0.32. We performed the same analysis for the PC-FSIN synapses and found N = 5.3, q = 35 pA, and pr for EPSC1 = 0.31 (Figure 3B). Our estimate for N was smaller but comparable to the estimate (6.9) at excitatory synapses in L2/3 of the mouse barrel cortex (Holler et al., 2021).

Low baseline occupancy of release sites and its increase during facilitation and post-tetanic augmentation

(A, B) Left, Representative EPSCs evoked by 5 and 40 Hz train stimulation (black, 5 Hz; purple, 40 Hz). Right, Variance-mean plots of EPSCs amplitude from averaged EPSCs recorded at PC- PC (A; n = 12, 17 for 5 and 40 Hz, respectively) and PC-FSIN (B; n = 8, 10) synapses. The data were fitted using multiple-probability fluctuation analysis (MPFA). Error bars are omitted for clarity. The 1st EPSC of 5 Hz train (broken line) was used to estimate the resting level of pocc. Filled circles were measured from post-tetanic augmented EPSCs (A, n = 12, 12, 9; B, n = 7, 7, 7 for 0.1, 0.2, 0.5 s IBIs, respectively). (C-E) Post-tetanic augmentation (PTA) experiments at PC-PC (top) and PC-FSIN (bottom) synapses. (C) Representative traces for EPSCs evoked by double 40 Hz train stimulations separated by 0.5 s. (D) Mean baseline-normalized amplitudes of EPSCs evoked by double 40 Hz trains at different inter-burst intervals (IBIs, 0.1, 0.2, 0.5, 1, 2, 5, 10 s). Upper, PC-PC synapse (n = 12, 12, 21, 11, 11, 16, 11 from short to long IBIs, respectively). Lower, PC-FSIN synapse (n = 10, 9, 8, 9, 9, 7, 9). (E) PTA time course. The baseline-normalized amplitudes of 1st EPSC from the 2nd train were plotted as a function of IBIs.

The pr estimate was clearly smaller than the possible minimum value for pv (0.81) estimated from the PPR data (Figure 2B). Given that pr is a product of pocc and pv (Neher, 2024), a pr smaller than pv indicates the partial occupancy of release sites in the resting state. If synaptic facilitation were mediated by the recruitment of new release sites, it would be accompanied by an increase in the variance of augmented EPSCs (Valera et al., 2012; Kobbersmed et al., 2020). However, the augmented EPSCs during facilitation followed the same parabola in the V-M plot (Figure 3A-B). Therefore, facilitation appears to be mediated by the progressive overfilling of a finite number of release sites beyond the basal occupancy. In the next section, we show that pv is close to unity based on high double failure rate upon paired pulse stimulation (Figure 3—figure supplement 2) and little effect of extracellular [Ca2+] ([Ca2+]o) elevation on baseline EPSC amplitudes (Figure 3—figure supplement 3). Assuming this, the baseline pr at the first pulse estimated from the fitted parabolic curve can be largely regarded as a baseline pocc (c.a. 30%), and the two-fold increase in EPSCs during facilitation can be attributed to an increase of up to 60% in pocc.

We investigated the time course of recovery from facilitation and post-tetanic augmentation (PTA) by applying dual-train stimulations (30 pulses at 40 Hz) separated by different inter- burst intervals (IBIs; Figure 3C-D). The initial EPSCs in the second train were enhanced by more than twofold, and followed by a strong PPD, suggesting that fusion probability stays high in the IBIs and thus synaptic transmission is carried by TS vesicles. The decay time course of PTA was characterized by two phases (initially fast and later slow; Figure 3E), which may reflect a two-phase Ca2+ decay or two distinct recovery processes that occur sequentially. The variance of potentiated EPSCs measured at IBIs of 0.1, 0.2, and 0.5 s was not different from that of EPSCs at the peak of facilitation (filled circles, Figure 3A-B). Given that the same RRP mediates both facilitation and augmentation and that post-tetanic EPSCs is carried by TS vesicles similar to the baseline EPSCs, this plot suggests that PTA, similar to short-term facilitation, is mediated by an increase in the TS vesicle occupancy.

High double failure rate and little effect of [Ca2+]o elevation on baseline EPSC amplitudes suggest that vesicular fusion probability is close to one

The vesicle dynamics of STP has been explained using a simple refilling model (Hosoi et al., 2007; Neher and Sakaba, 2008). This model assumes that the vesicles are reversibly docked to a limited number of release sites (N) with forward and reverse rate constants (denoted as k1 and b1, respectively). Under the framework of the simple refilling model (see Materials and Methods), the rate constant for the refilling of the RRP after depletion by the first pulse is the sum of the baseline k1 and b1. This is estimated as 23/s from the plot of PPR as a function of ISIs (Figure 3—figure supplement 2A). From the baseline occupancy of 0.3 and pocc = k1 / (k1 + b1), we estimated k1 to be 6.9/s.

When paired APs were applied with an ISI of 20 ms, the failure rate at 1st pulse, P(F1), was 10.6%, and the probability of two consecutive failures, P(F1, F2), was 6.2% (Figure 3—figure supplement 2A). From the equation P(F1) = (1 – pr)N, pr was calculated as 0.312 when N = 6, comparable to pr (= 0.32) estimated from the V-M analysis (Figure 3A). We calculated P(F1, F2) under the framework of a simple refilling model, with pv and k1 set as free parameters and other parameters set according to the following relationships: P(F1) = (1 – pr)N, pr = pv ·pocc, and pocc = k1 / (k1 + b1). The calculated values for P(F1, F2) and their difference from the observed value (6.2%) are shown in the plane of k1 vs. pv in Figure 3—figure supplement 2B-C. Given that k1 was greater than 5/s, the difference between the calculated and observed values of P(F1,F2) was minimal when pv = 1 (Figure 3—figure supplement 2C).

To find optimal values for k1 and pv that best explain observed probabilities shown in Figure 3—figure supplement 2Ac, a cost function was implemented as the sum of squared errors between observed and predicted probability values for the four combinations of success and failure of 1st and 2nd EPSCs as described in Materials and Methods. The minimum of the cost function (5.63 x 10-5) was found at k1 = 5.21/s and pv = 0.999, which predicted P11 = 41.5%, P10 = 47.9%, P01 = 4.93%, and P00 = 5.67% (subscript 1 and 0 denote success and failure, respectively).

To further validate that baseline pv is near to unity, we examined EPSC changes following elevation of [Ca²⁺]ₒ from 1.3 to 2.5 mM. According to the fourth-power relationship between synaptic responses and [Ca²⁺]ₒ (Equation 3 in (Dodge Jr and Rahamimoff, 1967)), a 3.24-fold increase in the EPSC amplitude was expected. However, the EPSC amplitude increased only 1.23-fold, a change that was not statistically significant (Figure 3—figure supplement 3). This minimal and sublinear response suggests that pv is already saturated at rest, limiting the dynamic range for further enhancement through increased calcium influx. A slight but significant reduction in PPR was also observed under high [Ca²⁺]ₒ. Although a reduction in PPR is often interpreted as an increase in pv, under conditions where pv is already high, it more likely reflects a slight increase in pocc or in the number of TS vesicles, consistent with the previous estimates (Lin et al., 2025).

Pharmacological manipulations reveal specific molecular identities underlying vesicle loading processes

Our results suggest that the baseline occupancy of TS vesicles is low and that the activity- dependent progressive overfilling of docking sites with TS vesicles is responsible for short- term facilitation and augmentation. To elucidate the specific molecular link between synaptic enhancement and Ca2+-dependent overfilling, we examined the effects of several drugs known to affect vesicle dynamics. We first examined the roles of phospholipase C (PLC) and diacylglycerol (DAG) by applying 5 μM U73122, a PLC inhibitor, or 20 μM 1-oleoyl-2-acetyl-sn-glycerol (OAG), a DAG analog. The STP at 40 Hz and PTA at 0.5-s interval were examined before and after applying each drug to the bath. For both drugs, the facilitation and augmentation were reduced (Figure 4A-B). U73122 did not affect the basal EPSC amplitude (EPSC1) but resulted in a stronger PPD, slower facilitation, and lower augmentation (Figure 4A). These effects of U73122 suggest the involvement of Ca2+-induced PLC activation in progressive overfilling. In contrast, OAG increased EPSC1 while maintaining a pronounced PPD (Figure 4B), indicating an increase in the TS vesicle pool size. Meanwhile, OAG slowed kSTF and reduced augmentation, indicating occlusion of activity-dependent synaptic facilitation. Phorbol esters and DAG are thought to accelerate priming through Munc13 activation (Rhee et al., 2002; Taschenberger et al., 2016; Aldahabi et al., 2022); they also shorten the bridges that interlink docked vesicles with the active zone (AZ) membrane (Papantoniou et al., 2023). Collectively, these suggest that PLC- and DAG-mediated signaling play a role in the overfilling of docking sites underlying short-term facilitation and augmentation.

Pharmacological experiments reveal specific molecular mechanisms underlying vesicle loading processes

(A-C) (a) Representative EPSC traces (left) and mean baseline-normalized EPSCs (right) evoked by double 40 Hz train stimulations separated by 0.5 s inter-burst interval (IBI) in control and in the presence of 5 μM U73122 (Aa, n = 7, orange), 20 μM OAG (Ba, n = 12, cyan) or 100 μM dynasore (Ca, n = 10, blue). (b) Mean values for baseline EPSCs (EPSC1, left) and augmentation (right) from the experiments shown in corresponding a panel. (Da) Representative EPSC traces evoked by paired pulses (left) and mean values for baseline EPSC amplitude (middle) and PPR (right) before and after applying 20 μM LatB (n = 6). (Db) Representative EPSC traces (left, upper) and average of normalized EPSCs (left, lower) evoked by double 40 Hz train stimulation separated by 0.5 s in control (n = 21) and in 20 μM LatB conditions (n = 16; pink). Right, Mean values for augmentation in control and LatB conditions. Gray symbols, individual data. All statistical data are represented as mean ± S.E.M., *, P<0.05; **, P<0.01; ***, P< 0.001; unpaired or paired t-test; n.s. = not significant.

Although the acceleration of Ca2+-dependent vesicle refilling supports rapid facilitation at 40 Hz, the later part of the delayed facilitation underwent a slight depression, which was more pronounced during the second burst (Figure 3D and Figure 4). Sustained synaptic transmission is limited by the spatial availability of docking sites because it is undermined by accumulation of exocytic remainings in the presynaptic AZ during HFS (Neher, 2010; Haucke et al., 2011). To verify this, we examined the effects of dynasore (100 μM), a specific inhibitor of endocytosis (Macia et al., 2006). The treatment with dynasore did not influence EPSC1 or the early phase of facilitation, but it exacerbated depression during the later phase and attenuated PTA (Figure 4C). This suggests that late synaptic transmission during the 40-Hz train is limited by site clearance.

Given the involvement of actin polymerization in the multiple steps of vesicle docking and recycling, we examined the effects of 20 μM latrunculin B (LatB), an inhibitor of actin polymerization. The effects of LatB were complicated: (1) EPSC1 was decreased, whereas PPR was increased (Figure 4Da) and (2) facilitation and augmentation were reduced (Figure 4Db). The latter effect may result from defects in vesicle replenishment rates, either through the impaired physical movement of vesicles or disrupted site clearance (Sakaba and Neher, 2003; Lee et al., 2012; Hallermann and Silver, 2013; Miki et al., 2016), as Ca2+-dependent vesicle refilling mediates facilitation and augmentation (Figure 2 and Figure 3). Meanwhile, the former effect might be attributed to a shift in equilibrium in the priming states towards a loosely docked state (LS) at rest. It is unknown whether actin affects vesicular fusogenicity in neurons, although it facilitates multiple steps of vesicle fusion by enhancing membrane tension in neuroendocrine cells (Wu and Chan, 2022).

Short-term facilitation at both types of local excitatory synapses in L2/3 is abolished by Syt7 knockdown

We examined whether Syt7 mediates facilitation at the PC-PC and PC-FSIN synapses in L2/3 of the mPFC. To this end, oChIEF and shRNA against Syt7 (shSyt7) were co-expressed in L2/3 PCs using IUE. Syt7 KD resulted in a complete loss of facilitation at all tested stimulation frequencies (Figure 5A-B). Short-term depression and paired-pulse depression were more pronounced as the stimulation frequency increased, and the frequency invariance of steady-state EPSC was abolished, implying little contribution of Ca2+-dependent vesicle recruitment (Figure 5C-F). The lack of facilitation could not be attributed to the failure of APs in presynaptic axons, as optical stimulation elicited reliable APs in Syt7 KD PCs during the 40-Hz train (Figure 1—figure supplement 2C). Syt7 KD had little effect on the basal properties, including the peak amplitude, rise/decay time, and time-to-peak, of single AP-induced EPSCs (Figure 1—figure supplement 3D-E). Moreover, Syt7 KD did not significantly affect the intrinsic properties of neurons (Figure 5—figure supplement 1).

STF at both types of local excitatory synapses is abolished by Syt7 KD

(A, B) Representative EPSC traces (left) and mean baseline-normalized amplitudes of EPSCs (right) evoked by 20-pulse trains at 5 to 40 Hz. STP was measured at PC-PC (A; n = 10, 9, 12, 9) and PC-FSIN (B; n = 9, 8, 10, 9) synapses, in which presynaptic Syt7 transcripts were depleted (Syt7 KD). Syt7 KD pyramidal cells are indicated as black triangles on the top. For comparison, STP in WT synapses is reproduced from Figure 1 (dotted lines). Same frequency color codes were used as in Figure 1. (C) Baseline-normalized amplitudes of steady-state EPSC (EPSCss; black symbols) and synaptic efficacy (EPSCss × f; red symbols) as a function of stimulation frequency (f) at PC-PC synapses. EPSCss was defined as the average of last 5 EPSC amplitudes from 20-pulse trains. (D) PPR as a function of inter-spike intervals (n = 10, 9, 12, 18, 26). (E) EPSCss and synaptic efficacy at PC-FSIN synapses. (F) PPR at PC-FSIN synapses (n = 9, 8, 10, 12, 24). Gray symbols, individual data.

To examine whether Syt7 KD had any nonautonomous effects on postsynaptic cells, we measured the quantal size (q) from asynchronous events in the presence of Sr2+ and found that the mean quantal size remained unaltered at Syt7 KD synapses (Figure 3—figure supplement 1C). This indicated that Syt7 did not influence postsynaptic parameters, including the density and kinetic properties of AMPARs. To test the specificity of shSyt7, we examined whether co-expression of the shRNA-resistant form of Syt7 restored synaptic facilitation in both excitatory synapse types. Presynaptic overexpression of shRNA-resistant Syt7 under the CAG promoter rescued the effect of shRNA at both PC-PC and PC-FSIN synapses (Figure 5— figure supplement 2). These results underscore the crucial role of presynaptic Syt7 in mediating short-term facilitation at local excitatory synapses in L2/3 of the mPFC.

Syt7 KD synapses exhibit complementary changes in the number of release sites and their vesicle occupancy

To estimate the quantal parameters in Syt7 KD synapses, we performed V-M analysis of EPSCs evoked by train stimulation (Figure 6A-B). Fitting a parabola to the V-M plot revealed higher baseline pocc in Syt7 KD synapses than in wild-type (WT) synapses for both synapse types (KD, 0.68 and 0.42; WT, 0.32 and 0.31 for PC-PC and PC-FSIN synapses, respectively). The pocc monotonously decreased during a train stimulation, as expected from short-term depression. Moreover, the number of release sites (N) was markedly lower in Syt7 KD synapses (N = 3.5) than in WT synapses (N = 5.3; Figure 3A). Next, we examined the PTA induced by the same protocol as in Figure 3 (two 40-Hz train stimulations separated by variable IBIs) at Syt7 KD synapses (Figure 6C). The 40 Hz train-induced synaptic depression quickly recovered during IBIs, but no PTA was observed at Syt7 KD synapses, indicating its crucial role in augmentation (Figure 6D-E).

Syt7 KD synapses exhibit complementary changes in the number of release sites and their vesicle occupancy

(A, B) Left, Representative traces of EPSCs evoked by 5 and 40 Hz train stimulations (black, 5 Hz; purple, 40 Hz). Right, Variance-mean plots of EPSCs amplitude from averaged EPSCs recorded at PC-PC (A; n = 9, 15, 12, 10, 9) and PC-FSIN (B; n = 4, 14, 10, 9, 8) synapses in which presynaptic Syt7 has been knocked down. The data were fitted using MPFA and error bars are omitted for clarity. The mean 1st EPSC amplitude of 5 Hz train (vertical broken line) was used for estimation of baseline pocc. Dashed parabolas indicate MPFA fits to variance-mean plot of WT synapses reproduced from Figure 3. (C-E) Recovery experiments at PC-PC (top) and PC- FSIN (bottom) synapses in which presynaptic Syt7 was knocked-down. (C) Representative EPSCs evoked by double 40 Hz train stimulations separated by 0.5 s. (D) Mean baseline- normalized amplitudes of EPSCs evoked by double 40 Hz trains at PC-PC (n = 12, 10, 9, 9, 10, 9, 8) and PC-FSIN (n = 10, 9, 10, 11, 8, 9, 10) synapses at different interburst intervals (IBIs). (E) Recovery time course. Baseline-normalized amplitudes of 1st EPSC from the 2nd burst were plotted as a function of various IBIs. Dotted lines indicate augmented EPSCs in the W reproduced from Figure 3.

TS vesicle recovery following depletion is accelerated by Syt7

Syt7 is recognized not only for its crucial role in facilitation, but also for mediating Ca2+- dependent replenishment of releasable vesicles (Liu et al., 2014; Tawfik et al., 2021). Given that facilitation at L2/3 excitatory synapses is driven by progressive overfilling (Figure 2 and Figure 3), the loss of facilitation at Syt7 KD synapses in the present study implies that Syt7 plays a key role in the Ca2+-dependent acceleration of refilling and overfilling of the docking sites with TS vesicles. To test this hypothesis, we examined the recovery kinetics of EPSCs after TS vesicles were depleted by 3 pulses at 40 Hz in wild-type (WT) and Syt7 KD synapses (Figure 7). Strong PPD was observed not only in the first burst, but also in the second burst. This suggested that EPSCs in the second burst were mediated by TS vesicles, similar to those in the first burst (Figure 7A, D). Therefore, EPSC1 in the second burst can be interpreted as a release from TS vesicles recovering during a given IBI. The recovery time course of EPSC1 in the second burst indicated full recovery of TS vesicles within 100 ms in both excitatory synapse types in WT rats (Figure 7C, F). However, this recovery was remarkably slower in Syt7 KD synapses, requiring 5 s for full recovery, indicating that TS vesicle recovery was greatly accelerated by Syt7. Overall, synaptic facilitation and augmentation at local excitatory synapses in L2/3 can be ascribed to progressive activity-dependent increases in the occupancy of TS vesicles (Figure 2 and Figure 3), in which Syt7 plays a key role.

Recovery of TS vesicles following depletion is accelerated by Syt7

(A-C). Recovery experiments at PC-PC synapses in WT (A) and Syt7 KD (B). Mean baseline- normalized amplitudes of EPSCs evoked by two consecutive 3-pulse 40 Hz trains in WT (n = 16, 13, 11, 11, 11, 9 from short to long IBIs, respectively) and KD (n = 13, 11, 12, 8, 10, 11) synapses at different IBIs (0.1, 0.2, 0.5, 1, 2, 5 s). Inset, representative traces of EPSCs evoked by two consecutive 3-pulse 40 Hz train stimulations separated by 0.2 s. (D-F). Recovery experiments at PC-FSIN synapses in WT (D) and Syt7 KD (E). Mean baseline-normalized amplitudes of EPSCs evoked by two consecutive 3-pulse 40 Hz trains in WT (n = 9, 11, 11, 9, 10, 13) and KD (n = 9, 12, 9, 7, 10, 9) synapses at different IBIs. Inset, representative traces of EPSCs evoked by two consecutive 3-pulse 40 Hz train stimulations separated by 0.2 s. (C, F) Recovery time course. Baseline-normalized amplitudes of 1st EPSC from the 2nd burst were plotted as a function of various IBIs.

Behavioral consequences of Syt7 KD in L2/3 PCs of the mPFC

We previously proposed that the disparity in STP between the PC-PC and PC-IN synapses results in activity-dependent changes in the ratio of synaptic weights at the PC-PC and PC-IN synapses (Jee/Jie). This may in turn have a profound influence on persistent activity in a recurrent network by biasing the E-I balance (Yoon et al., 2020). Syt7 KD not only eliminated facilitation, but also the activity-dependent increase in the Jee/Jie ratio in the L2/3 network (Figure 8—figure supplement 1). This suggests the possibility that Syt7 contributes to the persistent activity during working memory tasks in the L2/3 recurrent network. Trace fear conditioning (tFC) requires associative learning between an auditory cue (conditioned stimulus, CS) and temporally separate aversive events (unconditioned stimulus, US). The formation of trace fear memory requires the prelimbic area of the mPFC and hippocampus. mPFC activity during the trace interval is thought to temporarily hold CS information, enabling the network to associate temporally discontinuous events (i.e., temporal associative learning) (Gilmartin et al., 2013). Synaptic facilitation and augmentation are implicated in temporary memory retention in recurrent networks (Mongillo et al., 2008; Mongillo et al., 2012).

Given that Syt7 is crucial for synaptic facilitation (Figure 5) and augmentation (Figure 6) in prelimbic L2/3 recurrent excitatory synapses, we tested whether trace fear memory formation was impaired in rats with depleted Syt7 transcripts, specifically in the L2/3 PCs of the mPFC. Syt7-targeted shRNA (shSyt7) or scrambled shRNA (Scr) constructs were transfected bilaterally into the L2/3 PCs of the mPFC using IUE (Figure 8A). Both the Scr- transfected control and shSyt7-transfected rats underwent tFC training (Figure 8B). Consistent with Gilmartin et al. (2013), in which prefrontal persistent firing was optogenetically inhibited, WT and KD animals exhibited similar freezing behavior during this acquisition phase. The following day, the rats were subjected to a tone test in which trace fear memory formation was assessed from the freezing response to a tone alone in a distinct context. Compared to the control group, the shSyt7-transfected group exhibited significantly lower levels of freezing behavior in response to auditory cues during the first four trials, suggesting an impairment in trace memory formation (Figure 8C).

Behavioral effects of Syt7 deficiency in L2/3 PCs of the mPFC

(A) Top, Representative images showing bilateral expression of U6-GFP in L2/3 of PCs after IUE at E17.5. Scale bar: 1 mm, 100 μm. Bottom, Schematic of trace fear conditioning and extinction (tone test) protocol. (B) Freezing behavior of control (expressing scrambled shRNA, Scr) and Syt7 KD rats during acquisition of tFC. (C) Freezing ratio during tone tests on following days. Data are shown as average freezing during T1–T4, T5–T8, or T9–T12 (T, trials). The freezing on T1-T4 was significantly lower in KD rats suggesting that formation of trace memory was impaired in Syt7 KD rats (n = 10, 11 for Ctrl and KD, respectively; P = 0.0007, F(1, 19) = 16.43, two-way repeated measures ANOVA; P = 0.0064, 0.0211, 0.0064, 0.0064; Holm-Sidak test). (D) Representative images of c-Fos immunoreactivity in the prelimbic cortex of control or Syt7 KD rats 90 min after tFC acquisition. c-Fos (left, red, Cy5) and GAD67 (right, cyan, Cy3) were immunostained in the same brain slice expressing U6-GFP (middle; green). Scale bar, 50 μm. (E) Exemplar images of c-Fos positive neurons expressing GFP (top) or GAD67 (bottom) in control (left) or Syt7 KD rats (right). Scale bar, 50 μm. (F-H) Effects of Syt7 KD on c-Fos density (F) and percentage of c-Fos positive neurons co-labeled with GAD67 (G) or GFP (H) in L2/3 or L5 of prelimbic cortex (n = 9, 9 for Ctrl and KD, respectively). Open symbols, individual data. All statistical data are represented as mean ± S.E.M.; ****, P<0.0001; unpaired t-test.

The tone test was repeated the following day to test for extinction memory formation. Syt7 KD animals exhibited much less freezing to the CS (+1d) than did WT animals. Considering the Rescorla-Wagner model (Yau and McNally, 2023), our results suggest that the strength of CS- US associative memory was weaker in KD animals than in WT animals; thus, the acquisition of trace fear memory was impaired by Syt7 KD in the L2/3 PCs of the mPFC.

Next, we assessed the locomotor activity and anxiety levels of rats using the open field test (OFT) and elevated plus maze test (EPM). Scr controls and Syt7 KD rats showed similar exploratory patterns, displaying comparable total distance moved and time spent in the periphery of the OFT or the open/closed arms of the EPM (Figure 8—figure supplement 2). Previous studies have suggested that the acquisition of trace fear memory requires elevated trace activity in prelimbic PCs (Gilmartin et al., 2013). Given that c-Fos, an immediate early gene, is upregulated following patterned activities in neurons, its expression is considered a reliable indicator of recent neuronal activation (Fields et al., 1997; Guzowski et al., 2005). To investigate the effect of Syt7 on neuronal activity in vivo during tFC, animals were sacrificed at the time of peak c-Fos protein expression, approximately 90 min after completion of the behavioral task (Arime and Akiyama, 2017).

To elucidate the activation patterns of specific neuronal populations, we performed immunohistochemical co-labeling of c-Fos and GAD67 (a GABAergic interneuron marker) upon tFC acquisition. The density of c-Fos(+) cells and the percentage of c-Fos(+) cells among GAD67(+) cells in all prelimbic layers were significantly lower in Syt7 KD rats than in control rats (Figure 8F-G). Additionally, analysis of the proportion of active cells among transfected L2/3 PCs (c-Fos+/GFP+) revealed that Syt7 KD L2/3 PCs were significantly more active than Scr-transfected PCs (Figure 8H). This unexpected hyperactivity of Syt7 KD PCs may be attributed to the high reciprocal connectivity between PCs and FSINs in the neocortex (Holmgren et al., 2003; Otsuka and Kawaguchi, 2009), as PCs with short-term depression would transmit less input to FSINs, which in turn would exert less feedback inhibition on reciprocally connected PCs. Collectively, these findings show that the slow activation of vesicle refilling supports various forms of facilitation at excitatory synapses in PFC L2/3, which is essential for neuronal activity during temporal associative learning.

Discussion

The present study characterized STP at local excitatory synapses in the L2/3 network of the rat mPFC at physiological extracellular [Ca2+]. These synapses displayed initial strong depression and then delayed facilitation at 40 Hz, while monotonous slowly developing facilitation was observed at lower frequencies. Our study suggests that such delayed facilitation after a brief depression results from a high pv (Figure 2 and Figure 3 and Figure 3—figure supplement 2), along with Ca2+-dependent delayed acceleration of vesicle refilling and overfilling (Figure 2 and Figure 3). The two-fold increase in EPSC during 20 pulse train stimulations under the condition of such a high pv suggests a Ca2+-dependent increase in the vesicle replenishment rate during a train; this then leads to overfilling of release sites beyond their incomplete basal occupancy (Figure 3A-B). The high pv and strong PPD did not result from artificial bouton stimulation, AMPAR desensitization, or any potential distortion caused by optogenetic methods (Figure 1—figure supplements 4 and 5). Additionally, similar to short-term facilitation, PTA could be explained by an increase in the TS vesicle occupancy (Figure 3C-E), consistent with previous studies showing post-tetanic increases in the pocc and/or vesicle pool size at various types of synapses (Lee et al., 2010; Vandael et al., 2020; Tran et al., 2023; Silva et al., 2024). These findings suggest that release sites are partially occupied at rest and that a progressive overfilling of release sites with TS vesicles underlies facilitation and augmentation at intracortical L2/3 excitatory synapses.

Definition of vesicular release probability in the present and previous studies

Previous studies for post-tetanic augmentation (Stevens and Wesseling, 1999; Garcia-Perez and Wesseling, 2008) observed invariance of the RRP size after tetanic stimulation. In these studies, the RRP size was estimated by hypertonic sucrose solution or as the sum of EPSCs evoked 20 Hz/60 pulses train (denoted as ‘RRPhyper’). Because reluctant vesicles can be quickly converted to TS vesicles (16/s) and are released during a train (Lee et al., 2012), it is likely that the RRP size measured by these methods encompasses both LS and TS vesicles. In contrast, we assert high pv based on the observation of strong PPD, failure rates upon paired stimulations at ISI of 20 ms (Figure 2 and Figure 3—figure supplement 2). Given that single AP-induced vesicular release occurs from TS vesicles but not from LS vesicles, pv in the present study indicates the fusion probability of TS vesicles. From the same reasons, pocc denotes the occupancy of release sites by TS vesicles. Note that our study does not provide direct clue whether release sites are occupied by LS vesicles that are not tapped by a single AP, although an increase in the LS vesicle number may accelerate the recovery of TS vesicles. As suggested in Neher (2024), even if the number of LS plus TS vesicles are kept constant, an increase in pocc (occupancy by TS vesicles) would be interpreted as an increase in ‘vesicular release probability’ if it was measured based on RRPhyper.

By the way, it should be noted that increases in both pv and pocc may contribute to PTA. Post- tetanic potentiation (PTP) and phorbol esters had similar effects on baseline EPSCs (Taschenberger et al., 2016), and PDBu increased both the TS vesicle pool size (1.9-fold) and fusion probability of TS vesicles (1.3-fold) at the calyx synapses (Lin et al., 2025). Therefore, it is possible that PTP may increase both pocc and pv of TS vesicles, although it has not been yet directly tested. Moreover, acceleration of the latency in the hypertonicity-induced vesicle release raises a possibility for an increase in fusion probability of TS vesicles (Stevens and Wesseling, 1999; Garcia-Perez and Wesseling, 2008), although it is still controversial that the acceleration of the release latency represents a reduction in the activation energy barrier for vesicle fusion (Schotten et al., 2015).

Progressive overfilling of docking sites

Conventional models for facilitation assume an increase in pv after an AP of RRP vesicles, which had been in a low pv or reluctant state at rest (Dittman et al., 2000; Turecek et al., 2016). This facilitation model is unlikely to explain our results, because TS vesicles are predominantly released not only at rest but also in the facilitated state during or after train stimulation, as evidenced by the strong PPD (ISI, 25 ms) during a 5-Hz train (Figure 2D-E) and at different intervals after a 40-Hz/30 pulse trains (Figure 3D). Moreover, the PPD (ISI, 25 ms) remained pronounced during the early phase of facilitation (Figure 7), arguing against the possibility of slowly increasing the pv of reluctant vesicles. These results suggested that facilitation was mediated by an activity-dependent progressive increase in occupancy of TS vesicles (TS occupancy). Facilitation through an increase in TS occupancy has been termed frequency facilitation, differentiating it from paired pulse facilitation that relies on a residual calcium- dependent transient increase in the pv of releasable vesicles (Neher, 2024).

Assuming an increase in TS occupancy, to attain a two-fold increase in EPSC (Figure 1), the baseline occupancy of the docking sites by TS vesicles should be low, as estimated from the V-M plot (c.a. 30% in Figure 3). However, it is unclear whether the remaining docking sites that are not occupied by TS vesicles are truly vacant or are occupied by reluctant LS vesicles that are not released by the AP or AP trains. TS vesicles can be supplied by the transformation of LS to TS vesicles, both of which reside at the same docking site (LS/TS model; reviewed in Neher, 2024). Alternatively, TS vesicles can be supplied from a distinct replacement pool (RS/DS model; reviewed in Silva et al., 2021). Further studies, including ultrastructural analyses of active zones, are required to address this issue.

Given the crucial role of Syt7 in facilitation (Figure 5), the facilitation mechanisms are closely related to how Syt7 mediates synaptic facilitation. Based on the slow kinetics of Syt7, previous computational modeling studies have proposed that Syt7 may mediate short-term facilitation through an activity-dependent increase in pv (Turecek et al., 2016; Jackman and Regehr, 2017; Turecek et al., 2017; Norman et al., 2023). The present study, however, found that both facilitation and augmentation were mediated by overfilling of the release sites (Figure 2 and Figure 3). Moreover, Syt7-KD synapses displayed slower recovery of TS vesicles (Figure 7). These results suggest that Syt7 is essential for activity-dependent overfilling of docking sites with TS vesicles. Whereas the baseline EPSC was not altered (Figure 1—figure supplement 3), and complementary changes in the number of docking sites and their baseline occupancy were observed in Syt7 KD synapses (Figure 6). These results raise a possibility that Syt7 may play a role in providing additional vacant docking sites that can be overfilled during facilitation. It remains to be elucidated whether the decrease in the number of docking sites at Syt7 KD synapses is related to the subcellular localization of Syt7 to the plasma membrane of the active zone (Sugita et al., 2001; Vevea et al., 2021).

STP model in light of known Ca2+ binding kinetics of Syt7

Syt7 acts as an upstream Ca2+ sensor that facilitates activity-dependent replenishment (Liu et al., 2014; Bacaj et al., 2015; Chen et al., 2017; Tawfik et al., 2021). The present study shows that EPSCs are mediated by TS vesicles both at rest and during activity-dependent facilitation. Releasable vesicles are homogeneous, and Syt7-dependent overfilling is responsible for the activity-dependent enhancement of EPSCs. Thus, we tested whether slowly developing facilitation could be replicated within the framework of a simple refilling model based on the known Ca2+ binding kinetics of Syt7 (Brandt et al., 2012). To this end, we made following assumptions: (1) a Ca2+-dependent increase in the forward refilling rate (k1) requires a full Ca2+-bound form of Syt7, to which multiple Ca2+ ions can bind (two or three Ca2+ ions on each C2A and/or C2B domain) (Brandt et al., 2012; Chon et al., 2015; Voleti et al., 2017); (2) sequential Ca2+ binding steps are synergistic through a cooperativity factor, b; (3) Syt7 senses AP-induced local [Ca2+] similar to that estimated at the calyx of Held (a Gaussian function with a full width at half maximum value of 0.2 ms and a peak of 40 μM) (Wang et al., 2008); and (4) residual [Ca2+] follows a mono-exponential function with an amplitude of 1 μM and decay time constant of 50 ms (Jackson and Redman, 2003).

Under these assumptions and the framework of the simple refilling model, slow and delayed facilitation could be replicated by modeling k1 as k1 = k1,b + K1,max × [Syt7: n Ca2+], where [Syt7: n Ca2+] represents the fraction of fully Ca2+-bound Syt7 (Figure 9A-C). Syt7 was slowly activated over the course of the stimulus train, suggesting that the cooperative binding of multiple Ca2+ ions progressively increased the putative active form of Syt7 because of its slow membrane binding/unbinding kinetics. Nevertheless, the late phase of facilitation at high frequencies was predicted to be higher in the simulation than in the experiment. As expected, based on the notion that sustained release during HFS is limited by site clearance in the active zone (Figure 4C), the model/data discrepancy became more pronounced at higher frequencies. Incorporating endocytic terms into the model may mitigate late-phase discrepancies. Moreover, removal of the catalytic function of Syt7 in the model effectively accounted for the complete loss of facilitation observed in Syt7 KD (Figure 9D).

STP model in light of known Ca2+ binding kinetics of Syt7

(A-D) Left, Schematic of allosteric calcium binding to Syt7. The number of Ca2+ bound to Syt7 was denoted as # in ‘S#’ in the reaction scheme. kon = 7/μM/s, koff = 10/s. Middle, Simulated changes of k1 (black) in response to 5 Hz train of Ca2+ transients (light blue traces). The priming step of the simple refilling model was assumed to be catalyzed by full Ca2+-bound form of Syt7. Accordingly, Ca2+-dependent increase in k1 was calculated as K1,max multiplied by a fraction of full Ca2+-bound form of Syt7. We assumed that local [Ca2+]i(t) follows a Gaussian function: (1/σ √2π) exp[-(t-tp)2/ 2σ2], in which tp = 0.25 ms and σ = 0.085 ms. Right, Fits of the Syt7 model to the STP data. To fit this model to the STP data, K1,max was set to 300/s (A), 220/s (B), and 180/s (C). Cooperativity factor (b) was set to 0.35 (A), 0.2 (B), and 0.05 (C). k1 is set to be constant for Syt7 KD (D). Same frequency color codes were used as in Figure 1.

Conventional STF models involving cumulative increase in pv of reluctant vesicles

Conventional models for facilitation assume a post-AP residual Ca2+-dependent step increase in pv of RRP (Dittman et al., 2000) or reluctant vesicles (Turecek et al., 2016). Given that pv of TS vesicles is close to one in the present study, an increase in pv of TS vesicles cannot account for facilitation. The possibility for activity-dependent increase in fusion probability of LS vesicles (denoted as pv,LS) should be considered in two ways depending on whether LS and TS vesicles reside in distinct pools or in the same pool. Notably, strong PPD at short ISI implies that pv,LS is near zero at the resting state. Whereas LS vesicles do not contribute to baseline transmission, short-term facilitation (STF) may be mediated by cumulative increase in pv of LS vesicles that reside in a distinct pool. Because the increase in pv,LS during facilitation recruits new release sites (increase in N), the variance of EPSCs should become larger as stimulation frequency increases, resulting in upward deviation from a parabola in the V-M plane, as shown in recent studies (Valera et al., 2012; Kobbersmed et al., 2020). This prediction is not compatible with our results of V-M analysis (Figure 3), showing that EPSCs during STF fell on the same parabola regardless of stimulation frequencies. Therefore, it is unlikely that an increase in fusion probability of reluctant vesicles residing in a distinct release pool mediates STF in the present study. For the latter case, in which LS and TS vesicles occupy in the same release sites, it is hard to distinguish a step increase in fusion probability of LS vesicles from a conversion of LS vesicles to TS. Nevertheless, our results do not support the possibility for a gradual increase in pv,LS that occurs in parallel with STF. Strong PPD, indicative of high pv, was consistently found not only in the baseline (Figure 2 and Figure 2—figure supplement 1) but also during post-tetanic augmentation phase (Figure 3D) and even during the early development of facilitation (Figure 2D-E and Figure 7), arguing against the gradual increase in pv,LS. One may argue that STF may be mediated by a drastic step increase of pv,LS from zero to one, but it is not distinguishable from conversion of LS to TS vesicles.

Possible mechanisms underlying the depression-facilitation sequence

Cooperative binding of multiple Ca2+ ions to Syt7 predicts delayed activation and a progressive increase in vesicle refilling during train stimulations, resulting in delayed facilitation. Combining this with a high pv could replicate not only the slow facilitation at 5 Hz, but also the unique depression-facilitation sequence observed at 40 Hz (Figure 9). As simulated by Pulido and Marty (2018), delayed facilitation can be reproduced by a high pv in combination with a low baseline occupancy of replacement sites and subsequent Ca2+-dependent refilling. However, whether this model is also able to replicate slow facilitation at low frequencies remains to be tested. An alternative, but not exclusive, possibility for delayed facilitation is that the presynaptic global Ca2+ itself may undergo a slow buildup during a train due to the saturation of endogenous Ca2+-binding proteins.

Facilitation requires not only Syt7, but also activation of PLC-DAG signaling

To gain insight into the specific molecular mechanisms underlying the vesicle-loading processes, we examined the effects of pharmacological inhibitors targeting PLC and DAG (Figure 4). Both a PLC inhibitor and a DAG analog reduced facilitation and augmentation. This suggests that the progressive overfilling of docking sites with TS vesicles is supported by the PLC-DAG signaling pathway and probably downstream Munc13 (Rhee et al., 2002; Lou et al., 2008). Phorbol esters have been recently shown to increase the TS fraction of docked vesicles by enhancing vesicle priming without affecting Ca2+ transients or pv (Taschenberger et al., 2016; Aldahabi et al., 2022; Papantoniou et al., 2023; Aldahabi et al., 2024). In line with these findings, signaling involving PLC and DAG may support Ca2+-dependent overfilling of TS vesicles by enhancing molecular priming processes.

Roles of facilitation and augmentation in working memory and related tasks

Previous theoretical studies have proposed that working memory in a recurrent network can be maintained in the form of augmentation at synapses (Mongillo et al., 2008) and/or persistent activity of neurons in a memory ensemble (Mongillo et al., 2012). The latter can be supported by short-term facilitation of recurrent excitatory synapses or disparity in STP between excitatory and inhibitory synapses (Mongillo et al., 2012; Yoon et al., 2020). However, whether these short-term enhancements in synaptic efficacy contribute to working memory- related behaviors has not been examined. tFC is a temporal associative learning task requiring temporary retention of CS information during a trace period. The mPFC and hippocampus are crucial for tFC, and tFC is interfered with by a load of working memory and attention distraction, implying that the tFC shares the same neural substrate with other working memory tasks (Raybuck and Lattal, 2014). In the present study, tFC in WT and Syt7 KD animals revealed that Syt7 deficiency in the L2/3 PCs of the mPFC impaired the acquisition of trace fear memory and reduced c-Fos expression associated with tFC training (Figure 8).

The acquisition of trace fear memory was impaired by inhibition of persistent activity in mPFC during trace period (Gilmartin et al., 2013). The similar deficit observed in Syt7 KD animals is consistent with the hypothesis that STF provides bi-stable ensemble activity in a recurrent network (Mongillo et al., 2012). Nevertheless, alternative mechanisms may be responsible for the behavioral deficit. Not only recurrent network but also long-range loop between the mPFC and the mediodorsal (MD) thalamus play a critical role in maintaining persistent activity within the mPFC especially for a delay period longer than 10 s (Bolkan et al., 2017). Prefrontal L2/3 is heavily innervated by MD thalamus, and L2/3-PCs subsequently relay signals to L5 cortico-thalamic (CT) neurons (Collins et al., 2018). Given that L2/3 is an essential component of the PFC-thalamic loop, loss of STF at recurrent synapses between L2/3 PCs may lead to insufficient L2/3 inputs to L5 CT neurons and failure in the reverberant PFC-MD thalamic feedback loop. Therefore, not only L2/3 recurrent network but also its output to downstream network should be considered as a possible network mechanism underlying behavioral deficit caused by Syt7 KD L2/3.

In summary, activity-dependent synaptic facilitation and augmentation at prelimbic L2/3 recurrent excitatory synapses can be attributed to overfilling of release sites with TS vesicles. Given that the lack of short-term facilitation in L2/3 PCs impairs trace fear memory acquisition, temporary maintenance of CS information during trace period may be supported by an increase in the TS occupancy in presynaptic terminals of CS-representing neuronal ensembles.

Materials and Methods

Animals/Subjects

All experiments were carried out on Sprague-Dawley rats of either sex at p28-43. Rats were maintained in temperature-controlled rooms on a constant 12 h light/12 h dark cycle and were given access to water and food ad libitum. All animal procedures were approved and performed in accordance with the Institutional Animal Care and Use Committee at Seoul National University (SNU200522-1).

In utero electroporation (IUE)

Pregnant Sprague Dawley rats on embryonic day (E) 17.5 were deeply anesthetized with 5% (v/v) isoflurane for the duration of surgery. The uterine horns were exposed by laparotomy and wet with warm sterile phosphate-buffered saline (PBS). The plasmids (0.5-2 µg/µl) together with the 0.01% fast green dye were injected into either the left or the right lateral ventricle or bilateral ventricles of embryos through a thin glass capillary (Narishige PC-10). The embryo’s head was carefully held between the forceps-shaped electrodes of 10 mm diameter (CUY65010; NEPAGENE, Japan). Electroporation pulses (50 V for 50 ms) were delivered five times at intervals of 150 ms using a square-wave generator (ECM830; BEX, United States). The uterine horns were placed back into the abdominal cavity and abdominal muscles and skin were sutured. After the surgery, the animals were recovered under the infrared lamp. For knockdown experiments in slice electrophysiology, the shRNA constructs were co-electroporated with the pCAG-oChIEF-tdTomato, because of the consistent co-expression of the two constructs in a significant proportion of the transfected cells in vivo (Bony et al., 2013). Pups with robust fluorescence signals in the prefrontal cortex of both hemispheres were used for behavior experiments.

Preparation of vector constructs

For an optic stimulation, pCAG-oChIEF-tdTomato was transfected by IUE. Short hairpin RNA (shRNA) against Syt7 sequence (GATCTACCTGTCCTGGAAGAG) was expressed under the control of an U6 promoter of pAAV-U6-GFP vector (Cell Biolabs, #VPK-413). This shRNA was shown to effectively knockdown Syt7 in cell cultures and in vivo by previous reports from more than one research group (Bacaj et al., 2013; Li et al., 2017). Scrambled control sequence (TCGCATAGCGTATGCCGTT) was designed by shuffling the recognition region sequences and a BLAST analysis confirmed that the control shRNA sequence had no target gene. For rescue experiments, the cDNA for rat Syt7 (NM_021659) rendered resistant to the shRNA were inserted in the KD vector and their expression was driven by the CAG promoter; the vector also contained a self-cleaving P2A peptide sequence followed by EGFP, which ensures visualization of transfected cells. All the constructs above were verified by DNA sequencing. Plasmid DNA was purified and concentrated under endotoxin free conditions (Qiagen EndoFree Maxi Kit).

Slice preparation

Acute brain slices were prepared by using a VT1200S Vibratome (Leica Microsystems) after isoflurane anesthesia and decapitation. Coronal sections of the mPFC (300 μm in thickness) were cut in ice-cold artificial cerebrospinal fluid (aCSF) composed of (in mM) 125 NaCl, 3.2 KCl, 25 NaHCO3, 1.25 NaH2PO4, 20 D-glucose, 2 NaPyr, 1 MgCl2, and 1.3 CaCl2 (Yoon et al., 2020), bubbled with 95% O2 and CO2 (pH = 7.3; 310 mOsm). Slices were allowed to recover for a minimum of 30 minutes at 36 °C, and subsequently maintained at room temperature until recording. Composition of aCSF was identical throughout the preparation, incubation, and recording periods unless otherwise stated. For the recordings, the mPFC slices were continuously perfused with oxygenated aCSF at a flow rate of 1.5 ml/min, and the temperature was maintained at 31–32 °C using an in-line temperature controller (Warner Instruments).

Electrophysiology

Whole-cell patch-clamp recordings were performed from pyramidal cells and fast-spiking interneurons in layer 2/3 of the prelimbic mPFC. Neuronal subtypes were initially selected based on morphology under visual guidance by using an upright microscope equipped with infrared differential interference contrast (IR-DIC) optics (BX51WI, Olympus), and were further identified by their electrophysiological properties (see Figure 1—figure supplement 1; van Aerde and Feldmeyer, 2015; Tremblay et al., 2016). To analyze these intrinsic properties in current-clamp experiments, we measured the following parameters: (1) resting membrane potential (RMP), (2) input resistance (Rin), (3) sag ratio, and (4) the F-I curve, which plots firing frequencies (F) against the amplitude of injected currents (I). Both cell types exhibited a low sag ratio below 0.1, consistent with findings from other studies (van Aerde and Feldmeyer, 2015; Yoon et al., 2020). PCs and FSINs were easily distinguished by their unique firing frequency, Rin, spike adaptation, RMP and morphological characteristics including the shape of soma and the thickness of apical dendrites. For experiments involving oChIEF-expressing neurons in WT or Syt7 KD rats, whole-cell recordings were performed from the fluorescently labeled pyramidal neurons in layer 2/3 of the prelimbic cortex using suitable filter sets. Patch pipettes (tip resistance of 2–3 MΩ) were pulled from borosilicate glass and filled with (in mM) 130 K-gluconate, 8 KCl, 10 HEPES, 0.2 EGTA, 20 Na2Pcr, 0.3 Na2GTP, 4 MgATP. Intracellular solutions were adjusted to pH 7.25 and 300–310 mOsm. Neurons were voltage clamped at −78 mV and series resistance (Rs) were continuously monitored during the EPSCs recordings. Experiments were discarded before analysis if the Rs changed by 20% or larger deviation of baseline value. Measurements were not corrected for Rs compensation, bridge balance, or liquid junction potential. Electrophysiological data were low-pass-filtered at 1 kHz (Bessel) and acquired at 20 kHz using a Multiclamp 700B amplifier paired with Digidata 1440A digitizer and Clampex 10.2 software (Molecular Devices).

Synaptic stimulation

IUE resulted in sparse expression of channelrhodopsin (10-20%) (Figure 1A). Exploiting the sparse expression, we employed a minimal optical stimulation method to activate a single or very few excitatory synapses in layer 2/3. The optic minimal stimulation was performed by our published protocol with minor modifications (Yoon et al., 2020). This approach aimed to minimize the risk of synaptic contamination, which could arise from engaging extra terminals that were initially inactive (depolarized below threshold) but became active by temporal summation during high-frequency trains. It also helped prevent the buildup of depolarization at the synaptic terminals, which might influence the probability of release, as noted by (Jackman et al., 2014). To achieve this, we employed a collimated digital micromirror device (DMD)-coupled LED (Polygon400; Mightex Systems) to confine 470 nm blue light to a small area with a radius of approximately 3–8 μm (typically 3–4 μm), as measured at the focal plane of a 60× water immersion objective [numerical aperture (NA) = 1.0; LUMPlanFL, Olympus] (Figure 1—figure supplement 3A). To find the minimal stimulation area, the radius of illumination area was increased from 2 μm by 50% at each step (Figure 1—figure supplement 3A), and selected the smallest illumination area that elicited EPSCs. Setting the duration of illumination between 3 and 5 ms (typically 5 ms), photostimulation onto oChIEF-expressing cells reliably induced action potentials (APs) during 600-pulse trains (see Figure 1—figure supplement 2).

STP at various stimulus frequencies (5, 10, 20, and 40 Hz) was examined using 20-pulse trains. Trials of train stimulation were repeated with a prolonged interval of 30 s to avoid change in baseline properties. We observed no systematic changes in the amplitudes of the first EPSC of each train, nor did we find significant differences in the paired-pulse ratio (PPR) across stimulation trials in the same cell. For each stimulus frequency, approximately ten to twenty consecutive traces were collected and then averaged across trials to yield a mean EPSC trace. For recovery experiments, two consecutive trains of 3 or 30 stimuli at 40 Hz, with varying time intervals (0.1-10 s), were applied every 10 s or 60 s, respectively, at both excitatory synapses. For 600-stimulus trains at 10Hz, the trial was delivered only once for each cell to avoid possible after-effects caused by prolonged synaptic stimulation (Hirsch and Crepel, 1990). Prior to the stimulus trains, the EPSCs were normalized to the mean amplitude of the baseline. For the baseline measurements, paired pulses with various ISIs (20-200 ms) were delivered every 3 s (1/3 Hz) at least 20 times.

For all recordings, picrotoxin (PTX, 10 μM) was included in the recording solution to isolate glutamatergic responses. mPFC PC terminals were entirely glutamatergic as the application of 6-cyano-7-nitroquinoxaline-2,3-dione (CNQX, 20 μM), an AMPA receptor antagonist, completely abolished the monosynaptic events evoked by photostimulation. A low concentration of PTX was employed to minimize the relieving effects on spike suppression caused by blockade of GABAa receptors. We confirmed that no synaptic response was elicited by electrical stimulation in the presence of both 10 μM PTX and 20 μM CNQX.

In Figure 1—figure supplement 4A, 1 µM TTX (1 mM stock solution) and 0.1 mM 4-AP (500 mM stock solution) were added to the bath solution to block action potentials and restore presynaptic glutamate release, respectively (Figure 1—figure supplement 4). In Figure 1— figure supplement 5A, 50 µM of cyclothiazide, a positive allosteric modulator of the AMPA receptors, was included to minimize rapid desensitization of the postsynaptic receptors. In Figure 1—figure supplement 5B, 0.5 mM kynurenate was added to bath solution to prevent possible saturation of AMPA receptors, especially during maximal synaptic facilitation. For pharmacological manipulations of synaptic vesicle dynamics (Figure 4), drugs related to vesicle supply including U73122, OAG, LatB, EGTA-AM, and dynasore at indicated concentrations were added to the recording solution during baseline measurements or PTA/recovery experiments (IBI = 0.5 s). Chemicals were from Sigma-Aldrich, Merck, or Tocris.

Measurements of quantal size

Experiments for measuring asynchronous release were performed as previously described (Yoon et al., 2020). In brief, we replaced extracellular Ca2+ entirely with 2 mM Sr2+ to induce asynchronous release. A single stimulus was delivered at 3-second intervals over 100 repetitions, and asynchronous release events were detected during a 100-ms window starting 25 ms after stimulus onset for each trial. The same procedure was also executed under sham conditions (0% light intensity) to serve as a control, thus mitigating contamination from nonspecific spontaneous events and potential errors associated with in the detection algorithm. q was determined as the mean value from the first Gaussian function after fitting the EPSC distribution with either a single or double Gaussian function (Figure 3—figure supplement 1).

Estimation of Ca2+ kinetics at axonal boutons

Cytosolic [Ca2+] at single axonal boutons was estimated by employing the two-dye, dual- excitation wavelength technique as described by (Oheim et al., 1998). This method involves measuring the fluorescence ratio of a Ca2+-sensitive dye to that of a Ca2+-insensitive dye to correct for dye concentrations. We used Fluo-5F at concentrations of 150, 250, or 500 μM (dissociation constant, Kd = 2.3 μM) as the calcium indicator dye and Alexa Flour 555 (50 μM) as the reference dye. The fluorophores were from Invitrogen. Both dyes were co-loaded into presynaptic terminals of layer 2/3 pyramidal cells in mPFC through whole-cell patch clamp. Before data acquisition, the dyes were allowed to diffuse to terminal boutons for at least 30 minutes after the patch break-in, a duration sufficient for the equilibrium of dye concentrations between the soma and its axonal terminals.

Calcium imaging was performed by sequentially exciting Fluo-5F and AF555 at wavelengths of 473 and 559 nm, respectively, using a confocal laser-scanning microscope (FV1200; Olympus, Japan) equipped with a 60× water immersion objective (NA, 0.9; LUMPlanFl; Olympus, Tokyo, Japan). Once a single axonal bouton was identified, lines for scanning were drawn across boutons, oriented perpendicular to the axon (Figure 2—figure supplement 2A-B). After baseline fluorescence was measured, line scanning was proceeded while a single or train APs were induced by applying current pulses. Upon stimulation, green fluorescence but not red one increased along the vertical array of lines across the bouton (Figure 2—figure supplement 2B-C). Measurements were obtained under current-clamp conditions, with the resting potential regularly monitored. The scanning speed was adjusted to 600-700 Hz by controlling the line length. Line scan for measuring fluorescent transients in a bouton was repeated at least 3 times, and then averaged to enhance the signal-to-noise ratio. To mitigate photobleaching, minimal intensity of the laser power and maximal pinhole size were used. Emitted light was filtered through 490/590 or 575/675 bandpass filters before detection by photomultiplier tubes for green or red fluorescence, respectively.

Fluorescence measurements were converted to [Ca2+]i using the ratio of background- subtracted fluorescence (F = F0 - Fb) from two dyes. The fluorescence ratio (R = Fgreen/Fred) in data traces were converted to [Ca2+]i using the following equation: Calibration parameters were obtained using an in-cell calibration protocol (Helmchen, 2011), wherein the minimum (Rmin) and maximum ratio (Rmax) values were established through the use of intracellular solutions containing 10 mM EGTA or 10 mM [Ca2+], respectively.

When Ca2+ is extruded with a rate constant, γ, after an AP-induced short pulse of Ca2+ influx to a single compartment, [Ca2+]i initially rises to ACa and then undergoes a mono-exponential decay with a time constant (τCa). Under this framework, ACa and τCa depend on Ca2+ binding ratios of endo- and exogenous Ca2+ buffers (denoted as κS and κB, respectively) as following equations (Neher, 1995; Kim et al., 2003):

, where Δ[Ca2+]T is an increment in total [Ca2+] (free plus bound) elicited by an AP. Ca2+ binding ratio indicates the ratio of concentration changes in the Ca2+-buffer complex relative to free [Ca2+] change, and linearized κB is calculated from the dye concentration and the change from [Ca2+]1 to [Ca2+]2 as follows: The total dye concentration, [B]t, was taken as [Fluo-5F] in the patch pipette. With κB determined from Equation 4 and τCa from the decay of a Ca2+ transient, κS can be estimated by plotting τ versus κB and applying a linear fit to the plot, which gives the x-intercept as -(1 + κS). Once determined, κS can be used to calculate Δ[Ca2+]T.

Optimization of k1 and pv from failure rates

Let n0 = the number of docked SVs before arrival of a first AP; n1 = the number of docked SVs remaining after a first AP (i.e. n0 - n1 vesicles are released); n2 = the number of docked SVs just before arrival of a 2nd AP (ISI = 20 ms; Figure 3—figure supplement 2). If the number of release sites (N), forward and reverse rates for vesicle docking (k1 and b1, respectively), mean release site occupancy (pocc) and vesicular fusion probability upon an AP (pv) are given, we can predict P(n0, n1, n2), the probability for [n0, n1, n2], as following steps.

Probability for a synapse harbouring n0 docked SVs just before arrival of a first AP is calculated as:

Probability that n1 vesicles remain after a first AP when the initial number of docked SVs is n0 [P(n1|n0)] is calculated as:

When forward and backward rate constants are k1 and b1, respectively, probability that n2 vesicles are docked just before arrival of a 2nd AP when n1 SVs remain after a 1st AP (Δt = 20 ms) [P(n2 | n1)] is predicted as follows: Assuming that docking and undocking follow a Poisson process,

Combining probabilities calculated from above three steps, we can calculate P(n0, n1, n2): P(n0, n1, n2) = P(n0) x P(n1|n0) x P(n2|n1)

Once P(n0, n1, n2) is given, we can predict probability for double failure (P00) and probability for the 1st success and the 2nd failure (P10) as:

Probability for the 1st failure and the 2nd success (P01) and probability for double success (P11) were calculated as:

, in which P(F1) and P(S1) are probability for the 1st failure and the 1st success, respectively, and these are calculated as: To optimize k1 and pv, we used the Matlab routine, fmincon, in which the sum of squared errors (= observed value minus predicted value) was set as a cost function to be minimized. Lower and upper bounds for k1 and pv were set as [4, 8] and [0.8, 1], respectively. Release probability (pr) was calculated from the observed 1st failure rate, P(F1), as N was assumed to be six as shown in Figure 3. Other parameters (b1, and pocc) were set according to following relationships:

Calculation of probability for double failure

Double failure probability, P00, can be more easily predicted. By the Bayes rule, , where F1 and F2 means 1st and 2nd failure. Moreover, and n0 = n1 because of the 1st failure. Thus, the probability for two consecutive failures can be calculated as: P(F1) and N are observed values. For calculation in Figure 3—figure supplement 2, k1 and pv were set free variables. Other parameters were set under the following constrains: P(F1) = (1 – pr)N, pr = pv pocc, and pocc = k1 / (k1 + b1).

Behavioral tests

Trace fear conditioning

For trace fear conditioning (tFC) (Gilmartin et al., 2013), individual rats were placed in the training chamber (30.5 x 25.4 x 30.5 cm; Coulbourn Instruments) consisting of metal grid floor connected to an electrical stimulator (H10-11R-TC-SF, Coulbourn Instruments) for delivery of a footshock. After a 5 min habituation period, rats received five pairings of a 20 s tone conditioned stimulus (CS, 2.5 kHz, 75 dB), followed by empty 20s trace period and a unconditioned stimulus (US; 1 s, 0.7 mA footshock) with a pseudorandom inter- trial interval (ITI) of 240 ± 20 s. During this training session, rats learned to associate the auditory CS with the shock US. The next day, the rats were tested for memory of CS-US trace association (called tone test). The acquisition of fear learning, often indicated by the extent of freezing to the CS, was assessed in a novel chamber (a white hexagonal enclosure) in a dimly lit room. During this tone (or retrieval) test, rats were first allowed to explore the new context for 5 min (habituation), followed by twelve 20 s CS presentations with a variable ITI of 120 ± 20 s. This entire session was repeated on the next day for assessment of extinction learning. Identities of the chambers were also determined with the presence of distinct odors; the training and testing chambers were cleaned with 70% ethanol and 1% acetic acid respectively before and after each session. For all experimental sessions, the activity of animals in the chambers was recorded at 30 frames per second using the EthoVision XT (Noldus Information Technology, Wageningen, Netherlands), and stored as a video file. The freezing behavior, defined as behavioral immobility other than respiratory movement, was assessed using an open-source video analysis pipeline ezTrack (Pennington et al., 2019). The percentage of freezing was calculated as a total duration of freezing divided by the total duration of observation.

Elevated Plus Maze

Elevated plus maze (EPM) test was performed using a plus-shaped apparatus elevated 60 cm above the ground. The maze consisted of two open arms (50 x 10 x 0.5 cm (H)), two closed arms (50 x 10 x 40 cm (H)) and a center area (10 x 10 cm). Rats were individually placed on the center area, facing an open arm, and the path of the animal was recorded with a video camera for 5 min. We analyzed the total distance traveled for 5 min and time spent in each arm. Animals were tracked offline using the open-source tracking system ezTrack (Pennington et al., 2019).

Open field test

Open field exploration test (OFT) was performed in an open field apparatus in a dimly lit room. The open field consisted of a white plastic board (1.20 m in diameter) surrounded by white plastic walls (50 cm in height). Individual rats were placed in a center zone (0.6 m in diameter), and allowed to freely explore the apparatus for 10 min. Time spent in the center zone and total distance traveled were analyzed for initial 5 min. Animals were tracked offline using the open-source tracking system ezTrack (Pennington et al., 2019).

Histology and immunohistochemistry

At the end of the behavioral experiments, rats were anesthetized using isoflurane inhalation and transcardially perfused with PBS followed by 4% paraformaldehyde (PFA, T&I, Korea). Brains were extracted and post-fixed in 4% PFA solution for 24 h. Perfused brains were sliced at 100 μm for estimation of GFP expression or 50 μm for immunohistochemistry. For analysis of c-Fos expression, rats were returned to their home cages immediately after the training session and perfused approximately 90 minutes later (Arime and Akiyama, 2017). Brains were then cryopreserved, frozen, and sectioned coronally. Immunofluorescence staining was conducted on free-floating slices using a rabbit monoclonal anti c-fos (1:1000, Cell Signaling Technology, cat# 2250s) and mouse monoclonal anti-GAD67 (1:500, Millipore, cat# mab5406). Fluorescent secondary antibodies included Cy5-conjugated goat anti-rabbit IgG (1:1000, Abcam, cat# ab97077) and Cy3-conjugated goat anti-mouse IgG (1:1000, Abcam, cat# ab97035). Brain slices were permeabilized and blocked with PBS containing 0.5% Triton X-100 and 5% normal goat serum (NGS) for 1h at room temperature. Slices were incubated overnight at 4 °C with primary antibodies, followed by 4 h incubation with secondary antibodies at room temperature. Slices were then mounted with Antifading Mounting medium (Abcam, cat# ab104139). Confocal images were acquired using a Leica TCS SP8 microscope set to the same laser intensity and acquisition parameters to compare the immunosignals in sections from Scr and Syt7 KD groups. For cell counting and colocalization, fluorescent protein-positive cells were automatically quantified with spot-detection algorithms in Imaris 9.5 (Oxford instruments).

Data analysis

Data were analyzed and presented using ClampFit (Molecular Devices), Igor Pro (Wavemetrics), MATLAB (Mathworks), and Prism (GraphPad). Data points and error bars represent the mean ± standard error of the mean. All statistical analyses were conducted using two-tailed comparisons. Sample sizes and statistical tests for each group are detailed in the figure legends.

Numerical integration of the simple refilling model

This model posits reversible docking of vesicles at finite release sites (Nmax) with forward (k1) and reverse rate constants (b1) (Hosoi et al., 2007; Neher and Sakaba, 2008). According to this scheme, dn/dt = k1 u - b1 n - pv n δ(t - tAP), where u is the number of unoccupied sites (= Nmax - n); tAP is the timing of AP firing; and δ is the Dirac delta function. The pv was assumed to be unity (Figure 3—figure supplement 2). The sum of basal k1 (= k1,b) and b1 was be estimated as 23/s from the dependence PPR on ISIs (Figure 3—figure supplement 2Ab). From the baseline occupancy of 0.3 and pocc = k1,b / (k1,b + b1), we could estimate baseline k1 and b1 as 6.9/s, and 16.1/s, respectively. Ca2+ and Syt7-dependent increase of k1 was modeled as k1 = k1,b + K1,max [Syt7: n Ca2+], where [Syt7: n Ca2+] represents the fraction of full Ca2+-bound Syt7. For deterministic simulations, we time-integrated a set of differential equations for each model using Euler methods with a time step of 1 ms. All calculations were performed on the platform of Matlab (R2022b, Mathworks, USA).

eLife Assessment

This valuable work explores how synaptic activity encodes information during memory tasks. All reviewers agree that the work is of very high quality and that the methodological approach is praiseworthy. Although the experimental data support the possibility that phospholipase diacylglycerol signaling and synaptotagmin 7 (Syt7) dynamically regulate the vesicle pool required for presynaptic release, concerns remain that the central finding of paired-pulse depression at very short intervals may be more likely due to Ca²⁺ channel inactivation rather than vesicle pool depletion. Overall, this is a solid study although the results warrant consideration of alternative interpretations.

Reviewer #3 (Public review):

The central issue for evaluating the overfilling hypothesis is the identity of the mechanism that causes the very potent (>80% when inter pulse is 20 ms), but very quickly reverting (< 50 ms) paired pulse depression (Fig 1G, I). To summarize: the logic for overfilling at local cortical L2/3 synapses depends critically on the premise that probability of release (pv) for docked and fully primed vesicles is already close to 100%. If so, the reasoning goes, the only way to account for the potent short-term enhancement seen when stimulation is extended beyond 2 pulses would be by concluding that the readily releasable pool overfills. However, the conclusion that pv is close to 100% depends on the premise that the quickly reverting depression is caused by exocytosis dependent depletion of release sites, and the evidence for this is not strong in my opinion. Caution is especially reasonable given that similarly quickly reverting depression at Schaffer collateral synapses, which are morphologically similar, was previously shown to NOT depend on exocytosis (Dobrunz and Stevens 1997). Note that the authors of the 1997 study speculated that Ca2+-channel inactivation might be the cause, but did not rule out a wide variety of other types of mechanisms that have been discovered since, including the transient vesicle undocking/re-docking (and subsequent re-priming) reported by Kusick et al (2020), which seems to have the correct timing.

In an earlier round of review, I suggested raising extracellular Ca2+, to see if this would increase synaptic strength. This is a strong test of the authors' model because there is essentially no room for an increase in synaptic strength. The authors have now done experiments along these lines, but the result is not clear cut. On one hand, the new results suggest an increase in synaptic strength that is not compatible with the authors' model; technically the increase does not reach statistical significance, but, likely, this is only because the data set is small and the variation between experiments is large. Moreover, a more granular analysis of the individual experiments seems to raise more serious problems, even supporting the depletion-independent counter hypothesis to some extent. On the other hand, the increase in synaptic strength that is seen in the newly added experiments does seem to be less at local L2/3 cortical synapses compared to other types of synapses, measured by other groups, which goes in the general direction of supporting the critical premise that pv is unusually high at L2/3 cortical synapses. Overall, I am left wishing that the new data set were larger, and that reversal experiments had been included as explained in the specific points below.

Specific Points:

(1) One of the standard methods for distinguishing between depletion-dependent and depletion-independent depression mechanisms is by analyzing failures during paired pulses of minimal stimulation. The current study includes experiments along these lines showing that pv would have to be extremely close to 1 when Ca2+ is 1.25 mM to preserve the authors' model (Section "High double failure rate ..."). Lower values for pv are not compatible with their model because the k1 parameter already had to be pushed a bit beyond boundaries established by other types of experiments. The authors now report a mean increase in synaptic strength of 23% after raising Ca to 2.5 mM. The mean increase is not quite statistically significant, but this is likely because of the small sample size. I extracted a 95% confidence interval of [-4%, +60%] from their numbers, with a 92% probability that the mean value of the increase in the full population is > 5%. I used the 5% value as the greatest increase that the model could bear because 5% implies pv < 0.9 using the equation from Dodge and Rahamimoff referenced in the rebuttal. My conclusion from this is that the mean result, rather than supporting the model, actually undermines it to some extent. It would have likely taken 1 or 2 more experiments to get above the 95% confidence threshold for statistical significance, but this is ultimately an arbitrary cut off.

(2) The variation between experiments seems to be even more problematic, at least as currently reported. The plot in Figure 3-figure supplement 3 (left) suggests that the variation reflects true variation between synapses, not measurement error. And yet, synaptic strength increased almost 2-fold in 2 of the 8 experiments, which back extrapolates to pv < 0.2. If all of the depression is caused by depletion as assumed, these individuals would exhibit paired pulse facilitation, not depression. And yet, from what I can tell, the individuals depressed, possibly as much as the synapses with low sensitivity to Ca2+, arguing against the critical premise that depression equals depletion, and even arguing - to some extent - for the counter hypothesis that a component of the depression is caused by a mechanism that is independent of depletion. I would strongly recommend adding an additional plot that documents the relationship between the amount of increase in synaptic strength after increasing extracellular Ca2+ and the paired pulse ratio as this seems central.

(3) Decrease in PPR. The authors recognize that the decrease in the paired-pulse ratio after increasing Ca2+ seems problematic for the overfilling hypothesis by stating: "Although a reduction in PPR is often interpreted as an increase in pv, under conditions where pv is already high, it more likely reflects a slight increase in pocc or in the number of TS vesicles, consistent with the previous estimates (Lin et al., 2025)." I looked quickly, but did not immediately find an explanation in Lin et al 2025 involving an increase in pocc or number of TS vesicles, much less a reason to prefer this over the standard explanation that reduced PPR indicates an increase in pv. The authors should explain why the most straightforward interpretation is not the correct one in this particular case to avoid the appearance of cherry picking explanations to fit the hypothesis.

(4) The authors concede in the rebuttal that mean pv must be < 0.7, but I couldn't find any mention of this within the manuscript itself, nor any explanation for how the new estimate could be compatible with the value of > 0.99 in the section about failures.

(5) Although not the main point, comparisons to synapses in other brain regions reported in other studies might not be accurate without directly matching experiments. As it is, 2 of 8 synapses got weaker instead of stronger, hinting at possible rundown, but this cannot be assessed because reversibility was not evaluated. In addition, comparing axons with and without channel rhodopsins might be problematic because the channel rhodopsins might widen action potentials.

(6) Perhaps authors could double check with Schotten et al about whether PDBu does/does not decrease the latency between osmotic shock and transmitter release. This might be an interesting discrepancy, but my understanding is that Schotten et al didn't acquire information about latency because of how the experiments were designed.

(7) The authors state: "These data are difficult to reconcile with a model in which facilitation is mediated by Ca2+-dependent increases in pv." However, I believe that discarding the premise that depression is always caused by depletion would open up wide range of viable possibilities.

Author response:

The following is the authors’ response to the previous reviews

Reviewer #3 (Public review):

The central issue for evaluating the overfilling hypothesis is the identity of the mechanism that causes the very potent (>80% when inter pulse is 20 ms), but very quickly reverting (< 50 ms) paired pulse depression (Fig 1G, I). To summarize: the logic for overfilling at local cortical L2/3 synapses depends critically on the premise that probability of release (pv) for docked and fully primed vesicles is already close to 100%. If so, the reasoning goes, the only way to account for the potent short-term enhancement seen when stimulation is extended beyond 2 pulses would be by concluding that the readily releasable pool overfills. However, the conclusion that pv is close to 100% depends on the premise that the quickly reverting depression is caused by exocytosis dependent depletion of release sites, and the evidence for this is not strong in my opinion. Caution is especially reasonable given that similarly quickly reverting depression at Schaffer collateral synapses, which are morphologically similar, was previously shown to NOT depend on exocytosis (Dobrunz and Stevens 1997). Note that the authors of the 1997 study speculated that Ca2+-channel inactivation might be the cause, but did not rule out a wide variety of other types of mechanisms that have been discovered since, including the transient vesicle undocking/re-docking (and subsequent re-priming) reported by Kusick et al (2020), which seems to have the correct timing.

Thank you for your comments on an alternative possibility besides Ca2+ channel inactivation. Kusick et al. (2020) showed that transient destabilization of docked vesicle pool is recovered within 14 ms after stimulation. This rapid recovery implies that post-stimulation undocking events might be largely resolved before the 20 ms inter-stimulus interval (ISI) used in our paired-pulse ratio (PPR) experiments, arguing against the possibility that post-AP undocking/re-docking events significantly influence PPR measured at 20 ms ISI. Furthermore, Vevea et al. (2021) showed that post-stimulus undocking is facilitated in synaptotagmin-7 (Syt7) knockout synapses. In our study, Syt7 knockdown did not affect PPR at 20 ms ISI, suggesting that the undocking process described in Kusick et al. may not be a major contributor to the paired-pulse depression observed at 20 ms interval in our study. Therefore, it is unlikely that transient vesicle undocking primarily underlies the strong PPD at 20 ms ISI in our experiments. Taken together, the undocking/redocking dynamics reported by Kusick et al. are too rapid to affect PPR at 20 ms ISI, and our Syt7 knockdown data further argue against a significant role of this process in the PPD observed at 20 ms interval.

In an earlier round of review, I suggested raising extracellular Ca2+, to see if this would increase synaptic strength. This is a strong test of the authors' model because there is essentially no room for an increase in synaptic strength. The authors have now done experiments along these lines, but the result is not clear cut. On one hand, the new results suggest an increase in synaptic strength that is not compatible with the authors' model; technically the increase does not reach statistical significance, but, likely, this is only because the data set is small and the variation between experiments is large. Moreover, a more granular analysis of the individual experiments seems to raise more serious problems, even supporting the depletion-independent counter hypothesis to some extent. On the other hand, the increase in synaptic strength that is seen in the newly added experiments does seem to be less at local L2/3 cortical synapses compared to other types of synapses, measured by other groups, which goes in the general direction of supporting the critical premise that pv is unusually high at L2/3 cortical synapses. Overall, I am left wishing that the new data set were larger, and that reversal experiments had been included as explained in the specific points below.

Specific Points:

(1) One of the standard methods for distinguishing between depletion-dependent and depletion-independent depression mechanisms is by analyzing failures during paired pulses of minimal stimulation. The current study includes experiments along these lines showing that pv would have to be extremely close to 1 when Ca2+ is 1.25 mM to preserve the authors' model (Section "High double failure rate ..."). Lower values for pv are not compatible with their model because the k1 parameter already had to be pushed a bit beyond boundaries established by other types of experiments.

It should be noted that we did not arbitrarily pushed the k1 parameter beyond boundaries, but estimated the range of k1 based on the fast time constant for recovery from paired pulse depression as shown in Fig. 3-S2-Ab.

The authors now report a mean increase in synaptic strength of 23% after raising Ca to 2.5 mM. The mean increase is not quite statistically significant, but this is likely because of the small sample size. I extracted a 95% confidence interval of [-4%, +60%] from their numbers, with a 92% probability that the mean value of the increase in the full population is > 5%. I used the 5% value as the greatest increase that the model could bear because 5% implies pv < 0.9 using the equation from Dodge and Rahamimoff referenced in the rebuttal. My conclusion from this is that the mean result, rather than supporting the model, actually undermines it to some extent. It would have likely taken 1 or 2 more experiments to get above the 95% confidence threshold for statistical significance, but this is ultimately an arbitrary cut off.

Our key claim in Fig. 3-S3 is not the statistical non-significance of EPSC changes, but the small magnitude of the change (1.23-fold). This small increase is far less than the 3.24-fold increase predicted by the fourth-power relationship (D&R equation, Dodge & Rahamimoff, 1967), which would be valid under the conditions that the fusion probability of docked vesicles (pv) is not saturated. We do not believe that addition of new experiments would increase the magnitude of EPSC change as high as the Dodge & Rahamimoff equation predicts, even if more experiments (n) yielded a statistical significance. In other words, even a small but statistically significant EPSC changes would still contradict with what we expect from low pv synapses. It should be noted that our main point is the extent of EPSC increase induced by high external [Ca2+], not a p-value. In this regard, it is hard for us to accept the Reviewer’s request for larger sample size expecting lower p-value.

Although we agree to Reviewer’s assertion that our data may indicate a 92% probability for the high Ca2+ -induced EPSC increases by more than 5%, we do not agree to the Reviewer’s interpretation that the EPSC increase necessarily implies an increase in pv. We are sorry that we could not clearly understand the Reviewer’s inference that the 5% increase of EPSCs implies pv < 0.9. Please note that release probability (pr) is the product of pv and the occupancy of docked vesicles in an active zone (pocc). We imagine that this inference might be under the premise that pocc is constant irrespective of external [Ca2+]. Contrary to the Reviewer’s premise, Figure 2c in Kusick et al. (2020) showed that the number of docked SVs increased by c. a. 20% upon increasing external [Ca2+] to 2 mM. Moreover, Figure 7F in Lin et al. (2025) demonstrated that the number of TS vesicles, equivalent to pocc increased by 23% at high external [Ca2+]. These extents of pocc increases are similar to our magnitude of high external Ca2+ -induced increase in EPSC (1.23-fold). Of course, it is possible that both increase of pocc and pv contributed to the high [Ca2+]o-induced increase in EPSC. The low PPR and failure rate analysis, however, suggest that pv is already saturated in baseline conditions of 1.3 mM [Ca2+]o and thus it is more likely that an increase in pocc is primarily responsible for the 1.23-fold increase. Moreover, the 1.23-fold increase, does not match to the prediction of the D&R equation, which would be valid at synapses with low pv. Therefore, interpreting our observation (1.23-fold increase) as a slight increase in pocc is rather consistent with recent papers (Kusick et al.,2020; Lin et al., 2025) as well as our other results supporting the baseline saturation of pv as shown in Figure 2 and associated supplement figures (Fig. 2-S1 and Fig. 2-S2).

(2) The variation between experiments seems to be even more problematic, at least as currently reported. The plot in Figure 3-figure supplement 3 (left) suggests that the variation reflects true variation between synapses, not measurement error.

Note that there was a substantial variance in the number of docked or TS vesicles at baseline and its fold changes at high external Ca2+ condition in previous studies too (Lin et al., 2025; Kusick et al., 2020). Our study did not focus on the heterogeneity but on the mean dynamics of short-term plasticity at L2/3 recurrent synapses. Acknowledging this, the short-term plasticity of these synapses could be best explained by assuming that vesicular fusion probability (pv) is near to unity, and that release probability is regulated by pocc. In other words, even though pv is near to unity, synaptic strength can increase upon high external [Ca2+], if the baseline occupancy of release sites (pocc) is low and pocc is increased by high [Ca2+]. Lin et al. (2025) showed that high external [Ca2+] induces an increase in the number of TS vesicles (equivalent to pocc) by 23% at the calyx synapses. Different from our synapses, the baseline pv (denoted as pfusion in Lin et al., 2025) of the calyx synapse is not saturated (= 0.22) at 1.5 mM external [Ca2+], and thus the calyx synapses displayed 2.36-fold increase of EPSC at 2 mM external [Ca2+], to which increases in pocc as well as in pv (from 0.22 to 0.42) contributed. Therefore, the small increase in EPSC (= 23%) supports that pv is already saturated at L2/3 recurrent synapses.

And yet, synaptic strength increased almost 2-fold in 2 of the 8 experiments, which back extrapolates to pv < 0.2.

We are sorry that we could not understand the first comment in this paragraph. Could you explain in detail why two-fold increase implies pv < 0.2?

If all of the depression is caused by depletion as assumed, these individuals would exhibit paired pulse facilitation, not depression. And yet, from what I can tell, the individuals depressed, possibly as much as the synapses with low sensitivity to Ca2+, arguing against the critical premise that depression equals depletion, and even arguing - to some extent - for the counter hypothesis that a component of the depression is caused by a mechanism that is independent of depletion.

For the first statement in this paragraph, we imagine that ‘the depression’ means paired pulse depression (PPD). If so, we can not understand why depletion-dependent PPD should lead to PPF. If the paired pulse interval is too short for docked vesicles to be replenished, the first pulse-induced vesicle depletion would result in PPD. We are very sorry that we could not understand Reviewer’s subsequent inference, because we could not understand the first statement.

I would strongly recommend adding an additional plot that documents the relationship between the amount of increase in synaptic strength after increasing extracellular Ca2+ and the paired pulse ratio as this seems central.

We found no clear correlation of EPSC1 with PPR changes (ΔPPR) as shown in the figure below.

Plot of PPR changes as a function of EPSC1.

(3) Decrease in PPR. The authors recognize that the decrease in the paired-pulse ratio after increasing Ca2+ seems problematic for the overfilling hypothesis by stating: "Although a reduction in PPR is often interpreted as an increase in pv, under conditions where pv is already high, it more likely reflects a slight increase in pocc or in the number of TS vesicles, consistent with the previous estimates (Lin et al., 2025)."

We admit that there is a logical jump in our statement you mentioned here. We appreciate your comment. We re-wrote that part in the revised manuscript (line 285) as follows:

“Recent morphological and functional studies revealed that elevation of [Ca2+]o induces an increase in the number of TS or docked vesicles to a similar extent as our observation (Kusick et al., 2020; Lin et al., 2025), raising a possibility that an increase in pocc is responsible for the 1.23-fold increase in EPSC at high [Ca2+]o . A slight but significant reduction in PPR was observed under high [Ca2+]o too. An increase in pocc is thought to be associated with that in the baseline vesicle refilling rate. While PPR is always reduced by an increase in pv, the effects of refilling rate to PPR is complicated. For example, PPR can be reduced by both a decrease (Figure 2—figure supplement 1) and an increase (Lin et al., 2025) in the refilling rate induced by EGTA-AM and PDBu, respectively. Thus, the slight reduction in PPR is not contradictory to the possible contribution of pocc to the high [Ca2+]o effects.”

I looked quickly, but did not immediately find an explanation in Lin et al 2025 involving an increase in pocc or number of TS vesicles, much less a reason to prefer this over the standard explanation that reduced PPR indicates an increase in pv.

Fig. 7F of Lin et al. (2025) shows an 1.23-fold increase in the number of TS vesicles by high external [Ca2+]. The same figure (Fig. 7E) in Lin et al. (2025) also shows a two-fold increase of pfusion (equivalent to pv in our study) by high external [Ca2+] (from 0.22 to 0.42,). Because pocc is the occupancy of TS vesicles in a limited number of slots in an active zone, the fold change in the number of TS vesicles should be similar to that of pocc.

The authors should explain why the most straightforward interpretation is not the correct one in this particular case to avoid the appearance of cherry picking explanations to fit the hypothesis.

The results of Lin et al. (2025) indicate that high external [Ca2+] induces a milder increase in pocc (23%) compared to pv (190%) at the calyx synapses. Because the extent of pocc increase is much smaller than that of pv and multiple lines of evidence in our study support that the baseline pv is already saturated, we raised a possibility that an increase in pocc would primarily contribute to the unexpectedly low increase of EPSC at 2.5 mM [Ca2+]o. As mentioned above, our interpretation is also consistent with the EM study of Kusick et al. (2020). Nevertheless, the reduction of PPR at 2.5 mM Ca2+ seems to support an increase in pv, arguing against this possibility. On the other hand, because pocc = k1/(k1+b1) under the simple vesicle refilling model (Fig. 3-S2Aa), a change in pocc should associate with changes in k1 and/or b1. While PPR is always reduced by an increase in pv, the effects of refilling rate to PPR is complicated. For example, despite that EGTA-AM would not increase pv, it reduced PPR probably through reducing refilling rate (Fig. 2-S1). On the contrary, PDBu is thought to increase k1 because it induces two-fold increase of pocc (Fig. 7L of Lin et al., 2025). Such a marked increase of pocc, rather than pv, seems to be responsible for the PDBu-induced marked reduction of PPR (Fig. 7I of Lin et al., 2025), because PDBu induced only a slight increase in pv (Fig. 7K of Lin et al., 2025). Therefore, the slight reduction of PPR is not contradictory to our interpretation that an increase in pocc might be responsible for the slight increase in EPSC induced by high [Ca2+]o.

(4) The authors concede in the rebuttal that mean pv must be < 0.7, but I couldn't find any mention of this within the manuscript itself, nor any explanation for how the new estimate could be compatible with the value of > 0.99 in the section about failures.

We have never stated in the rebuttal or elsewhere that the mean pv must be < 0.7. On the contrary, both of our manuscript and previous rebuttals consistently argued that the baseline pv is already saturated, based on our observations including low PPR, tight coupling, high double failure rate and the minimal effect of external Ca2+ elevation.

(5) Although not the main point, comparisons to synapses in other brain regions reported in other studies might not be accurate without directly matching experiments.

Please understand that it not trivial to establish optimal experimental settings for studying other synapses using the same methods employed in the study. We think that it should be performed in a separate study. Furthermore, we have already shown in the manuscript that action potentials (APs) evoked by oChIEF activation occur in a physiologically natural manner, and the STP induced by these oChIEF-evoked APs is indistinguishable from the STP elicited by APs evoked by dual-patch electrical stimulation. Therefore, we believe that our use of optogenetic stimulation did not introduce any artificial bias in measuring STP.

As it is, 2 of 8 synapses got weaker instead of stronger, hinting at possible rundown, but this cannot be assessed because reversibility was not evaluated. In addition, comparing axons with and without channel rhodopsins might be problematic because the channel rhodopsins might widen action potentials.

We continuously monitored series resistance and baseline EPSC amplitude throughout the experiments. The figure below shows the mean time course of EPSCs at two different [Ca2+]o. As it shows, we observed no tendency for run-down of EPSCs during experiments. If any, such recordings were discarded from analysis. In addition, please understand that there is a substantial variance in the number of docked vesicles at both baseline and high external Ca2+ (Lin et al., 2025; Kusick et al., 2020) as well as short-term dynamics of EPSCs at our synapses.

Time course of normalized amplitudes of the first EPSCs during paired-pulse stimulation at 20 ms ISI in control and in the elevated external Ca2+ (n = 8).

(6) Perhaps authors could double check with Schotten et al about whether PDBu does/does not decrease the latency between osmotic shock and transmitter release. This might be an interesting discrepancy, but my understanding is that Schotten et al didn't acquire information about latency because of how the experiments were designed.

Schotten et al. (2015) directly compared experimental and simulation data for hypertonicity-induced vesicle release. They showed a pronounced acceleration of the latency as the tonicity increases (Fig. 2-S2), but this tonicity-dependent acceleration was not reproduced by reducing the activation energy barrier for fusion (ΔEa) in their simulations (Fig. 2-S1). Thus, the authors mentioned that an unknown compensatory mechanism counteracting the osmotic perturbation might be responsible for the tonicity-dependent changes in the latency. Importantly, their modeling demonstrated that reducing ΔEa, which would correspond to increasing pv results in larger peak amplitudes and shorter time-to-peak, but did not accelerate the latency. Therefore, there is currently no direct explanation for the notion that PDBu or similar manipulations shorten latency via an increase in pv.

(7) The authors state: "These data are difficult to reconcile with a model in which facilitation is mediated by Ca2+-dependent increases in pv." However, I believe that discarding the premise that depression is always caused by depletion would open up wide range of viable possibilities.

We hope that Reviewer understands the reasons why we reached the conclusion that the baseline pv is saturated at our synapses. First of all, strong paired pulse depression (PPD) cannot be attributed to Ca2+ channel inactivation because Ca2+ influx at the axon terminal remained constant during 40 Hz train stimulation (Fig.2 -S2). Moreover, even if Ca2+ channel inactivation is responsible for the strong PPD, this view cannot explain the delayed facilitation that emerges subsequent pulses (third EPSC and so on) in the 40 Hz train stimulation (Fig. 1-4), because Ca2+ channel inactivation gradually accumulates during train stimulations as directly shown by Wykes et al. (2007) in chromaffin cells. Secondly, the strong PPD and very fast recovery from PPD indicates very fast refilling rate constant (k1). Under this high k1, the failure rates were best explained by pv close to unity. Thirdly, the extent of EPSC increase induced by high external Ca2+ was much smaller than other synapses such as calyx synapses at which pv is not saturated (Lin et al., 2025), and rather similar to the increases in pocc estimated at calyx synapses or the EM study (Kusick et al., 2020; Lin et al., 2025).

Reference

Wykes et al. (2007). Differential regulation of endogenous N-and P/Q-type Ca2+ channel inactivation by Ca2+/calmodulin impacts on their ability to support exocytosis in chromaffin cells. Journal of Neuroscience, 27(19), 5236-5248.

Reviewer #3 (Recommendations for the authors):

I continue to think that measuring changes in synaptic strength when raising extracellular Ca2+ is a good experiment for evaluating the overfilling hypothesis. Future experiments would be better if the authors would include reversibility criteria to rule out rundown, etc. Also, comparisons to other types of synapses would be stronger if the same experimenter did the experiments at both types of synapses.

We observed no systemic tendency for run-down of EPSCs during these experiments (Author response image 2). Furthermore, the observed variability is well within the expected variance range in the number of docked vesicles at both baseline and high external Ca²⁺ (Lin et al., 2025; Kusick et al., 2020) and reflects biological variability rather than experimental artifact. Therefore, we believe that additional reversibility experiments are not warranted. However, we are open to further discussion if the Reviewer has specific methodological concerns not resolved by our present data.

For the second issue, as mentioned above, we think that studying at other synapse types should be done in a separate study.

Version 4

Abstract

Abstract

Short-term facilitation of recurrent excitatory synapses within the cortical network has been proposed to support persistent activity during working memory tasks, yet the underlying mechanisms remain poorly understood. We characterized short-term plasticity at the local excitatory synapses in layer 2/3 of the rat medial prefrontal cortex and studied its presynaptic mechanisms. Low-frequency stimulation induced slowly developing facilitation, whereas high-frequency stimulation initially induced strong depression followed by rapid facilitation. This non-monotonic delayed facilitation after a brief depression resulted from a high vesicular fusion probability and slow activation of Ca2+-dependent vesicle replenishment, which led to the overfilling of release sites beyond their basal occupancy. Pharmacological and gene knockdown (KD) experiments revealed that the facilitation was mediated by phospholipase C/diacylglycerol signaling and synaptotagmin 7 (Syt7). Notably, Syt7 KD abolished facilitation and slowed the refilling rate of vesicles with high fusion probability. Furthermore, Syt7 deficiency in layer 2/3 pyramidal neurons impaired the acquisition of trace fear memory and reduced c-Fos activity. In conclusion, Ca2+- and Syt7-dependent overfilling of release sites mediates synaptic facilitation at layer 2/3 recurrent excitatory synapses and contributes to temporal associative learning.

Introduction

Synapses in the mammalian central nervous system undergo activity-dependent modulation of synaptic weight during repetitive stimulation, known as short-term synaptic plasticity (STP). STP is mediated by presynaptic mechanisms that regulate the release probability (pr), size of the readily releasable pool (RRP), and kinetics of vesicle replenishment of RRP (Zucker and Regehr, 2002; Jackman and Regehr, 2017; Neher and Brose, 2018). Short-term depression has been observed at synapses with a high baseline pr, primarily because of the rapid depletion of releasable vesicles during high-frequency stimulation (HFS) (Lin et al., 2022). Conversely, vesicle depletion is less pronounced at synapses with a low baseline pr, and cumulative increases in pr during HFS are considered to mediate short-term facilitation (Rozov et al., 2001; Pan and Zucker, 2009; Aldahabi et al., 2024). Central synapses exhibit a wide spectrum of STP depending on different combinations of presynaptic factors mediating short-term depression and facilitation.

Previously, facilitation was modeled as an increase in pr after an action potential (AP) and its slow decay during the inter-spike interval (ISI), resulting in a cumulative increase in pr during HFS (Markram et al., 1998; Dittman et al., 2000). Recent studies have suggested that the number of docking (or release) sites in an active zone is limited and only partially occupied by releasable vesicles in the resting states (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Therefore, pr is determined by the vesicular fusion probability (pv) and release site occupancy (pocc). However, distinguishing the individual contributions of pv and pocc to pr is challenging (Silva et al., 2021; Neher, 2024). It is also unclear whether facilitation is mediated by activity-dependent increases in the pv, pocc, or both.

Two recent models for vesicle priming and release view the ‘priming’ as a two-step reversible process that leads to a release-ready state. The release-ready vesicles could be vesicles in a tightly docked state (TS) or those occupying specialized docking sites (DS), and are supplied from vesicles in a loosely docked state (LS) or those occupying a distinct replacement sites (RS), which are called LS/TS model (Aldahabi et al., 2024; Neher, 2024) and RS/DS model (Miki et al., 2016; Silva et al., 2021), respectively. Common to both models is the assumption that release-ready vesicles constitute only a fraction of all docked vesicles under resting conditions and that this fraction is up- and down-regulated by synaptic activity.

Synaptotagmin-7 (Syt7) was recently found to play a crucial role in facilitation at different types of central synapses (Jackman et al., 2016; Chen et al., 2017; Martinetti et al., 2022). In addition to its role in facilitation, Syt7 accelerates refilling of the RRP after depletion and mediates asynchronous release (Bacaj et al., 2013; Liu et al., 2014; Jackman et al., 2016; Tawfik et al., 2021). Syt7-dependent synaptic facilitation is interpreted as an activity-dependent increase in pv (Turecek et al., 2016; Jackman and Regehr, 2017; Turecek et al., 2017; Norman et al., 2023). However, it is unclear whether Syt7 contributes to facilitation through a Ca2+-dependent increase in pocc (i.e., overfilling). Overfilling has been proposed as a mechanism supporting facilitation at the Drosophila neuromuscular junction (Kobbersmed et al., 2020), yet whether this process also occurs at the mammalian central synapses mediated by Syt7 has not been studied.

Each of layer 2/3 (L2/3) and layer 5 (L5) of neocortex displays intralaminar excitatory synapses between pyramidal cells comprising a recurrent network (Holmgren et al., 2003; Thomson and Lamy, 2007). Short-term facilitation at recurrent excitatory synapses in a cortical network may play a crucial role in maintaining working memory (Mongillo et al., 2008; Mongillo et al., 2012; Hansel and Mato, 2013). Although L2/3 pyramidal cells (PCs) represent task-relevant information during a delay period in working memory tasks (Fujisawa et al., 2008; Ozdemir et al., 2020), most studies on local excitatory recurrent synapses in the neocortex have focused on L5 exhibiting short-term depression (Reyes and Sakmann, 1999; Hempel et al., 2000; Yoon et al., 2020). Further, the synaptic properties of excitatory synapses in L2/3 remain poorly understood.

In the present study, we found that HFS of local excitatory synapses in L2/3 of the medial prefrontal cortex (mPFC) induced initial strong depression followed by delayed facilitation, whereas low-frequency stimulation induced slow monotonous facilitation. This unique form of STP could be explained by 1) high pv of RRP vesicles at rest and throughout a train stimulation; 2) the low resting vesicular occupancy of RRP and its activity-dependent increase. These conditions led to delayed facilitation through progressive overfilling of release sites during HFS with high pv vesicles beyond the resting occupancy. Syt7 knockdown (KD) in L2/3 PCs completely abolished these synaptic enhancements by inhibiting the Ca2+-dependent increase in release site occupancy. Finally, behavioral tests for trace fear conditioning showed that Syt7 KD impaired the acquisition of trace fear memory and reduced c-Fos expression in the mPFC. Collectively, these results support that Ca2+- and Syt7-dependent overfilling of release sites mediates synaptic facilitation at L2/3 recurrent excitatory synapses and contributes to temporal associative learning.

Results

Frequency dependence of STP at local excitatory synapses in L2/3 of the prelimbic cortex

We examined STP profiles at recurrent excitatory synapses between L2/3 PCs and at synapses from L2/3-PCs to fast-spiking interneurons (FSINs) in the prelimbic area of the rat mPFC. The PCs and FSINs were identified based on their distinct intrinsic properties and morphologies (Figure 1—figure supplement 1). To stimulate presynaptic PCs, L2/3 PCs were transfected with a plasmid encoding oChIEF using in utero electroporation (IUE; Figure 1A; Lee et al., 2024). Photo-stimulation of the oChIEF-expressing cell soma at 40 Hz consistently evoked single APs with high fidelity throughout the 600-pulse trains (Figure 1—figure supplement 2A-B), which was consistent with little use-dependent inactivation of oChIEF (Lin et al., 2009; Hass and Glickfeld, 2016).

Frequency-dependence of short-term synaptic plasticity at prelimbic L2/3 excitatory synapses.

(A) In utero electroporation following injection of the plasmid (CAG-oChIEF-tdTomato) into the ventricle of an embryo (E17.5). (B) Representative images showing specific expression of oChIEF-tdTomato in L2/3 pyramidal cells after IUE. The red fluorescence of tdTomato clearly visualizes oChIEF-expressing cell bodies in L2/3 and axons in L5. Scale bar: 1 mm, 100 µm, 50 µm from left to right. (C) Recording schematic showing photostimulation of oChIEF-expressing axon fibers of transfected PCs (filled triangles) and a whole-cell recording from non-transfected PCs (empty triangles) or FSINs (empty circles). A collimated DMD-coupled LED was used to confine the area of excitation (typically 3–4 μm in diameter) to a small region (blue circle) near the soma. (D, E) Representative traces for EPSCs averaged over 10 trials at each frequency (left) and average amplitudes of baseline-normalized EPSCs (right) during 20-pulse trains at frequencies from 5 to 40 Hz at PC-PC (D; n = 12, 9, 21, 10 cells for 5, 10, 20, 40 Hz, respectively) and PC-FSIN (E; n = 8, 9, 10, 10) synapses. Each data point was normalized to the average of the first EPSC. (F) Baseline-normalized amplitudes of steady-state EPSCs (EPSCss; black symbols) and synaptic efficacy (EPSCss × f; red symbols) as a function of stimulation frequency (f) at PC–PC synapses. EPSCss was measured from the average of last 5 EPSCs from the 20-pulse trains (n = 12, 9, 21, 10). (G) Paired pulse ratio (PPR) as a function of inter-spike intervals (n = 25, 9, 25, 22, 43 for 200, 100, 50, 25, 20 ms ISI, respectively) at PC-PC synapses. (H) EPSCss and synaptic efficacy at PC-FSIN (n = 8, 9, 10, 10). (I) PPR at PC-FSIN (n = 8, 9, 10, 15, 34). Gray symbols, individual data.

Given that IUE transfected approximately 10-20% of L2/3 PCs (Figure 1A), we made a whole-cell patch on a non-transfected L2/3 PC or an FSIN (Figure 1C). To stimulate a minimal number of excitatory axon fibers, EPSCs were evoked using minimal optical stimulation (Figure 1—figure supplement 3A; 470 nm, 3-4 μm in radius and 3-5 ms in duration). EPSCs evoked by minimal stimulation were examined to determine the uniformity of the kinetic properties (Figure 1—figure supplement 3B-C). Moreover, we reassessed the kinetics of EPSCs during trains post hoc to ensure uniformity. EPSCs were evoked by 20-pulse trains at 4 different frequencies (5, 10, 20, and 40 Hz) to cover the range of firing rates of PCs observed in vivo during working memory tasks (Baeg et al., 2003). Excitatory synapses from PCs onto both other PCs and FSINs (PC–PC and PC–FSIN) showed strong (∼2-fold enhancement) and lasting (several seconds) short-term facilitation at 20 Hz and lower frequencies (Figure 1D-E). The PC-PC synapses exhibited stronger facilitation than did the PC-FSIN synapses.

Steady-state EPSCs at the PC-PC synapses were not significantly different from 5 to 20 Hz (Figure 1F). Frequency invariance at the PC-PC synapses suggests that Ca2+-dependent synaptic facilitation may counteract vesicle depletion in a frequency-dependent manner (Turecek et al., 2017; Lin et al., 2022). Despite facilitation at all frequencies, the paired pulse ratio (PPR) rapidly decreased as the ISI decreased (Figure 1G, I). At 40 Hz stimulation, both PC-PC and PC-FSIN synapses initially underwent strong depression then followed by facilitation in a few stimuli (Figure 1Dd, Ed). Such delayed facilitation after paired pulse depression (PPD) has not been previously observed in central synapses except for a slight increase in EPSC at the cerebellar synapses under artificially depolarized conditions (Pulido and Marty, 2018).

We confirmed that optically evoked EPSCs were strictly dependent on AP generation (Figure 1—figure supplement 4A), suggesting that EPSCs are evoked by optical stimulation of presynaptic axon fibers rather than by direct depolarization of presynaptic boutons. Additionally, dual patch-clamp techniques showed that the STP of optically evoked EPSCs did not differ from the electrically evoked STP of EPSCs at the PC-FSIN synapses (Figure 1—figure supplement 4B). Finally, we confirmed that AMPA receptor (AMPAR) desensitization was not responsible for PPD at 40 Hz, ruling out the possible effects of postsynaptic factors on STP (Figure 1—figure supplement 5A), and that AMPAR saturation was not significant during facilitation (Figure 1—figure supplement 5B).

Delayed facilitation results from slow activation of Ca2+-dependent vesicle replenishment at a constantly high vesicular fusion probability

The strong PPD at 40 Hz suggests a high vesicular fusion probability (pv) and likely tight coupling of the releasable vesicles to a Ca2+ source. To test this hypothesis, we examined the effects of EGTA, a slow calcium chelator, on synaptic transmission at the PC-PC synapses. The baseline EPSC amplitude (EPSC1) was not affected by incubation of the slice in the bath solution containing 50 μM EGTA-AM for longer than 20 min, supporting tight coupling. Moreover, the PPR was decreased to 0.19, and subsequent facilitation slowed but was still distinct (Figure 2A-B). These findings indicate that the initial pv should be higher than 0.81 considering vesicle recruitment during ISI. The effects of EGTA-AM were confirmed in the same cell by measuring the baseline EPSC and PPR before and after applying EGTA-AM to the bath. EPSC1 remained unchanged, while PPR was reduced by half (Figure 2—figure supplement 1).

Delayed facilitation results from slow activation of Ca2+-dependent vesicle replenishment at a constantly high vesicular fusion probability.

(A) Representative traces (left) and mean baseline-normalized amplitudes (right) of EPSCs evoked by 30-pulse trains at 40 Hz in control (n = 21; black) and in the presence of 50 μM EGTA-AM (n = 14; green). (B) Mean values for the first EPSC amplitude (EPSC1, left) and PPR (right) from the experiments displayed in (A). Gray symbols, individual data. (C) Plot of rate constants for short-term facilitation (kSTF) as a function of stimulation frequency (fstim), showing a linear relationship. The linear regression line (black) is shown fitted to kSTF values, estimated from Figure 1. (D) Representative EPSC traces (left) and average of baseline-normalized EPSCs (right) evoked by 12-pulse stimulation at 5 Hz, followed by 40 Hz 7-pulse train (n = 12). Note that slowly developing facilitation was converted to rapid facilitation after strong PPD. (E) Mean values for PPR at 40 Hz. The baseline PPR was reproduced from Figure 1G and the PPR during 5 Hz train was calculated as (13th EPSC) / (12th EPSC). Gray symbols, individual data. All statistical data are represented as mean ± S.E.M.; n.s. = not significant; **, P<0.01; unpaired t-test.

Notably, the application of EGTA-AM markedly slowed facilitation at 40 Hz (Figure 2A, C). The frequency invariance of steady-state EPSCs (Figure 1F) and the slower facilitation at 40 Hz in the presence of EGTA (Figure 2A) imply a Ca2+-dependent increase in the refilling rate of the RRP and/or an increase in the pv of reluctant vesicles in light of previous studies (Hosoi et al., 2007; Liu et al., 2014; Ritzau-Jost et al., 2014; Turecek et al., 2016; Ritzau-Jost et al., 2018; Kusick et al., 2020). Moreover, facilitation was accelerated at higher stimulation frequencies.

The plot of the rate constant for the increase in EPSCs (denoted as kSTF, the rate constant for the development of short-term facilitation) as a function of the stimulation frequency (fstim) revealed a linear relationship (Figure 2C). kSTF in the presence of EGTA-AM was distinctly below this relationship, supporting the Ca2+-dependent acceleration of facilitation. kSTF was also accelerated when the stimulation frequency was increased from 5 Hz to 40 Hz (Figure 2D; 1.27 ± 0.18/s to 24.27 ± 3.44/s). Notably, the second 40-Hz train stimulation led to strong PPD, followed by accelerated facilitation (Figure 2D-E). The strong PPD suggests that the vesicular fusion probability kept high during the 5-Hz stimulation, arguing against the possibility for an increase in pv. Both two-step sequential priming models, LS/TS and RS/DS models, are suitable to interpret our data and we will refer to the different populations of vesicles as TS and LS vesicles, respectively (Taschenberger et al., 2016; Neher and Brose, 2018; Lin et al., 2022; Neher, 2024). Within this framework, vesicles are released from TS with high pv both in the resting state and during facilitation, and the delayed facilitation is seen as a result of Ca2+-dependent progressive increase in the occupancy of TS vesicles on docking sites.

Presynaptic calcium measurements at axonal boutons of L2/3 pyramidal cells

To test contribution of Ca2+ channel regulation to STP (Dobrunz et al., 1997; Jackman and Regehr, 2017), we estimated total amount of Ca2+ increments during 40 Hz AP trains at axonal boutons of L2/3 PCs in the mPFC. To this end, we measured AP-evoked Ca2+ transients (AP-CaTs) using two-dye ratiometry techniques, and estimated calcium binding ratio from AP-CaTs (Figure 2—figure supplement 2A-C). Measuring the amplitudes and decay time constants of AP-CaTs at different concentrations of Fluo-5F (150, 250 or 500 μM), and plotting them as a function of calcium binding ratio of Fluo-5F (κB), we estimated the amplitude of AP-CaTs (Aca = 1.16 μM), Ca2+ decay time constant (τc = 43 ms) in the absence exogenous Ca2+ buffer, and endogenous static Ca2+ binding ratio (κS = 96.4 or 77.3 estimated from Figure 2—figure supplement 2D-E, respectively; See Materials and Methods). The resting [Ca2+] ([Ca2+]rest) was measured as 50 nM (Figure 2—figure supplement 2F). Estimating the total Ca2+ increments for the first four AP-CaTs in a 40 Hz train according to Equation 2 (Materials and Methods), we found no evidence for inactivation or facilitation of calcium influx during a 40 Hz train (Figure 2—figure supplement 2G-I), arguing against the possibility that Ca2+ channel inactivation and facilitation contribute to strong PPD and delayed facilitation during a 40 Hz train.

Release sites have low baseline occupancy and these are increased during facilitation and post-tetanic augmentation

Given that pv was greater than 0.8, the two-fold increase in EPSC during 20 pulse train stimulations could not be explained by an increase in pv. Recent studies have shown that vesicle release sites (N) are limited in number and are partially occupied at rest (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Under such a high pv, two-fold synaptic facilitation should be attributed to an increase in pocc. The vesicle replenishment rate is accelerated during HFS in a Ca2+-dependent manner (Figure 2A-C; Hosoi et al., 2007). This may lead to the overfilling of release sites beyond their basal vesicular occupancy, which in turn may increase EPSC under high pv. To test this hypothesis, we determined the quantal parameters at the L2/3 local excitatory synapses. We first measured the quantal size from asynchronous release events after replacing extracellular Ca2+ with Sr2+, as previously described (Yoon et al., 2020) (Figure 3—figure supplement 1A-B). The mean and coefficient of variance of the quantal sizes were 24 pA and 0.25 at PC-PC synapses, respectively, and were 30 pA and 0.31 at PC-FSIN synapses (Figure 3—figure supplement 1C). Next, we performed a variance-mean (V-M) analysis of phasic EPSCs at excitatory synapses during train stimulation at 5 Hz and 40 Hz (Scheuss et al., 2002). Applying multiple-probability fluctuation analysis (MPFA) using intra-site quantal variability of 0.2 (Valera et al., 2012), the V-M plot was best fitted by the number of release sites (N) of 5.3 and the quantal size (q) of 25 pA at the PC-PC synapses (Figure 3A). The pr for EPSC1 was calculated as 0.32. We performed the same analysis for the PC-FSIN synapses and found N = 5.3, q = 35 pA, and pr for EPSC1 = 0.31 (Figure 3B). Our estimate for N was smaller but comparable to the estimate (6.9) at excitatory synapses in L2/3 of the mouse barrel cortex (Holler et al., 2021).

Low baseline occupancy of release sites and its increase during facilitation and post-tetanic augmentation.

(A, B) Left, Representative EPSCs evoked by 5 and 40 Hz train stimulation (black, 5 Hz; purple, 40 Hz). Right, Variance-mean plots of EPSCs amplitude from averaged EPSCs recorded at PC-PC (A; n = 12, 17 for 5 and 40 Hz, respectively) and PC-FSIN (B; n = 8, 10) synapses. The data were fitted using multiple-probability fluctuation analysis (MPFA). Error bars are omitted for clarity. The 1st EPSC of 5 Hz train (broken line) was used to estimate the resting level of pocc. Filled circles were measured from post-tetanic augmented EPSCs (A, n = 12, 12, 9; B, n = 7, 7, 7 for 0.1, 0.2, 0.5 s IBIs, respectively). (C-E) Post-tetanic augmentation (PTA) experiments at PC-PC (top) and PC-FSIN (bottom) synapses. (C) Representative traces for EPSCs evoked by double 40 Hz train stimulations separated by 0.5 s. (D) Mean baseline-normalized amplitudes of EPSCs evoked by double 40 Hz trains at different inter-burst intervals (IBIs, 0.1, 0.2, 0.5, 1, 2, 5, 10 s). Upper, PC-PC synapse (n = 12, 12, 21, 11, 11, 16, 11 from short to long IBIs, respectively). Lower, PC-FSIN synapse (n = 10, 9, 8, 9, 9, 7, 9). (E) PTA time course. The baseline-normalized amplitudes of 1st EPSC from the 2nd train were plotted as a function of IBIs.

The pr estimate was clearly smaller than the possible minimum value for pv (0.81) estimated from the PPR data (Figure 2B). Given that pr is a product of pocc and pv (Neher, 2024), a pr smaller than pv indicates the partial occupancy of release sites in the resting state. If synaptic facilitation were mediated by the recruitment of new release sites, it would be accompanied by an increase in the variance of augmented EPSCs (Valera et al., 2012; Kobbersmed et al., 2020). However, the augmented EPSCs during facilitation followed the same parabola in the V-M plot (Figure 3A-B). Therefore, facilitation appears to be mediated by the progressive overfilling of a finite number of release sites beyond the basal occupancy. In the next section, we show that pv is close to unity based on high double failure rate upon paired pulse stimulation (Figure 3—figure supplement 2) and little effect of extracellular [Ca2+] ([Ca2+]o) elevation on baseline EPSC amplitudes (Figure 3—figure supplement 3). Assuming this, the baseline pr at the first pulse estimated from the fitted parabolic curve can be largely regarded as a baseline pocc (c.a. 30%), and the two-fold increase in EPSCs during facilitation can be attributed to an increase of up to 60% in pocc.

We investigated the time course of recovery from facilitation and post-tetanic augmentation (PTA) by applying dual-train stimulations (30 pulses at 40 Hz) separated by different inter-burst intervals (IBIs; Figure 3C-D). The initial EPSCs in the second train were enhanced by more than twofold, and followed by a strong PPD, suggesting that fusion probability stays high in the IBIs and thus synaptic transmission is carried by TS vesicles. The decay time course of PTA was characterized by two phases (initially fast and later slow; Figure 3E), which may reflect a two-phase Ca2+ decay or two distinct recovery processes that occur sequentially. The variance of potentiated EPSCs measured at IBIs of 0.1, 0.2, and 0.5 s was not different from that of EPSCs at the peak of facilitation (filled circles, Figure 3A-B). Given that the same RRP mediates both facilitation and augmentation and that post-tetanic EPSCs is carried by TS vesicles similar to the baseline EPSCs, this plot suggests that PTA, similar to short-term facilitation, is mediated by an increase in the TS vesicle occupancy.

High double failure rate and little effect of [Ca2+]o elevation on baseline EPSC amplitudes suggest that vesicular fusion probability is close to one

The vesicle dynamics of STP has been explained using a simple refilling model (Hosoi et al., 2007; Neher and Sakaba, 2008). This model assumes that the vesicles are reversibly docked to a limited number of release sites (N) with forward and reverse rate constants (denoted as k1 and b1, respectively). Under the framework of the simple refilling model (see Materials and Methods), the rate constant for the refilling of the RRP after depletion by the first pulse is the sum of the baseline k1 and b1. This is estimated as 23/s from the plot of PPR as a function of ISIs (Figure 3—figure supplement 2A). From the baseline occupancy of 0.3 and pocc = k1 / (k1 + b1), we estimated k1 to be 6.9/s.

When paired APs were applied with an ISI of 20 ms, the failure rate at 1st pulse, P(F1), was 10.6%, and the probability of two consecutive failures, P(F1, F2), was 6.2% (Figure 3—figure supplement 2A). From the equation P(F1) = (1 – pr)N, pr was calculated as 0.312 when N = 6, comparable to pr (= 0.32) estimated from the V-M analysis (Figure 3A). We calculated P(F1, F2) under the framework of a simple refilling model, with pv and k1 set as free parameters and other parameters set according to the following relationships: P(F1) = (1 – pr)N, pr = pv · pocc, and pocc = k1 / (k1 + b1). The calculated values for P(F1, F2) and their difference from the observed value (6.2%) are shown in the plane of k1 vs. pv in Figure 3—figure supplement 2B-C. Given that k1 was greater than 5/s, the difference between the calculated and observed values of P(F1,F2) was minimal when pv = 1 (Figure 3—figure supplement 2C).

To find optimal values for k1 and pv that best explain observed probabilities shown in Figure 3—figure supplement 2Ac, a cost function was implemented as the sum of squared errors between observed and predicted probability values for the four combinations of success and failure of 1st and 2nd EPSCs as described in Materials and Methods. The minimum of the cost function (5.63 x 10-5) was found at k1 = 5.21/s and pv = 0.999, which predicted P11 = 41.5%, P10 = 47.9%, P01 = 4.93%, and P00 = 5.67% (subscript 1 and 0 denote success and failure, respectively).

To further validate that baseline pv is near to unity, we examined EPSC changes following elevation of [Ca²⁺]ₒ from 1.3 to 2.5 mM. According to the fourth-power relationship between synaptic responses and [Ca²⁺]ₒ (Equation 3 in (Dodge Jr and Rahamimoff, 1967)), a 3.24-fold increase in the EPSC amplitude was expected. However, the EPSC amplitude increased only 1.23-fold in average, a change that was not statistically significant (Figure 3—figure supplement 3). This small response suggests that pv is already saturated at rest, limiting the dynamic range for further enhancement through increased calcium influx. Recent morphological and functional studies revealed that elevation of [Ca2+]o induces an increase in the number of TS or docked vesicles to a similar extent as our observation (Kusick et al., 2020; Lin et al., 2025), raising a possibility that an increase in pocc is responsible for the 1.23-fold increase in EPSC at high [Ca2+]o. A slight but significant reduction in PPR was observed under high [Ca²⁺]ₒ too. An increase in pocc is thought to be associated with that in the baseline vesicle refilling rate. While PPR is always reduced by an increase in pv, the effects of refilling rate to PPR is complicated. For example, PPR can be reduced by both a decrease (Figure 2—figure supplement 1) and an increase (Lin et al., 2025) in the refilling rate induced by EGTA-AM and PDBu, respectively. Thus, the slight reduction in PPR is not contradictory to the possible contribution of pocc to the high [Ca2+]o effects.

Pharmacological manipulations reveal specific molecular identities underlying vesicle loading processes

Our results suggest that the baseline occupancy of TS vesicles is low and that the activity-dependent progressive overfilling of docking sites with TS vesicles is responsible for short-term facilitation and augmentation. To elucidate the specific molecular link between synaptic enhancement and Ca2+-dependent overfilling, we examined the effects of several drugs known to affect vesicle dynamics. We first examined the roles of phospholipase C (PLC) and diacylglycerol (DAG) by applying 5 µM U73122, a PLC inhibitor, or 20 µM 1-oleoyl-2-acetyl-sn-glycerol (OAG), a DAG analog. The STP at 40 Hz and PTA at 0.5-s interval were examined before and after applying each drug to the bath. For both drugs, the facilitation and augmentation were reduced (Figure 4A-B). U73122 did not affect the basal EPSC amplitude (EPSC1) but resulted in a stronger PPD, slower facilitation, and lower augmentation (Figure 4A). These effects of U73122 suggest the involvement of Ca2+-induced PLC activation in progressive overfilling. In contrast, OAG increased EPSC1 while maintaining a pronounced PPD (Figure 4B), indicating an increase in the TS vesicle pool size. Meanwhile, OAG slowed kSTF and reduced augmentation, indicating occlusion of activity-dependent synaptic facilitation. Phorbol esters and DAG are thought to accelerate priming through Munc13 activation (Rhee et al., 2002; Taschenberger et al., 2016; Aldahabi et al., 2022); they also shorten the bridges that interlink docked vesicles with the active zone (AZ) membrane (Papantoniou et al., 2023). Collectively, these suggest that PLC- and DAG-mediated signaling play a role in the overfilling of docking sites underlying short-term facilitation and augmentation.

Pharmacological experiments reveal specific molecular mechanisms underlying vesicle loading processes.

(A-C) (a) Representative EPSC traces (left) and mean baseline-normalized EPSCs (right) evoked by double 40 Hz train stimulations separated by 0.5 s inter-burst interval (IBI) in control and in the presence of 5 μM U73122 (Aa, n = 7, orange), 20 μM OAG (Ba, n = 12, cyan) or 100 μM dynasore (Ca, n = 10, blue). (b) Mean values for baseline EPSCs (EPSC1, left) and augmentation (right) from the experiments shown in corresponding a panel. (Da) Representative EPSC traces evoked by paired pulses (left) and mean values for baseline EPSC amplitude (middle) and PPR (right) before and after applying 20 μM LatB (n = 6). (Db) Representative EPSC traces (left, upper) and average of normalized EPSCs (left, lower) evoked by double 40 Hz train stimulation separated by 0.5 s in control (n = 21) and in 20 μM LatB conditions (n = 16; pink). Right, Mean values for augmentation in control and LatB conditions. Gray symbols, individual data. All statistical data are represented as mean ± S.E.M., *, P<0.05; **, P<0.01; ***, P< 0.001; unpaired or paired t-test; n.s. = not significant.

Although the acceleration of Ca2+-dependent vesicle refilling supports rapid facilitation at 40 Hz, the later part of the delayed facilitation underwent a slight depression, which was more pronounced during the second burst (Figure 3D and Figure 4). Sustained synaptic transmission is limited by the spatial availability of docking sites because it is undermined by accumulation of exocytic remainings in the presynaptic AZ during HFS (Neher, 2010; Haucke et al., 2011). To verify this, we examined the effects of dynasore (100 µM), a specific inhibitor of endocytosis (Macia et al., 2006). The treatment with dynasore did not influence EPSC1 or the early phase of facilitation, but it exacerbated depression during the later phase and attenuated PTA (Figure 4C). This suggests that late synaptic transmission during the 40-Hz train is limited by site clearance.

Given the involvement of actin polymerization in the multiple steps of vesicle docking and recycling, we examined the effects of 20 µM latrunculin B (LatB), an inhibitor of actin polymerization. The effects of LatB were complicated: (1) EPSC1 was decreased, whereas PPR was increased (Figure 4Da) and (2) facilitation and augmentation were reduced (Figure 4Db). The latter effect may result from defects in vesicle replenishment rates, either through the impaired physical movement of vesicles or disrupted site clearance (Sakaba and Neher, 2003; Lee et al., 2012; Hallermann and Silver, 2013; Miki et al., 2016), as Ca2+-dependent vesicle refilling mediates facilitation and augmentation (Figure 2 and Figure 3). Meanwhile, the former effect might be attributed to a shift in equilibrium in the priming states towards a loosely docked state (LS) at rest. It is unknown whether actin affects vesicular fusogenicity in neurons, although it facilitates multiple steps of vesicle fusion by enhancing membrane tension in neuroendocrine cells (Wu and Chan, 2022).

Short-term facilitation at both types of local excitatory synapses in L2/3 is abolished by Syt7 knockdown

We examined whether Syt7 mediates facilitation at the PC-PC and PC-FSIN synapses in L2/3 of the mPFC. To this end, oChIEF and shRNA against Syt7 (shSyt7) were co-expressed in L2/3 PCs using IUE. Syt7 KD resulted in a complete loss of facilitation at all tested stimulation frequencies (Figure 5A-B). Short-term depression and paired-pulse depression were more pronounced as the stimulation frequency increased, and the frequency invariance of steady-state EPSC was abolished, implying little contribution of Ca2+-dependent vesicle recruitment (Figure 5C-F). The lack of facilitation could not be attributed to the failure of APs in presynaptic axons, as optical stimulation elicited reliable APs in Syt7 KD PCs during the 40-Hz train (Figure 1—figure supplement 2C). Syt7 KD had little effect on the basal properties, including the peak amplitude, rise/decay time, and time-to-peak, of single AP-induced EPSCs (Figure 1—figure supplement 3D-E). Moreover, Syt7 KD did not significantly affect the intrinsic properties of neurons (Figure 5—figure supplement 1).

STF at both types of local excitatory synapses is abolished by Syt7 KD.

(A, B) Representative EPSC traces (left) and mean baseline-normalized amplitudes of EPSCs (right) evoked by 20-pulse trains at 5 to 40 Hz. STP was measured at PC-PC (A; n = 10, 9, 12, 9) and PC-FSIN (B; n = 9, 8, 10, 9) synapses, in which presynaptic Syt7 transcripts were depleted (Syt7 KD). Syt7 KD pyramidal cells are indicated as black triangles on the top. For comparison, STP in WT synapses is reproduced from Figure 1 (dotted lines). Same frequency color codes were used as in Figure 1. (C) Baseline-normalized amplitudes of steady-state EPSC (EPSCss; black symbols) and synaptic efficacy (EPSCss × f; red symbols) as a function of stimulation frequency (f) at PC-PC synapses. EPSCss was defined as the average of last 5 EPSC amplitudes from 20-pulse trains. (D) PPR as a function of inter-spike intervals (n = 10, 9, 12, 18, 26). (E) EPSCss and synaptic efficacy at PC-FSIN synapses. (F) PPR at PC-FSIN synapses (n = 9, 8, 10, 12, 24). Gray symbols, individual data.

To examine whether Syt7 KD had any nonautonomous effects on postsynaptic cells, we measured the quantal size (q) from asynchronous events in the presence of Sr2+ and found that the mean quantal size remained unaltered at Syt7 KD synapses (Figure 3—figure supplement 1C). This indicated that Syt7 did not influence postsynaptic parameters, including the density and kinetic properties of AMPARs. To test the specificity of shSyt7, we examined whether co-expression of the shRNA-resistant form of Syt7 restored synaptic facilitation in both excitatory synapse types. Presynaptic overexpression of shRNA-resistant Syt7 under the CAG promoter rescued the effect of shRNA at both PC-PC and PC-FSIN synapses (Figure 5—figure supplement 2). These results underscore the crucial role of presynaptic Syt7 in mediating short-term facilitation at local excitatory synapses in L2/3 of the mPFC.

Syt7 KD synapses exhibit complementary changes in the number of release sites and their vesicle occupancy

To estimate the quantal parameters in Syt7 KD synapses, we performed V-M analysis of EPSCs evoked by train stimulation (Figure 6A-B). Fitting a parabola to the V-M plot revealed higher baseline pocc in Syt7 KD synapses than in wild-type (WT) synapses for both synapse types (KD, 0.68 and 0.42; WT, 0.32 and 0.31 for PC-PC and PC-FSIN synapses, respectively). The pocc monotonously decreased during a train stimulation, as expected from short-term depression. Moreover, the number of release sites (N) was markedly lower in Syt7 KD synapses (N = 3.5) than in WT synapses (N = 5.3; Figure 3A). Next, we examined the PTA induced by the same protocol as in Figure 3 (two 40-Hz train stimulations separated by variable IBIs) at Syt7 KD synapses (Figure 6C). The 40 Hz train-induced synaptic depression quickly recovered during IBIs, but no PTA was observed at Syt7 KD synapses, indicating its crucial role in augmentation (Figure 6D-E).

Syt7 KD synapses exhibit complementary changes in the number of release sites and their vesicle occupancy.

(A, B) Left, Representative traces of EPSCs evoked by 5 and 40 Hz train stimulations (black, 5 Hz; purple, 40 Hz). Right, Variance-mean plots of EPSCs amplitude from averaged EPSCs recorded at PC-PC (A; n = 9, 15, 12, 10, 9) and PC-FSIN (B; n = 4, 14, 10, 9, 8) synapses in which presynaptic Syt7 has been knocked down. The data were fitted using MPFA and error bars are omitted for clarity. The mean 1st EPSC amplitude of 5 Hz train (vertical broken line) was used for estimation of baseline pocc. Dashed parabolas indicate MPFA fits to variance-mean plot of WT synapses reproduced from Figure 3. (C-E) Recovery experiments at PC-PC (top) and PC-FSIN (bottom) synapses in which presynaptic Syt7 was knocked-down. (C) Representative EPSCs evoked by double 40 Hz train stimulations separated by 0.5 s. (D) Mean baseline-normalized amplitudes of EPSCs evoked by double 40 Hz trains at PC-PC (n = 12, 10, 9, 9, 10, 9, 8) and PC-FSIN (n = 10, 9, 10, 11, 8, 9, 10) synapses at different interburst intervals (IBIs). (E) Recovery time course. Baseline-normalized amplitudes of 1st EPSC from the 2nd burst were plotted as a function of various IBIs. Dotted lines indicate augmented EPSCs in the WT reproduced from Figure 3.

TS vesicle recovery following depletion is accelerated by Syt7

Syt7 is recognized not only for its crucial role in facilitation, but also for mediating Ca2+-dependent replenishment of releasable vesicles (Liu et al., 2014; Tawfik et al., 2021). Given that facilitation at L2/3 excitatory synapses is driven by progressive overfilling (Figure 2 and Figure 3), the loss of facilitation at Syt7 KD synapses in the present study implies that Syt7 plays a key role in the Ca2+-dependent acceleration of refilling and overfilling of the docking sites with TS vesicles. To test this hypothesis, we examined the recovery kinetics of EPSCs after TS vesicles were depleted by 3 pulses at 40 Hz in wild-type (WT) and Syt7 KD synapses (Figure 7). Strong PPD was observed not only in the first burst, but also in the second burst. This suggested that EPSCs in the second burst were mediated by TS vesicles, similar to those in the first burst (Figure 7A, D). Therefore, EPSC1 in the second burst can be interpreted as a release from TS vesicles recovering during a given IBI. The recovery time course of EPSC1 in the second burst indicated full recovery of TS vesicles within 100 ms in both excitatory synapse types in WT rats (Figure 7C, F). However, this recovery was remarkably slower in Syt7 KD synapses, requiring 5 s for full recovery, indicating that TS vesicle recovery was greatly accelerated by Syt7. Overall, synaptic facilitation and augmentation at local excitatory synapses in L2/3 can be ascribed to progressive activity-dependent increases in the occupancy of TS vesicles (Figure 2 and Figure 3), in which Syt7 plays a key role.

Recovery of TS vesicles following depletion is accelerated by Syt7.

(A-C). Recovery experiments at PC-PC synapses in WT (A) and Syt7 KD (B). Mean baseline-normalized amplitudes of EPSCs evoked by two consecutive 3-pulse 40 Hz trains in WT (n = 16, 13, 11, 11, 11, 9 from short to long IBIs, respectively) and KD (n = 13, 11, 12, 8, 10, 11) synapses at different IBIs (0.1, 0.2, 0.5, 1, 2, 5 s). Inset, representative traces of EPSCs evoked by two consecutive 3-pulse 40 Hz train stimulations separated by 0.2 s. (D-F). Recovery experiments at PC-FSIN synapses in WT (D) and Syt7 KD (E). Mean baseline-normalized amplitudes of EPSCs evoked by two consecutive 3-pulse 40 Hz trains in WT (n = 9, 11, 11, 9, 10, 13) and KD (n = 9, 12, 9, 7, 10, 9) synapses at different IBIs. Inset, representative traces of EPSCs evoked by two consecutive 3-pulse 40 Hz train stimulations separated by 0.2 s. (C, F) Recovery time course. Baseline-normalized amplitudes of 1st EPSC from the 2nd burst were plotted as a function of various IBIs.

Behavioral consequences of Syt7 KD in L2/3 PCs of the mPFC

We previously proposed that the disparity in STP between the PC-PC and PC-IN synapses results in activity-dependent changes in the ratio of synaptic weights at the PC-PC and PC-IN synapses (Jee/Jie). This may in turn have a profound influence on persistent activity in a recurrent network by biasing the E-I balance (Yoon et al., 2020). Syt7 KD not only eliminated facilitation, but also the activity-dependent increase in the Jee/Jie ratio in the L2/3 network (Figure 8—figure supplement 1). This suggests the possibility that Syt7 contributes to the persistent activity during working memory tasks in the L2/3 recurrent network. Trace fear conditioning (tFC) requires associative learning between an auditory cue (conditioned stimulus, CS) and temporally separate aversive events (unconditioned stimulus, US). The formation of trace fear memory requires the prelimbic area of the mPFC and hippocampus. mPFC activity during the trace interval is thought to temporarily hold CS information, enabling the network to associate temporally discontinuous events (i.e., temporal associative learning) (Gilmartin et al., 2013). Synaptic facilitation and augmentation are implicated in temporary memory retention in recurrent networks (Mongillo et al., 2008; Mongillo et al., 2012).

Given that Syt7 is crucial for synaptic facilitation (Figure 5) and augmentation (Figure 6) in prelimbic L2/3 recurrent excitatory synapses, we tested whether trace fear memory formation was impaired in rats with depleted Syt7 transcripts, specifically in the L2/3 PCs of the mPFC. Syt7-targeted shRNA (shSyt7) or scrambled shRNA (Scr) constructs were transfected bilaterally into the L2/3 PCs of the mPFC using IUE (Figure 8A). Both the Scr-transfected control and shSyt7-transfected rats underwent tFC training (Figure 8B). Consistent with Gilmartin et al. (2013), in which prefrontal persistent firing was optogenetically inhibited, WT and KD animals exhibited similar freezing behavior during this acquisition phase. The following day, the rats were subjected to a tone test in which trace fear memory formation was assessed from the freezing response to a tone alone in a distinct context. Compared to the control group, the shSyt7-transfected group exhibited significantly lower levels of freezing behavior in response to auditory cues during the first four trials, suggesting an impairment in trace memory formation (Figure 8C).

Behavioral effects of Syt7 deficiency in L2/3 PCs of the mPFC.

(A) Top, Representative images showing bilateral expression of U6-GFP in L2/3 of PCs after IUE at E17.5. Scale bar: 1 mm, 100 μm. Bottom, Schematic of trace fear conditioning and extinction (tone test) protocol. (B) Freezing behavior of control (expressing scrambled shRNA, Scr) and Syt7 KD rats during acquisition of tFC. (C) Freezing ratio during tone tests on following days. Data are shown as average freezing during T1–T4, T5–T8, or T9–T12 (T, trials). The freezing on T1-T4 was significantly lower in KD rats suggesting that formation of trace memory was impaired in Syt7 KD rats (n = 10, 11 for Ctrl and KD, respectively; P = 0.0007, F(1, 19) = 16.43, two-way repeated measures ANOVA; P = 0.0064, 0.0211, 0.0064, 0.0064; Holm-Sidak test). (D) Representative images of c-Fos immunoreactivity in the prelimbic cortex of control or Syt7 KD rats 90 min after tFC acquisition. c-Fos (left, red, Cy5) and GAD67 (right, cyan, Cy3) were immunostained in the same brain slice expressing U6-GFP (middle; green). Scale bar, 50 μm. (E) Exemplar images of c-Fos positive neurons expressing GFP (top) or GAD67 (bottom) in control (left) or Syt7 KD rats (right). Scale bar, 50 μm. (F-H) Effects of Syt7 KD on c-Fos density (F) and percentage of c-Fos positive neurons co-labeled with GAD67 (G) or GFP (H) in L2/3 or L5 of prelimbic cortex (n = 9, 9 for Ctrl and KD, respectively). Open symbols, individual data. All statistical data are represented as mean ± S.E.M.; ****, P<0.0001; unpaired t-test.

The tone test was repeated the following day to test for extinction memory formation. Syt7 KD animals exhibited much less freezing to the CS (+1d) than did WT animals. Considering the Rescorla-Wagner model (Yau and McNally, 2023), our results suggest that the strength of CS-US associative memory was weaker in KD animals than in WT animals; thus, the acquisition of trace fear memory was impaired by Syt7 KD in the L2/3 PCs of the mPFC.

Next, we assessed the locomotor activity and anxiety levels of rats using the open field test (OFT) and elevated plus maze test (EPM). Scr controls and Syt7 KD rats showed similar exploratory patterns, displaying comparable total distance moved and time spent in the periphery of the OFT or the open/closed arms of the EPM (Figure 8—figure supplement 2). Previous studies have suggested that the acquisition of trace fear memory requires elevated trace activity in prelimbic PCs (Gilmartin et al., 2013). Given that c-Fos, an immediate early gene, is upregulated following patterned activities in neurons, its expression is considered a reliable indicator of recent neuronal activation (Fields et al., 1997; Guzowski et al., 2005). To investigate the effect of Syt7 on neuronal activity in vivo during tFC, animals were sacrificed at the time of peak c-Fos protein expression, approximately 90 min after completion of the behavioral task (Arime and Akiyama, 2017).

To elucidate the activation patterns of specific neuronal populations, we performed immunohistochemical co-labeling of c-Fos and GAD67 (a GABAergic interneuron marker) upon tFC acquisition. The density of c-Fos(+) cells and the percentage of c-Fos(+) cells among GAD67(+) cells in all prelimbic layers were significantly lower in Syt7 KD rats than in control rats (Figure 8F-G). Additionally, analysis of the proportion of active cells among transfected L2/3 PCs (c-Fos+/GFP+) revealed that Syt7 KD L2/3 PCs were significantly more active than Scr-transfected PCs (Figure 8H). This unexpected hyperactivity of Syt7 KD PCs may be attributed to the high reciprocal connectivity between PCs and FSINs in the neocortex (Holmgren et al., 2003; Otsuka and Kawaguchi, 2009), as PCs with short-term depression would transmit less input to FSINs, which in turn would exert less feedback inhibition on reciprocally connected PCs. Collectively, these findings show that the slow activation of vesicle refilling supports various forms of facilitation at excitatory synapses in PFC L2/3, which is essential for neuronal activity during temporal associative learning.

Discussion

The present study characterized STP at local excitatory synapses in the L2/3 network of the rat mPFC at physiological extracellular [Ca2+]. These synapses displayed initial strong depression and then delayed facilitation at 40 Hz, while monotonous slowly developing facilitation was observed at lower frequencies. Our study suggests that such delayed facilitation after a brief depression results from a high pv (Figure 2 and Figure 3 and Figure 3—figure supplement 2), along with Ca2+-dependent delayed acceleration of vesicle refilling and overfilling (Figure 2 and Figure 3). The two-fold increase in EPSC during 20 pulse train stimulations under the condition of such a high pv suggests a Ca2+-dependent increase in the vesicle replenishment rate during a train; this then leads to overfilling of release sites beyond their incomplete basal occupancy (Figure 3A-B). The high pv and strong PPD did not result from artificial bouton stimulation, AMPAR desensitization, or any potential distortion caused by optogenetic methods (Figure 1—figure supplements 4 and 5). Additionally, similar to short-term facilitation, PTA could be explained by an increase in the TS vesicle occupancy (Figure 3C-E), consistent with previous studies showing post-tetanic increases in the pocc and/or vesicle pool size at various types of synapses (Lee et al., 2010; Vandael et al., 2020; Tran et al., 2023; Silva et al., 2024). These findings suggest that release sites are partially occupied at rest and that a progressive overfilling of release sites with TS vesicles underlies facilitation and augmentation at intracortical L2/3 excitatory synapses.

Definition of vesicular release probability in the present and previous studies

Previous studies for post-tetanic augmentation (Stevens and Wesseling, 1999; Garcia-Perez and Wesseling, 2008) observed invariance of the RRP size after tetanic stimulation. In these studies, the RRP size was estimated by hypertonic sucrose solution or as the sum of EPSCs evoked 20 Hz/60 pulses train (denoted as ‘RRPhyper’). Because reluctant vesicles can be quickly converted to TS vesicles (16/s) and are released during a train (Lee et al., 2012), it is likely that the RRP size measured by these methods encompasses both LS and TS vesicles. In contrast, we assert high pv based on the observation of strong PPD, failure rates upon paired stimulations at ISI of 20 ms (Figure 2 and Figure 3—figure supplement 2). Given that single AP-induced vesicular release occurs from TS vesicles but not from LS vesicles, pv in the present study indicates the fusion probability of TS vesicles. From the same reasons, pocc denotes the occupancy of release sites by TS vesicles. Note that our study does not provide direct clue whether release sites are occupied by LS vesicles that are not tapped by a single AP, although an increase in the LS vesicle number may accelerate the recovery of TS vesicles. As suggested in Neher (2024), even if the number of LS plus TS vesicles are kept constant, an increase in pocc (occupancy by TS vesicles) would be interpreted as an increase in ‘vesicular release probability’ if it was measured based on RRPhyper.

By the way, it should be noted that increases in both pv and pocc may contribute to PTA. Post-tetanic potentiation (PTP) and phorbol esters had similar effects on baseline EPSCs (Taschenberger et al., 2016), and PDBu increased both the TS vesicle pool size (1.9-fold) and fusion probability of TS vesicles (1.3-fold) at the calyx synapses (Lin et al., 2025). Therefore, it is possible that PTP may increase both pocc and pv of TS vesicles, although it has not been yet directly tested. Moreover, acceleration of the latency in the hypertonicity-induced vesicle release raises a possibility for an increase in fusion probability of TS vesicles (Stevens and Wesseling, 1999; Garcia-Perez and Wesseling, 2008), although it is still controversial that the acceleration of the release latency represents a reduction in the activation energy barrier for vesicle fusion (Schotten et al., 2015).

Progressive overfilling of docking sites

Conventional models for facilitation assume an increase in pv after an AP of RRP vesicles, which had been in a low pv or reluctant state at rest (Dittman et al., 2000; Turecek et al., 2016). This facilitation model is unlikely to explain our results, because TS vesicles are predominantly released not only at rest but also in the facilitated state during or after train stimulation, as evidenced by the strong PPD (ISI, 25 ms) during a 5-Hz train (Figure 2D-E) and at different intervals after a 40-Hz/30 pulse trains (Figure 3D). Moreover, the PPD (ISI, 25 ms) remained pronounced during the early phase of facilitation (Figure 7), arguing against the possibility of slowly increasing the pv of reluctant vesicles. These results suggested that facilitation was mediated by an activity-dependent progressive increase in occupancy of TS vesicles (TS occupancy). Facilitation through an increase in TS occupancy has been termed frequency facilitation, differentiating it from paired pulse facilitation that relies on a residual calcium-dependent transient increase in the pv of releasable vesicles (Neher, 2024).

Assuming an increase in TS occupancy, to attain a two-fold increase in EPSC (Figure 1), the baseline occupancy of the docking sites by TS vesicles should be low, as estimated from the V-M plot (c.a. 30% in Figure 3). However, it is unclear whether the remaining docking sites that are not occupied by TS vesicles are truly vacant or are occupied by reluctant LS vesicles that are not released by the AP or AP trains. TS vesicles can be supplied by the transformation of LS to TS vesicles, both of which reside at the same docking site (LS/TS model; reviewed in Neher, 2024). Alternatively, TS vesicles can be supplied from a distinct replacement pool (RS/DS model; reviewed in Silva et al., 2021). Further studies, including ultrastructural analyses of active zones, are required to address this issue.

Given the crucial role of Syt7 in facilitation (Figure 5), the facilitation mechanisms are closely related to how Syt7 mediates synaptic facilitation. Based on the slow kinetics of Syt7, previous computational modeling studies have proposed that Syt7 may mediate short-term facilitation through an activity-dependent increase in pv (Turecek et al., 2016; Jackman and Regehr, 2017; Turecek et al., 2017; Norman et al., 2023). The present study, however, found that both facilitation and augmentation were mediated by overfilling of the release sites (Figure 2 and Figure 3). Moreover, Syt7-KD synapses displayed slower recovery of TS vesicles (Figure 7). These results suggest that Syt7 is essential for activity-dependent overfilling of docking sites with TS vesicles. Whereas the baseline EPSC was not altered (Figure 1—figure supplement 3), and complementary changes in the number of docking sites and their baseline occupancy were observed in Syt7 KD synapses (Figure 6). These results raise a possibility that Syt7 may play a role in providing additional vacant docking sites that can be overfilled during facilitation. It remains to be elucidated whether the decrease in the number of docking sites at Syt7 KD synapses is related to the subcellular localization of Syt7 to the plasma membrane of the active zone (Sugita et al., 2001; Vevea et al., 2021).

STP model in light of known Ca2+ binding kinetics of Syt7

Syt7 acts as an upstream Ca2+ sensor that facilitates activity-dependent replenishment (Liu et al., 2014; Bacaj et al., 2015; Chen et al., 2017; Tawfik et al., 2021). The present study shows that EPSCs are mediated by TS vesicles both at rest and during activity-dependent facilitation. Releasable vesicles are homogeneous, and Syt7-dependent overfilling is responsible for the activity-dependent enhancement of EPSCs. Thus, we tested whether slowly developing facilitation could be replicated within the framework of a simple refilling model based on the known Ca2+ binding kinetics of Syt7 (Brandt et al., 2012). To this end, we made following assumptions: (1) a Ca2+-dependent increase in the forward refilling rate (k1) requires a full Ca2+-bound form of Syt7, to which multiple Ca2+ ions can bind (two or three Ca2+ ions on each C2A and/or C2B domain) (Brandt et al., 2012; Chon et al., 2015; Voleti et al., 2017); (2) sequential Ca2+ binding steps are synergistic through a cooperativity factor, b; (3) Syt7 senses AP-induced local [Ca2+] similar to that estimated at the calyx of Held (a Gaussian function with a full width at half maximum value of 0.2 ms and a peak of 40 μM) (Wang et al., 2008); and (4) residual [Ca2+] follows a mono-exponential function with an amplitude of 1 μM and decay time constant of 50 ms (Jackson and Redman, 2003).

Under these assumptions and the framework of the simple refilling model, slow and delayed facilitation could be replicated by modeling k1 as k1 = k1,b + K1,max × [Syt7: n Ca2+], where [Syt7: n Ca2+] represents the fraction of fully Ca2+-bound Syt7 (Figure 9A-C). Syt7 was slowly activated over the course of the stimulus train, suggesting that the cooperative binding of multiple Ca2+ ions progressively increased the putative active form of Syt7 because of its slow membrane binding/unbinding kinetics. Nevertheless, the late phase of facilitation at high frequencies was predicted to be higher in the simulation than in the experiment. As expected, based on the notion that sustained release during HFS is limited by site clearance in the active zone (Figure 4C), the model/data discrepancy became more pronounced at higher frequencies. Incorporating endocytic terms into the model may mitigate late-phase discrepancies. Moreover, removal of the catalytic function of Syt7 in the model effectively accounted for the complete loss of facilitation observed in Syt7 KD (Figure 9D).

STP model in light of known Ca2+ binding kinetics of Syt7.

(A-D) Left, Schematic of allosteric calcium binding to Syt7. The number of Ca2+ bound to Syt7 was denoted as # in ‘S#’ in the reaction scheme. kon = 7/μM/s, koff = 10/s. Middle, Simulated changes of k1 (black) in response to 5 Hz train of Ca2+ transients (light blue traces). The priming step of the simple refilling model was assumed to be catalyzed by full Ca2+-bound form of Syt7. Accordingly, Ca2+-dependent increase in k1 was calculated as K1,max multiplied by a fraction of full Ca2+-bound form of Syt7. We assumed that local [Ca2+]i(t) follows a Gaussian function: (1/σ √2π) exp[-(t-tp)2/ 2σ2], in which tp = 0.25 ms and σ = 0.085 ms. Right, Fits of the Syt7 model to the STP data. To fit this model to the STP data, K1,max was set to 300/s (A), 220/s (B), and 180/s (C). Cooperativity factor (b) was set to 0.35 (A), 0.2 (B), and 0.05 (C). k1 is set to be constant for Syt7 KD (D). Same frequency color codes were used as in Figure 1.

Conventional STF models involving cumulative increase in pv of reluctant vesicles

Conventional models for facilitation assume a post-AP residual Ca2+-dependent step increase in pv of RRP (Dittman et al., 2000) or reluctant vesicles (Turecek et al., 2016). Given that pv of TS vesicles is close to one in the present study, an increase in pv of TS vesicles cannot account for facilitation. The possibility for activity-dependent increase in fusion probability of LS vesicles (denoted as pv,LS) should be considered in two ways depending on whether LS and TS vesicles reside in distinct pools or in the same pool. Notably, strong PPD at short ISI implies that pv,LS is near zero at the resting state. Whereas LS vesicles do not contribute to baseline transmission, short-term facilitation (STF) may be mediated by cumulative increase in pv of LS vesicles that reside in a distinct pool. Because the increase in pv,LS during facilitation recruits new release sites (increase in N), the variance of EPSCs should become larger as stimulation frequency increases, resulting in upward deviation from a parabola in the V-M plane, as shown in recent studies (Valera et al., 2012; Kobbersmed et al., 2020). This prediction is not compatible with our results of V-M analysis (Figure 3), showing that EPSCs during STF fell on the same parabola regardless of stimulation frequencies. Therefore, it is unlikely that an increase in fusion probability of reluctant vesicles residing in a distinct release pool mediates STF in the present study. For the latter case, in which LS and TS vesicles occupy in the same release sites, it is hard to distinguish a step increase in fusion probability of LS vesicles from a conversion of LS vesicles to TS. Nevertheless, our results do not support the possibility for a gradual increase in pv,LS that occurs in parallel with STF. Strong PPD, indicative of high pv, was consistently found not only in the baseline (Figure 2 and Figure 2—figure supplement 1) but also during post-tetanic augmentation phase (Figure 3D) and even during the early development of facilitation (Figure 2D-E and Figure 7), arguing against the gradual increase in pv,LS. One may argue that STF may be mediated by a drastic step increase of pv,LS from zero to one, but it is not distinguishable from conversion of LS to TS vesicles.

Possible mechanisms underlying the depression-facilitation sequence

Cooperative binding of multiple Ca2+ ions to Syt7 predicts delayed activation and a progressive increase in vesicle refilling during train stimulations, resulting in delayed facilitation. Combining this with a high pv could replicate not only the slow facilitation at 5 Hz, but also the unique depression-facilitation sequence observed at 40 Hz (Figure 9). As simulated by Pulido and Marty (2018), delayed facilitation can be reproduced by a high pv in combination with a low baseline occupancy of replacement sites and subsequent Ca2+-dependent refilling. However, whether this model is also able to replicate slow facilitation at low frequencies remains to be tested. An alternative, but not exclusive, possibility for delayed facilitation is that the presynaptic global Ca2+ itself may undergo a slow buildup during a train due to the saturation of endogenous Ca2+-binding proteins.

Facilitation requires not only Syt7, but also activation of PLC-DAG signaling

To gain insight into the specific molecular mechanisms underlying the vesicle-loading processes, we examined the effects of pharmacological inhibitors targeting PLC and DAG (Figure 4). Both a PLC inhibitor and a DAG analog reduced facilitation and augmentation. This suggests that the progressive overfilling of docking sites with TS vesicles is supported by the PLC-DAG signaling pathway and probably downstream Munc13 (Rhee et al., 2002; Lou et al., 2008). Phorbol esters have been recently shown to increase the TS fraction of docked vesicles by enhancing vesicle priming without affecting Ca2+ transients or pv (Taschenberger et al., 2016; Aldahabi et al., 2022; Papantoniou et al., 2023; Aldahabi et al., 2024). In line with these findings, signaling involving PLC and DAG may support Ca2+-dependent overfilling of TS vesicles by enhancing molecular priming processes.

Roles of facilitation and augmentation in working memory and related tasks

Previous theoretical studies have proposed that working memory in a recurrent network can be maintained in the form of augmentation at synapses (Mongillo et al., 2008) and/or persistent activity of neurons in a memory ensemble (Mongillo et al., 2012). The latter can be supported by short-term facilitation of recurrent excitatory synapses or disparity in STP between excitatory and inhibitory synapses (Mongillo et al., 2012; Yoon et al., 2020). However, whether these short-term enhancements in synaptic efficacy contribute to working memory-related behaviors has not been examined. tFC is a temporal associative learning task requiring temporary retention of CS information during a trace period. The mPFC and hippocampus are crucial for tFC, and tFC is interfered with by a load of working memory and attention distraction, implying that the tFC shares the same neural substrate with other working memory tasks (Raybuck and Lattal, 2014). In the present study, tFC in WT and Syt7 KD animals revealed that Syt7 deficiency in the L2/3 PCs of the mPFC impaired the acquisition of trace fear memory and reduced c-Fos expression associated with tFC training (Figure 8).

The acquisition of trace fear memory was impaired by inhibition of persistent activity in mPFC during trace period (Gilmartin et al., 2013). The similar deficit observed in Syt7 KD animals is consistent with the hypothesis that STF provides bi-stable ensemble activity in a recurrent network (Mongillo et al., 2012). Nevertheless, alternative mechanisms may be responsible for the behavioral deficit. Not only recurrent network but also long-range loop between the mPFC and the mediodorsal (MD) thalamus play a critical role in maintaining persistent activity within the mPFC especially for a delay period longer than 10 s (Bolkan et al., 2017). Prefrontal L2/3 is heavily innervated by MD thalamus, and L2/3-PCs subsequently relay signals to L5 cortico-thalamic (CT) neurons (Collins et al., 2018). Given that L2/3 is an essential component of the PFC-thalamic loop, loss of STF at recurrent synapses between L2/3 PCs may lead to insufficient L2/3 inputs to L5 CT neurons and failure in the reverberant PFC-MD thalamic feedback loop. Therefore, not only L2/3 recurrent network but also its output to downstream network should be considered as a possible network mechanism underlying behavioral deficit caused by Syt7 KD L2/3.

In summary, activity-dependent synaptic facilitation and augmentation at prelimbic L2/3 recurrent excitatory synapses can be attributed to overfilling of release sites with TS vesicles. Given that the lack of short-term facilitation in L2/3 PCs impairs trace fear memory acquisition, temporary maintenance of CS information during trace period may be supported by an increase in the TS occupancy in presynaptic terminals of CS-representing neuronal ensembles.

Materials and Methods

Animals/Subjects

All experiments were carried out on Sprague-Dawley rats of either sex at p28-43. Rats were maintained in temperature-controlled rooms on a constant 12 h light/12 h dark cycle and were given access to water and food ad libitum. All animal procedures were approved and performed in accordance with the Institutional Animal Care and Use Committee at Seoul National University (SNU200522-1).

In utero electroporation (IUE)

Pregnant Sprague Dawley rats on embryonic day (E) 17.5 were deeply anesthetized with 5% (v/v) isoflurane for the duration of surgery. The uterine horns were exposed by laparotomy and wet with warm sterile phosphate-buffered saline (PBS). The plasmids (0.5-2 µg/µl) together with the 0.01% fast green dye were injected into either the left or the right lateral ventricle or bilateral ventricles of embryos through a thin glass capillary (Narishige PC-10). The embryo’s head was carefully held between the forceps-shaped electrodes of 10 mm diameter (CUY65010; NEPAGENE, Japan). Electroporation pulses (50 V for 50 ms) were delivered five times at intervals of 150 ms using a square-wave generator (ECM830; BEX, United States). The uterine horns were placed back into the abdominal cavity and abdominal muscles and skin were sutured. After the surgery, the animals were recovered under the infrared lamp. For knockdown experiments in slice electrophysiology, the shRNA constructs were co-electroporated with the pCAG-oChIEF-tdTomato, because of the consistent co-expression of the two constructs in a significant proportion of the transfected cells in vivo (Bony et al., 2013). Pups with robust fluorescence signals in the prefrontal cortex of both hemispheres were used for behavior experiments.

Preparation of vector constructs

For an optic stimulation, pCAG-oChIEF-tdTomato was transfected by IUE. Short hairpin RNA (shRNA) against Syt7 sequence (GATCTACCTGTCCTGGAAGAG) was expressed under the control of an U6 promoter of pAAV-U6-GFP vector (Cell Biolabs, #VPK-413). This shRNA was shown to effectively knockdown Syt7 in cell cultures and in vivo by previous reports from more than one research group (Bacaj et al., 2013; Li et al., 2017). Scrambled control sequence (TCGCATAGCGTATGCCGTT) was designed by shuffling the recognition region sequences and a BLAST analysis confirmed that the control shRNA sequence had no target gene. For rescue experiments, the cDNA for rat Syt7 (NM_021659) rendered resistant to the shRNA were inserted in the KD vector and their expression was driven by the CAG promoter; the vector also contained a self-cleaving P2A peptide sequence followed by EGFP, which ensures visualization of transfected cells. All the constructs above were verified by DNA sequencing. Plasmid DNA was purified and concentrated under endotoxin free conditions (Qiagen EndoFree Maxi Kit).

Slice preparation

Acute brain slices were prepared by using a VT1200S Vibratome (Leica Microsystems) after isoflurane anesthesia and decapitation. Coronal sections of the mPFC (300 μm in thickness) were cut in ice-cold artificial cerebrospinal fluid (aCSF) composed of (in mM) 125 NaCl, 3.2 KCl, 25 NaHCO3, 1.25 NaH2PO4, 20 D-glucose, 2 NaPyr, 1 MgCl2, and 1.3 CaCl2 (Yoon et al., 2020), bubbled with 95% O2 and CO2 (pH = 7.3; 310 mOsm). Slices were allowed to recover for a minimum of 30 minutes at 36 °C, and subsequently maintained at room temperature until recording. Composition of aCSF was identical throughout the preparation, incubation, and recording periods unless otherwise stated. For the recordings, the mPFC slices were continuously perfused with oxygenated aCSF at a flow rate of 1.5 ml/min, and the temperature was maintained at 31–32 °C using an in-line temperature controller (Warner Instruments).

Electrophysiology

Whole-cell patch-clamp recordings were performed from pyramidal cells and fast-spiking interneurons in layer 2/3 of the prelimbic mPFC. Neuronal subtypes were initially selected based on morphology under visual guidance by using an upright microscope equipped with infrared differential interference contrast (IR-DIC) optics (BX51WI, Olympus), and were further identified by their electrophysiological properties (see Figure 1—figure supplement 1; van Aerde and Feldmeyer, 2015; Tremblay et al., 2016). To analyze these intrinsic properties in current-clamp experiments, we measured the following parameters: (1) resting membrane potential (RMP), (2) input resistance (Rin), (3) sag ratio, and (4) the F-I curve, which plots firing frequencies (F) against the amplitude of injected currents (I). Both cell types exhibited a low sag ratio below 0.1, consistent with findings from other studies (van Aerde and Feldmeyer, 2015; Yoon et al., 2020). PCs and FSINs were easily distinguished by their unique firing frequency, Rin, spike adaptation, RMP and morphological characteristics including the shape of soma and the thickness of apical dendrites. For experiments involving oChIEF-expressing neurons in WT or Syt7 KD rats, whole-cell recordings were performed from the fluorescently labeled pyramidal neurons in layer 2/3 of the prelimbic cortex using suitable filter sets. Patch pipettes (tip resistance of 2–3 MΩ) were pulled from borosilicate glass and filled with (in mM) 130 K-gluconate, 8 KCl, 10 HEPES, 0.2 EGTA, 20 Na2Pcr, 0.3 Na2GTP, 4 MgATP. Intracellular solutions were adjusted to pH 7.25 and 300–310 mOsm. Neurons were voltage clamped at −78 mV and series resistance (Rs) were continuously monitored during the EPSCs recordings. Experiments were discarded before analysis if the Rs changed by 20% or larger deviation of baseline value. Measurements were not corrected for Rs compensation, bridge balance, or liquid junction potential. Electrophysiological data were low-pass-filtered at 1 kHz (Bessel) and acquired at 20 kHz using a Multiclamp 700B amplifier paired with Digidata 1440A digitizer and Clampex 10.2 software (Molecular Devices).

Synaptic stimulation

IUE resulted in sparse expression of channelrhodopsin (10-20%) (Figure 1A). Exploiting the sparse expression, we employed a minimal optical stimulation method to activate a single or very few excitatory synapses in layer 2/3. The optic minimal stimulation was performed by our published protocol with minor modifications (Yoon et al., 2020). This approach aimed to minimize the risk of synaptic contamination, which could arise from engaging extra terminals that were initially inactive (depolarized below threshold) but became active by temporal summation during high-frequency trains. It also helped prevent the buildup of depolarization at the synaptic terminals, which might influence the probability of release, as noted by (Jackman et al., 2014). To achieve this, we employed a collimated digital micromirror device (DMD)-coupled LED (Polygon400; Mightex Systems) to confine 470 nm blue light to a small area with a radius of approximately 3–8 μm (typically 3–4 μm), as measured at the focal plane of a 60× water immersion objective [numerical aperture (NA) = 1.0; LUMPlanFL, Olympus] (Figure 1—figure supplement 3A). To find the minimal stimulation area, the radius of illumination area was increased from 2 μm by 50% at each step (Figure 1—figure supplement 3A), and selected the smallest illumination area that elicited EPSCs. Setting the duration of illumination between 3 and 5 ms (typically 5 ms), photostimulation onto oChIEF-expressing cells reliably induced action potentials (APs) during 600-pulse trains (see Figure 1—figure supplement 2).

STP at various stimulus frequencies (5, 10, 20, and 40 Hz) was examined using 20-pulse trains. Trials of train stimulation were repeated with a prolonged interval of 30 s to avoid change in baseline properties. We observed no systematic changes in the amplitudes of the first EPSC of each train, nor did we find significant differences in the paired-pulse ratio (PPR) across stimulation trials in the same cell. For each stimulus frequency, approximately ten to twenty consecutive traces were collected and then averaged across trials to yield a mean EPSC trace. For recovery experiments, two consecutive trains of 3 or 30 stimuli at 40 Hz, with varying time intervals (0.1-10 s), were applied every 10 s or 60 s, respectively, at both excitatory synapses. For 600-stimulus trains at 10Hz, the trial was delivered only once for each cell to avoid possible after-effects caused by prolonged synaptic stimulation (Hirsch and Crepel, 1990). Prior to the stimulus trains, the EPSCs were normalized to the mean amplitude of the baseline. For the baseline measurements, paired pulses with various ISIs (20-200 ms) were delivered every 3 s (1/3 Hz) at least 20 times.

For all recordings, picrotoxin (PTX, 10 μM) was included in the recording solution to isolate glutamatergic responses. mPFC PC terminals were entirely glutamatergic as the application of 6-cyano-7-nitroquinoxaline-2,3-dione (CNQX, 20 μM), an AMPA receptor antagonist, completely abolished the monosynaptic events evoked by photostimulation. A low concentration of PTX was employed to minimize the relieving effects on spike suppression caused by blockade of GABAa receptors. We confirmed that no synaptic response was elicited by electrical stimulation in the presence of both 10 μM PTX and 20 μM CNQX.

In Figure 1—figure supplement 4A, 1 µM TTX (1 mM stock solution) and 0.1 mM 4-AP (500 mM stock solution) were added to the bath solution to block action potentials and restore presynaptic glutamate release, respectively (Figure 1—figure supplement 4). In Figure 1— figure supplement 5A, 50 µM of cyclothiazide, a positive allosteric modulator of the AMPA receptors, was included to minimize rapid desensitization of the postsynaptic receptors. In Figure 1—figure supplement 5B, 0.5 mM kynurenate was added to bath solution to prevent possible saturation of AMPA receptors, especially during maximal synaptic facilitation. For pharmacological manipulations of synaptic vesicle dynamics (Figure 4), drugs related to vesicle supply including U73122, OAG, LatB, EGTA-AM, and dynasore at indicated concentrations were added to the recording solution during baseline measurements or PTA/recovery experiments (IBI = 0.5 s). Chemicals were from Sigma-Aldrich, Merck, or Tocris.

Measurements of quantal size

Experiments for measuring asynchronous release were performed as previously described (Yoon et al., 2020). In brief, we replaced extracellular Ca2+ entirely with 2 mM Sr2+ to induce asynchronous release. A single stimulus was delivered at 3-second intervals over 100 repetitions, and asynchronous release events were detected during a 100-ms window starting 25 ms after stimulus onset for each trial. The same procedure was also executed under sham conditions (0% light intensity) to serve as a control, thus mitigating contamination from nonspecific spontaneous events and potential errors associated with in the detection algorithm. q was determined as the mean value from the first Gaussian function after fitting the EPSC distribution with either a single or double Gaussian function (Figure 3—figure supplement 1).

Estimation of Ca2+ kinetics at axonal boutons

Cytosolic [Ca2+] at single axonal boutons was estimated by employing the two-dye, dual-excitation wavelength technique as described by (Oheim et al., 1998). This method involves measuring the fluorescence ratio of a Ca2+-sensitive dye to that of a Ca2+-insensitive dye to correct for dye concentrations. We used Fluo-5F at concentrations of 150, 250, or 500 μM (dissociation constant, Kd = 2.3 μM) as the calcium indicator dye and Alexa Flour 555 (50 μM) as the reference dye. The fluorophores were from Invitrogen. Both dyes were co-loaded into presynaptic terminals of layer 2/3 pyramidal cells in mPFC through whole-cell patch clamp. Before data acquisition, the dyes were allowed to diffuse to terminal boutons for at least 30 minutes after the patch break-in, a duration sufficient for the equilibrium of dye concentrations between the soma and its axonal terminals.

Calcium imaging was performed by sequentially exciting Fluo-5F and AF555 at wavelengths of 473 and 559 nm, respectively, using a confocal laser-scanning microscope (FV1200; Olympus, Japan) equipped with a 60× water immersion objective (NA, 0.9; LUMPlanFl; Olympus, Tokyo, Japan). Once a single axonal bouton was identified, lines for scanning were drawn across boutons, oriented perpendicular to the axon (Figure 2—figure supplement 2A-B). After baseline fluorescence was measured, line scanning was proceeded while a single or train APs were induced by applying current pulses. Upon stimulation, green fluorescence but not red one increased along the vertical array of lines across the bouton (Figure 2—figure supplement 2B-C). Measurements were obtained under current-clamp conditions, with the resting potential regularly monitored. The scanning speed was adjusted to 600-700 Hz by controlling the line length. Line scan for measuring fluorescent transients in a bouton was repeated at least 3 times, and then averaged to enhance the signal-to-noise ratio. To mitigate photobleaching, minimal intensity of the laser power and maximal pinhole size were used. Emitted light was filtered through 490/590 or 575/675 bandpass filters before detection by photomultiplier tubes for green or red fluorescence, respectively.

Fluorescence measurements were converted to [Ca2+]i using the ratio of background-subtracted fluorescence (F = F0 - Fb) from two dyes. The fluorescence ratio (R = Fgreen/Fred) in data traces were converted to [Ca2+]i using the following equation: Calibration parameters were obtained using an in-cell calibration protocol (Helmchen, 2011), wherein the minimum (Rmin) and maximum ratio (Rmax) values were established through the use of intracellular solutions containing 10 mM EGTA or 10 mM [Ca2+], respectively.

When Ca2+ is extruded with a rate constant, γ, after an AP-induced short pulse of Ca2+ influx to a single compartment, [Ca2+]i initially rises to ACa and then undergoes a mono-exponential decay with a time constant (τCa). Under this framework, ACa and τCa depend on Ca2+ binding ratios of endo- and exogenous Ca2+ buffers (denoted as κS and κB, respectively) as following equations (Neher, 1995; Kim et al., 2003):

, where Δ[Ca2+]T is an increment in total [Ca2+] (free plus bound) elicited by an AP. Ca2+ binding ratio indicates the ratio of concentration changes in the Ca2+-buffer complex relative to free [Ca2+] change, and linearized κB is calculated from the dye concentration and the change from [Ca2+]1 to [Ca2+]2 as follows: The total dye concentration, [B]t, was taken as [Fluo-5F] in the patch pipette. With κB determined from Equation 4 and τCa from the decay of a Ca2+ transient, κS can be estimated by plotting τ versus κB and applying a linear fit to the plot, which gives the x-intercept as -(1 + κS). Once determined, κS can be used to calculate Δ[Ca2+]T.

Optimization of k1 and pv from failure rates

Let n0 = the number of docked SVs before arrival of a first AP; n1 = the number of docked SVs remaining after a first AP (i.e. n0 - n1 vesicles are released); n2 = the number of docked SVs just before arrival of a 2nd AP (ISI = 20 ms; Figure 3—figure supplement 2). If the number of release sites (N), forward and reverse rates for vesicle docking (k1 and b1, respectively), mean release site occupancy (pocc) and vesicular fusion probability upon an AP (pv) are given, we can predict P(n0, n1, n2), the probability for [n0, n1, n2], as following steps.

Probability for a synapse harbouring n0 docked SVs just before arrival of a first AP is calculated as:

Probability that n1 vesicles remain after a first AP when the initial number of docked SVs is n0 [P(n1|n0)] is calculated as:

When forward and backward rate constants are k1 and b1, respectively, probability that n2 vesicles are docked just before arrival of a 2nd AP when n1 SVs remain after a 1st AP (Δt = 20 ms) [P(n2 | n1)] is predicted as follows: Assuming that docking and undocking follow a Poisson process,

Combining probabilities calculated from above three steps, we can calculate P(n0, n1, n2):

Once P(n0, n1, n2) is given, we can predict probability for double failure (P00) and probability for the 1st success and the 2nd failure (P10) as:

Probability for the 1st failure and the 2nd success (P01) and probability for double success (P11) were calculated as:

, in which P(F1) and P(S1) are probability for the 1st failure and the 1st success, respectively, and these are calculated as: To optimize k1 and pv, we used the Matlab routine, fmincon, in which the sum of squared errors (= observed value minus predicted value) was set as a cost function to be minimized. Lower and upper bounds for k1 and pv were set as [4, 8] and [0.8, 1], respectively. Release probability (pr) was calculated from the observed 1st failure rate, P(F1), as N was assumed to be six as shown in Figure 3. Other parameters (b1, and pocc) were set according to following relationships:

Calculation of probability for double failure

Double failure probability, P00, can be more easily predicted. By the Bayes rule, , where F1 and F2 means 1st and 2nd failure. Moreover, and n0 = n1 because of the 1st failure. Thus, the probability for two consecutive failures can be calculated as: P(F1) and N are observed values. For calculation in Figure 3—figure supplement 2, k1 and pv were set free variables. Other parameters were set under the following constrains: P(F1) = (1 – pr)N, pr = pv pocc, and pocc = k1 / (k1 + b1).

Behavioral tests

Trace fear conditioning

For trace fear conditioning (tFC) (Gilmartin et al., 2013), individual rats were placed in the training chamber (30.5 x 25.4 x 30.5 cm; Coulbourn Instruments) consisting of metal grid floor connected to an electrical stimulator (H10-11R-TC-SF, Coulbourn Instruments) for delivery of a footshock. After a 5 min habituation period, rats received five pairings of a 20 s tone conditioned stimulus (CS, 2.5 kHz, 75 dB), followed by empty 20s trace period and a unconditioned stimulus (US; 1 s, 0.7 mA footshock) with a pseudorandom inter- trial interval (ITI) of 240 ± 20 s. During this training session, rats learned to associate the auditory CS with the shock US. The next day, the rats were tested for memory of CS-US trace association (called tone test). The acquisition of fear learning, often indicated by the extent of freezing to the CS, was assessed in a novel chamber (a white hexagonal enclosure) in a dimly lit room. During this tone (or retrieval) test, rats were first allowed to explore the new context for 5 min (habituation), followed by twelve 20 s CS presentations with a variable ITI of 120 ± 20 s. This entire session was repeated on the next day for assessment of extinction learning. Identities of the chambers were also determined with the presence of distinct odors; the training and testing chambers were cleaned with 70% ethanol and 1% acetic acid respectively before and after each session. For all experimental sessions, the activity of animals in the chambers was recorded at 30 frames per second using the EthoVision XT (Noldus Information Technology, Wageningen, Netherlands), and stored as a video file. The freezing behavior, defined as behavioral immobility other than respiratory movement, was assessed using an open-source video analysis pipeline ezTrack (Pennington et al., 2019). The percentage of freezing was calculated as a total duration of freezing divided by the total duration of observation.

Elevated Plus Maze

Elevated plus maze (EPM) test was performed using a plus-shaped apparatus elevated 60 cm above the ground. The maze consisted of two open arms (50 x 10 x 0.5 cm (H)), two closed arms (50 x 10 x 40 cm (H)) and a center area (10 x 10 cm). Rats were individually placed on the center area, facing an open arm, and the path of the animal was recorded with a video camera for 5 min. We analyzed the total distance traveled for 5 min and time spent in each arm. Animals were tracked offline using the open-source tracking system ezTrack (Pennington et al., 2019).

Open field test

Open field exploration test (OFT) was performed in an open field apparatus in a dimly lit room. The open field consisted of a white plastic board (1.20 m in diameter) surrounded by white plastic walls (50 cm in height). Individual rats were placed in a center zone (0.6 m in diameter), and allowed to freely explore the apparatus for 10 min. Time spent in the center zone and total distance traveled were analyzed for initial 5 min. Animals were tracked offline using the open-source tracking system ezTrack (Pennington et al., 2019).

Histology and immunohistochemistry

At the end of the behavioral experiments, rats were anesthetized using isoflurane inhalation and transcardially perfused with PBS followed by 4% paraformaldehyde (PFA, T&I, Korea). Brains were extracted and post-fixed in 4% PFA solution for 24 h. Perfused brains were sliced at 100 µm for estimation of GFP expression or 50 µm for immunohistochemistry. For analysis of c-Fos expression, rats were returned to their home cages immediately after the training session and perfused approximately 90 minutes later (Arime and Akiyama, 2017). Brains were then cryopreserved, frozen, and sectioned coronally. Immunofluorescence staining was conducted on free-floating slices using a rabbit monoclonal anti c-fos (1:1000, Cell Signaling Technology, cat# 2250s) and mouse monoclonal anti-GAD67 (1:500, Millipore, cat# mab5406). Fluorescent secondary antibodies included Cy5-conjugated goat anti-rabbit IgG (1:1000, Abcam, cat# ab97077) and Cy3-conjugated goat anti-mouse IgG (1:1000, Abcam, cat# ab97035). Brain slices were permeabilized and blocked with PBS containing 0.5% Triton X-100 and 5% normal goat serum (NGS) for 1h at room temperature. Slices were incubated overnight at 4 °C with primary antibodies, followed by 4 h incubation with secondary antibodies at room temperature. Slices were then mounted with Antifading Mounting medium (Abcam, cat# ab104139). Confocal images were acquired using a Leica TCS SP8 microscope set to the same laser intensity and acquisition parameters to compare the immunosignals in sections from Scr and Syt7 KD groups. For cell counting and colocalization, fluorescent protein-positive cells were automatically quantified with spot-detection algorithms in Imaris 9.5 (Oxford instruments).

Data analysis

Data were analyzed and presented using ClampFit (Molecular Devices), Igor Pro (Wavemetrics), MATLAB (Mathworks), and Prism (GraphPad). Data points and error bars represent the mean ± standard error of the mean. All statistical analyses were conducted using two-tailed comparisons. Sample sizes and statistical tests for each group are detailed in the figure legends.

Numerical integration of the simple refilling model

This model posits reversible docking of vesicles at finite release sites (Nmax) with forward (k1) and reverse rate constants (b1) (Hosoi et al., 2007; Neher and Sakaba, 2008). According to this scheme, dn/dt = k1 u - b1 n - pv n δ(t - tAP), where u is the number of unoccupied sites (= Nmax - n); tAP is the timing of AP firing; and δ is the Dirac delta function. The pv was assumed to be unity (Figure 3—figure supplement 2). The sum of basal k1 (= k1,b) and b1 was be estimated as 23/s from the dependence PPR on ISIs (Figure 3—figure supplement 2Ab). From the baseline occupancy of 0.3 and pocc = k1,b / (k1,b + b1), we could estimate baseline k1 and b1 as 6.9/s, and 16.1/s, respectively. Ca2+ and Syt7-dependent increase of k1 was modeled as k1 = k1,b + K1,max [Syt7: n Ca2+], where [Syt7: n Ca2+] represents the fraction of full Ca2+-bound Syt7. For deterministic simulations, we time-integrated a set of differential equations for each model using Euler methods with a time step of 1 ms. All calculations were performed on the platform of Matlab (R2022b, Mathworks, USA).

eLife Assessment

This valuable work explores how synaptic activity encodes information during memory tasks. All reviewers agree that the work is of very high quality and that the methodological approach is praiseworthy. Although the experimental data support the possibility that phospholipase diacylglycerol signaling and synaptotagmin 7 (Syt7) dynamically regulate the vesicle pool required for presynaptic release, a concern remains that the central finding of paired-pulse depression at very short intervals could be due to a mechanism that does not depend on exocytosis, such as Ca²⁺ channel inactivation, rather than vesicle pool depletion. Overall, this is a solid study although the results still warrant consideration of alternative interpretations.

Reviewer #3 (Public review):

To summarize: The authors' overfilling hypothesis depends crucially on the premise that the very-quickly reverting paired-pulse depression seen after unusually short rest intervals of << 50 ms is caused by depletion of release sites whereas Dobrunz and Stevens (1997) concluded that the cause was some other mechanism that does not involve depletion. The authors now include experiments where switching extracellular Ca2+ from 1.2 to 2.5 mM increases synaptic strength on average, but not by as much as at other synapse types. They contend that the result supports the depletion hypothesis. I didn't agree because the model used to generate the hypothesis had no room for any increase at all, and because a more granular analysis revealed a mixed population with a subset where: (a) synaptic strength increased by as much as at standard synapses; and yet (b) the quickly reverting depression for the subset was the same as the overall population.

The authors raise the possibility of additional experiments, and I do think this could clarify things if they pre-treat with EGTA as I recommended initially. They've already shown they can do this routinely, and it would allow them to elegantly distinguish between pv and pocc explanations for both the increases in synaptic strength and the decreases in the paired pulse ratio upon switching Ca2+ to 2.5 mM. Plus/minus EGTA pre-treatment trials could be interleaved and done blind with minimal additional effort.

Showing reversibility would be a great addition too, because, in our experience, this does not always happen in whole-cell recordings in ex-vivo tissue even when electrical properties do not change. If the goal is to show that L2/3 synapses are less sensitive to changes in Ca2+ compared to other synapse types - which is interesting but a bit off point - then I would additionally include a positive control, done by the same person with the same equipment, at one of those other synapse types using the same kind of presynaptic stimulation (i.e. ChRs).

Specific points (quotations are from the Authors' rebuttal)

(1) Regarding the Author response image 1, I was instead suggesting a plot of PPR in 1.2 mM Ca2+ versus the relative increase in synaptic strength in 2.5 versus in 1.2 mM. This continues to seem relevant.

(2) "Could you explain in detail why two-fold increase implies pv < 0.2?"

a. start with power((2.5/(1 + (2.5/K1) + 1/2.97)),4) = 2*power((1.3/(1 + (1.3/K1) + 1/2.97)),4);

b. solve for K1 (this turns out to be 0.48);

c. then implement the premise that pv -> 1.0 when Ca2+ is high by calculating Max = power((C/(1 + (C/K1) + 1/2.97)),4) where C is [Ca] -> infinity.

d. pv when [Ca] = 1.3. mM must then be power((1.3/(1 + (1.3/K1) + 1/2.97)),4)/Max, which is <0.2.

Note that modern updates of Dodge and Rahamimoff typically include a parameter that prevents pv from approaching 1.0; this is the gamma parameter in the versions from Neher group.

(3) "If so, we can not understand why depletion-dependent PPD should lead to PPF."

When PPD is caused by depletion and pv < 0.2, the number of occupied release sites should not be decreased by more than one-fifth at the second stimulus so, without facilitation, PPR should be > 0.8. The EGTA results then indicate there should be strong facilitation, driving PPR to something like 1.2 with conservative assumptions. And yet, a value of < 0.4 is measured, which is a large miss.

(4) Despite the authors' suggestion to the contrary, I continue to think there is a substantial chance that Ca2+-channel inactivation is the mechanism underlying the very quickly reverting paired-pulse depression. However, this is only one example of a non-depletion mechanism among many, with the main point being that any non-depletion mechanism would undercut the reasoning for overfilling. And, this is what Dobrunz and Stevens claimed to show; that the mechanism - whatever it is - does not involve depletion. The most effective way to address this would be affirmative experiments showing that the quickly reverting depression is caused by depletion after all. Attempting to prove that Ca2+-channel inactivation does not occur does not seem like a worthwhile strategy because it would not address the many other possibilities.

(5) True that Kusick et al. observed morphological re-docking, but then vesicles would have to re-prime and Mahfooz et al. (2016) showed that re-priming would have to be slower than 110 ms (at least during heavy use at calyx of Held).

Author response:

The following is the authors’ response to the previous reviews

Public Reviews:

Reviewer #3 (Public review):

To summarize: The authors' overfilling hypothesis depends crucially on the premise that the very quickly reverting paired-pulse depression seen after unusually short rest intervals of << 50 ms is caused by depletion of release sites whereas Dobrunz and Stevens (1997) concluded that the cause was some other mechanism that does not involve depletion on. The authors now include experiments where switching extracellular Ca2+ from 1.2 to 2.5 mM increases synaptic strength on average, but not by as much as at other synapse types. They contend that the result supports the depletion on hypothesis. I didn't agree because the model used to generate the hypothesis had no room for any increase at all, and because a more granular analysis revealed a mixed population with a subset where: (a) synaptic strength increased by as much as at standard synapses; and yet (b) the quickly reverting depression for the subset was the same as the overall population.

The authors raise the possibility of additional experiments, and I do think this could clarify things if they pre-treat with EGTA as I recommended initially. They've already shown they can do this routinely, and it would allow them to elegantly distinguish between pv and pocc explanations for both the increases in synaptic strength and the decreases in the paired pulse ratio upon switching Ca2+ to 2.5 mM. Plus/minus EGTA pre-treatment trials could be interleaved and done blind with minimal additional effort.

Showing reversibility would be a great addition too, because, in our experience, this does not always happen in whole-cell recordings in ex-vivo tissue even when electrical properties do not change. If the goal is to show that L2/3 synapses are less sensitive to changes in Ca2+ compared to other synapse types - which is interesting but a bit off point - then I would additionally include a positive control, done by the same person with the same equipment, at one of those other synapse types using the same kind of presynaptic stimulation (i.e. ChRs).

Specific points (quotations are from the Authors' rebuttal)

(1) Regarding the Author response image 1, I was instead suggesting a plot of PPR in 1.2 mM Ca2+ versus the relative increase in synaptic strength in 2.5 versus in 1.2 mM. This continues to seem relevant.

Complying with your suggestion, we studied the effects of external [Ca2+] ([Ca2+]o) after pre-incubating the slice in aCSF containing 50 μM EGTA-AM, and added the results as Figure 3—figure supplement 3C-D. Elevation of ([Ca2+]o) from 1.3 to 2.5 mM produced no significant change in either baseline EPSC amplitude or PPR, supporting that the pv is already saturated at 1.3 mM [Ca2+]o and implying that the modest Ca2+ dependence of baseline EPSCs and PPR in the absence of EGTA (Figure 3—figure supplement 3A-B) is mediated by the change in baseline vesicular occupancy of release sites (pocc) rather than fusion probability of docked vesicles (pv).

We found some correlation of high Ca2+-induced relative increase in synaptic strength with the PPR at low Ca2+ (Author response image 1-A). But this correlation was abolished by pre-incubating the slices in EGTA-AM too (Author response image 1-B). It should be noted that high PPR does not always mean low pv. For example, when the replenishment is equal between high and low baseline pocc synapses, the PPR would be higher at low pocc synapses than that at high pocc synapses, even if pv is close to unity. Therefore, high baseline release probability (Pr), whatever it is attributed to high pv or high pocc, can result in low PPR, considering that Pr = pocc x pv.

As we have already mentioned in our previous letter, the relationship of PPR with refilling rate is complicated and can be bidirectional, whereas an increase in pv always results in a reduction of PPR. For example, PPR can be reduced by both a decrease and an increase in the refilling rate (Figure 2— figure supplement 1 and Lin et al., 2025). Therefore, the PPR analysis alone is insufficient to differentiate the contributions of pv and pocc Thanks to your suggestion, we could resolve this ambiguity by the EGTA-AM pre-incubation study (Figure 3—figure supplement 3C-D).

Plot of PPR at low [Ca2+]o (1.

3 mM) as a function of the baseline EPSC at high [Ca2+]o (2.5 mM) normalized to that at low [Ca2+]o measured at recurrent excitatory synapses in L2/3 of the prelimbic cortex under the conditions without EGTA-AM (A) and after pre-incubating the slices in EGTA-AM (50 μM) (B)

(2) "Could you explain in detail why two-fold increase implies pv < 0.2?"

(a) start with power((2.5/(1 + (2.5/K1) + 1/2.97)),4) = 2*power((1.3/(1 + (1.3/K1) + 1/2.97)),4);

(b) solve for K1 (this turns out to be 0.48);

(c) then implement the premise that pv -> 1.0 when Ca2+ is high by calculating Max = power((C/(1 + (C/K1) + 1/2.97)),4) where C is [Ca] -> infinity.

(d) pv when [Ca] = 1.3. mM must then be power((1.3/(1 + (1.3/K1) + 1/2.97)),4)/Max, which is <0.2. Note that modern updates of Dodge and Rahamimoff typically include a parameter that prevents pv from approaching 1.0; this is the gamma parameter in the versions from Neher group.

Thank you very much for your kind explanation. This interpretation, however, based on the premise that pv is not saturated at low[Ca2+]o, and that Pr = pv. In the present study, however, we presented multiple convergent lines of evidence supporting that pv is already saturated at 1.3 mM [Ca2+]o as follows: (1) little effect of EGTA-AM on the baseline EPSCs (Figure 2—figure supplement 1); (2) high double failure rates (Figure 3—figure supplement 2); (3) little effect of high [Ca2+]o on baseline EPSC (Figure 3—figure supplement 3). Therefore, our results suggest that the classical Dodge-Rahamimoff fourth-power relationship can not be applied to estimate pv at the L2/3 recurrent excitatory synapses.

(3) "If so, we can not understand why depletion-dependent PPD should lead to PPF." When PPD is caused by depletion and pv < 0.2, the number of occupied release sites should not be decreased by more than one-filth at the second stimulus so, without facilitation, PPR should be > 0.8. The EGTA results then indicate there should be strong facilitation, driving PPR to something like 1.2 with conservative assumptions. And yet, a value of < 0.4 is measured, which is a large miss.

As mentioned above, the framework used for inferring that pv < 0.2, the Dodge-Rahamimoff equation, is not applicable to our experimental system. Consequently, the subsequent deduction— that depletion-dependent PPD should logically lead to PPF—is based on a model that does not compatible with aforementioned multiple convergent lines of evidence, which supports high pv rather than the low pv facilitation model.

(4) Despite the authors' suggestion to the contrary, I continue to think there is a substantial chance that Ca2+-channel inactivation is the mechanism underlying the very quickly reverting paired-pulse depression. However, this is only one example of a non-depletion mechanism among many, with the main point being that any non-depletion mechanism would undercut the reasoning for overfilling. And, this is what Dobrunz and Stevens claimed to show; that the mechanism - whatever it is - does not involve depletion. The most effective way to address this would be affirmative experiments showing that the quickly reverting depression is caused by depletion after all. Attempting to prove that Ca2+channel inactivation does not occur does not seem like a worthwhile strategy because it would not address the many other possibilities.

We have systematically ruled out alternative possibilities that may underlie the strong PPD observed at our synapses and demonstrated that it arises from high pv-induced vesicle depletion through multiple independent lines of evidence. First, we excluded (1) AMPAR desensitization or saturation (Figure 1—figure supplement 5), (2) Ca2+ channel inactivation (Figure 2—figure supplement 2), (3) channelrhodopsin inactivation (Figure 1—figure supplement 2), (4) artificial bouton stimulation (Figure 1—figure supplement 4), and (5) transient vesicle undocking (Figure 5; addressed in our previous rebuttal). Second, EGTA-AM experiments (Figure 2, Figure 2—figure supplement 1) revealed that release sites are tightly coupled to Ca2+ channels, and that EGTA further exacerbates PPD. Third, we validated high baseline pv through analysis of double failure rates (Figure 3—figure supplement 2). Fourth, the minimal increase in baseline EPSCs upon elevation of external [Ca2+] (Figure 3—figure supplement 3) further supports that baseline pv is already saturated at low [Ca2+]o. Additionally, to further validate our hypothesis, we performed the specific experiment suggested by the reviewer. We have now added EGTA pre-incubation experiments (Figure 3—figure supplement 3C-D) and have revised the manuscript. Specifically, when slices were pre-incubated with 50 μM EGTA-AM, elevation of extracellular [Ca2+] from 1.3 to 2.5 mM produced no significant change in either baseline EPSC amplitude or PPR, strongly supporting that the high [Ca2+]o effects in the absence of EGTA are primarily mediated by changes in pocc rather than pv

(5) True that Kusick et al. observed morphological re-docking, but then vesicles would have to re-prime and Mahfooz et al. (2016) showed that re-priming would have to be slower than 110 ms (at least during heavy use at calyx of Held).

As previously discussed, Kusick et al. (2020) demonstrated that the transient destabilization of the docked vesicle pool recovers very rapidly within 14 ms after stimulation. This implies that any posts stimulation undocking events are likely recovered before the 20 ms ISI used in our PPR experiments. Consequently, transient undocking/re-docking events are unlikely to significantly influence the PPR measured at this interval. Furthermore, regarding the slow re-priming kinetics (>100 ms) reported by Mahfooz et al. (2016) and Kusick et al., (2020), our 20 ms ISI effectively falls into a me window that avoids the potential confounds of both processes: it is long enough for the rapid morphological recovery (~14 ms) of docked vesicles to occur, yet too short for the slow re-priming process to make a substantial contribution. Furthermore, Vevea et al. (2021) showed that post-stimulus undocking is facilitated in synaptotagmin-7 (Syt7) knockout synapses. In our study, however, Syt7 knockdown did not affect PPR at 20 ms ISI, suggesting that the undocking process described in Kusick et al. (2020) is not a major contributor to the PPD observed at 20 ms intervals in our experiments. Therefore, we conclude that the 20 ms ISI used in our experiments falls within a me window that is influenced neither by the rapid undocking (<14 ms) reported nor by the slow re-priming process (>100 ms).

Version 5

Abstract

Short-term facilitation of recurrent excitatory synapses within the cortical network has been proposed to support persistent activity during working memory tasks, yet the underlying mechanisms remain poorly understood. We characterized short-term plasticity at the local excitatory synapses in layer 2/3 of the rat medial prefrontal cortex and studied its presynaptic mechanisms. Low-frequency stimulation induced slowly developing facilitation, whereas high-frequency stimulation initially induced strong depression followed by rapid facilitation. This non-monotonic delayed facilitation after a brief depression resulted from a high vesicular fusion probability and slow activation of Ca2+-dependent vesicle replenishment, which led to the overfilling of release sites beyond their basal occupancy. Pharmacological and gene knockdown (KD) experiments revealed that the facilitation was mediated by phospholipase C/diacylglycerol signaling and synaptotagmin 7 (Syt7). Notably, Syt7 KD abolished facilitation and slowed the refilling rate of vesicles with high fusion probability. Furthermore, Syt7 deficiency in layer 2/3 pyramidal neurons impaired the acquisition of trace fear memory and reduced c-Fos activity. In conclusion, Ca2+- and Syt7-dependent overfilling of release sites mediates synaptic facilitation at layer 2/3 recurrent excitatory synapses and contributes to temporal associative learning.

Introduction

Synapses in the mammalian central nervous system undergo activity-dependent modulation of synaptic weight during repetitive stimulation, known as short-term synaptic plasticity (STP). STP is mediated by presynaptic mechanisms that regulate the release probability (pr), size of the readily releasable pool (RRP), and kinetics of vesicle replenishment of RRP (Zucker and Regehr, 2002; Jackman and Regehr, 2017; Neher and Brose, 2018). Short-term depression has been observed at synapses with a high baseline pr, primarily because of the rapid depletion of releasable vesicles during high-frequency stimulation (HFS) (Lin et al., 2022). Conversely, vesicle depletion is less pronounced at synapses with a low baseline pr, and cumulative increases in pr during HFS are considered to mediate short-term facilitation (Rozov et al., 2001; Pan and Zucker, 2009; Aldahabi et al., 2024). Central synapses exhibit a wide spectrum of STP depending on different combinations of presynaptic factors mediating short-term depression and facilitation.

Previously, facilitation was modeled as an increase in pr after an action potential (AP) and its slow decay during the inter-spike interval (ISI), resulting in a cumulative increase in pr during HFS (Markram et al., 1998; Dittman et al., 2000). Recent studies have suggested that the number of docking (or release) sites in an active zone is limited and only partially occupied by releasable vesicles in the resting states (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Therefore, pr is determined by the vesicular fusion probability (pv) and release site occupancy (pocc). However, distinguishing the individual contributions of pv and pocc to pr is challenging (Silva et al., 2021; Neher, 2024). It is also unclear whether facilitation is mediated by activity-dependent increases in the pv, pocc, or both.

Two recent models for vesicle priming and release view the ‘priming’ as a two-step reversible process that leads to a release-ready state. The release-ready vesicles could be vesicles in a tightly docked state (TS) or those occupying specialized docking sites (DS), and are supplied from vesicles in a loosely docked state (LS) or those occupying a distinct replacement sites (RS), which are called LS/TS model (Aldahabi et al., 2024; Neher, 2024) and RS/DS model (Miki et al., 2016; Silva et al., 2021), respectively. Common to both models is the assumption that release-ready vesicles constitute only a fraction of all docked vesicles under resting conditions and that this fraction is up- and down-regulated by synaptic activity.

Synaptotagmin-7 (Syt7) was recently found to play a crucial role in facilitation at different types of central synapses (Jackman et al., 2016; Chen et al., 2017; Martinetti et al., 2022). In addition to its role in facilitation, Syt7 accelerates refilling of the RRP after depletion and mediates asynchronous release (Bacaj et al., 2013; Liu et al., 2014; Jackman et al., 2016; Tawfik et al., 2021). Syt7-dependent synaptic facilitation is interpreted as an activity-dependent increase in pv (Turecek et al., 2016; Jackman and Regehr, 2017; Turecek et al., 2017; Norman et al., 2023). However, it is unclear whether Syt7 contributes to facilitation through a Ca2+-dependent increase in pocc (i.e., overfilling). Overfilling has been proposed as a mechanism supporting facilitation at the Drosophila neuromuscular junction (Kobbersmed et al., 2020), yet whether this process also occurs at the mammalian central synapses mediated by Syt7 has not been studied.

Each of layer 2/3 (L2/3) and layer 5 (L5) of neocortex displays intralaminar excitatory synapses between pyramidal cells comprising a recurrent network (Holmgren et al., 2003; Thomson and Lamy, 2007). Short-term facilitation at recurrent excitatory synapses in a cortical network may play a crucial role in maintaining working memory (Mongillo et al., 2008; Mongillo et al., 2012; Hansel and Mato, 2013). Although L2/3 pyramidal cells (PCs) represent task-relevant information during a delay period in working memory tasks (Fujisawa et al., 2008; Ozdemir et al., 2020), most studies on local excitatory recurrent synapses in the neocortex have focused on L5 exhibiting short-term depression (Reyes and Sakmann, 1999; Hempel et al., 2000; Yoon et al., 2020). Further, the synaptic properties of excitatory synapses in L2/3 remain poorly understood.

In the present study, we found that HFS of local excitatory synapses in L2/3 of the medial prefrontal cortex (mPFC) induced initial strong depression followed by delayed facilitation, whereas low-frequency stimulation induced slow monotonous facilitation. This unique form of STP could be explained by 1) high pv of RRP vesicles at rest and throughout a train stimulation; 2) the low resting vesicular occupancy of RRP and its activity-dependent increase. These conditions led to delayed facilitation through progressive overfilling of release sites during HFS with high pv vesicles beyond the resting occupancy. Syt7 knockdown (KD) in L2/3 PCs completely abolished these synaptic enhancements by inhibiting the Ca2+-dependent increase in release site occupancy. Finally, behavioral tests for trace fear conditioning showed that Syt7 KD impaired the acquisition of trace fear memory and reduced c-Fos expression in the mPFC. Collectively, these results support that Ca2+- and Syt7-dependen overfilling of release sites mediates synaptic facilitation at L2/3 recurrent excitatory synapses and contributes to temporal associative learning.

Results

Frequency dependence of STP at local excitatory synapses in L2/3 of the prelimbic cortex

We examined STP profiles at recurrent excitatory synapses between L2/3 PCs and at synapses from L2/3-PCs to fast-spiking interneurons (FSINs) in the prelimbic area of the rat mPFC. The PCs and FSINs were identified based on their distinct intrinsic properties and morphologies (Figure 1—figure supplement 1). To stimulate presynaptic PCs, L2/3 PCs were transfected with a plasmid encoding oChIEF using in utero electroporation (IUE; Figure 1A; Lee et al., 2024). Photo-stimulation of the oChIEF-expressing cell soma at 40 Hz consistently evoked single APs with high fidelity throughout the 600-pulse trains (Figure 1—figure supplement 2A-B), which was consistent with little use-dependent inactivation of oChIEF (Lin et al., 2009; Hass and Glickfeld, 2016).

Frequency-dependence of short-term synaptic plasticity at prelimbic L2/3 excitatory synapses.

(A) In utero electroporation following injection of the plasmid (CAG-oChIEF-tdTomato) into the ventricle of an embryo (E17.5). (B) Representative images showing specific expression of oChIEF-tdTomato in L2/3 pyramidal cells after IUE. The red fluorescence of tdTomato clearly visualizes oChIEF-expressing cell bodies in L2/3 and axons in L5. Scale bar: 1 mm, 100 μm, 50 μm from left to right. (C) Recording schematic showing photostimulation of oChIEF-expressing axon fibers of transfected PCs (filled triangles) and a whole-cell recording from non-transfected PCs (empty triangles) or FSINs (empty circles). A collimated DMD-coupled LED was used to confine the area of excitation (typically 3–4 μm in diameter) to a small region (blue circle) near the soma. (D, E) Representative traces for EPSCs averaged over 10 trials at each frequency (left) and average amplitudes of baseline-normalized EPSCs (right) during 20-pulse trains at frequencies from 5 to 40 Hz at PC-PC (D; n = 12, 9, 21, 10 cells for 5, 10, 20, 40 Hz, respectively) and PC-FSIN (E; n = 8, 9, 10, 10) synapses. Each data point was normalized to the average of the first EPSC. (F) Baseline-normalized amplitudes of steady-state EPSCs (EPSCss; black symbols) and synaptic efficacy (EPSCss × f; red symbols) as a function of stimulation frequency (f) at PC–PC synapses. EPSCss was measured from the average of last 5 EPSCs from the 20-pulse trains (n = 12, 9, 21, 10). (G) Paired pulse ratio (PPR) as a function of inter-spike intervals (n = 25, 9, 25, 22, 43 for 200, 100, 50, 25, 20 ms ISI, respectively) at PC-PC synapses. (H) EPSCss and synaptic efficacy at PC-FSIN (n = 8, 9, 10, 10). (I) PPR at PC-FSIN (n = 8, 9, 10, 15, 34). Gray symbols, individual data.

Given that IUE transfected approximately 10-20% of L2/3 PCs (Figure 1A), we made a whole-cell patch on a non-transfected L2/3 PC or an FSIN (Figure 1C). To stimulate a minimal number of excitatory axon fibers, EPSCs were evoked using minimal optical stimulation (Figure 1—figure supplement 3A; 470 nm, 3-4 μm in radius and 3-5 ms in duration). EPSCs evoked by minimal stimulation were examined to determine the uniformity of the kinetic properties (Figure 1—figure supplement 3B-C). Moreover, we reassessed the kinetics of EPSCs during trains post hoc to ensure uniformity. EPSCs were evoked by 20-pulse trains at 4 different frequencies (5, 10, 20, and 40 Hz) to cover the range of firing rates of PCs observed in vivo during working memory tasks (Baeg et al., 2003). Excitatory synapses from PCs onto both other PCs and FSINs (PC–PC and PC–FSIN) showed strong (∼2-fold enhancement) and lasting (several seconds) short-term facilitation at 20 Hz and lower frequencies (Figure 1D-E). The PC-PC synapses exhibited stronger facilitation than did the PC-FSIN synapses.

Steady-state EPSCs at the PC-PC synapses were not significantly different from 5 to 20 Hz (Figure 1F). Frequency invariance at the PC-PC synapses suggests that Ca2+-dependent synaptic facilitation may counteract vesicle depletion in a frequency-dependent manner (Turecek et al., 2017; Lin et al., 2022). Despite facilitation at all frequencies, the paired pulse ratio (PPR) rapidly decreased as the ISI decreased (Figure 1G, I). At 40 Hz stimulation, both PC-PC and PC-FSIN synapses initially underwent strong depression then followed by facilitation in a few stimuli (Figure 1Dd, Ed). Such delayed facilitation after paired pulse depression (PPD) has not been previously observed in central synapses except for a slight increase in EPSC at the cerebellar synapses under artificially depolarized conditions (Pulido and Marty, 2018).

We confirmed that optically evoked EPSCs were strictly dependent on AP generation (Figure 1—figure supplement 4A), suggesting that EPSCs are evoked by optical stimulation of presynaptic axon fibers rather than by direct depolarization of presynaptic boutons. Additionally, dual patch-clamp techniques showed that the STP of optically evoked EPSCs did not differ from the electrically evoked STP of EPSCs at the PC-FSIN synapses (Figure 1—figure supplement 4B). Finally, we confirmed that AMPA receptor (AMPAR) desensitization was not responsible for PPD at 40 Hz, ruling out the possible effects of postsynaptic factors on STP (Figure 1—figure supplement 5A), and that AMPAR saturation was not significant during facilitation (Figure 1—figure supplement 5B).

Delayed facilitation results from slow activation of Ca2+-dependent vesicle replenishment at a constantly high vesicular fusion probability

The strong PPD at 40 Hz suggests a high vesicular fusion probability (pv) and likely tight coupling of the releasable vesicles to a Ca2+ source. To test this hypothesis, we examined the effects of EGTA, a slow calcium chelator, on synaptic transmission at the PC-PC synapses. The baseline EPSC amplitude (EPSC1) was not affected by incubation of the slice in the bath solution containing 50 μM EGTA-AM for longer than 20 min, supporting tight coupling. Moreover, the PPR was decreased to 0.19, and subsequent facilitation slowed but was still distinct (Figure 2A-B). These findings indicate that the initial pv should be higher than 0.81 considering vesicle recruitment during ISI. The effects of EGTA-AM were confirmed in the same cell by measuring the baseline EPSC and PPR before and after applying EGTA-AM to the bath. EPSC1 remained unchanged, while PPR was reduced by half (Figure 2—figure supplement 1).

Delayed facilitation results from slow activation of Ca2+-dependent vesicle replenishment at a constantly high vesicular fusion probability

(A) Representative traces (left) and mean baseline-normalized amplitudes (right) of EPSCs evoked by 30-pulse trains at 40 Hz in control (n = 21; black) and in the presence of 50 μM EGTA-AM (n = 14; green). (B) Mean values for the first EPSC amplitude (EPSC1, left) and PPR (right) from the experiments displayed in (A). Gray symbols, individual data. (C) Plot of rate constants for short-term facilitation (kSTF) as a function of stimulation frequency (fstim), showing a linear relationship. The linear regression line (black) is shown fitted to kSTF values, estimated from Figure 1. (D) Representative EPSC traces (left) and average of baseline-normalized EPSCs (right) evoked by 12-pulse stimulation at 5 Hz, followed by 40 Hz 7-pulse train (n = 12). Note that slowly developing facilitation was converted to rapid facilitation after strong PPD. (E) Mean values for PPR at 40 Hz. The baseline PPR was reproduced from Figure 1G and the PPR during 5 Hz train was calculated as (13th EPSC) / (12th EPSC). Gray symbols, individual data. All statistical data are represented as mean ± S.E.M.; n.s. = not significant; **, P<0.01; unpaired t-test.

Notably, the application of EGTA-AM markedly slowed facilitation at 40 Hz (Figure 2A, C). The frequency invariance of steady-state EPSCs (Figure 1F) and the slower facilitation at 40 Hz in the presence of EGTA (Figure 2A) imply a Ca2+-dependent increase in the refilling rate of the RRP and/or an increase in the pv of reluctant vesicles in light of previous studies (Hosoi et al., 2007; Liu et al., 2014; Ritzau-Jost et al., 2014; Turecek et al., 2016; Ritzau-Jost et al., 2018; Kusick et al., 2020). Moreover, facilitation was accelerated at higher stimulation frequencies. The plot of the rate constant for the increase in EPSCs (denoted as kSTF, the rate constant for the development of short-term facilitation) as a function of the stimulation frequency (fstim) revealed a linear relationship (Figure 2C). kSTF in the presence of EGTA-AM was distinctly below this relationship, supporting the Ca2+-dependent acceleration of facilitation.

kSTF was also accelerated when the stimulation frequency was increased from 5 Hz to 40 Hz (Figure 2D; 1.27 ± 0.18/s to 24.27 ± 3.44/s). Notably, the second 40-Hz train stimulation led to strong PPD, followed by accelerated facilitation (Figure 2D-E). The strong PPD suggests that the vesicular fusion probability kept high during the 5-Hz stimulation, arguing against the possibility for an increase in pv. Both two-step sequential priming models, LS/TS and RS/DS models, are suitable to interpret our data and we will refer to the different populations of vesicles as TS and LS vesicles, respectively (Taschenberger et al., 2016; Neher and Brose, 2018; Lin et al., 2022; Neher, 2024). Within this framework, vesicles are released from TS with high pv both in the resting state and during facilitation, and the delayed facilitation is seen as a result of Ca2+-dependent progressive increase in the occupancy of TS vesicles on docking sites.

Presynaptic calcium measurements at axonal boutons of L2/3 pyramidal cells

To test contribution of Ca2+ channel regulation to STP (Dobrunz et al., 1997; Jackman and Regehr, 2017), we estimated total amount of Ca2+ increments during 40 Hz AP trains at axonal boutons of L2/3 PCs in the mPFC. To this end, we measured AP-evoked Ca2+ transients (AP-CaTs) using two-dye ratiometry techniques, and estimated calcium binding ratio from AP-CaTs (Figure 2—figure supplement 2A-C). Measuring the amplitudes and decay time constants of AP-CaTs at different concentrations of Fluo-5F (150, 250 or 500 μM), and plotting them as a function of calcium binding ratio of Fluo-5F (κB), we estimated the amplitude of AP-CaTs (ACa = 1.16 μM), Ca2+ decay time constant (τc = 43 ms) in the absence exogenous Ca2+ buffer, and endogenous static Ca2+ binding ratio (κS = 96.4 or 77.3 estimated from Figure 2—figure supplement 2D-E, respectively; See Materials and Methods). The resting [Ca2+] ([Ca2+]rest) was measured as 50 nM (Figure 2—figure supplement 2F). Estimating the total Ca2+ increments for the first four AP-CaTs in a 40 Hz train according to Equation 2 (Materials and Methods), we found no evidence for inactivation or facilitation of calcium influx during a 40 Hz train (Figure 2—figure supplement 2G-I), arguing against the possibility that Ca2+ channel inactivation and facilitation contribute to strong PPD and delayed facilitation during a 40 Hz train.

Release sites have low baseline occupancy and these are increased during facilitation and post-tetanic augmentation

Given that pv was greater than 0.8, the two-fold increase in EPSC during 20 pulse train stimulations could not be explained by an increase in pv. Recent studies have shown that vesicle release sites (N) are limited in number and are partially occupied at rest (Miki et al., 2016; Pulido and Marty, 2018; Malagon et al., 2020; Lin et al., 2022). Under such a high pv, two-fold synaptic facilitation should be attributed to an increase in pocc. The vesicle replenishment rate is accelerated during HFS in a Ca2+-dependent manner (Figure 2A-C; Hosoi et al., 2007). This may lead to the overfilling of release sites beyond their basal vesicular occupancy, which in turn may increase EPSC under high pv. To test this hypothesis, we determined the quantal parameters at the L2/3 local excitatory synapses. We first measured the quantal size from asynchronous release events after replacing extracellular Ca2+ with Sr2+, as previously described (Yoon et al., 2020) (Figure 3—figure supplement 1A-B). The mean and coefficient of variance of the quantal sizes were 24 pA and 0.25 at PC-PC synapses, respectively, and were 30 pA and 0.31 at PC-FSIN synapses (Figure 3—figure supplement 1C).

Next, we performed a variance-mean (V-M) analysis of phasic EPSCs at excitatory synapses during train stimulation at 5 Hz and 40 Hz (Scheuss et al., 2002). Applying multiple-probability fluctuation analysis (MPFA) using intra-site quantal variability of 0.2 (Valera et al., 2012), the V-M plot was best fitted by the number of release sites (N) of 5.3 and the quantal size (q) of 25 pA at the PC-PC synapses (Figure 3A). The pr for EPSC1 was calculated as 0.32. We performed the same analysis for the PC-FSIN synapses and found N = 5.3, q = 35 pA, and pr for EPSC1 = 0.31 (Figure 3B). Our estimate for N was smaller but comparable to the estimate (6.9) at excitatory synapses in L2/3 of the mouse barrel cortex (Holler et al., 2021).

Low baseline occupancy of release sites and its increase during facilitation and post-tetanic augmentation

(A, B) Left, Representative EPSCs evoked by 5 and 40 Hz train stimulation (black, 5 Hz; purple, 40 Hz). Right, Variance-mean plots of EPSCs amplitude from averaged EPSCs recorded at PC-PC (A; n = 12, 17 for 5 and 40 Hz, respectively) and PC-FSIN (B; n = 8, 10) synapses. The data were fitted using multiple-probability fluctuation analysis (MPFA). Error bars are omitted for clarity. The 1st EPSC of 5 Hz train (broken line) was used to estimate the resting level of pocc. Filled circles were measured from post-tetanic augmented EPSCs (A, n = 12, 12, 9; B, n = 7, 7, 7 for 0.1, 0.2, 0.5 s IBIs, respectively). (C-E) Post-tetanic augmentation (PTA) experiments at PC-PC (top) and PC-FSIN (bottom) synapses. (C) Representative traces for EPSCs evoked by double 40 Hz train stimulations separated by 0.5 s. (D) Mean baseline-normalized amplitudes of EPSCs evoked by double 40 Hz trains at different inter-burst intervals (IBIs, 0.1, 0.2, 0.5, 1, 2, 5, 10 s). Upper, PC-PC synapse (n = 12, 12, 21, 11, 11, 16, 11 from short to long IBIs, respectively). Lower, PC-FSIN synapse (n = 10, 9, 8, 9, 9, 7, 9). (E) PTA time course. The baseline-normalized amplitudes of 1st EPSC from the 2nd train were plotted as a function of IBIs.

The pr estimate was clearly smaller than the possible minimum value for pv (0.81) estimated from the PPR data (Figure 2B). Given that pr is a product of pocc and pv (Neher, 2024), a pr smaller than pv indicates the partial occupancy of release sites in the resting state. If synaptic facilitation were mediated by the recruitment of new release sites, it would be accompanied by an increase in the variance of augmented EPSCs (Valera et al., 2012; Kobbersmed et al., 2020). However, the augmented EPSCs during facilitation followed the same parabola in the V-M plot (Figure 3A-B). Therefore, facilitation appears to be mediated by the progressive overfilling of a finite number of release sites beyond the basal occupancy. In the next section, we show that pv is close to unity based on high double failure rate upon paired pulse stimulation (Figure 3—figure supplement 2) and little effect of extracellular [Ca2+] ([Ca2+]o) elevation on baseline EPSC amplitudes (Figure 3—figure supplement 3). Assuming this, the baseline pr at the first pulse estimated from the fitted parabolic curve can be largely regarded as a baseline pocc (c.a. 30%), and the two-fold increase in EPSCs during facilitation can be attributed to an increase of up to 60% in pocc.

We investigated the time course of recovery from facilitation and post-tetanic augmentation (PTA) by applying dual-train stimulations (30 pulses at 40 Hz) separated by different inter-burst intervals (IBIs; Figure 3C-D). The initial EPSCs in the second train were enhanced by more than twofold, and followed by a strong PPD, suggesting that fusion probability stays high in the IBIs and thus synaptic transmission is carried by TS vesicles. The decay time course of PTA was characterized by two phases (initially fast and later slow; Figure 3E), which may reflect a two-phase Ca2+ decay or two distinct recovery processes that occur sequentially. The variance of potentiated EPSCs measured at IBIs of 0.1, 0.2, and 0.5 s was not different from that of EPSCs at the peak of facilitation (filled circles, Figure 3A-B). Given that the same RRP mediates both facilitation and augmentation and that post-tetanic EPSCs is carried by TS vesicles similar to the baseline EPSCs, this plot suggests that PTA, similar to short-term facilitation, is mediated by an increase in the TS vesicle occupancy.

High double failure rate and little effect of [Ca2+]o elevation on baseline EPSC amplitudes suggest that vesicular fusion probability is close to one

The vesicle dynamics of STP has been explained using a simple refilling model (Hosoi et al., 2007; Neher and Sakaba, 2008). This model assumes that the vesicles are reversibly docked to a limited number of release sites (N) with forward and reverse rate constants (denoted as k1 and b1, respectively). Under the framework of the simple refilling model (see Materials and Methods), the rate constant for the refilling of the RRP after depletion by the first pulse is the sum of the baseline k1 and b1. This is estimated as 23/s from the plot of PPR as a function of ISIs (Figure 3—figure supplement 2A). From the baseline occupancy of 0.3 and pocc = k1 / (k1 + b1), we estimated k1 to be 6.9/s.

When paired APs were applied with an ISI of 20 ms, the failure rate at 1st pulse, P(F1), was 10.6%, and the probability of two consecutive failures, P(F1, F2), was 6.2% (Figure 3—figure supplement 2A). From the equation P(F1) = (1 – pr)N, pr was calculated as 0.312 when N = 6, comparable to pr (= 0.32) estimated from the V-M analysis (Figure 3A). We calculated P(F1, F2) under the framework of a simple refilling model, with pv and k1 set as free parameters and other parameters set according to the following relationships: P(F1) = (1 – pr)N, pr = pv·pocc, and pocc = k1 / (k1 + b1). The calculated values for P(F1, F2) and their difference from the observed value (6.2%) are shown in the plane of k1 vs. pv in Figure 3—figure supplement 2B-C. Given that k1 was greater than 5/s, the difference between the calculated and observed values of P(F1,F2) was minimal when pv = 1 (Figure 3—figure supplement 2C).

To find optimal values for k1 and pv that best explain observed probabilities shown in Figure 3—figure supplement 2Ac, a cost function was implemented as the sum of squared errors between observed and predicted probability values for the four combinations of success and failure of 1st and 2nd EPSCs as described in Materials and Methods. The minimum of the cost function (5.63 x 10-5) was found at k1 = 5.21/s and pv = 0.999, which predicted P11 = 41.5%, P10 = 47.9%, P01 = 4.93%, and P00 = 5.67% (subscript 1 and 0 denote success and failure, respectively).

To further validate that baseline pv is near to unity, we examined EPSC changes following elevation of [Ca²⁺]ₒ from 1.3 to 2.5 mM. According to the fourth-power relationship between synaptic responses and [Ca²⁺]ₒ (Equation 3 in Dodge Jr and Rahamimoff, 1967), a 3.24-fold increase in the EPSC amplitude was expected. However, the EPSC amplitude increased only 1.23-fold in average, a change that was not statistically significant (Figure 3—figure supplement 3A-B). This small response suggests that pv is already saturated at rest, limiting the dynamic range for further enhancement through increased calcium influx. Recent morphological and functional studies revealed that elevation of [Ca2+]o induces an increase in the number of TS or docked vesicles to a similar extent as our observation (Kusick et al., 2020; Lin et al., 2025), raising a possibility that an increase in pocc is responsible for the 1.23-fold increase in EPSC at high [Ca2+]o. A slight but significant reduction in PPR was observed under high [Ca²⁺]ₒ too. An increase in pocc is thought to be associated with that in the baseline vesicle refilling rate. While PPR is always reduced by an increase in pv, the effects of refilling rate to PPR is complicated. For example, PPR can be reduced by both a decrease (Figure 2—figure supplement 1) and an increase (Lin et al., 2025) in the refilling rate induced by EGTA-AM and PDBu, respectively. Thus, the slight reduction in PPR is not contradictory to the possible contribution of pocc to the high [Ca2+]o effects. Notably, when slices were pre-incubated with 50 μM EGTA-AM, elevating extracellular [Ca2+] from 1.3 to 2.5 mM produced no significant change in either baseline EPSC amplitude or PPR (Figure 3—figure supplement 3C-D), supporting that the modest Ca2+-dependence of baseline EPSCs and PPR in the absence of EGTA is primarily mediated by a change in pocc rather than pv.

Pharmacological manipulations reveal specific molecular identities underlying vesicle loading processes

Our results suggest that the baseline occupancy of TS vesicles is low and that the activity-dependent progressive overfilling of docking sites with TS vesicles is responsible for short-term facilitation and augmentation. To elucidate the specific molecular link between synaptic enhancement and Ca2+-dependent overfilling, we examined the effects of several drugs known to affect vesicle dynamics. We first examined the roles of phospholipase C (PLC) and diacylglycerol (DAG) by applying 5 μM U73122, a PLC inhibitor, or 20 μM 1-oleoyl-2-acetyl-sn-glycerol (OAG), a DAG analog. The STP at 40 Hz and PTA at 0.5-s interval were examined before and after applying each drug to the bath. For both drugs, the facilitation and augmentation were reduced (Figure 4A-B). U73122 did not affect the basal EPSC amplitude (EPSC1) but resulted in a stronger PPD, slower facilitation, and lower augmentation (Figure 4A). These effects of U73122 suggest the involvement of Ca2+-induced PLC activation in progressive overfilling. In contrast, OAG increased EPSC1 while maintaining a pronounced PPD (Figure 4B), indicating an increase in the TS vesicle pool size. Meanwhile, OAG slowed kSTF and reduced augmentation, indicating occlusion of activity-dependent synaptic facilitation. Phorbol esters and DAG are thought to accelerate priming through Munc13 activation (Rhee et al., 2002; Taschenberger et al., 2016; Aldahabi et al., 2022); they also shorten the bridges that interlink docked vesicles with the active zone (AZ) membrane (Papantoniou et al., 2023). Collectively, these suggest that PLC- and DAG-mediated signaling play a role in the overfilling of docking sites underlying short-term facilitation and augmentation.

Pharmacological experiments reveal specific molecular mechanisms underlying vesicle loading processes

(A-C) (a) Representative EPSC traces (left) and mean baseline-normalized EPSCs (right) evoked by double 40 Hz train stimulations separated by 0.5 s inter-burst interval (IBI) in control and in the presence of 5 μM U73122 (Aa, n = 7, orange), 20 μM OAG (Ba, n = 12, cyan) or 100 μM dynasore (Ca, n = 10, blue). (b) Mean values for baseline EPSCs (EPSC1, left) and augmentation (right) from the experiments shown in corresponding a panel. (Da) Representative EPSC traces evoked by paired pulses (left) and mean values for baseline EPSC amplitude (middle) and PPR (right) before and after applying 20 μM LatB (n = 6). (Db) Representative EPSC traces (left, upper) and average of normalized EPSCs (left, lower) evoked by double 40 Hz train stimulation separated by 0.5 s in control (n = 21) and in 20 μM LatB conditions (n = 16; pink). Right, Mean values for augmentation in control and LatB conditions. Gray symbols, individual data. All statistical data are represented as mean ± S.E.M., *, P<0.05; **, P<0.01; ***, P<0.001; unpaired or paired t-test; n.s. = not significant.

Although the acceleration of Ca2+-dependent vesicle refilling supports rapid facilitation at 40 Hz, the later part of the delayed facilitation underwent a slight depression, which was more pronounced during the second burst (Figure 3D and Figure 4). Sustained synaptic transmission is limited by the spatial availability of docking sites because it is undermined by accumulation of exocytic remainings in the presynaptic AZ during HFS (Neher, 2010; Haucke et al., 2011). To verify this, we examined the effects of dynasore (100 μM), a specific inhibitor of endocytosis (Macia et al., 2006). The treatment with dynasore did not influence EPSC1 or the early phase of facilitation, but it exacerbated depression during the later phase and attenuated PTA (Figure 4C). This suggests that late synaptic transmission during the 40-Hz train is limited by site clearance.

Given the involvement of actin polymerization in the multiple steps of vesicle docking and recycling, we examined the effects of 20 μM latrunculin B (LatB), an inhibitor of actin polymerization. The effects of LatB were complicated: (1) EPSC1 was decreased, whereas PPR was increased (Figure 4Da) and (2) facilitation and augmentation were reduced (Figure 4Db). The latter effect may result from defects in vesicle replenishment rates, either through the impaired physical movement of vesicles or disrupted site clearance (Sakaba and Neher, 2003; Lee et al., 2012; Hallermann and Silver, 2013; Miki et al., 2016), as Ca2+-dependent vesicle refilling mediates facilitation and augmentation (Figure 2 and Figure 3). Meanwhile, the former effect might be attributed to a shift in equilibrium in the priming states towards a loosely docked state (LS) at rest. It is unknown whether actin affects vesicular fusogenicity in neurons, although it facilitates multiple steps of vesicle fusion by enhancing membrane tension in neuroendocrine cells (Wu and Chan, 2022).

Short-term facilitation at both types of local excitatory synapses in L2/3 is abolished by Syt7 knockdown

We examined whether Syt7 mediates facilitation at the PC-PC and PC-FSIN synapses in L2/3 of the mPFC. To this end, oChIEF and shRNA against Syt7 (shSyt7) were co-expressed in L2/3 PCs using IUE. Syt7 KD resulted in a complete loss of facilitation at all tested stimulation frequencies (Figure 5A-B). Short-term depression and paired-pulse depression were more pronounced as the stimulation frequency increased, and the frequency invariance of steady-state EPSC was abolished, implying little contribution of Ca2+-dependent vesicle recruitment (Figure 5C-F). The lack of facilitation could not be attributed to the failure of APs in presynaptic axons, as optical stimulation elicited reliable APs in Syt7 KD PCs during the 40-Hz train (Figure 1—figure supplement 2C). Syt7 KD had little effect on the basal properties, including the peak amplitude, rise/decay time, and time-to-peak, of single AP-induced EPSCs (Figure 1—figure supplement 3D-E). Moreover, Syt7 KD did not significantly affect the intrinsic properties of neurons (Figure 5—figure supplement 1).

STF at both types of local excitatory synapses is abolished by Syt7 KD

(A, B) Representative EPSC traces (left) and mean baseline-normalized amplitudes of EPSCs (right) evoked by 20-pulse trains at 5 to 40 Hz. STP was measured at PC-PC (A; n = 10, 9, 12, 9) and PC-FSIN (B; n = 9, 8, 10, 9) synapses, in which presynaptic Syt7 transcripts were depleted (Syt7 KD). Syt7 KD pyramidal cells are indicated as black triangles on the top. For comparison, STP in WT synapses is reproduced from Figure 1 (dotted lines). Same frequency color codes were used as in Figure 1. (C) Baseline-normalized amplitudes of steady-state EPSC (EPSCss; black symbols) and synaptic efficacy (EPSCss × f; red symbols) as a function of stimulation frequency (f) at PC-PC synapses. EPSCss was defined as the average of last 5 EPSC amplitudes from 20-pulse trains. (D) PPR as a function of inter-spike intervals (n = 10, 9, 12, 18, 26). (E) EPSCss and synaptic efficacy at PC-FSIN synapses. (F) PPR at PC-FSIN synapses (n = 9, 8, 10, 12, 24). Gray symbols, individual data.

To examine whether Syt7 KD had any nonautonomous effects on postsynaptic cells, we measured the quantal size (q) from asynchronous events in the presence of Sr2+ and found that the mean quantal size remained unaltered at Syt7 KD synapses (Figure 3—figure supplement 1C). This indicated that Syt7 did not influence postsynaptic parameters, including the density and kinetic properties of AMPARs. To test the specificity of shSyt7, we examined whether co-expression of the shRNA-resistant form of Syt7 restored synaptic facilitation in both excitatory synapse types. Presynaptic overexpression of shRNA-resistant Syt7 under the CAG promoter rescued the effect of shRNA at both PC-PC and PC-FSIN synapses (Figure 5—figure supplement 2). These results underscore the crucial role of presynaptic Syt7 in mediating short-term facilitation at local excitatory synapses in L2/3 of the mPFC.

Syt7 KD synapses exhibit complementary changes in the number of release sites and their vesicle occupancy

To estimate the quantal parameters in Syt7 KD synapses, we performed V-M analysis of EPSCs evoked by train stimulation (Figure 6A-B). Fitting a parabola to the V-M plot revealed higher baseline pocc in Syt7 KD synapses than in wild-type (WT) synapses for both synapse types (KD, 0.68 and 0.42; WT, 0.32 and 0.31 for PC-PC and PC-FSIN synapses, respectively). The pocc monotonously decreased during a train stimulation, as expected from short-term depression. Moreover, the number of release sites (N) was markedly lower in Syt7 KD synapses (N = 3.5) than in WT synapses (N = 5.3; Figure 3A). Next, we examined the PTA induced by the same protocol as in Figure 3 (two 40-Hz train stimulations separated by variable IBIs) at Syt7 KD synapses (Figure 6C). The 40 Hz train-induced synaptic depression quickly recovered during IBIs, but no PTA was observed at Syt7 KD synapses, indicating its crucial role in augmentation (Figure 6D-E).

Syt7 KD synapses exhibit complementary changes in the number of release sites and their vesicle occupancy

(A,B) Left, Representative traces of EPSCs evoked by 5 and 40 Hz train stimulations (black, 5 Hz; purple, 40 Hz). Right, Variance-mean plots of EPSCs amplitude from averaged EPSCs recorded at PC-PC (A; n = 9, 15, 12, 10, 9) and PC-FSIN (B; n = 4, 14, 10, 9, 8) synapses in which presynaptic Syt7 has been knocked down. The data were fitted using MPFA and error bars are omitted for clarity. The mean 1st EPSC amplitude of 5 Hz train (vertical broken line) was used for estimation of baseline pocc. Dashed parabolas indicate MPFA fits to variance-mean plot of WT synapses reproduced from Figure 3. (C-E) Recovery experiments at PC-PC (top) and PC-FSIN (bottom) synapses in which presynaptic Syt7 was knocked-down. (C) Representative EPSCs evoked by double 40 Hz train stimulations separated by 0.5 s. (D) Mean baseline-normalized amplitudes of EPSCs evoked by double 40 Hz trains at PC-PC (n = 12, 10, 9, 9, 10, 9, 8) and PC-FSIN (n = 10, 9, 10, 11, 8, 9, 10) synapses at different interburst intervals (IBIs). (E) Recovery time course. Baseline-normalized amplitudes of 1st EPSC from the 2nd burst were plotted as a function of various IBIs. Dotted lines indicate augmented EPSCs in the WT reproduced from Figure 3.

TS vesicle recovery following depletion is accelerated by Syt7

Syt7 is recognized not only for its crucial role in facilitation, but also for mediating Ca2+-dependent replenishment of releasable vesicles (Liu et al., 2014; Tawfik et al., 2021). Given that facilitation at L2/3 excitatory synapses is driven by progressive overfilling (Figure 2 and Figure 3), the loss of facilitation at Syt7 KD synapses in the present study implies that Syt7 plays a key role in the Ca2+-dependent acceleration of refilling and overfilling of the docking sites with TS vesicles. To test this hypothesis, we examined the recovery kinetics of EPSCs after TS vesicles were depleted by 3 pulses at 40 Hz in wild-type (WT) and Syt7 KD synapses (Figure 7). Strong PPD was observed not only in the first burst, but also in the second burst. This suggested that EPSCs in the second burst were mediated by TS vesicles, similar to those in the first burst (Figure 7A, D). Therefore, EPSC1 in the second burst can be interpreted as a release from TS vesicles recovering during a given IBI. The recovery time course of EPSC1 in the second burst indicated full recovery of TS vesicles within 100 ms in both excitatory synapse types in WT rats (Figure 7C, F). However, this recovery was remarkably slower in Syt7 KD synapses, requiring 5 s for full recovery, indicating that TS vesicle recovery was greatly accelerated by Syt7. Overall, synaptic facilitation and augmentation at local excitatory synapses in L2/3 can be ascribed to progressive activity-dependent increases in the occupancy of TS vesicles (Figure 2 and Figure 3), in which Syt7 plays a key role.

Recovery of TS vesicles following depletion is accelerated by Syt7

(A-C). Recovery experiments at PC-PC synapses in WT (A) and Syt7 KD (B). Mean baseline-normalized amplitudes of EPSCs evoked by two consecutive 3-pulse 40 Hz trains in WT (n = 16, 13, 11, 11, 11, 9 from short to long IBIs, respectively) and KD (n = 13, 11, 12, 8, 10, 11) synapses at different IBIs (0.1, 0.2, 0.5, 1, 2, 5 s). Inset, representative traces of EPSCs evoked by two consecutive 3-pulse 40 Hz train stimulations separated by 0.2 s. (D-F). Recovery experiments at PC-FSIN synapses in WT (D) and Syt7 KD (E). Mean baseline-normalized amplitudes of EPSCs evoked by two consecutive 3-pulse 40 Hz trains in WT (n = 9, 11, 11, 9, 10, 13) and KD (n = 9, 12, 9, 7, 10, 9) synapses at different IBIs. Inset, representative traces of EPSCs evoked by two consecutive 3-pulse 40 Hz train stimulations separated by 0.2 s. (C, F) Recovery time course. Baseline-normalized amplitudes of 1st EPSC from the 2nd burst were plotted as a function of various IBIs.

Behavioral consequences of Syt7 KD in L2/3 PCs of the mPFC

We previously proposed that the disparity in STP between the PC-PC and PC-IN synapses results in activity-dependent changes in the ratio of synaptic weights at the PC-PC and PC-IN synapses (Jee/Jie). This may in turn have a profound influence on persistent activity in a recurrent network by biasing the E-I balance (Yoon et al., 2020). Syt7 KD not only eliminated facilitation, but also the activity-dependent increase in the Jee/Jie ratio in the L2/3 network (Figure 8—figure supplement 1). This suggests the possibility that Syt7 contributes to the persistent activity during working memory tasks in the L2/3 recurrent network. Trace fear conditioning (tFC) requires associative learning between an auditory cue (conditioned stimulus, CS) and temporally separate aversive events (unconditioned stimulus, US). The formation of trace fear memory requires the prelimbic area of the mPFC and hippocampus. mPFC activity during the trace interval is thought to temporarily hold CS information, enabling the network to associate temporally discontinuous events (i.e., temporal associative learning) (Gilmartin et al., 2013). Synaptic facilitation and augmentation are implicated in temporary memory retention in recurrent networks (Mongillo et al., 2008; Mongillo et al., 2012).

Given that Syt7 is crucial for synaptic facilitation (Figure 5) and augmentation (Figure 6) in prelimbic L2/3 recurrent excitatory synapses, we tested whether trace fear memory formation was impaired in rats with depleted Syt7 transcripts, specifically in the L2/3 PCs of the mPFC. Syt7-targeted shRNA (shSyt7) or scrambled shRNA (Scr) constructs were transfected bilaterally into the L2/3 PCs of the mPFC using IUE (Figure 8A). Both the Scr-transfected control and shSyt7-transfected rats underwent tFC training (Figure 8B). Consistent with Gilmartin et al. (2013), in which prefrontal persistent firing was optogenetically inhibited, WT and KD animals exhibited similar freezing behavior during this acquisition phase. The following day, the rats were subjected to a tone test in which trace fear memory formation was assessed from the freezing response to a tone alone in a distinct context. Compared to the control group, the shSyt7-transfected group exhibited significantly lower levels of freezing behavior in response to auditory cues during the first four trials, suggesting an impairment in trace memory formation (Figure 8C).

Behavioral effects of Syt7 deficiency in L2/3 PCs of the mPFC

(A) Top, Representative images showing bilateral expression of U6-GFP in L2/3 of PCs after IUE at E17.5. Scale bar: 1 mm, 100 μm. Bottom, Schematic of trace fear conditioning and extinction (tone test) protocol. (B) Freezing behavior of control (expressing scrambled shRNA, Scr) and Syt7 KD rats during acquisition of tFC. (C) Freezing ratio during tone tests on following days. Data are shown as average freezing during T1–T4, T5–T8, or T9–T12 (T, trials). The freezing on T1-T4 was significantly lower in KD rats suggesting that formation of trace memory was impaired in Syt7 KD rats (n = 10, 11 for Ctrl and KD, respectively; P = 0.0007, F(1, 19) = 16.43, two-way repeated measures ANOVA; P = 0.0064, 0.0211, 0.0064, 0.0064; Holm-Sidak test). (D) Representative images of c-Fos immunoreactivity in the prelimbic cortex of control or Syt7 KD rats 90 min after tFC acquisition. c-Fos (left, red, Cy5) and GAD67 (right, cyan, Cy3) were immunostained in the same brain slice expressing U6-GFP (middle; green). Scale bar, 50 μm. (E) Exemplar images of c-Fos positive neurons expressing GFP (top) or GAD67 (bottom) in control (left) or Syt7 KD rats (right). Scale bar, 50 μm. (F-H) Effects of Syt7 KD on c-Fos density (F) and percentage of c-Fos positive neurons co-labeled with GAD67 (G) or GFP (H) in L2/3 or L5 of prelimbic cortex (n = 9, 9 for Ctrl and KD, respectively). Open symbols, individual data. All statistical data are represented as mean ± S.E.M.; ****, P<0.0001; unpaired t-test.

The tone test was repeated the following day to test for extinction memory formation. Syt7 KD animals exhibited much less freezing to the CS (+1d) than did WT animals. Considering the Rescorla-Wagner model (Yau and McNally, 2023), our results suggest that the strength of CS-US associative memory was weaker in KD animals than in WT animals; thus, the acquisition of trace fear memory was impaired by Syt7 KD in the L2/3 PCs of the mPFC.

Next, we assessed the locomotor activity and anxiety levels of rats using the open field test (OFT) and elevated plus maze test (EPM). Scr controls and Syt7 KD rats showed similar exploratory patterns, displaying comparable total distance moved and time spent in the periphery of the OFT or the open/closed arms of the EPM (Figure 8—figure supplement 2). Previous studies have suggested that the acquisition of trace fear memory requires elevated trace activity in prelimbic PCs (Gilmartin et al., 2013). Given that c-Fos, an immediate early gene, is upregulated following patterned activities in neurons, its expression is considered a reliable indicator of recent neuronal activation (Fields et al., 1997; Guzowski et al., 2005). To investigate the effect of Syt7 on neuronal activity in vivo during tFC, animals were sacrificed at the time of peak c-Fos protein expression, approximately 90 min after completion of the behavioral task (Arime and Akiyama, 2017).

To elucidate the activation patterns of specific neuronal populations, we performed immunohistochemical co-labeling of c-Fos and GAD67 (a GABAergic interneuron marker) upon tFC acquisition. The density of c-Fos(+) cells and the percentage of c-Fos(+) cells among GAD67(+) cells in all prelimbic layers were significantly lower in Syt7 KD rats than in control rats (Figure 8F-G). Additionally, analysis of the proportion of active cells among transfected L2/3 PCs (c-Fos+/GFP+) revealed that Syt7 KD L2/3 PCs were significantly more active than Scr-transfected PCs (Figure 8H). This unexpected hyperactivity of Syt7 KD PCs may be attributed to the high reciprocal connectivity between PCs and FSINs in the neocortex (Holmgren et al., 2003; Otsuka and Kawaguchi, 2009), as PCs with short-term depression would transmit less input to FSINs, which in turn would exert less feedback inhibition on reciprocally connected PCs. Collectively, these findings show that the slow activation of vesicle refilling supports various forms of facilitation at excitatory synapses in PFC L2/3, which is essential for neuronal activity during temporal associative learning.

Discussion

The present study characterized STP at local excitatory synapses in the L2/3 network of the rat mPFC at physiological extracellular [Ca2+]. These synapses displayed initial strong depression and then delayed facilitation at 40 Hz, while monotonous slowly developing facilitation was observed at lower frequencies. Our study suggests that such delayed facilitation after a brief depression results from a high pv (Figure 2 and Figure 3 and Figure 3—figure supplement 2), along with Ca2+-dependent delayed acceleration of vesicle refilling and overfilling (Figure 2 and Figure 3). The two-fold increase in EPSC during 20 pulse train stimulations under the condition of such a high pv suggests a Ca2+-dependent increase in the vesicle replenishment rate during a train; this then leads to overfilling of release sites beyond their incomplete basal occupancy (Figure 3A-B). The high pv and strong PPD did not result from artificial bouton stimulation, AMPAR desensitization, or any potential distortion caused by optogenetic methods (Figure 1—figure supplements 4 and 5). Additionally, similar to short-term facilitation, PTA could be explained by an increase in the TS vesicle occupancy (Figure 3C-E), consistent with previous studies showing post-tetanic increases in the pocc and/or vesicle pool size at various types of synapses (Lee et al., 2010; Vandael et al., 2020; Tran et al., 2023; Silva et al., 2024). These findings suggest that release sites are partially occupied at rest and that a progressive overfilling of release sites with TS vesicles underlies facilitation and augmentation at intracortical L2/3 excitatory synapses.

Definition of vesicular release probability in the present and previous studies

Previous studies for post-tetanic augmentation (Stevens and Wesseling, 1999; Garcia-Perez and Wesseling, 2008) observed invariance of the RRP size after tetanic stimulation. In these studies, the RRP size was estimated by hypertonic sucrose solution or as the sum of EPSCs evoked 20 Hz/60 pulses train (denoted as ‘RRPhyper’). Because reluctant vesicles can be quickly converted to TS vesicles (16/s) and are released during a train (Lee et al., 2012), it is likely that the RRP size measured by these methods encompasses both LS and TS vesicles. In contrast, we assert high pv based on the observation of strong PPD, failure rates upon paired stimulations at ISI of 20 ms (Figure 2 and Figure 3—figure supplement 2). Given that single AP-induced vesicular release occurs from TS vesicles but not from LS vesicles, pv in the present study indicates the fusion probability of TS vesicles. From the same reasons, pocc denotes the occupancy of release sites by TS vesicles. Note that our study does not provide direct clue whether release sites are occupied by LS vesicles that are not tapped by a single AP, although an increase in the LS vesicle number may accelerate the recovery of TS vesicles. As suggested in Neher (2024), even if the number of LS plus TS vesicles are kept constant, an increase in pocc (occupancy by TS vesicles) would be interpreted as an increase in ‘vesicular release probability’ if it was measured based on RRPhyper.

By the way, it should be noted that increases in both pv and pocc may contribute to PTA. Post-tetanic potentiation (PTP) and phorbol esters had similar effects on baseline EPSCs (Taschenberger et al., 2016), and PDBu increased both the TS vesicle pool size (1.9-fold) and fusion probability of TS vesicles (1.3-fold) at the calyx synapses (Lin et al., 2025). Therefore, it is possible that PTP may increase both pocc and pv of TS vesicles, although it has not been yet directly tested. Moreover, acceleration of the latency in the hypertonicity-induced vesicle release raises a possibility for an increase in fusion probability of TS vesicles (Stevens and Wesseling, 1999; Garcia-Perez and Wesseling, 2008), although it is still controversial that the acceleration of the release latency represents a reduction in the activation energy barrier for vesicle fusion (Schotten et al., 2015).

Progressive overfilling of docking sites

Conventional models for facilitation assume an increase in pv after an AP of RRP vesicles, which had been in a low pv or reluctant state at rest (Dittman et al., 2000; Turecek et al., 2016). This facilitation model is unlikely to explain our results, because TS vesicles are predominantly released not only at rest but also in the facilitated state during or after train stimulation, as evidenced by the strong PPD (ISI, 25 ms) during a 5-Hz train (Figure 2D-E) and at different intervals after a 40-Hz/30 pulse trains (Figure 3D). Moreover, the PPD (ISI, 25 ms) remained pronounced during the early phase of facilitation (Figure 7), arguing against the possibility of slowly increasing the pv of reluctant vesicles. These results suggested that facilitation was mediated by an activity-dependent progressive increase in occupancy of TS vesicles (TS occupancy). Facilitation through an increase in TS occupancy has been termed frequency facilitation, differentiating it from paired pulse facilitation that relies on a residual calcium-dependent transient increase in the pv of releasable vesicles (Neher, 2024).

Assuming an increase in TS occupancy, to attain a two-fold increase in EPSC (Figure 1), the baseline occupancy of the docking sites by TS vesicles should be low, as estimated from the V-M plot (c.a. 30% in Figure 3). However, it is unclear whether the remaining docking sites that are not occupied by TS vesicles are truly vacant or are occupied by reluctant LS vesicles that are not released by the AP or AP trains. TS vesicles can be supplied by the transformation of LS to TS vesicles, both of which reside at the same docking site (LS/TS model; reviewed in Neher, 2024). Alternatively, TS vesicles can be supplied from a distinct replacement pool (RS/DS model; reviewed in Silva et al., 2021). Further studies, including ultrastructural analyses of active zones, are required to address this issue.

Given the crucial role of Syt7 in facilitation (Figure 5), the facilitation mechanisms are closely related to how Syt7 mediates synaptic facilitation. Based on the slow kinetics of Syt7, previous computational modeling studies have proposed that Syt7 may mediate short-term facilitation through an activity-dependent increase in pv (Turecek et al., 2016; Jackman and Regehr, 2017; Turecek et al., 2017; Norman et al., 2023). The present study, however, found that both facilitation and augmentation were mediated by overfilling of the release sites (Figure 2 and Figure 3). Moreover, Syt7-KD synapses displayed slower recovery of TS vesicles (Figure 7). These results suggest that Syt7 is essential for activity-dependent overfilling of docking sites with TS vesicles. Whereas the baseline EPSC was not altered (Figure 1—figure supplement 3), and complementary changes in the number of docking sites and their baseline occupancy were observed in Syt7 KD synapses (Figure 6). These results raise a possibility that Syt7 may play a role in providing additional vacant docking sites that can be overfilled during facilitation. It remains to be elucidated whether the decrease in the number of docking sites at Syt7 KD synapses is related to the subcellular localization of Syt7 to the plasma membrane of the active zone (Sugita et al., 2001; Vevea et al., 2021).

STP model in light of known Ca2+ binding kinetics of Syt7

Syt7 acts as an upstream Ca2+ sensor that facilitates activity-dependent replenishment (Liu et al., 2014; Bacaj et al., 2015; Chen et al., 2017; Tawfik et al., 2021). The present study shows that EPSCs are mediated by TS vesicles both at rest and during activity-dependent facilitation. Releasable vesicles are homogeneous, and Syt7-dependent overfilling is responsible for the activity-dependent enhancement of EPSCs. Thus, we tested whether slowly developing facilitation could be replicated within the framework of a simple refilling model based on the known Ca2+ binding kinetics of Syt7 (Brandt et al., 2012). To this end, we made following assumptions: (1) a Ca2+-dependent increase in the forward refilling rate (k1) requires a full Ca2+-bound form of Syt7, to which multiple Ca2+ ions can bind (two or three Ca2+ ions on each C2A and/or C2B domain) (Brandt et al., 2012; Chon et al., 2015; Voleti et al., 2017); (2) sequential Ca2+ binding steps are synergistic through a cooperativity factor, b; (3) Syt7 senses AP-induced local [Ca2+] similar to that estimated at the calyx of Held (a Gaussian function with a full width at half maximum value of 0.2 ms and a peak of 40 μM) (Wang et al., 2008); and (4) residual [Ca2+] follows a mono-exponential function with an amplitude of 1 μM and decay time constant of 50 ms (Jackson and Redman, 2003).

Under these assumptions and the framework of the simple refilling model, slow and delayed facilitation could be replicated by modeling k1 as k1 = k1,b + K1,max × [Syt7: n Ca2+], where [Syt7: n Ca2+] represents the fraction of fully Ca2+-bound Syt7 (Figure 9A-C). Syt7 was slowly activated over the course of the stimulus train, suggesting that the cooperative binding of multiple Ca2+ ions progressively increased the putative active form of Syt7 because of its slow membrane binding/unbinding kinetics. Nevertheless, the late phase of facilitation at high frequencies was predicted to be higher in the simulation than in the experiment. As expected, based on the notion that sustained release during HFS is limited by site clearance in the active zone (Figure 4C), the model/data discrepancy became more pronounced at higher frequencies. Incorporating endocytic terms into the model may mitigate late-phase discrepancies. Moreover, removal of the catalytic function of Syt7 in the model effectively accounted for the complete loss of facilitation observed in Syt7 KD (Figure 9D).

STP model in light of known Ca2+ binding kinetics of Syt7

(A-D) Left, Schematic of allosteric calcium binding to Syt7. The number of Ca2+ bound to Syt7 was denoted as # in ‘S#’ in the reaction scheme. kon = 7/μM/s, koff = 10/s. Middle, Simulated changes of k1 (black) in response to 5 Hz train of Ca2+ transients (light blue traces). The priming step of the simple refilling model was assumed to be catalyzed by full Ca2+-bound form of Syt7. Accordingly, Ca2+-dependent increase in k1 was calculated as K1,max multiplied by a fraction of full Ca2+-bound form of Syt7. We assumed that local [Ca2+]i(t) follows a Gaussian function: (1/σ√2π) exp[-(t-tp)2/ 2σ2], in which tp = 0.25 ms and σ = 0.085 ms. Right, Fits of the Syt7 model to the STP data. To fit this model to the STP data, K1,max was set to 300/s (A), 220/s (B), and 180/s (C). Cooperativity factor (b) was set to 0.35 (A), 0.2 (B), and 0.05 (C). k1 is set to be constant for Syt7 KD (D). Same frequency color codes were used as in Figure 1.

Conventional STF models involving cumulative increase in pv of reluctant vesicles

Conventional models for facilitation assume a post-AP residual Ca2+-dependent step increase in pv of RRP (Dittman et al., 2000) or reluctant vesicles (Turecek et al., 2016). Given that pv of TS vesicles is close to one in the present study, an increase in pv of TS vesicles cannot account for facilitation. The possibility for activity-dependent increase in fusion probability of LS vesicles (denoted as pv,LS) should be considered in two ways depending on whether LS and TS vesicles reside in distinct pools or in the same pool. Notably, strong PPD at short ISI implies that pv,LS is near zero at the resting state. Whereas LS vesicles do not contribute to baseline transmission, short-term facilitation (STF) may be mediated by cumulative increase in pv of LS vesicles that reside in a distinct pool. Because the increase in pv,LS during facilitation recruits new release sites (increase in N), the variance of EPSCs should become larger as stimulation frequency increases, resulting in upward deviation from a parabola in the V-M plane, as shown in recent studies (Valera et al., 2012; Kobbersmed et al., 2020). This prediction is not compatible with our results of V-M analysis (Figure 3), showing that EPSCs during STF fell on the same parabola regardless of stimulation frequencies. Therefore, it is unlikely that an increase in fusion probability of reluctant vesicles residing in a distinct release pool mediates STF in the present study. For the latter case, in which LS and TS vesicles occupy in the same release sites, it is hard to distinguish a step increase in fusion probability of LS vesicles from a conversion of LS vesicles to TS. Nevertheless, our results do not support the possibility for a gradual increase in pv,LS that occurs in parallel with STF. Strong PPD, indicative of high pv, was consistently found not only in the baseline (Figure 2 and Figure 2—figure supplement 1) but also during post-tetanic augmentation phase (Figure 3D) and even during the early development of facilitation (Figure 2D-E and Figure 7), arguing against the gradual increase in pv,LS. One may argue that STF may be mediated by a drastic step increase of pv,LS from zero to one, but it is not distinguishable from conversion of LS to TS vesicles.

Possible mechanisms underlying the depression-facilitation sequence

Cooperative binding of multiple Ca2+ ions to Syt7 predicts delayed activation and a progressive increase in vesicle refilling during train stimulations, resulting in delayed facilitation. Combining this with a high pv could replicate not only the slow facilitation at 5 Hz, but also the unique depression-facilitation sequence observed at 40 Hz (Figure 9). As simulated by Pulido and Marty (2018), delayed facilitation can be reproduced by a high pv in combination with a low baseline occupancy of replacement sites and subsequent Ca2+-dependent refilling. However, whether this model is also able to replicate slow facilitation at low frequencies remains to be tested. An alternative, but not exclusive, possibility for delayed facilitation is that the presynaptic global Ca2+ itself may undergo a slow buildup during a train due to the saturation of endogenous Ca2+-binding proteins.

Facilitation requires not only Syt7, but also activation of PLC-DAG signaling

To gain insight into the specific molecular mechanisms underlying the vesicle-loading processes, we examined the effects of pharmacological inhibitors targeting PLC and DAG (Figure 4). Both a PLC inhibitor and a DAG analog reduced facilitation and augmentation. This suggests that the progressive overfilling of docking sites with TS vesicles is supported by the PLC-DAG signaling pathway and probably downstream Munc13 (Rhee et al., 2002; Lou et al., 2008). Phorbol esters have been recently shown to increase the TS fraction of docked vesicles by enhancing vesicle priming without affecting Ca2+ transients or pv (Taschenberger et al., 2016; Aldahabi et al., 2022; Papantoniou et al., 2023; Aldahabi et al., 2024). In line with these findings, signaling involving PLC and DAG may support Ca2+-dependent overfilling of TS vesicles by enhancing molecular priming processes.

Roles of facilitation and augmentation in working memory and related tasks

Previous theoretical studies have proposed that working memory in a recurrent network can be maintained in the form of augmentation at synapses (Mongillo et al., 2008) and/or persistent activity of neurons in a memory ensemble (Mongillo et al., 2012). The latter can be supported by short-term facilitation of recurrent excitatory synapses or disparity in STP between excitatory and inhibitory synapses (Mongillo et al., 2012; Yoon et al., 2020). However, whether these short-term enhancements in synaptic efficacy contribute to working memory-related behaviors has not been examined. tFC is a temporal associative learning task requiring temporary retention of CS information during a trace period. The mPFC and hippocampus are crucial for tFC, and tFC is interfered with by a load of working memory and attention distraction, implying that the tFC shares the same neural substrate with other working memory tasks (Raybuck and Lattal, 2014). In the present study, tFC in WT and Syt7 KD animals revealed that Syt7 deficiency in the L2/3 PCs of the mPFC impaired the acquisition of trace fear memory and reduced c-Fos expression associated with tFC training (Figure 8).

The acquisition of trace fear memory was impaired by inhibition of persistent activity in mPFC during trace period (Gilmartin et al., 2013). The similar deficit observed in Syt7 KD animals is consistent with the hypothesis that STF provides bi-stable ensemble activity in a recurrent network (Mongillo et al., 2012). Nevertheless, alternative mechanisms may be responsible for the behavioral deficit. Not only recurrent network but also long-range loop between the mPFC and the mediodorsal (MD) thalamus play a critical role in maintaining persistent activity within the mPFC especially for a delay period longer than 10 s (Bolkan et al., 2017). Prefrontal L2/3 is heavily innervated by MD thalamus, and L2/3-PCs subsequently relay signals to L5 cortico-thalamic (CT) neurons (Collins et al., 2018). Given that L2/3 is an essential component of the PFC-thalamic loop, loss of STF at recurrent synapses between L2/3 PCs may lead to insufficient L2/3 inputs to L5 CT neurons and failure in the reverberant PFC-MD thalamic feedback loop. Therefore, not only L2/3 recurrent network but also its output to downstream network should be considered as a possible network mechanism underlying behavioral deficit caused by Syt7 KD L2/3.

In summary, activity-dependent synaptic facilitation and augmentation at prelimbic L2/3 recurrent excitatory synapses can be attributed to overfilling of release sites with TS vesicles. Given that the lack of short-term facilitation in L2/3 PCs impairs trace fear memory acquisition, temporary maintenance of CS information during trace period may be supported by an increase in the TS occupancy in presynaptic terminals of CS-representing neuronal ensembles.

Materials and methods

Animals/Subjects

All experiments were carried out on Sprague-Dawley rats of either sex at p28-43. Rats were maintained in temperature-controlled rooms on a constant 12 h light/12 h dark cycle and were given access to water and food ad libitum. All animal procedures were approved and performed in accordance with the Institutional Animal Care and Use Committee at Seoul National University (SNU200522-1).

In utero electroporation (IUE)

Pregnant Sprague Dawley rats on embryonic day (E) 17.5 were deeply anesthetized with 5% (v/v) isoflurane for the duration of surgery. The uterine horns were exposed by laparotomy and wet with warm sterile phosphate-buffered saline (PBS). The plasmids (0.5-2 µg/µl) together with the 0.01% fast green dye were injected into either the left or the right lateral ventricle or bilateral ventricles of embryos through a thin glass capillary (Narishige PC-10). The embryo’s head was carefully held between the forceps-shaped electrodes of 10 mm diameter (CUY65010; NEPAGENE, Japan). Electroporation pulses (50 V for 50 ms) were delivered five times at intervals of 150 ms using a square-wave generator (ECM830; BEX, United States). The uterine horns were placed back into the abdominal cavity and abdominal muscles and skin were sutured. After the surgery, the animals were recovered under the infrared lamp. For knockdown experiments in slice electrophysiology, the shRNA constructs were co-electroporated with the pCAG-oChIEF-tdTomato, because of the consistent co-expression of the two constructs in a significant proportion of the transfected cells in vivo (Bony et al., 2013). Pups with robust fluorescence signals in the prefrontal cortex of both hemispheres were used for behavior experiments.

Preparation of vector constructs

For an optic stimulation, pCAG-oChIEF-tdTomato was transfected by IUE. Short hairpin RNA (shRNA) against Syt7 sequence (GATCTACCTGTCCTGGAAGAG) was expressed under the control of an U6 promoter of pAAV-U6-GFP vector (Cell Biolabs, #VPK-413). This shRNA was shown to effectively knockdown Syt7 in cell cultures and in vivo by previous reports from more than one research group (Bacaj et al., 2013; Li et al., 2017). Scrambled control sequence (TCGCATAGCGTATGCCGTT) was designed by shuffling the recognition region sequences and a BLAST analysis confirmed that the control shRNA sequence had no target gene. For rescue experiments, the cDNA for rat Syt7 (NM_021659) rendered resistant to the shRNA were inserted in the KD vector and their expression was driven by the CAG promoter; the vector also contained a self-cleaving P2A peptide sequence followed by EGFP, which ensures visualization of transfected cells. All the constructs above were verified by DNA sequencing. Plasmid DNA was purified and concentrated under endotoxin free conditions (Qiagen EndoFree Maxi Kit).

Slice preparation

Acute brain slices were prepared by using a VT1200S Vibratome (Leica Microsystems) after isoflurane anesthesia and decapitation. Coronal sections of the mPFC (300 μm in thickness) were cut in ice-cold artificial cerebrospinal fluid (aCSF) composed of (in mM) 125 NaCl, 3.2 KCl, 25 NaHCO3, 1.25 NaH2PO4, 20 D-glucose, 2 NaPyr, 1 MgCl2, and 1.3 CaCl2 (Yoon et al., 2020), bubbled with 95% O2 and CO2 (pH = 7.3; 310 mOsm). Slices were allowed to recover for a minimum of 30 minutes at 36 °C, and subsequently maintained at room temperature until recording. Composition of aCSF was identical throughout the preparation, incubation, and recording periods unless otherwise stated. For the recordings, the mPFC slices were continuously perfused with oxygenated aCSF at a flow rate of 1.5 ml/min, and the temperature was maintained at 31–32 °C using an in-line temperature controller (Warner Instruments).

Electrophysiology

Whole-cell patch-clamp recordings were performed from pyramidal cells and fast-spiking interneurons in layer 2/3 of the prelimbic mPFC. Neuronal subtypes were initially selected based on morphology under visual guidance by using an upright microscope equipped with infrared differential interference contrast (IR-DIC) optics (BX51WI, Olympus), and were further identified by their electrophysiological properties (see Figure 1—figure supplement 1; van Aerde and Feldmeyer, 2015; Tremblay et al., 2016). To analyze these intrinsic properties in current-clamp experiments, we measured the following parameters: (1) resting membrane potential (RMP), (2) input resistance (Rin), (3) sag ratio, and (4) the F-I curve, which plots firing frequencies (F) against the amplitude of injected currents (I). Both cell types exhibited a low sag ratio below 0.1, consistent with findings from other studies (van Aerde and Feldmeyer, 2015; Yoon et al., 2020). PCs and FSINs were easily distinguished by their unique firing frequency, Rin, spike adaptation, RMP and morphological characteristics including the shape of soma and the thickness of apical dendrites. For experiments involving oChIEF-expressing neurons in WT or Syt7 KD rats, whole-cell recordings were performed from the fluorescently labeled pyramidal neurons in layer 2/3 of the prelimbic cortex using suitable filter sets. Patch pipettes (tip resistance of 2–3 MΩ) were pulled from borosilicate glass and filled with (in mM) 130 K-gluconate, 8 KCl, 10 HEPES, 0.2 EGTA, 20 Na2Pcr, 0.3 Na2GTP, 4 MgATP. Intracellular solutions were adjusted to pH 7.25 and 300–310 mOsm. Neurons were voltage clamped at −78 mV and series resistance (Rs) were continuously monitored during the EPSCs recordings. Experiments were discarded before analysis if the Rs changed by 20% or larger deviation of baseline value. Measurements were not corrected for Rs compensation, bridge balance, or liquid junction potential. Electrophysiological data were low-pass-filtered at 1 kHz (Bessel) and acquired at 20 kHz using a Multiclamp 700B amplifier paired with Digidata 1440A digitizer and Clampex 10.2 software (Molecular Devices).

Synaptic stimulation

IUE resulted in sparse expression of channelrhodopsin (10-20%) (Figure 1A). Exploiting the sparse expression, we employed a minimal optical stimulation method to activate a single or very few excitatory synapses in layer 2/3. The optic minimal stimulation was performed by our published protocol with minor modifications (Yoon et al., 2020). This approach aimed to minimize the risk of synaptic contamination, which could arise from engaging extra terminals that were initially inactive (depolarized below threshold) but became active by temporal summation during high-frequency trains. It also helped prevent the buildup of depolarization at the synaptic terminals, which might influence the probability of release, as noted by (Jackman et al., 2014). To achieve this, we employed a collimated digital micromirror device (DMD)-coupled LED (Polygon400; Mightex Systems) to confine 470 nm blue light to a small area with a radius of approximately 3–8 μm (typically 3–4 μm), as measured at the focal plane of a 60× water immersion objective [numerical aperture (NA) = 1.0; LUMPlanFL, Olympus] (Figure 1—figure supplement 3A). To find the minimal stimulation area, the radius of illumination area was increased from 2 μm by 50% at each step (Figure 1—figure supplement 3A), and selected the smallest illumination area that elicited EPSCs. Setting the duration of illumination between 3 and 5 ms (typically 5 ms), photostimulation onto oChIEF-expressing cells reliably induced action potentials (APs) during 600-pulse trains (see Figure 1—figure supplement 2).

STP at various stimulus frequencies (5, 10, 20, and 40 Hz) was examined using 20-pulse trains. Trials of train stimulation were repeated with a prolonged interval of 30 s to avoid change in baseline properties. We observed no systematic changes in the amplitudes of the first EPSC of each train, nor did we find significant differences in the paired-pulse ratio (PPR) across stimulation trials in the same cell. For each stimulus frequency, approximately ten to twenty consecutive traces were collected and then averaged across trials to yield a mean EPSC trace. For recovery experiments, two consecutive trains of 3 or 30 stimuli at 40 Hz, with varying time intervals (0.1-10 s), were applied every 10 s or 60 s, respectively, at both excitatory synapses. For 600-stimulus trains at 10Hz, the trial was delivered only once for each cell to avoid possible after-effects caused by prolonged synaptic stimulation (Hirsch and Crepel, 1990). Prior to the stimulus trains, the EPSCs were normalized to the mean amplitude of the baseline. For the baseline measurements, paired pulses with various ISIs (20-200 ms) were delivered every 3 s (1/3 Hz) at least 20 times.

For all recordings, picrotoxin (PTX, 10 μM) was included in the recording solution to isolate glutamatergic responses. mPFC PC terminals were entirely glutamatergic as the application of 6-cyano-7-nitroquinoxaline-2,3-dione (CNQX, 20 μM), an AMPA receptor antagonist, completely abolished the monosynaptic events evoked by photostimulation. A low concentration of PTX was employed to minimize the relieving effects on spike suppression caused by blockade of GABAa receptors. We confirmed that no synaptic response was elicited by electrical stimulation in the presence of both 10 μM PTX and 20 μM CNQX.

In Figure 1—figure supplement 4A, 1 µM TTX (1 mM stock solution) and 0.1 mM 4-AP (500 mM stock solution) were added to the bath solution to block action potentials and restore presynaptic glutamate release, respectively (Figure 1—figure supplement 4). In Figure 1—figure supplement 5A, 50 µM of cyclothiazide, a positive allosteric modulator of the AMPA receptors, was included to minimize rapid desensitization of the postsynaptic receptors. In Figure 1—figure supplement 5B, 0.5 mM kynurenate was added to bath solution to prevent possible saturation of AMPA receptors, especially during maximal synaptic facilitation. For pharmacological manipulations of synaptic vesicle dynamics (Figure 4), drugs related to vesicle supply including U73122, OAG, LatB, EGTA-AM, and dynasore at indicated concentrations were added to the recording solution during baseline measurements or PTA/recovery experiments (IBI = 0.5 s). Chemicals were from Sigma-Aldrich, Merck, or Tocris.

Measurements of quantal size

Experiments for measuring asynchronous release were performed as previously described (Yoon et al., 2020). In brief, we replaced extracellular Ca2+ entirely with 2 mM Sr2+ to induce asynchronous release. A single stimulus was delivered at 3-second intervals over 100 repetitions, and asynchronous release events were detected during a 100-ms window starting 25 ms after stimulus onset for each trial. The same procedure was also executed under sham conditions (0% light intensity) to serve as a control, thus mitigating contamination from nonspecific spontaneous events and potential errors associated with in the detection algorithm. q was determined as the mean value from the first Gaussian function after fitting the EPSC distribution with either a single or double Gaussian function (Figure 3—figure supplement 1).

Estimation of Ca2+ kinetics at axonal boutons

Cytosolic [Ca2+] at single axonal boutons was estimated by employing the two-dye, dual-excitation wavelength technique as described by (Oheim et al., 1998). This method involves measuring the fluorescence ratio of a Ca2+-sensitive dye to that of a Ca2+-insensitive dye to correct for dye concentrations. We used Fluo-5F at concentrations of 150, 250, or 500 μM (dissociation constant, Kd = 2.3 μM) as the calcium indicator dye and Alexa Flour 555 (50 μM) as the reference dye. The fluorophores were from Invitrogen. Both dyes were co-loaded into presynaptic terminals of layer 2/3 pyramidal cells in mPFC through whole-cell patch clamp. Before data acquisition, the dyes were allowed to diffuse to terminal boutons for at least 30 minutes after the patch break-in, a duration sufficient for the equilibrium of dye concentrations between the soma and its axonal terminals.

Calcium imaging was performed by sequentially exciting Fluo-5F and AF555 at wavelengths of 473 and 559 nm, respectively, using a confocal laser-scanning microscope (FV1200; Olympus, Japan) equipped with a 60× water immersion objective (NA, 0.9; LUMPlanFl; Olympus, Tokyo, Japan). Once a single axonal bouton was identified, lines for scanning were drawn across boutons, oriented perpendicular to the axon (Figure 2—figure supplement 2A-B). After baseline fluorescence was measured, line scanning was proceeded while a single or train APs were induced by applying current pulses. Upon stimulation, green fluorescence but not red one increased along the vertical array of lines across the bouton (Figure 2—figure supplement 2B-C). Measurements were obtained under current-clamp conditions, with the resting potential regularly monitored. The scanning speed was adjusted to 600-700 Hz by controlling the line length. Line scan for measuring fluorescent transients in a bouton was repeated at least 3 times, and then averaged to enhance the signal-to-noise ratio. To mitigate photobleaching, minimal intensity of the laser power and maximal pinhole size were used. Emitted light was filtered through 490/590 or 575/675 bandpass filters before detection by photomultiplier tubes for green or red fluorescence, respectively.

Fluorescence measurements were converted to [Ca2+]i using the ratio of background-subtracted fluorescence (F = F0 - Fb) from two dyes. The fluorescence ratio (R = Fgreen/Fred) in data traces were converted to [Ca2+]i using the following equation: Calibration parameters were obtained using an in-cell calibration protocol (Helmchen, 2011), wherein the minimum (Rmin) and maximum ratio (Rmax) values were established through the use of intracellular solutions containing 10 mM EGTA or 10 mM [Ca2+], respectively.

When Ca2+ is extruded with a rate constant, γ, after an AP-induced short pulse of Ca2+ influx to a single compartment, [Ca2+]i initially rises to ACa and then undergoes a mono-exponential decay with a time constant (τCa). Under this framework, ACa and τCa depend on Ca2+ binding ratios of endo- and exogenous Ca2+ buffers (denoted as κS and κB, respectively) as following equations (Neher, 1995; Kim et al., 2003):

, where Δ[Ca2+]T is an increment in total [Ca2+] (free plus bound) elicited by an AP. Ca2+ binding ratio indicates the ratio of concentration changes in the Ca2+-buffer complex relative to free [Ca2+] change, and linearized κB is calculated from the dye concentration and the change from [Ca2+]1 to [Ca2+]2 as follows: The total dye concentration, [B]t, was taken as [Fluo-5F] in the patch pipette. With κB determined from Equation 4 and τCa from the decay of a Ca2+ transient, κS can be estimated by plotting τ versus κB and applying a linear fit to the plot, which gives the x-intercept as -(1 + κS). Once determined, κS can be used to calculate Δ[Ca2+]T.

Optimization of k1 and pv from failure rates

Let n0 = the number of docked SVs before arrival of a first AP; n1 = the number of docked SVs remaining after a first AP (i.e. n0 - n1 vesicles are released); n2 = the number of docked SVs just before arrival of a 2nd AP (ISI = 20 ms; Figure 3—figure supplement 2). If the number of release sites (N), forward and reverse rates for vesicle docking (k1 and b1, respectively), mean release site occupancy (pocc) and vesicular fusion probability upon an AP (pv) are given, we can predict P(n0, n1, n2), the probability for [n0, n1, n2], as following steps.

1) Probability for a synapse harbouring n0 docked SVs just before arrival of a first AP is calculated as: 2) Probability that n1 vesicles remain after a first AP when the initial number of docked SVs is n0 [P(n1|n0)] is calculated as: 3) When forward and backward rate constants are k1 and b1, respectively, probability that n2 vesicles are docked just before arrival of a 2nd AP when n1 SVs remain after a 1st AP (Δt = 20 ms) [P(n2 | n1)] is predicted as follows: Assuming that docking and undocking follow a Poisson process, , where B = 1 – exp(-b1 Δt); F = 1 – exp(-k1 Δt); i = n2 – n1; o1 = N-n1.

4) Combining probabilities calculated from above three steps, we can calculate P(n0, n1, n2):

5) Once P(n0, n1, n2) is given, we can predict probability for double failure (P00) and probability for the 1st success and the 2nd failure (P10) as:

6) Probability for the 1st failure and the 2nd success (P01) and probability for double success (P11) were calculated as: , in which P(F1) and P(S1) are probability for the 1st failure and the 1st success, respectively, and these are calculated as: To optimize k1 and pv, we used the Matlab routine, fmincon, in which the sum of squared errors (= observed value minus predicted value) was set as a cost function to be minimized. Lower and upper bounds for k1 and pv were set as [4, 8] and [0.8, 1], respectively. Release probability (pr) was calculated from the observed 1st failure rate, P(F1), as N was assumed to be six as shown in Figure 3. Other parameters (b1, and pocc) were set according to following relationships:

Calculation of probability for double failure

Double failure probability, P00, can be more easily predicted. By the Bayes rule, , where F1 and F2 means 1st and 2nd failure, because n0 = n1 after the 1st failure. P(F1) and N are observed values. For calculation in Figure 3—figure supplement 2, k1 and pv were set free variables. Other parameters were set under the following constrains: P(F1) = (1 – pr)N, pr = pv pocc, and pocc = k1 / (k1 + b1).

Behavioral tests

Trace fear conditioning

For trace fear conditioning (tFC) (Gilmartin et al., 2013), individual rats were placed in the training chamber (30.5 x 25.4 x 30.5 cm; Coulbourn Instruments) consisting of metal grid floor connected to an electrical stimulator (H10-11R-TC-SF, Coulbourn Instruments) for delivery of a footshock. After a 5 min habituation period, rats received five pairings of a 20 s tone conditioned stimulus (CS, 2.5 kHz, 75 dB), followed by empty 20s trace period and a unconditioned stimulus (US; 1 s, 0.7 mA footshock) with a pseudorandom inter-trial interval (ITI) of 240 ± 20 s. During this training session, rats learned to associate the auditory CS with the shock US. The next day, the rats were tested for memory of CS-US trace association (called tone test). The acquisition of fear learning, often indicated by the extent of freezing to the CS, was assessed in a novel chamber (a white hexagonal enclosure) in a dimly lit room. During this tone (or retrieval) test, rats were first allowed to explore the new context for 5 min (habituation), followed by twelve 20 s CS presentations with a variable ITI of 120 ± 20 s. This entire session was repeated on the next day for assessment of extinction learning. Identities of the chambers were also determined with the presence of distinct odors; the training and testing chambers were cleaned with 70% ethanol and 1% acetic acid respectively before and after each session. For all experimental sessions, the activity of animals in the chambers was recorded at 30 frames per second using the EthoVision XT (Noldus Information Technology, Wageningen, Netherlands), and stored as a video file. The freezing behavior, defined as behavioral immobility other than respiratory movement, was assessed using an open-source video analysis pipeline ezTrack (Pennington et al., 2019). The percentage of freezing was calculated as a total duration of freezing divided by the total duration of observation.

Elevated Plus Maze

Elevated plus maze (EPM) test was performed using a plus-shaped apparatus elevated 60 cm above the ground. The maze consisted of two open arms (50 x 10 x 0.5 cm (H)), two closed arms (50 x 10 x 40 cm (H)) and a center area (10 x 10 cm). Rats were individually placed on the center area, facing an open arm, and the path of the animal was recorded with a video camera for 5 min. We analyzed the total distance traveled for 5 min and time spent in each arm. Animals were tracked offline using the open-source tracking system ezTrack (Pennington et al., 2019).

Open field test

Open field exploration test (OFT) was performed in an open field apparatus in a dimly lit room. The open field consisted of a white plastic board (1.20 m in diameter) surrounded by white plastic walls (50 cm in height). Individual rats were placed in a center zone (0.6 m in diameter), and allowed to freely explore the apparatus for 10 min. Time spent in the center zone and total distance traveled were analyzed for initial 5 min. Animals were tracked offline using the open-source tracking system ezTrack (Pennington et al., 2019).

Histology and immunohistochemistry

At the end of the behavioral experiments, rats were anesthetized using isoflurane inhalation and transcardially perfused with PBS followed by 4% paraformaldehyde (PFA, T&I, Korea). Brains were extracted and post-fixed in 4% PFA solution for 24 h. Perfused brains were sliced at 100 μm for estimation of GFP expression or 50 μm for immunohistochemistry. For analysis of c-Fos expression, rats were returned to their home cages immediately after the training session and perfused approximately 90 minutes later (Arime and Akiyama, 2017). Brains were then cryopreserved, frozen, and sectioned coronally. Immunofluorescence staining was conducted on free-floating slices using a rabbit monoclonal anti c-fos (1:1000, Cell Signaling Technology, cat# 2250s) and mouse monoclonal anti-GAD67 (1:500, Millipore, cat# mab5406). Fluorescent secondary antibodies included Cy5-conjugated goat anti-rabbit IgG (1:1000, Abcam, cat# ab97077) and Cy3-conjugated goat anti-mouse IgG (1:1000, Abcam, cat# ab97035). Brain slices were permeabilized and blocked with PBS containing 0.5% Triton X-100 and 5% normal goat serum (NGS) for 1h at room temperature. Slices were incubated overnight at 4 °C with primary antibodies, followed by 4 h incubation with secondary antibodies at room temperature. Slices were then mounted with Antifading Mounting medium (Abcam, cat# ab104139). Confocal images were acquired using a Leica TCS SP8 microscope set to the same laser intensity and acquisition parameters to compare the immunosignals in sections from Scr and Syt7 KD groups. For cell counting and colocalization, fluorescent protein-positive cells were automatically quantified with spot-detection algorithms in Imaris 9.5 (Oxford instruments).

Data analysis

Data were analyzed and presented using ClampFit (Molecular Devices), Igor Pro (Wavemetrics), MATLAB (Mathworks), and Prism (GraphPad). Data points and error bars represent the mean ± standard error of the mean. All statistical analyses were conducted using two-tailed comparisons. Sample sizes and statistical tests for each group are detailed in the figure legends.

Numerical integration of the simple refilling model

This model posits reversible docking of vesicles at finite release sites (Nmax) with forward (k1) and reverse rate constants (b1) (Hosoi et al., 2007; Neher and Sakaba, 2008). According to this scheme, dn/dt = k1 u - b1 n - pv n δ(t - tAP), where u is the number of unoccupied sites (= Nmax - n); tAP is the timing of AP firing; and δ is the Dirac delta function. The pv was assumed to be unity (Figure 3—figure supplement 2). The sum of basal k1 (= k1,b) and b1 was be estimated as 23/s from the dependence PPR on ISIs (Figure 3—figure supplement 2Ab). From the baseline occupancy of 0.3 and pocc = k1,b / (k1,b + b1), we could estimate baseline k1 and b1 as 6.9/s, and 16.1/s, respectively. Ca2+- and Syt7-dependent increase of k1 was modeled as k1 = k1,b + K1,max [Syt7: n Ca2+], where [Syt7: n Ca2+] represents the fraction of full Ca2+-bound Syt7. For deterministic simulations, we time-integrated a set of differential equations for each model using Euler methods with a time step of 1 ms. All calculations were performed on the platform of Matlab (R2022b, Mathworks, USA).

Material availability statement

All data generated or analyzed during this study are included in the manuscript and supporting files. Source data files are provided in this paper. Matlab codes for the analyses and simulation are presented in this paper.